
Glosses have three parts.

\exdisplay
\it
|\begingl|\par
preamble\par
interlinear gloss\par
free translation\par
|\endgl|
\xe

The preamble and free translation parts are optional.  The interlinear
gloss can be omitted if there is a preamble.\footnote{%
Contrary to first impressions, this actually makes sense in some contexts.}
The preamble has the form |\glpreamble| \dots\ |//|.  The free translation
has the form |\glft| \dots\ |//|.

The interlinear gloss has the form:

\exdisplay
\it
|\gla| \dots\ |//|\par
|\gl|levelname \dots |//|\par
|\gl|levelname \dots |//|\par
|\gl|levelname \dots |//|\par
|\gl|levelname \dots |//|\par
\xe
The |\gla| line is obligatory, provided of course that there is an
interlinear gloss, must come first and cannot be repeated.  The level
names pre, ft, a, b, and c are predefined.  The user can define other
level names (other than the predefined level names).  Other level
names in the interlinear gloss can be repeated as many times as the
user desires.  Only the a level is obligatory.

Levels b and c, and user defined levels, have two associated
parameters: |everygl|{\it levelname} and |abovegl|{\it
levelname\/}|skip|.  The value of |everygl|{\it levelname} is inserted
when the level is processed, allowing the user to control the font
used on that level.  The other parameter should be self-explanatory.
No skip is inserted above the a-level, but |everygla| is a parameter.
Additionally, there are parameters |everyglpreamble|,
|belowglpreambleskip|, |everyglft|, and |aboveglftskip|, which should
all be self-explanatory at this point.


\input pstricks-add
\ex
\def\TOP{\pnode(3.4in,1.5ex)}%
\def\BOT{\pnode(0,-.5ex)}%
\begingl[glwidth=3.2in]
\glpreamble \TOP{A1}Um-\"asudda' h\"am yan i taotao ni si Juan
ilek-\~na nu guahu mal\"agu' gui
asudd\"a'-\~na.\BOT{B1}//
\gla \TOP{A2}Um-\"asudda' h\"am yan [ i taotao [ {\it O\/} ni si Juan
ilek-\~na nu guahu [ mal\"agu' gui [ asudd\"a'-\~na {\it
t\/} ] ] ] ] @ .//
\glb agr-meet we with the person Op Comp the Juan say-agr Obl me
agr.want he {\it WH\/}[obl].meet-agr\BOT{B2}//
\glft \TOP{A3}``I met the person who Juan told me he wanted to
meet.''\BOT{B3}//
\endgl
\SpecialCoor
\psbrace[ref=lC](A1|B1)(A1){ preamble}
\psbrace[ref=lC](A1|B2)(A2){ interlinear gloss}
\psbrace[ref=lC](A1|B3)(A3){ free translation}
\xe


