
\vfil
\eject
\global\pageno=1
\section Introduction

Many of the needs of linguists who wish to produce
typographically attractive papers using \Tex\ or \LaTex\
are not specific to linguistic papers.  There are therefore many
macro packages which deal with tables of contents, references,
section headings, font selection, indexing, etc. But linguistics
does have some special typographic needs. I addressed two of
these with the macro packages {\sl PST-JTree\/} and {\sl
PST-ASR\/}, which typeset syntactic trees and autosegmental
representations. \ExPex\ addresses the main remaining special
\Tex\ need in linguistics: formatting examples, examples with
multiple parts, glosses, and the like, and referring to examples
and parts of examples. The name comes from the two central
macros, |\ex| and |\pex|, used to typeset examples and examples
with labeled parts.

{\sl PST-JTree\/} and {\sl PST-ASR\/} rely heavily on Hendri
Adriaens' \XKeyVal\ package, which has become the standard for
\PSTricks\ based macro packages.  Although \ExPex\ is not based
on \PSTricks\relax, it does handle parameterization the same
way that {\sl PST-JTree\/} and {\sl PST-ASR\/} do. When
{\sl expex.tex} is loaded, it immediately checks to see
whether {\sl xkeyval.tex} has already been loaded.  If not,
it does so.

The goal in writing a macro package for general use is to make it
simple to use if only simple things need to be done, but powerful
enough so that users who have complex needs can get those needs
satisfied if they are willing to deal with the complexities that
complex needs inevitably involve.  If you think there are simple
things that are not simple to do, or complex things that cannot
be done, please write to me at {\sl j.frampton@neu.edu}.  The
\ExPex\ macros have evolved over the last 15 or so years
and like anything which evolves, various features of the current
state may have more to do with history than with optimal design.
Please let me know about departures from optimal design.  Perhaps
the next version can be improved.

This User's Guide begins with four examples which demonstrate
\ExPex\ in action.  It serves as a ``A Quick Guide to {\sl
Expex\/}''.  Each page gives some code at the top, with the
product of this code below. Its main purpose is to give a sense
of how \ExPex\ works, so that curious readers have some
basis for determining whether they want to proceed with the
details.  It is possible to begin to use \ExPex\ solely on
the basis of the four demo pages and learn the more subtle
capabilities as needed.  A quick survey of the index and the table
of contents should give you some idea of what is available, if
you need it.

\subsection  Changes

The most recent previous version of \ExPex\ is version 4.1d. That
version  will still be available on CTAN, zipped as {\it
expex41d.zip}. Aside from bug fixes, the changes in this version are:
\count255=0
\def\Item{\smallskip\advance\count255 by 1 \item{\the\count255.}}

\Item{The gloss parameter |abovemoreglskip| has been deprecated (declared
obsolete and slated to be eliminated) for some time.  It has now been
removed from \ExPex.  The function it served is now served in a more
general way by |extraglskip|.}

\Item{An entirely new way of coding glosses, called the {\it nlevel
style\/} has been introduced.  See Section \getref{nlevelsec} for the
many details.  This style makes it much easier to keep track of words
and their associated glosses, particularly when long glosses need to
be typeset.}

\Item{A new way is introduced for including unglossed items (brackets,
for instance) among the items which are glossed.  The exceptional
gloss items `[' and `]' are now deprecated as exceptional gloss items.
Consequentially, the parameters |everybrack|, |glbrackwordsep|, and
|glbrackbracksep| have also been deprecated.  All these deprecated
items will be removed from \ExPex\ at some time in the future.}

\Item{Control over vertical spacing in glosses has been streamlined.
Two little used parameters |autoglskip| (which was rarely used outside
the default setting) and |gllineskip| have been eliminated. Their
functions have been taken over by |extraglskip| and strut insertion.
It is possible that this may create some problems with backwards
compatibility.  If there are problems, either use Version 4.1d or
contact me directly to see if there is an easy solution to the
problem.}

\Item{Glosses now break over pages, provided they are full width.}

\Item{Center alignment is now an option for glosses.  Glosses of words
can be centered under the word they gloss.}

\Item{\ExPex\ example numbering is now compatible with the \LaTex\
|\includeonly| mechanism.}

\subsection LaTex/Tex cooperation

\ExPex\ is designed to be used by either {\sl Plain Tex\/} or \LaTex\
users.  \LaTex\ users need to say |\usepackage{expex}| and \Tex\ users
|\input expex|.  All of the code for the examples in this
documentation should run equally well in either system, subject
to the notes below.

\subsubsection Note to \LaTex\ users

|\it| (now a deprecated \LaTex\ command) is used in a few places.
In case \ExPex\ detects that \LaTex\ is being used, it executes
|\let\it=\itshape|.

\subsubsection Note to \Tex\ users

Three macros are used in the examples in this documentation which
are defined in \LaTex\ but not in {\it Plain Tex}: |\footnotesize|,
|\sc|, and |\textsc|.  Assuming, for example, that text is set in
10pt computer modern, the following would suffice for all the
examples in this documentation.

\codedisplay
\font\eightrm=cmr8
\font\eightit=cmti8
\def\footnotesize{\eightrm \let\it=\eightit \baselineskip=9pt}
\font\tensc=cmcsc10
\let\sc=\tensc
\def\textsc#1{{\sc #1}}
|endcodedisplay

Most {\sl Plain Tex\/} users will have other fonts and much more
general size changing macros at their disposal.  The code above
is barely sufficient to handle the examples in this is
documentation, but it will do the job. For what it is worth, this
documentation was typeset using {\it Plain Tex}. |\twelvepoint| and
|\tenpoint| were defined modeled on pages 414--415 in the {\sl
TeXbook\/} and |\let\footnotesize=\tenpoint| was executed to
define |\footnotesize|.  The running text is \textdim{12 pt}.
|\sc| was defined by |\font\twelvesc=cmcsc10 scaled\magstep1|%
\footnote{cmcsc10 scaled was used rather than
cmcsc12 because Postscript fonts for cmcsc10 are more readily
available than those for cmcsc12.}
and |\let\sc=\twelvesc|%
\footnote{Small caps are used only in text size}%
.

\subsection Acknowledgements

Many participants in the {\sl Ling-Tex\/} discussion group have
contributed to the development of \ExPex\relax, either by posing good
questions, solving problems, or providing informed discussion of
desirable features.  In particular, I thank Stephen Anderson, Mario
Bisiada, Noah Constant, James Crippen, Alexis Dimitriadis, Claude
Dionne, Kevin Donnely, Antonio Fortin, Jeremy Hammond, Daniel Harbor,
Joshua Jensen, Don Killian, Joost Kremers, John Lyon, Alan Munn,
Christos Vlachos, and Natalie Weber.





