\documentclass[a4paper,10pt]{article}
\usepackage[utf8x]{inputenc}
\usepackage{tipa}
\usepackage{expex}
\usepackage[british]{babel}
\usepackage[nodayofweek]{datetime}
\usepackage[usenames,dvipsnames]{color}

%\renewcommand{\rmdefault}{phv}  % Uncomment to get sans-serif text.

%Thanks to Manas Tungare (http://manas.tungare.name/software/latex) for these settings.
\setlength{\paperwidth}{210mm}
\setlength{\paperheight}{297mm}

\setlength{\textwidth}{160mm}
\setlength{\textheight}{247mm}

\setlength{\evensidemargin}{1in}
\setlength{\oddsidemargin}{0in}
\setlength{\topmargin}{-0.5in}

%\lingset{everygla=\bf,everyglb=\it,everyglc=\rm}
%\definelingstyle{glabc}{glstyle=multilevel,everygl=,everygla=\bf,everyglb=\tenpoint,aboveglbskip=.3ex,everyglc=\tenpoint,aboveglcskip=-.3ex,everyglft=\it,aboveglftskip=1ex}
\definelingstyle{gergl}
{glstyle=wrap,
glspace=1em,
everygl=\openup0.5ex,
everyglword=\openup-0.7ex,
everygla=\bf,
everyglb=\it,
everyglc=\it,
aboveglcskip=0pt,
glhangindent=0pt}  % This gives a flush left margin; remove it if you want the normal hanging indent, but remember to retain the brace and remove the preceding comma.

% To get hanging indents that put the speaker out to one side, add this to the above, but remember  to retain the brace and remove the preceding comma.
%glhangindent=43pt

%opening
\title{stammers4.cha}
\author{}
\date{}

\begin{document}

\maketitle

% Change the default heading from ``Abstract'' - note that this command must be given after \begin{document}
\renewcommand{\abstractname}{Conversation details}

%\begin{abstract}
%\textbf{Participants}: Sofía (SOF) is an adult female, and Kevin (KEV) is an adult male.
%\textbf{Background}: SOF and KEV are neighbors who have lived in the same neighborhood for several years. They are sitting in the back porch of KEV’s house, which has a view out to a lake, and converse about plants and their neighborhood.
%\textbf{Duration}: 33 min 52 sec.
%\textbf{Date}: 2 March 2008.
%\textbf{Transcriber}: M. Carmen Parafita Couto, Prifysgol Bangor University
%\end{abstract}

\rule{\linewidth}{0.2mm}\\[0.5cm]

\ex
\begingl[lingstyle=gergl]
\glchat a pwy sy (y)n dod i weld y sioe ? //
\glsurface ELE:  a pwy sy yn dod i weld y sioe ?  //
\glauto \%aut  and{\scriptsize .CONJ} who{\scriptsize .PRON} which\_is{\scriptsize .V.3SP.PRES.REL} stative{\scriptsize .STAT} come{\scriptsize .V.INFIN} to{\scriptsize .PREP} see{\scriptsize .V.INFIN+SM} the{\scriptsize .DET.DEF} show{\scriptsize .N.F.SG}   //
\glmanual \%gls  and who be{\scriptsize .PRES.REL} PRT come{\scriptsize .NONFIN} to see{\scriptsize .NONFIN} DET show   //
\gleng and who's coming to see the show? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat rhieni . //
\glsurface ALN:  rhieni .  //
\glauto \%aut  parents{\scriptsize .N.M.PL}   //
\glmanual \%gls  parents   //
\gleng parents. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae [?] sioe gyhoeddus . //
\glsurface ALN:  mae sioe gyhoeddus .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} show{\scriptsize .N.F.SG} public{\scriptsize .ADJ+SM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} show public   //
\gleng it's a public show. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ o\_k@s:cym\&eng [=! quietly] . //
\glsurface ELE:  o\_k$^{C}_{E}$ .  //
\glauto \%aut  OK{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i meddwl fy(dd) Andrew@s:cym\&eng o\_gwmpas ar\_gyfer y sioe . //
\glsurface ALN:  dw i meddwl fydd Andrew$^{C}_{E}$ o\_gwmpas ar\_gyfer y sioe .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} think{\scriptsize .V.INFIN} be{\scriptsize .V.3S.FUT+SM} name around{\scriptsize .ADV} for{\scriptsize .PREP} the{\scriptsize .DET.DEF} show{\scriptsize .N.F.SG}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} think{\scriptsize .NONFIN} be{\scriptsize .3S.FUT} Andrew around for DET show   //
\gleng I think Andrew will be around for the show. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a fydd Anna@s:cym\&eng \# a [?] Bev@s:cym\&eng yn dod i\_fewn \#\# yn ystod y dydd a gyda (y)r nos dw i (y)n meddwl . //
\glsurface ALN:  a fydd Anna$^{C}_{E}$ a Bev$^{C}_{E}$ yn dod i\_fewn yn ystod y dydd a gyda yr nos dw i yn meddwl .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.FUT+SM} name and{\scriptsize .CONJ} name stative{\scriptsize .STAT} come{\scriptsize .V.INFIN} in{\scriptsize .ADV} in{\scriptsize .PREP} range{\scriptsize .N.F.SG} the{\scriptsize .DET.DEF} day{\scriptsize .N.M.SG} and{\scriptsize .CONJ} with{\scriptsize .PREP} the{\scriptsize .DET.DEF} night{\scriptsize .N.F.SG} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} stative{\scriptsize .STAT} think{\scriptsize .V.INFIN}   //
\glmanual \%gls  and be{\scriptsize .3S.FUT} Anna and Bev PRT come{\scriptsize .NONFIN} in in range DET day and with DET night be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT think{\scriptsize .NONFIN}   //
\gleng and Anna and Bev will be coming in during the day and in the evening I think. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng [=! quietly] . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat um@s:cym\&eng \# dw i (dd)im yn mynd i\_fewn . //
\glsurface ALN:  um$^{C}_{E}$ dw i ddim yn mynd i\_fewn .  //
\glauto \%aut  um{\scriptsize .IM} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} go{\scriptsize .V.INFIN} in{\scriptsize .ADV}   //
\glmanual \%gls  IM be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} NEG PRT go{\scriptsize .NONFIN} in   //
\gleng I'm not going in. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond maen nhw (y)n wneud o nos Fawrth a nos Fercher . //
\glsurface ALN:  ond maen nhw yn wneud o nos Fawrth a nos Fercher .  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.3P.PRES.SPOKEN} they{\scriptsize .PRON.3P} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM} from{\scriptsize .PREP} night{\scriptsize .N.F.SG} March{\scriptsize .N.M.SG+SM[or]Mars.N.M.SG+SM} and{\scriptsize .CONJ} night{\scriptsize .N.F.SG} Mercury{\scriptsize .N.F.SG+SM}   //
\glmanual \%gls  but be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL} PRT do{\scriptsize .NONFIN} PRON{\scriptsize .3SM} night Tuesday and night Wednesday   //
\gleng but they're doing it Tuesday night and Wednesday night. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng right@s:cym\&eng ! //
\glsurface ELE:  oh$^{C}_{E}$ right$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM} right{\scriptsize .ADJ}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$dw i$>$ [/] dw i heb gynnig yn hun . //
\glsurface ALN:  dw i dw i heb gynnig yn hun .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} without{\scriptsize .PREP} offer{\scriptsize .V.INFIN+SM} in{\scriptsize .PREP} self{\scriptsize .PRON.SG}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} without offer{\scriptsize .NONFIN} POSS{\scriptsize .1S} self   //
\gleng I've not offered myself. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i heb fod yn ymwneud â (y)r sioe a deud y gwir . //
\glsurface ALN:  dw i heb fod yn ymwneud â yr sioe a deud y gwir .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} without{\scriptsize .PREP} be{\scriptsize .V.INFIN+SM} stative{\scriptsize .STAT} concern{\scriptsize .V.INFIN} with{\scriptsize .PREP} the{\scriptsize .DET.DEF} show{\scriptsize .N.F.SG} and{\scriptsize .CONJ} say{\scriptsize .V.INFIN.SPOKEN} the{\scriptsize .DET.DEF} true{\scriptsize .ADJ}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} without be{\scriptsize .NONFIN} PRT appertain{\scriptsize .NONFIN} with DET show PRT say{\scriptsize .NONFIN} DET truth   //
\gleng I've not been involved with the show, to tell you the truth. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$dw i (we)di$>$ [/] dw i (we)di wneud ychydig bach o waith canu efo nhw . //
\glsurface ALN:  dw i wedi dw i wedi wneud ychydig bach o waith canu efo nhw .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} after{\scriptsize .PREP} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} after{\scriptsize .PREP} make{\scriptsize .V.INFIN+SM} a\_little{\scriptsize .QUAN} small{\scriptsize .ADJ} of{\scriptsize .PREP} work{\scriptsize .N.M.SG+SM} sing{\scriptsize .V.INFIN} with{\scriptsize .PREP} they{\scriptsize .PRON.3P}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT{\scriptsize .PAST} be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT{\scriptsize .PAST} do{\scriptsize .NONFIN} a\_bit little of work singing with PRON{\scriptsize .3PL}   //
\gleng I've done a little bit of singing work with them. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na fedri di mynd i bob\_dim na fedri ? //
\glsurface ELE:  na fedri di mynd i bob\_dim na fedri ?  //
\glauto \%aut  neg{\scriptsize .PRT} be{\scriptsize .V.2S.PRES+SM} you{\scriptsize .PRON.2S+SM} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} everything{\scriptsize .PRON} neg{\scriptsize .PRT[or]who\_not.PRON.REL.NEG[or](n)or.CONJ[or]than.CONJ} be{\scriptsize .V.2S.PRES+SM}   //
\glmanual \%gls  NEG can{\scriptsize .2S.NONPAST} PRON{\scriptsize .2S} go{\scriptsize .NONFIN} to everything NEG can{\scriptsize .2S.NONPAST}   //
\gleng you can't go to everything, can you? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng na . //
\glsurface ALN:  oh$^{C}_{E}$ na .  //
\glauto \%aut  oh{\scriptsize .IM} neg{\scriptsize .PRT}   //
\glmanual \%gls  IM no   //
\gleng oh, no. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a dw i (we)di deud +. . . //
\glsurface ALN:  a dw i wedi deud .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} after{\scriptsize .PREP} say{\scriptsize .V.INFIN.SPOKEN}   //
\glmanual \%gls  and be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT{\scriptsize .PAST} say{\scriptsize .NONFIN}   //
\gleng and I've said \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat timod $<$wnes i$>$ [/] wnes i ofyn i Anna@s:cym\&eng heddiw +"/ . //
\glsurface ALN:  timod wnes i wnes i ofyn i Anna$^{C}_{E}$ heddiw .  //
\glauto \%aut  know{\scriptsize .V.2S.PRES.SPOKEN} do{\scriptsize .V.1S.PAST.SPOKEN+SM} I{\scriptsize .PRON.1S} do{\scriptsize .V.1S.PAST.SPOKEN+SM} I{\scriptsize .PRON.1S} ask{\scriptsize .V.INFIN+SM} to{\scriptsize .PREP} name today{\scriptsize .ADV}   //
\glmanual \%gls  know{\scriptsize .2S} do{\scriptsize .1S.PAST} PRON{\scriptsize .1S} do{\scriptsize .1S.PAST} PRON{\scriptsize .1S} ask{\scriptsize .NONFIN} to Anna today   //
\gleng you know, I asked Anna today: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" wyt ti angen fi ar\_gyfer y sioe ? //
\glsurface ALN:  wyt ti angen fi ar\_gyfer y sioe ?  //
\glauto \%aut  be{\scriptsize .V.2S.PRES} you{\scriptsize .PRON.2S} need{\scriptsize .N.M.SG} I{\scriptsize .PRON.1S+SM} for{\scriptsize .PREP} the{\scriptsize .DET.DEF} show{\scriptsize .N.F.SG}   //
\glmanual \%gls  be{\scriptsize .2S.PRES} PRON{\scriptsize .2S} need PRON{\scriptsize .1S} for DET show   //
\gleng "do you need me for the show?" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" well@s:cym\&eng na . //
\glsurface ALN:  well$^{C}_{E}$ na .  //
\glauto \%aut  well{\scriptsize .ADV} neg{\scriptsize .PRT}   //
\glmanual \%gls  well no   //
\gleng "well, no" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat meddai hi +" . //
\glsurface ALN:  meddai hi .  //
\glauto \%aut  say{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  say{\scriptsize .3S.PAST} PRON{\scriptsize .3SF}   //
\gleng she said. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng . //
\glsurface ELE:  yeah$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat timod (dy)dy hi ddim \# angen ei pherswadio bod angen +/ . //
\glsurface ALN:  timod dydy hi ddim angen ei pherswadio bod angen .  //
\glauto \%aut  know{\scriptsize .V.2S.PRES.SPOKEN} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} she{\scriptsize .PRON.F.3S} not{\scriptsize .ADV+SM} need{\scriptsize .N.M.SG} her{\scriptsize .ADJ.POSS} persuade{\scriptsize .V.INFIN+AM} be{\scriptsize .V.INFIN} need{\scriptsize .N.M.SG}   //
\glmanual \%gls  know{\scriptsize .2S} be{\scriptsize .3S.PRES.NEG} PRON{\scriptsize .3SF} NEG need POSS{\scriptsize .3SF} persuade{\scriptsize .NONFIN} be{\scriptsize .NONFIN} need   //
\gleng you know, she doesn't need to be persuaded that it needs \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na na ! //
\glsurface ELE:  na na !  //
\glauto \%aut  neg{\scriptsize .PRT} neg{\scriptsize .PRT}   //
\glmanual \%gls  no no   //
\gleng oh, no! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, chwech o bobl pan ellith [?] dau wneud o . //
\glsurface ALN:  chwech o bobl pan ellith dau wneud o .  //
\glauto \%aut  six{\scriptsize .NUM} of{\scriptsize .PREP} people{\scriptsize .N.F.SG+SM} when{\scriptsize .CONJ} be\_able{\scriptsize .V.3S.FUT.SPOKEN+SM} two{\scriptsize .NUM.M} make{\scriptsize .V.INFIN+SM} he{\scriptsize .PRON.M.3S.SPOKEN}   //
\glmanual \%gls  six of people when can{\scriptsize .3S.CONDIT} two{\scriptsize .M} do{\scriptsize .NONFIN} PRON{\scriptsize .3SM}   //
\gleng  \dots six people when two could do it. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na na . //
\glsurface ELE:  na na .  //
\glauto \%aut  neg{\scriptsize .PRT} neg{\scriptsize .PRT}   //
\glmanual \%gls  no no   //
\gleng no. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (dy)dy (ddi)m yn licio siarad efo rhieni ar y dechrau . //
\glsurface ALN:  dydy ddim yn licio siarad efo rhieni ar y dechrau .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} like{\scriptsize .V.INFIN} talk{\scriptsize .V.INFIN} with{\scriptsize .PREP} parents{\scriptsize .N.M.PL} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} beginning{\scriptsize .N.M.SG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES.NEG} NEG PRT like{\scriptsize .NONFIN} talk{\scriptsize .NONFIN} with parents on DET start   //
\gleng she doesn't like talking to parents at the start. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond wneith Andrew@s:cym\&eng wneud hynna so@s:cym\&eng mae hwnna (y)n iawn . //
\glsurface ALN:  ond wneith Andrew$^{C}_{E}$ wneud hynna so$^{C}_{E}$ mae hwnna yn iawn .  //
\glauto \%aut  but{\scriptsize .CONJ} do{\scriptsize .V.3S.FUT+SM} name make{\scriptsize .V.INFIN+SM} that{\scriptsize .PRON.SP.SPOKEN} so{\scriptsize .ADV} be{\scriptsize .V.3S.PRES} that{\scriptsize .PRON.M.SG.SPOKEN} stative{\scriptsize .STAT} OK{\scriptsize .ADV}   //
\glmanual \%gls  but do{\scriptsize .3S.NONPAST} Andrew do{\scriptsize .NONFIN} that so be{\scriptsize .3S.PRES} that PRT alright   //
\gleng but Andrew will that, so that's ok. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$dyn nhw$>$ [?] (dd)im yn broblem@s:cym\&eng felly . //
\glsurface ALN:  dyn nhw ddim yn broblem$^{C}_{E}$ felly .  //
\glauto \%aut  man{\scriptsize .N.M.SG[or]draw.V.3S.PRES+SM} they{\scriptsize .PRON.3P} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} problem{\scriptsize .N.SG+SM} so{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3PL.PRES.NEG} PRON{\scriptsize .3PL} NEG PRT problem thus   //
\gleng so they're no problem. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na . //
\glsurface ELE:  na .  //
\glauto \%aut  neg{\scriptsize .PRT}   //
\glmanual \%gls  no   //
\gleng no. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (y)dy o am sgwennu um@s:cym\&eng \# sioe ? //
\glsurface ELE:  ydy o am sgwennu um$^{C}_{E}$ sioe ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} for{\scriptsize .PREP} write{\scriptsize .V.INFIN} um{\scriptsize .IM} show{\scriptsize .N.F.SG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} for write{\scriptsize .NONFIN} IM show   //
\gleng is he going to write a show? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=deep\_breath mae (y)n sôn am sgwennu sioe \# um@s:cym\&eng +. . . //
\glsurface ALN:  mae yn sôn am sgwennu sioe um$^{C}_{E}$ .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} mention{\scriptsize .V.INFIN} for{\scriptsize .PREP} write{\scriptsize .V.INFIN} show{\scriptsize .N.F.SG} um{\scriptsize .IM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRT talk{\scriptsize .NONFIN} about write{\scriptsize .NONFIN} show IM   //
\gleng he's talking about writing a show, um \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&o mae [/] mae [//] yndy . //
\glsurface ALN:  mae mae yndy .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} be{\scriptsize .V.3S.PRES} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES}   //
\gleng yes //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae o mynd i sgwennu sioe . //
\glsurface ALN:  mae o mynd i sgwennu sioe .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} write{\scriptsize .V.INFIN} show{\scriptsize .N.F.SG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} go{\scriptsize .NONFIN} to write{\scriptsize .NONFIN} show   //
\gleng he's going to write a show. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat pryd fydd o (y)n barod +/ . //
\glsurface ALN:  pryd fydd o yn barod .  //
\glauto \%aut  when{\scriptsize .INT} be{\scriptsize .V.3S.FUT+SM} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} ready{\scriptsize .ADJ+SM}   //
\glmanual \%gls  when be{\scriptsize .3S.FUT} PRON{\scriptsize .3SM} PRT ready   //
\gleng when will it be ready \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat erbyn Dolig (fe)lly neu erbyn (y)r ha ? //
\glsurface ELE:  erbyn Dolig felly neu erbyn yr ha ?  //
\glauto \%aut  by{\scriptsize .PREP} name so{\scriptsize .ADV} or{\scriptsize .CONJ} by{\scriptsize .PREP} the{\scriptsize .DET.DEF} and{\scriptsize .CONJ+H}   //
\glmanual \%gls  by Christmas thus or by summer    //
\gleng by Christmas or by summer? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat erbyn (y)r haf . //
\glsurface ALN:  erbyn yr haf .  //
\glauto \%aut  by{\scriptsize .PREP} the{\scriptsize .DET.DEF} summer{\scriptsize .N.M.SG}   //
\glmanual \%gls  by DET summer   //
\gleng by summer. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng iawn (fe)lly . //
\glsurface ELE:  oh$^{C}_{E}$ iawn felly .  //
\glauto \%aut  oh{\scriptsize .IM} OK{\scriptsize .ADV} so{\scriptsize .ADV}   //
\glmanual \%gls  IM alright thus   //
\gleng oh, fine, then. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae [?] (y)n dipyn o amser . //
\glsurface ELE:  mae yn dipyn o amser .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} little{\scriptsize .bit.N.M.SG+SM} of{\scriptsize .PREP} time{\scriptsize .N.M.SG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRT a\_bit of time   //
\gleng there's a fair bit of time. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond mae (we)di gofyn i Brian@s:cym\&eng $<$wneud yr$>$ [//] \# sgwennu cerddoriaeth ar ei gyfer o . //
\glsurface ALN:  ond mae wedi gofyn i Brian$^{C}_{E}$ wneud yr sgwennu cerddoriaeth ar ei gyfer o .  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} after{\scriptsize .PREP} ask{\scriptsize .V.INFIN} to{\scriptsize .PREP} name make{\scriptsize .V.INFIN+SM} the{\scriptsize .DET.DEF} write{\scriptsize .V.INFIN} music{\scriptsize .N.F.SG} on{\scriptsize .PREP} his{\scriptsize .ADJ.POSS} direction{\scriptsize .N.M.SG+SM} he{\scriptsize .PRON.M.3S.SPOKEN}   //
\glmanual \%gls  but be{\scriptsize .3S.PRES} PRT{\scriptsize .PAST} ask{\scriptsize .NONFIN} to Brian do{\scriptsize .NONFIN} DET write{\scriptsize .NONFIN} music for POSS{\scriptsize .3SM} sake PRON{\scriptsize .3SM}   //
\gleng but he's asked Brian to do the. . . to write the music for it. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmhm@s:cym\&eng . //
\glsurface ELE:  mmhm$^{C}_{E}$ .  //
\glauto \%aut  mmhm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a $<$mae Brian@s:cym\&eng (y)n$>$ [/] \#\# mae Brian@s:cym\&eng (y)n gweithio (y)n ofnadwy o galed . //
\glsurface ALN:  a mae Brian$^{C}_{E}$ yn mae Brian$^{C}_{E}$ yn gweithio yn ofnadwy o galed .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} name in{\scriptsize .PREP} be{\scriptsize .V.3S.PRES} name stative{\scriptsize .STAT} work{\scriptsize .V.INFIN} stative{\scriptsize .STAT} terrible{\scriptsize .ADJ} of{\scriptsize .PREP} hard{\scriptsize .ADJ+SM}   //
\glmanual \%gls  and be{\scriptsize .3S.PRES} Brian PRT be{\scriptsize .3S.PRES} Brian PRT work{\scriptsize .NONFIN} PRT awful of hard   //
\gleng and Brian works incredibly hard. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae baratoi o (y)n ofnadwy o drylwyr . //
\glsurface ALN:  mae baratoi o yn ofnadwy o drylwyr .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} prepare{\scriptsize .V.INFIN+SM} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} terrible{\scriptsize .ADJ} of{\scriptsize .PREP} thorough{\scriptsize .ADJ+SM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} prepare{\scriptsize .NONFIN} PRON{\scriptsize .3SM} PRT awful of thorough   //
\gleng he prepares incredibly thoroughly. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng ella fydd o (y)n siwtio fo felly \# bod o (y)n hwyr yn cyrraedd . //
\glsurface ELE:  oh$^{C}_{E}$ ella fydd o yn siwtio fo felly bod o yn hwyr yn cyrraedd .  //
\glauto \%aut  oh{\scriptsize .IM} maybe{\scriptsize .ADV.SPOKEN} be{\scriptsize .V.3S.FUT+SM} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} suit{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S} so{\scriptsize .ADV} be{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} late{\scriptsize .ADJ} stative{\scriptsize .STAT} arrive{\scriptsize .V.INFIN}   //
\glmanual \%gls  IM maybe be{\scriptsize .3S.FUT} PRON{\scriptsize .3SM} PRT suit{\scriptsize .NONFIN} PRON{\scriptsize .3SM} thus be{\scriptsize .NONFIN} PRON{\scriptsize .3SM} late PRT arrive{\scriptsize .NONFIN}    //
\gleng oh, maybe that will suit him, then, that it's late arriving. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dy(dy) [/] $<$(dy)dy o (ddi)m$>$ [/] $<$(dy)dy o (ddi)m$>$ [//] (dy)dy o ddim yn ryw hapus o\_gwbl \# bod y peth yn hwyr yn cyrraedd de ? //
\glsurface ALN:  dydy dydy o ddim dydy o ddim dydy o ddim yn ryw hapus o\_gwbl bod y peth yn hwyr yn cyrraedd de ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} he{\scriptsize .PRON.M.3S.SPOKEN} not{\scriptsize .ADV+SM} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} he{\scriptsize .PRON.M.3S.SPOKEN} not{\scriptsize .ADV+SM} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} he{\scriptsize .PRON.M.3S.SPOKEN} not{\scriptsize .ADV+SM} in{\scriptsize .PREP} some{\scriptsize .PREQ+SM} happy{\scriptsize .ADJ} at\_all{\scriptsize .ADV} be{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} thing{\scriptsize .N.M.SG} stative{\scriptsize .STAT} late{\scriptsize .ADJ} stative{\scriptsize .STAT} arrive{\scriptsize .V.INFIN} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES.NEG} be{\scriptsize .3S.PRES.NEG} PRON{\scriptsize .3SM} NEG be{\scriptsize .3S.PRES.NEG} PRON{\scriptsize .3SM} NEG be{\scriptsize .3S.PRES.NEG} PRON{\scriptsize .3SM} NEG PRT some happy at\_all be{\scriptsize .NONFIN} DET thing PRT late PRT arrive{\scriptsize .NONFIN} TAG   //
\gleng he is not at all happy that it's late coming, right. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae o (y)n licio \# gwybod \# mis a hanner o flaen llaw be mae o (y)n wneud . //
\glsurface ALN:  mae o yn licio gwybod mis a hanner o flaen llaw be mae o yn wneud .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} like{\scriptsize .V.INFIN} know{\scriptsize .V.INFIN} month{\scriptsize .N.M.SG} and{\scriptsize .CONJ} half{\scriptsize .N.M.SG} of{\scriptsize .PREP} front{\scriptsize .N.M.SG+SM} hand{\scriptsize .N.F.SG} what{\scriptsize .INT} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT like{\scriptsize .NONFIN} know{\scriptsize .NONFIN} month and half of front hand what be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT do{\scriptsize .NONFIN}   //
\gleng he likes to know a month and a half ahead what he's doing. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ ++ paratoi . //
\glsurface ELE:  paratoi .  //
\glauto \%aut  prepare{\scriptsize .V.INFIN}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wnest di sôn $<$(wr)tho fo$>$ [?] heddiw ella bod (y)na dŷ yn Bethel@s:cym\&eng ? //
\glsurface ELE:  wnest di sôn wrtho fo heddiw ella bod yna dŷ yn Bethel$^{C}_{E}$ ?  //
\glauto \%aut  do{\scriptsize .V.2S.PAST.SPOKEN+SM} you{\scriptsize .PRON.2S+SM} mention{\scriptsize .V.INFIN} to\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S} today{\scriptsize .ADV} maybe{\scriptsize .ADV.SPOKEN} be{\scriptsize .V.INFIN} there{\scriptsize .ADV} house{\scriptsize .N.M.SG+SM} in{\scriptsize .PREP} name   //
\glmanual \%gls  do{\scriptsize .2S.PAST} PRON{\scriptsize .2S} mention{\scriptsize .NONFIN} to{\scriptsize .3SM} PRON{\scriptsize .3SM} today perhaps be{\scriptsize .NONFIN} there house in Bethel   //
\gleng did you tell him today perhaps there's a house in Bethel? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat do . //
\glsurface ALN:  do .  //
\glauto \%aut  yes{\scriptsize .ADV.PAST}   //
\glmanual \%gls  yes   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat do . //
\glsurface ALN:  do .  //
\glauto \%aut  yes{\scriptsize .ADV.PAST}   //
\glmanual \%gls  yes   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd o (y)n deud bod o rhy bell allan . //
\glsurface ALN:  oedd o yn deud bod o rhy bell allan .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN} be{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} too{\scriptsize .ADJ} far{\scriptsize .ADJ+SM} out{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} PRT say{\scriptsize .NONFIN} be{\scriptsize .NONFIN} PRON{\scriptsize .3SM} too far out   //
\gleng he was saying it's too far out. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat rhy bell ffordd (y)na . //
\glsurface ALN:  rhy bell ffordd yna .  //
\glauto \%aut  too{\scriptsize .ADJ} far{\scriptsize .ADJ+SM} road{\scriptsize .N.F.SG} there{\scriptsize .ADV}   //
\glmanual \%gls  too far way there   //
\gleng too far that way. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=dental\_click oh@s:cym\&eng [=! sighs] . //
\glsurface ELE:  oh$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae isio tŷ \#\# un\_ai ym Mangor@s:cym\&eng neu i (y)r cyfeiriad yna . //
\glsurface ALN:  mae isio tŷ un\_ai ym Mangor$^{C}_{E}$ neu i yr cyfeiriad yna .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} want{\scriptsize .N.M.SG} house{\scriptsize .N.M.SG} either{\scriptsize .ADV} in{\scriptsize .PREP} name or{\scriptsize .CONJ} to{\scriptsize .PREP} the{\scriptsize .DET.DEF} direction{\scriptsize .N.M.SG} there{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} want house either in Bangor or to DET direction there   //
\gleng he wants a house either in Bangor or in that direction. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat i ochrau \# Aber@s:cym\&eng +/ . //
\glsurface ALN:  i ochrau Aber$^{C}_{E}$ .  //
\glauto \%aut  to{\scriptsize .PREP} sides{\scriptsize .N.F.PL} name   //
\glmanual \%gls  to sides Aber   //
\gleng around Aber \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat Llandudno@s:cym\&eng [?] . //
\glsurface ELE:  Llandudno$^{C}_{E}$ .  //
\glauto \%aut  name   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng oes pam ? //
\glsurface ELE:  oh$^{C}_{E}$ oes pam ?  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES.INDEF[or]age.N.F.SG} why?.ADV   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} why   //
\gleng oh, yes, why? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, Llanfairfechan@s:cym\&eng . //
\glsurface ALN:  Llanfairfechan$^{C}_{E}$ .  //
\glauto \%aut  name   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well@s:cym\&eng \# mae hi yn gobeithio codi gwaith +/ . //
\glsurface ALN:  well$^{C}_{E}$ mae hi yn gobeithio codi gwaith .  //
\glauto \%aut  well{\scriptsize .ADV} be{\scriptsize .V.3S.PRES} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} hope{\scriptsize .V.INFIN} lift{\scriptsize .V.INFIN} work{\scriptsize .N.M.SG}   //
\glmanual \%gls  well be{\scriptsize .3S.PRES} PRON{\scriptsize .3SF} PRT hope{\scriptsize .NONFIN} raise{\scriptsize .NONFIN} work   //
\gleng well she's hoping to get work \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mwy am Gonwy@s:cym\&eng ffordd (y)na ? //
\glsurface ELE:  mwy am Gonwy$^{C}_{E}$ ffordd yna ?  //
\glauto \%aut  more{\scriptsize .ADJ.COMP} for{\scriptsize .PREP} name road{\scriptsize .N.F.SG} there{\scriptsize .ADV}   //
\glmanual \%gls  more about Conwy way there   //
\gleng more around Conwy way? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, mwy am ochrau Conwy@s:cym\&eng . //
\glsurface ALN:  mwy am ochrau Conwy$^{C}_{E}$ .  //
\glauto \%aut  more{\scriptsize .ADJ.COMP} for{\scriptsize .PREP} sides{\scriptsize .N.F.PL} name   //
\glmanual \%gls  more about sides Conwy   //
\gleng more around about Conwy. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=dental\_click ah@s:cym\&eng . //
\glsurface ELE:  ah$^{C}_{E}$ .  //
\glauto \%aut  ah{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dyna (y)dy ei bwriad hi . //
\glsurface ALN:  dyna ydy ei bwriad hi .  //
\glauto \%aut  that{\scriptsize .is.ADV} be{\scriptsize .V.3S.PRES} her{\scriptsize .ADJ.POSS} intention{\scriptsize .N.M.SG} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  there be{\scriptsize .3S.PRES} POSS{\scriptsize .3SF} intention PRON{\scriptsize .3SF}   //
\gleng that's her plan. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat so@s:cym\&eng wnaeth o ddeud diolch ["] am +. . . //
\glsurface ALN:  so$^{C}_{E}$ wnaeth o ddeud diolch am .  //
\glauto \%aut  so{\scriptsize .ADV} do{\scriptsize .V.3S.PAST+SM} of{\scriptsize .PREP} say{\scriptsize .V.INFIN.SPOKEN+SM} thank{\scriptsize .V.INFIN} for{\scriptsize .PREP}   //
\glmanual \%gls  so do{\scriptsize .3S.PAST} PRON{\scriptsize .3SM} say{\scriptsize .NONFIN} thank\_you for   //
\gleng so he said thank you for \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae just@s:cym\&eng yn y lle anghywir . //
\glsurface ALN:  mae just$^{C}_{E}$ yn y lle anghywir .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} just{\scriptsize .ADV[or]just.ADJ} in{\scriptsize .PREP} the{\scriptsize .DET.DEF} place{\scriptsize .N.M.SG} incorrect{\scriptsize .ADJ}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} just in DET place incorrect   //
\gleng it's just in the wrong place. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng xxx +// . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be (y)dy um@s:cym\&eng +/ . //
\glsurface ELE:  be ydy um$^{C}_{E}$ .  //
\glauto \%aut  what{\scriptsize .INT} be{\scriptsize .V.3S.PRES} um{\scriptsize .IM}   //
\glmanual \%gls  what be{\scriptsize .3S.PRES} IM   //
\gleng what's \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat xxx phone@s:cym\&eng cario [?] fi (y)n +. . . //
\glsurface ALN:  phone$^{C}_{E}$ cario fi yn .  //
\glauto \%aut  phone{\scriptsize .N.SG} carry{\scriptsize .V.INFIN} I{\scriptsize .PRON.1S+SM} yn{\scriptsize .PRT}   //
\glmanual \%gls  phone carrying PRON{\scriptsize .1S} PRT   //
\gleng xxx my mobile phone \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng right@s:cym\&eng be dan ni (y)n recordio fan (a)cw ? //
\glsurface ELE:  oh$^{C}_{E}$ right$^{C}_{E}$ be dan ni yn recordio fan acw ?  //
\glauto \%aut  oh{\scriptsize .IM} right{\scriptsize .ADJ} what{\scriptsize .INT} be{\scriptsize .V.1P.PRES.SPOKEN} we{\scriptsize .PRON.1P} stative{\scriptsize .STAT} record{\scriptsize .V.INFIN} place{\scriptsize .N.MF.SG+SM} over{\scriptsize .there.ADV}   //
\glmanual \%gls  IM right what be{\scriptsize .1PL.PRES} PRON{\scriptsize .1PL} PRT record{\scriptsize .NONFIN} place there   //
\gleng oh, right, what are we recording over there? [or, "what, are we recording there?"] //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hwnna (y)n troi . //
\glsurface ELE:  hwnna yn troi .  //
\glauto \%aut  that{\scriptsize .PRON.M.SG.SPOKEN} stative{\scriptsize .STAT} turn{\scriptsize .V.INFIN}   //
\glmanual \%gls  that PRT turn{\scriptsize .NONFIN}   //
\gleng that's turning. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat u:m@s:cym\&eng \# dw (ddi)m yn gwybod . //
\glsurface ALN:  um$^{C}_{E}$ dw ddim yn gwybod .  //
\glauto \%aut  um{\scriptsize .IM} be{\scriptsize .V.1S.PRES.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN}   //
\glmanual \%gls  IM be{\scriptsize .1S.PRES} NEG PRT know{\scriptsize .NONFIN}   //
\gleng um, don't know. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat Charlie@s:cym\&eng sy recordio rywbeth debyg \&=sniff . //
\glsurface ALN:  Charlie$^{C}_{E}$ sy recordio rywbeth debyg .  //
\glauto \%aut  name which\_is{\scriptsize .V.3SP.PRES.REL} record{\scriptsize .V.INFIN} something{\scriptsize .N.M.SG+SM} similar{\scriptsize .ADJ+SM}   //
\glmanual \%gls  Charlie be{\scriptsize .PRES.REL} record{\scriptsize .NONFIN} something likely   //
\gleng Charlie's recording something, probably. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$awn ni$>$ [?] sbïo wedyn . //
\glsurface ALN:  awn ni sbïo wedyn .  //
\glauto \%aut  go{\scriptsize .V.1P.PRES} we{\scriptsize .PRON.1P} look{\scriptsize .V.INFIN} afterwards{\scriptsize .ADV}   //
\glmanual \%gls  go{\scriptsize .1PL.NONPAST} PRON{\scriptsize .1PL} look{\scriptsize .NONFIN} after   //
\gleng we'll have a look afterwards. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng ! //
\glsurface ALN:  hmm$^{C}_{E}$ !  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$yeah@s:cym\&eng o'n i$>$ [?] (y)n deud efo (y)r piano@s:cym\&eng gynnau oedd um@s:cym\&eng +. . . //
\glsurface ELE:  yeah$^{C}_{E}$ o'n i yn deud efo yr piano$^{C}_{E}$ gynnau oedd um$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV} unk to{\scriptsize .PREP} stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN} with{\scriptsize .PREP} the{\scriptsize .DET.DEF} piano{\scriptsize .N.SG} light{\scriptsize .V.INFIN+SM} be{\scriptsize .V.3S.IMPERF} um{\scriptsize .IM}   //
\glmanual \%gls  yeah be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} PRT say{\scriptsize .NONFIN} with DET piano earlier be{\scriptsize .3S.IMP} IM   //
\gleng yeah, I was saying with the piano earlier, um \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o'n i angen Charlie@s:cym\&eng benderfynu be oedd o (y)n wneud . //
\glsurface ELE:  o'n i angen Charlie$^{C}_{E}$ benderfynu be oedd o yn wneud .  //
\glauto \%aut  unk to{\scriptsize .PREP} need{\scriptsize .N.M.SG} name decide{\scriptsize .V.INFIN+SM} what{\scriptsize .INT} be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} need Charlie decide{\scriptsize .NONFIN} what be{\scriptsize .3S.IMP} PRON{\scriptsize .3S} PRT do{\scriptsize .NONFIN}   //
\gleng I needed Charlie to decide what he was doing. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a mae (we)di penderfynu bod am wneud y darnau \# sy ganddo fo de ? //
\glsurface ELE:  a mae wedi penderfynu bod am wneud y darnau sy ganddo fo de ?  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} after{\scriptsize .PREP} decide{\scriptsize .V.INFIN} be{\scriptsize .V.INFIN} for{\scriptsize .PREP} make{\scriptsize .V.INFIN+SM} the{\scriptsize .DET.DEF} pieces{\scriptsize .N.M.PL} which\_is{\scriptsize .V.3SP.PRES.REL} with\_him{\scriptsize .PREP+PRON.M.3S[or]with\_him.PREP+PRON+SM} he{\scriptsize .PRON.M.3S} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  and be{\scriptsize .3S.PRES} PRT{\scriptsize .PAST} decide{\scriptsize .NONFIN} be{\scriptsize .NONFIN} for do{\scriptsize .NONFIN} DET pieces be{\scriptsize .PRES.REL} with{\scriptsize .3SM} PRON{\scriptsize .3SM} TAG   //
\gleng and he's decided he's going to do the pieces he's got, right? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie . //
\glsurface ALN:  ie .  //
\glauto \%aut  yes{\scriptsize .ADV}   //
\glmanual \%gls  yes   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a: bod ni angen ymarfer [?] deg munud bob dydd . //
\glsurface ELE:  a bod ni angen ymarfer deg munud bob dydd .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.INFIN} we{\scriptsize .PRON.1P} need{\scriptsize .N.M.SG} exercise{\scriptsize .N.F.SG} ten{\scriptsize .NUM[or]fair.ADJ+SM} minute{\scriptsize .N.M.SG} each{\scriptsize .PREQ+SM} day{\scriptsize .N.M.SG}   //
\glmanual \%gls  and be{\scriptsize .NONFIN} PRON{\scriptsize .1PL} need practise{\scriptsize .NONFIN} ten minute every day   //
\gleng and that we need to practise ten minutes every day. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a ei er@s:cym\&eng \#\# benderfyniad o neu ateb o erbyn diwedd oedd deud +"/ . //
\glsurface ELE:  a ei er$^{C}_{E}$ benderfyniad o neu ateb o erbyn diwedd oedd deud .  //
\glauto \%aut  and{\scriptsize .CONJ} his{\scriptsize .ADJ.POSS} er{\scriptsize .IM} decision{\scriptsize .N.M.SG+SM} he{\scriptsize .PRON.M.3S.SPOKEN} or{\scriptsize .CONJ} answer{\scriptsize .N.M.SG[or]answer.V.3S.PRES[or]answer.V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} by{\scriptsize .PREP} end{\scriptsize .N.M.SG} be{\scriptsize .V.3S.IMPERF} say{\scriptsize .V.INFIN.SPOKEN}   //
\glmanual \%gls  and POSS{\scriptsize .3SM} IM decision PRON{\scriptsize .3SM} or answer PRON{\scriptsize .3SM} by end be{\scriptsize .3S.IMP} say{\scriptsize .NONFIN}   //
\gleng and his decision, or his answer in the end was to say: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" dw (ddi)m yn gwybod os fyddwch chi (y)n licio hyn Mam@s:cym\&eng ond +. . . //
\glsurface ELE:  dw ddim yn gwybod os fyddwch chi yn licio hyn Mam$^{C}_{E}$ ond .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN} if{\scriptsize .CONJ} be{\scriptsize .V.2P.FUT+SM} you{\scriptsize .PRON.2P} stative{\scriptsize .STAT} like{\scriptsize .V.INFIN} this{\scriptsize .PRON.SP} name but{\scriptsize .CONJ}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} NEG PRT know{\scriptsize .NONFIN} if be{\scriptsize .2PL.FUT} PRON{\scriptsize .2PL} PRT like{\scriptsize .NONFIN} this Mam but   //
\gleng "I don't know if you'll like this, Mam, but \dots " //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat meddai fo +" . //
\glsurface ELE:  meddai fo .  //
\glauto \%aut  say{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S}   //
\glmanual \%gls  say{\scriptsize .3S.PAST} PRON{\scriptsize .3SM}   //
\gleng he said. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" um@s:cym\&eng beth am roid \#\# deg ceiniog yr wsnos i mi am ymarfer ? //
\glsurface ELE:  um$^{C}_{E}$ beth am roid deg ceiniog yr wsnos i mi am ymarfer ?  //
\glauto \%aut  um{\scriptsize .IM} what{\scriptsize .INT} for{\scriptsize .PREP} give{\scriptsize .V.0.IMPERF+SM} ten{\scriptsize .NUM} penny{\scriptsize .N.F.SG} the{\scriptsize .DET.DEF} week{\scriptsize .N.F.SG} to{\scriptsize .PREP} I{\scriptsize .PRON.1S} for{\scriptsize .PREP} practice{\scriptsize .V.INFIN}   //
\glmanual \%gls  IM what about give{\scriptsize .NONFIN} ten penny DET week to PRON{\scriptsize .1S} for practising   //
\gleng "how about giving me ten pence a week for practising?" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ah@s:cym\&eng ! //
\glsurface ALN:  ah$^{C}_{E}$ !  //
\glauto \%aut  ah{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat doniol ! //
\glsurface ALN:  doniol !  //
\glauto \%aut  funny{\scriptsize .ADJ}   //
\glmanual \%gls  funny   //
\gleng funny! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ a wnes innau ddeud +"/ . //
\glsurface ELE:  a wnes innau ddeud .  //
\glauto \%aut  and{\scriptsize .CONJ} do{\scriptsize .V.1S.PAST.SPOKEN+SM} I{\scriptsize .also.PRON} say{\scriptsize .V.INFIN.SPOKEN+SM}   //
\glmanual \%gls  and do{\scriptsize .1S.PAST} PRON{\scriptsize .1S} say{\scriptsize .NONFIN}   //
\gleng and I said: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" well@s:cym\&eng \# os wnei di deg munud \# gofalus a canolbwyntio am y ddeg munud de gei di ddeg ceiniog \# bob diwrnod . //
\glsurface ELE:  well$^{C}_{E}$ os wnei di deg munud gofalus a canolbwyntio am y ddeg munud de gei di ddeg ceiniog bob diwrnod .  //
\glauto \%aut  well{\scriptsize .ADV} if{\scriptsize .CONJ} do{\scriptsize .V.2S.PRES+SM} you{\scriptsize .PRON.2S+SM} ten{\scriptsize .NUM} minute{\scriptsize .N.M.SG} careful{\scriptsize .ADJ} and{\scriptsize .CONJ} concentrate{\scriptsize .V.INFIN} for{\scriptsize .PREP} the{\scriptsize .DET.DEF} ten{\scriptsize .NUM+SM} minute{\scriptsize .N.M.SG} isn't\_it{\scriptsize .IM[or]south.N.M.SG[or]right.N.M.SG[or]tea.N.M.SG+SM} quay{\scriptsize .N.M.SG+SM} you{\scriptsize .PRON.2S+SM} ten{\scriptsize .NUM+SM} penny{\scriptsize .N.F.SG} each{\scriptsize .PREQ+SM} day{\scriptsize .N.M.SG}   //
\glmanual \%gls  well if do{\scriptsize .2S.NONPAST} PRON{\scriptsize .2S} ten minute careful and concentrate{\scriptsize .NONFIN} for DET ten minute TAG get{\scriptsize .2S.NONPAST} PRON{\scriptsize .2S} ten penny every day   //
\gleng "well, if you do ten minutes carefully, and concentrate for the ten minutes, right, you can have ten pence every day." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat sy (y)n golygu bod (y)na chwe\_deg ceiniog am pres pocket@s:cym\&eng wedyn does ? //
\glsurface ELE:  sy yn golygu bod yna chwe\_deg ceiniog am pres pocket$^{C}_{E}$ wedyn does ?  //
\glauto \%aut  which\_is{\scriptsize .V.3SP.PRES.REL} stative{\scriptsize .STAT} edit{\scriptsize .V.INFIN} be{\scriptsize .V.INFIN} there{\scriptsize .ADV} sixty{\scriptsize .NUM} penny{\scriptsize .N.F.SG} for{\scriptsize .PREP} money{\scriptsize .N.M.SG} pocket{\scriptsize .N.SG} afterwards{\scriptsize .ADV} be{\scriptsize .V.3S.PRES.INDEF.NEG}   //
\glmanual \%gls  be{\scriptsize .PRES.REL} PRT mean{\scriptsize .NONFIN} be{\scriptsize .NONFIN} there sixty penny for money pocket then be{\scriptsize .3S.PRES.NEG}   //
\gleng which means there's sixty pence for pocket money then, isn't there? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oes . //
\glsurface ALN:  oes .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.INDEF}   //
\glmanual \%gls  be{\scriptsize .3S.PRES}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=sniff (y)dy o (y)n iawn ? //
\glsurface ELE:  ydy o yn iawn ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} OK{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT alright   //
\gleng is that ok? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ac oedd (y)n meddwl bod hynna (y)n fargain@s:cym\&eng oedd ? //
\glsurface ALN:  ac oedd yn meddwl bod hynna yn fargain$^{C}_{E}$ oedd ?  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} stative{\scriptsize .STAT} think{\scriptsize .V.INFIN} be{\scriptsize .V.INFIN} that{\scriptsize .PRON.SP.SPOKEN} stative{\scriptsize .STAT} bargain{\scriptsize .N.SG+SM} be{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  and be{\scriptsize .3S.IMP} PRT think{\scriptsize .NONFIN} be{\scriptsize .NONFIN} that PRT bargain be{\scriptsize .3S.IMP}   //
\gleng and he thought that was a bargain, did he? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng (we)di gwirioni ! //
\glsurface ELE:  oh$^{C}_{E}$ wedi gwirioni !  //
\glauto \%aut  oh{\scriptsize .IM} after{\scriptsize .PREP} dote{\scriptsize .V.INFIN}   //
\glmanual \%gls  IM PRT{\scriptsize .PAST} become\_silly{\scriptsize .NONFIN}   //
\gleng oh, he was thrilled! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a mae (y)n golygu \# sti \# bod o (y)n ddeg munud \#\# lle mae (y)n canolbwyntio am y deg munud (y)na (he)fyd de ? //
\glsurface ELE:  a mae yn golygu sti bod o yn ddeg munud lle mae yn canolbwyntio am y deg munud yna hefyd de ?  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} edit{\scriptsize .V.INFIN} you\_know{\scriptsize .IM} be{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} ten{\scriptsize .NUM+SM} minute{\scriptsize .N.M.SG} where{\scriptsize .INT[or]place.N.M.SG} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} concentrate{\scriptsize .V.INFIN} for{\scriptsize .PREP} the{\scriptsize .DET.DEF} ten{\scriptsize .NUM} minute{\scriptsize .N.M.SG} there{\scriptsize .ADV} also{\scriptsize .ADV} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  and be{\scriptsize .3S.PRES} PRT mean{\scriptsize .NONFIN} know{\scriptsize .2S} be{\scriptsize .NONFIN} PRON{\scriptsize .3SM} PRT ten minute where be{\scriptsize .3S.PRES} PRT concentrate{\scriptsize .NONFIN} for DET ten minute there TAG    //
\gleng and it means, you know, that it's ten minutes where he's concentrating for those ten minutes, too, right? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng yeah@s:cym\&eng . //
\glsurface ALN:  yeah$^{C}_{E}$ yeah$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV} yeah{\scriptsize .ADV}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (we)dyn dwn (dd)im parith o . //
\glsurface ELE:  wedyn dwn ddim parith o .  //
\glauto \%aut  afterwards{\scriptsize .ADV} know{\scriptsize .V.1S.PRES.NEG} not{\scriptsize .ADV+SM} pair{\scriptsize .V.3S.FUT.SPOKEN} he{\scriptsize .PRON.M.3S.SPOKEN}   //
\glmanual \%gls  then know{\scriptsize .1S.NONPAST.NEG} NEG last{\scriptsize .3S.NONPAST} PRON{\scriptsize .3SM}   //
\gleng then I don't know if it will last. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae $<$(rh)aid ni drio$>$ [?] rywbeth . //
\glsurface ELE:  mae rhaid ni drio rywbeth .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} necessity{\scriptsize .N.M.SG} we{\scriptsize .PRON.1P} try{\scriptsize .V.INFIN+SM} something{\scriptsize .N.M.SG+SM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} necessity PRON{\scriptsize .1PL} try{\scriptsize .NONFIN} something   //
\gleng we have to try something. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat achos oedd hi (we)di cael llond bol heno de braidd ? //
\glsurface ELE:  achos oedd hi wedi cael llond bol heno de braidd ?  //
\glauto \%aut  cause{\scriptsize .N.M.SG} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} after{\scriptsize .PREP} have{\scriptsize .V.INFIN} fullness{\scriptsize .N.M.SG} belly{\scriptsize .N.M.SG} tonight{\scriptsize .ADV} isn't\_it{\scriptsize .IM[or]south.N.M.SG[or]right.N.M.SG[or]tea.N.M.SG+SM} rather{\scriptsize .ADV}   //
\glmanual \%gls  because be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT{\scriptsize .PAST} have{\scriptsize .NONFIN} full belly tonight TAG somewhat   //
\gleng cause she was pretty fed up tonight, you know? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wnaeth hi fynegi do ? //
\glsurface ALN:  wnaeth hi fynegi do ?  //
\glauto \%aut  do{\scriptsize .V.3S.PAST+SM} she{\scriptsize .PRON.F.3S} express{\scriptsize .V.INFIN+SM} yes{\scriptsize .ADV.PAST}   //
\glmanual \%gls  do{\scriptsize .3S.PAST} PRON{\scriptsize .3SF} express{\scriptsize .NONFIN} yes   //
\gleng she expressed it, didn't she? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wnaeth hi fynegi right@s:cym\&eng gry . //
\glsurface ELE:  wnaeth hi fynegi right$^{C}_{E}$ gry .  //
\glauto \%aut  do{\scriptsize .V.3S.PAST+SM} she{\scriptsize .PRON.F.3S} express{\scriptsize .V.INFIN+SM} right{\scriptsize .ADJ} strong{\scriptsize .ADJ+SM}   //
\glmanual \%gls  do{\scriptsize .3S.PAST} PRON{\scriptsize .3SF} express{\scriptsize .NONFIN} right strong   //
\gleng she expressed it quite stongly. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o'n i (y)n meddwl oedd right@s:cym\&eng dda really@s:eng . //
\glsurface ELE:  o'n i yn meddwl oedd right$^{C}_{E}$ dda really$^{E}$ .  //
\glauto \%aut  unk to{\scriptsize .PREP} stative{\scriptsize .STAT} think{\scriptsize .V.INFIN} be{\scriptsize .V.3S.IMPERF} right{\scriptsize .ADJ} good{\scriptsize .ADJ+SM} real{\scriptsize .ADJ+LY}   //
\glmanual \%gls  be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} PRT think{\scriptsize .NONFIN} be{\scriptsize .3S.IMP} right good really   //
\gleng I thought it was quite good, really. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat da iawn . //
\glsurface ALN:  da iawn .  //
\glauto \%aut  good{\scriptsize .ADJ} very{\scriptsize .ADV}   //
\glmanual \%gls  good very   //
\gleng very good. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat um@s:cym\&eng \#\# yn\_de ? //
\glsurface ELE:  um$^{C}_{E}$ yn\_de ?  //
\glauto \%aut  um{\scriptsize .IM} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  IM TAG   //
\gleng isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng . //
\glsurface ALN:  yeah$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat timod \#\# um@s:cym\&eng +. . . //
\glsurface ELE:  timod um$^{C}_{E}$ .  //
\glauto \%aut  know{\scriptsize .V.2S.PRES.SPOKEN} um{\scriptsize .IM}   //
\glmanual \%gls  know{\scriptsize .2S} IM   //
\gleng you know, um \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$oh@s:cym\&eng mae (y)n$>$ [?] braf bod o (y)n dod o rywle arall heblaw ni dydy ? //
\glsurface ALN:  oh$^{C}_{E}$ mae yn braf bod o yn dod o rywle arall heblaw ni dydy ?  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} fine{\scriptsize .ADJ} be{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} come{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} somewhere{\scriptsize .N.M.SG+SM} other{\scriptsize .ADJ} without{\scriptsize .PREP} we{\scriptsize .PRON.1P} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} PRT fine be{\scriptsize .NONFIN} PRON{\scriptsize .3SM} PRT come{\scriptsize .NONFIN} from somewhere other besides PRON{\scriptsize .1PL} be{\scriptsize .3S.PRES.NEG}   //
\gleng oh it's nice for it to be coming from somewhere else other than us, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ yndy yndy . //
\glsurface ELE:  yndy yndy .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.EMPH} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES}   //
\gleng yes, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond mae (y)r \# darnau yn dod i ben \# +/ . //
\glsurface ELE:  ond mae yr darnau yn dod i ben .  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} the{\scriptsize .DET.DEF} pieces{\scriptsize .N.M.PL} stative{\scriptsize .STAT} come{\scriptsize .V.INFIN} to{\scriptsize .PREP} head{\scriptsize .N.M.SG+SM}   //
\glmanual \%gls  but be{\scriptsize .3S.PRES} DET pieces PRT come{\scriptsize .NONFIN} to head   //
\gleng but the pieces are ?getting there. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yndyn . //
\glsurface ALN:  yndyn .  //
\glauto \%aut  be{\scriptsize .V.3P.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3PL.PRES}   //
\gleng they are. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat so@s:cym\&eng $<$fydd raid iddo$>$ [?] ddysgu set@s:cym\&eng newydd os na (y)dy (y)n gwybod (r)hein . //
\glsurface ALN:  so$^{C}_{E}$ fydd raid iddo ddysgu set$^{C}_{E}$ newydd os na ydy yn gwybod rhein .  //
\glauto \%aut  so{\scriptsize .ADV} be{\scriptsize .V.3S.FUT+SM} necessity{\scriptsize .N.M.SG+SM} to\_him{\scriptsize .PREP+PRON.M.3S} teach{\scriptsize .V.INFIN+SM} set{\scriptsize .SV.INFIN} new{\scriptsize .ADJ} if{\scriptsize .CONJ} neg{\scriptsize .PRT} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN} these{\scriptsize .PRON}   //
\glmanual \%gls  so be{\scriptsize .3S.FUT} necessity to{\scriptsize .3SM} learn{\scriptsize .NONFIN} set new if NEG be{\scriptsize .3S.PRES} PRT know{\scriptsize .NONFIN} these   //
\gleng so he'll have to learn a new set if he doesn't know these. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ mis Mawrth de ? //
\glsurface ELE:  mis Mawrth de ?  //
\glauto \%aut  month{\scriptsize .N.M.SG} March{\scriptsize .N.M.SG} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  month March TAG   //
\gleng in March, right? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ basai $<$raid (i)ddo$>$ [?] ddechrau o: (y)r dechrau . //
\glsurface ELE:  basai raid iddo ddechrau o yr dechrau .  //
\glauto \%aut  be{\scriptsize .V.3S.PLUPERF.SPOKEN} necessity{\scriptsize .N.M.SG+SM} to\_him{\scriptsize .PREP+PRON.M.3S} begin{\scriptsize .V.INFIN+SM} of{\scriptsize .PREP} the{\scriptsize .DET.DEF} beginning{\scriptsize .N.M.SG}   //
\glmanual \%gls  be{\scriptsize .3S.CONDIT} necessity to{\scriptsize .3SM} start{\scriptsize .NONFIN} from DET start   //
\gleng he'd have to start right from the beginning. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat allai fod hynna mynd i\_lawr yn dda ! //
\glsurface ALN:  allai fod hynna mynd i\_lawr yn dda !  //
\glauto \%aut  be\_able{\scriptsize .V.3S.IMPERF+SM} be{\scriptsize .V.INFIN+SM} that{\scriptsize .PRON.SP.SPOKEN} go{\scriptsize .V.INFIN} down{\scriptsize .ADV} stative{\scriptsize .STAT} good{\scriptsize .ADJ+SM}   //
\glmanual \%gls  can{\scriptsize .3S.NONPAST} be{\scriptsize .NONFIN} that go{\scriptsize .NONFIN} down PRT good   //
\gleng that might go down well! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=sigh . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie $<$wedyn os (y)dy o (y)n cael mae (y)n sure@s:cym\&eng ddylai hi gael (fe)lly dylai$>$ [?] ? //
\glsurface ELE:  ie wedyn os ydy o yn cael mae yn sure$^{C}_{E}$ ddylai hi gael felly dylai ?  //
\glauto \%aut  yes{\scriptsize .ADV} afterwards{\scriptsize .ADV} if{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} have{\scriptsize .V.INFIN} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} sure{\scriptsize .ADJ} ought{\scriptsize .V.3S.IMPERF+SM} she{\scriptsize .PRON.F.3S} have{\scriptsize .V.INFIN+SM} so{\scriptsize .ADV} ought{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  yes after if be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT get{\scriptsize .NONFIN} be{\scriptsize .3S.PRES} PRT sure should{\scriptsize .3S.CONDIT} PRON{\scriptsize .3SF} get{\scriptsize .NONFIN} thus should{\scriptsize .3S.NONPAST}   //
\gleng yes, so if he gets it, she should get it too, shouldn't she, probably. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng dylai ! //
\glsurface ALN:  oh$^{C}_{E}$ dylai !  //
\glauto \%aut  oh{\scriptsize .IM} ought{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  IM should{\scriptsize .3S.CONDIT}   //
\gleng oh, she should! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond $<$be dan ni (y)n$>$ [//] \#\#\# sut dan ni (y)n er@s:cym\&eng +. . . //
\glsurface ALN:  ond be dan ni yn sut dan ni yn er$^{C}_{E}$ .  //
\glauto \%aut  but{\scriptsize .CONJ} what{\scriptsize .INT} be{\scriptsize .V.1P.PRES.SPOKEN} we{\scriptsize .PRON.1P} yn{\scriptsize .PRT} how{\scriptsize .INT} be{\scriptsize .V.1P.PRES.SPOKEN} we{\scriptsize .PRON.1P} yn{\scriptsize .PRT} er{\scriptsize .IM}   //
\glmanual \%gls  but what be{\scriptsize .1PL.PRES} PRON{\scriptsize .1PL} PRT how be{\scriptsize .1PL.PRES} PRON{\scriptsize .1PL} PRT IM   //
\gleng but what are we. . . how are we, er \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat piano@s:cym\&eng [=! quietly] . //
\glsurface ELE:  piano$^{C}_{E}$ .  //
\glauto \%aut  piano{\scriptsize .N.SG}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng \# yeah@s:cym\&eng o\_k@s:cym\&eng yeah@s:cym\&eng . //
\glsurface ALN:  yeah$^{C}_{E}$ yeah$^{C}_{E}$ o\_k$^{C}_{E}$ yeah$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV} yeah{\scriptsize .ADV} OK{\scriptsize .IM} yeah{\scriptsize .ADV}   //
\glmanual \%gls        //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng ? //
\glsurface ELE:  yeah$^{C}_{E}$ ?  //
\glauto \%aut  yeah{\scriptsize .ADV}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat duh@s:cym\&eng yeah@s:cym\&eng wrth\_gwrs ! //
\glsurface ALN:  duh$^{C}_{E}$ yeah$^{C}_{E}$ wrth\_gwrs !  //
\glauto \%aut  duh{\scriptsize .IM} yeah{\scriptsize .ADV} of\_course{\scriptsize .ADV}   //
\glmanual \%gls  IM yeah of\_course   //
\gleng duh, yeah, of course! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ achos am [/] \# am yr gwely mae (y)n cael yr llall de ? //
\glsurface ELE:  achos am am yr gwely mae yn cael yr llall de ?  //
\glauto \%aut  cause{\scriptsize .N.M.SG} for{\scriptsize .PREP} for{\scriptsize .PREP} the{\scriptsize .DET.DEF} bed{\scriptsize .N.M.SG} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} have{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} other{\scriptsize .PRON} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  because for for DET bed be{\scriptsize .3S.PRES} PRT get{\scriptsize .NONFIN} DET other TAG   //
\gleng because the other one he gets is for bed, right? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng yeah@s:cym\&eng ! //
\glsurface ALN:  yeah$^{C}_{E}$ yeah$^{C}_{E}$ !  //
\glauto \%aut  yeah{\scriptsize .ADV} yeah{\scriptsize .ADV}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wedyn um@s:cym\&eng +. . . //
\glsurface ELE:  wedyn um$^{C}_{E}$ .  //
\glauto \%aut  afterwards{\scriptsize .ADV} um{\scriptsize .IM}   //
\glmanual \%gls  after IM   //
\gleng then, um \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well@s:cym\&eng gawn ni weld sut eith o de ? //
\glsurface ALN:  well$^{C}_{E}$ gawn ni weld sut eith o de ?  //
\glauto \%aut  well{\scriptsize .ADV} get{\scriptsize .V.1P.PRES+SM} we{\scriptsize .PRON.1P} see{\scriptsize .V.INFIN+SM} how{\scriptsize .INT} go{\scriptsize .V.ES.PRES} he{\scriptsize .PRON.M.3S.SPOKEN[or]from.PREP[or]of.PREP} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  well have{\scriptsize .1PL.NONPAST} PRON{\scriptsize .1PL} see{\scriptsize .NONPAST} how go{\scriptsize .3S.NONPAST} PRON{\scriptsize .3SM} TAG   //
\gleng well, we'll see how it goes, won't we? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae gen i deimlad wneith o ymarfer y drums yn haws . //
\glsurface ALN:  mae gen i deimlad wneith o ymarfer y drums yn haws .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} by{\scriptsize .PREP.SPOKEN} I{\scriptsize .PRON.1S} feeling{\scriptsize .N.M.SG+SM} do{\scriptsize .V.3S.FUT+SM} he{\scriptsize .PRON.M.3S.SPOKEN} practice{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} unk stative{\scriptsize .STAT} easier{\scriptsize .ADJ}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} with{\scriptsize .1S} PRON{\scriptsize .1S} feeling do{\scriptsize .3S.NONPAST} PRON{\scriptsize .3SM} practise{\scriptsize .NONFIN} DET drums PRT easier   //
\gleng I've a feeling he'll practise the drums more easily. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be mae (y)n debygol o wneud efo fo dydd Gwener ta ? //
\glsurface ELE:  be mae yn debygol o wneud efo fo dydd Gwener ta ?  //
\glauto \%aut  what{\scriptsize .INT} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} likely{\scriptsize .ADJ+SM[or]probable.ADJ+SM} of{\scriptsize .PREP} make{\scriptsize .V.INFIN+SM} with{\scriptsize .PREP} he{\scriptsize .PRON.M.3S} day{\scriptsize .N.M.SG} name or{\scriptsize .CONJ}   //
\glmanual \%gls  what be{\scriptsize .3S.PRES} PRT likely of do{\scriptsize .NONFIN} with PRON{\scriptsize .3SM} day Friday then   //
\gleng what's he likely to do with him on Friday, then? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dim syniad ! //
\glsurface ALN:  dim syniad !  //
\glauto \%aut  not{\scriptsize .ADV} idea{\scriptsize .N.M.SG}   //
\glmanual \%gls  NEG idea   //
\gleng no idea! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat right@s:cym\&eng ! //
\glsurface ELE:  right$^{C}_{E}$ !  //
\glauto \%aut  right{\scriptsize .ADJ}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i (we)di deud bod fi (y)n awyddus iawn iddo fo ddarllen . //
\glsurface ALN:  dw i wedi deud bod fi yn awyddus iawn iddo fo ddarllen .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} after{\scriptsize .PREP} say{\scriptsize .V.INFIN.SPOKEN} be{\scriptsize .V.INFIN} I{\scriptsize .PRON.1S+SM} stative{\scriptsize .STAT} eager{\scriptsize .ADJ} very{\scriptsize .ADV} to\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S} read{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT{\scriptsize .PAST} say{\scriptsize .NONFIN} be{\scriptsize .NONFIN} PRON{\scriptsize .1S} PRT keen very for{\scriptsize .3SM} PRON{\scriptsize .3SM} read{\scriptsize .NONFIN}   //
\gleng I've said that I'm very keen for him to read. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a mae [/] mae Dean@s:cym\&eng yn dallt hynny . //
\glsurface ALN:  a mae mae Dean$^{C}_{E}$ yn dallt hynny .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} be{\scriptsize .V.3S.PRES} name stative{\scriptsize .STAT} understand{\scriptsize .V.INFIN.SPOKEN} that{\scriptsize .PRON.SP}   //
\glmanual \%gls  and be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES} Dean PRT understand{\scriptsize .NONFIN} that   //
\gleng and Dean understands that. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$mae o (y)n$>$ [//] mae (y)n cerddor . //
\glsurface ALN:  mae o yn mae yn cerddor .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} in{\scriptsize .PREP} be{\scriptsize .V.3S.PRES} yn{\scriptsize .PRT} musician{\scriptsize .N.M.SG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT be{\scriptsize .3S.PRES} PRT musician   //
\gleng he's a musician. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat um@s:cym\&eng +. . . //
\glsurface ALN:  um$^{C}_{E}$ .  //
\glauto \%aut  um{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae wneud y ddau yndy ? //
\glsurface ELE:  mae wneud y ddau yndy ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} make{\scriptsize .V.INFIN+SM} the{\scriptsize .DET.DEF} two{\scriptsize .NUM.M+SM} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} do{\scriptsize .NONFIN} DET two be{\scriptsize .3S.PRES}   //
\gleng he does both, doesn't he? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng yndy yndy yndy ! //
\glsurface ALN:  oh$^{C}_{E}$ yndy yndy yndy !  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES.EMPH} be{\scriptsize .V.3S.PRES.EMPH} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES}   //
\gleng oh, yes, yes, yes! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmhm@s:cym\&eng . //
\glsurface ELE:  mmhm$^{C}_{E}$ .  //
\glauto \%aut  mmhm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng mae wneud gradd yn fa(n) (y)ma . //
\glsurface ALN:  oh$^{C}_{E}$ mae wneud gradd yn fan yma .  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES} make{\scriptsize .V.INFIN+SM} degree{\scriptsize .N.MF.SG} stative{\scriptsize .STAT} place{\scriptsize .N.MF.SG+SM} here{\scriptsize .ADV}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} do{\scriptsize .NONFIN} degree in place here   //
\gleng oh he's doing a degree here. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng dyna mae wneud ! //
\glsurface ELE:  oh$^{C}_{E}$ dyna mae wneud !  //
\glauto \%aut  oh{\scriptsize .IM} that{\scriptsize .is.ADV} be{\scriptsize .V.3S.PRES} make{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  IM there be{\scriptsize .3S.PRES} do{\scriptsize .NONFIN}   //
\gleng oh, that's what he does! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng yeah@s:cym\&eng yeah@s:cym\&eng sorry@s:cym\&eng ! //
\glsurface ELE:  yeah$^{C}_{E}$ yeah$^{C}_{E}$ yeah$^{C}_{E}$ sorry$^{C}_{E}$ !  //
\glauto \%aut  yeah{\scriptsize .ADV} yeah{\scriptsize .ADV} yeah{\scriptsize .ADV} sorry{\scriptsize .ADJ}   //
\glmanual \%gls        //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o'n i (dd)im yn cofio . //
\glsurface ELE:  o'n i ddim yn cofio .  //
\glauto \%aut  unk to{\scriptsize .PREP} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} remember{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} NEG PRT remember{\scriptsize .NONFIN}   //
\gleng I'd forgotten. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae (y)n dod o gefndir \#\# cerddorion . //
\glsurface ALN:  mae yn dod o gefndir cerddorion .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} come{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} background{\scriptsize .N.M.SG+SM} musician{\scriptsize .N.M.PL}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRT come{\scriptsize .NONFIN} from background musicians   //
\gleng he comes from a background of musicians. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yndy yndy . //
\glsurface ELE:  yndy yndy .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.EMPH} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES}   //
\gleng yes, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat sy (y)n dallt \# mae raid chdi ddarllen . //
\glsurface ALN:  sy yn dallt mae raid chdi ddarllen .  //
\glauto \%aut  which\_is{\scriptsize .V.3SP.PRES.REL} stative{\scriptsize .STAT} understand{\scriptsize .V.INFIN.SPOKEN} be{\scriptsize .V.3S.PRES} necessity{\scriptsize .N.M.SG+SM} you{\scriptsize .PRON} read{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  be{\scriptsize .PRES.REL} PRT understand{\scriptsize .NONFIN} be{\scriptsize .3S.PRES} necessity PRON{\scriptsize .2S} read{\scriptsize .NONFIN}   //
\gleng that understands that you have to read. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng yeah@s:cym\&eng . //
\glsurface ELE:  yeah$^{C}_{E}$ yeah$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV} yeah{\scriptsize .ADV}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (dy)dy o (dd)im just@s:cym\&eng yn +"/ . //
\glsurface ALN:  dydy o ddim just$^{C}_{E}$ yn .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} he{\scriptsize .PRON.M.3S.SPOKEN} not{\scriptsize .ADV+SM} just{\scriptsize .ADV[or]just.ADJ} yn{\scriptsize .PRT}   //
\glmanual \%gls  be{\scriptsize .3S.PRES.NEG} PRON{\scriptsize .3SM} NEG just PRT   //
\gleng he's not just like: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" oh@s:cym\&eng yeah@s:cym\&eng just@s:cym\&eng chwarae dy offeryn . //
\glsurface ALN:  oh$^{C}_{E}$ yeah$^{C}_{E}$ just$^{C}_{E}$ chwarae dy offeryn .  //
\glauto \%aut  oh{\scriptsize .IM} yeah{\scriptsize .ADV} just{\scriptsize .ADV[or]just.ADJ} game{\scriptsize .N.M.SG[or]play.V.2S.IMPER[or]play.V.INFIN} your{\scriptsize .ADJ.POSS} instrument{\scriptsize .N.M.SG}   //
\glmanual \%gls  IM yeah just play{\scriptsize .2S.IMPER} POSS{\scriptsize .2S} instrument   //
\gleng "oh, yeah, just play your instrument." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" a fydd bob\_dim yn great@s:cym\&eng ! //
\glsurface ALN:  a fydd bob\_dim yn great$^{C}_{E}$ !  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.FUT+SM} everything{\scriptsize .PRON} stative{\scriptsize .STAT} great{\scriptsize .ADJ}   //
\glmanual \%gls  and be{\scriptsize .3S.FUT} everything PRT great   //
\gleng "and everything'll be great!" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat y lol yna . //
\glsurface ALN:  y lol yna .  //
\glauto \%aut  the{\scriptsize .DET.DEF} nonsense{\scriptsize .N.F.SG} there{\scriptsize .ADV}   //
\glmanual \%gls  DET nonsense there   //
\gleng that nonsense. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng hey@s:cym\&eng ti (dd)im (we)di cael ateb gyn Ed@s:cym\&eng am yr um@s:cym\&eng \# caster@s:cym\&eng (y)na ar\_gyfer y piano@s:cym\&eng ? //
\glsurface ELE:  hmm$^{C}_{E}$ hey$^{C}_{E}$ ti ddim wedi cael ateb gyn Ed$^{C}_{E}$ am yr um$^{C}_{E}$ caster$^{C}_{E}$ yna ar\_gyfer y piano$^{C}_{E}$ ?  //
\glauto \%aut  hmm{\scriptsize .IM} hey{\scriptsize .IM} you{\scriptsize .PRON.2S} nothing{\scriptsize .N.M.SG+SM[or]not.ADV+SM} after{\scriptsize .PREP} have{\scriptsize .V.INFIN} answer{\scriptsize .V.INFIN} with{\scriptsize .PREP[or]before.PREP+SM} name for{\scriptsize .PREP} the{\scriptsize .DET.DEF} um{\scriptsize .IM} caster{\scriptsize .N.SG} there{\scriptsize .ADV} for{\scriptsize .PREP} the{\scriptsize .DET.DEF} piano{\scriptsize .N.SG}   //
\glmanual \%gls  IM hey PRON{\scriptsize .2S} NEG PRT{\scriptsize .PAST} get{\scriptsize .NONFIN} answer with Ed about DET IM caster there for DET piano   //
\gleng hey, you haven't had an answer off Ed about that caster for the piano? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=slurp ddim hyd yn hyn . //
\glsurface ALN:  ddim hyd yn hyn .  //
\glauto \%aut  not{\scriptsize .ADV+SM} length{\scriptsize .N.M.SG} in{\scriptsize .PREP} this{\scriptsize .PRON.SP}   //
\glmanual \%gls  NEG until PRT this   //
\gleng not so far. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat nos Iau mae (y)n wneud yr \# gig@s:cym\&eng (y)na yeah@s:cym\&eng \# yn Gaernarfon@s:cym\&eng \# Galeri@s:cym\&eng ? //
\glsurface ELE:  nos Iau mae yn wneud yr gig$^{C}_{E}$ yna yeah$^{C}_{E}$ yn Gaernarfon$^{C}_{E}$ Galeri$^{C}_{E}$ ?  //
\glauto \%aut  night{\scriptsize .N.F.SG} Jupiter{\scriptsize .N.M.SG} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM} the{\scriptsize .DET.DEF} gig{\scriptsize .N.SG[or]cig.N.SG+SM} there{\scriptsize .ADV} yeah{\scriptsize .ADV} in{\scriptsize .PREP} name name   //
\glmanual \%gls  night Thursday be{\scriptsize .3S.PRES} PRT do{\scriptsize .NONFIN} DET gig there yeah in Caernarfon Galeri   //
\gleng Thursday night s/he's doing this gig, yeah, in Caernarfon, Galeri? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie . //
\glsurface ALN:  ie .  //
\glauto \%aut  yes{\scriptsize .ADV}   //
\glmanual \%gls  yes   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dan ni (dd)im yn rhydd na(g) (y)dan ? //
\glsurface ALN:  dan ni ddim yn rhydd nag ydan ?  //
\glauto \%aut  be{\scriptsize .V.1P.PRES.SPOKEN} we{\scriptsize .PRON.1P} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} free{\scriptsize .ADJ[or]give.V.3S.PRES} than{\scriptsize .CONJ} be{\scriptsize .V.1P.PRES.SPOKEN}   //
\glmanual \%gls  be{\scriptsize .1PL.PRES} PRON{\scriptsize .1PL} NEG PRT free NEG be{\scriptsize .1PL.PRES}   //
\gleng we're not free, are we? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng +. . . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (ba)sai (y)n doniol just@s:cym\&eng mynd i chwerthin ! //
\glsurface ALN:  basai yn doniol just$^{C}_{E}$ mynd i chwerthin !  //
\glauto \%aut  be{\scriptsize .V.3S.PLUPERF.SPOKEN} stative{\scriptsize .STAT} funny{\scriptsize .ADJ} just{\scriptsize .ADV[or]just.ADJ} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} laugh{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .3S.CONDIT} PRT funny just go{\scriptsize .NONFIN} to laugh{\scriptsize .NONFIN}   //
\gleng it'd be funny to just go for a laugh! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ well@s:cym\&eng na \&wǝ mae bosib de ? //
\glsurface ELE:  well$^{C}_{E}$ na mae bosib de ?  //
\glauto \%aut  well{\scriptsize .ADV} neg{\scriptsize .PRT} be{\scriptsize .V.3S.PRES} possible{\scriptsize .ADJ+SM} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  well no be{\scriptsize .3S.PRES} possible TAG   //
\gleng well, no, it is possible, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ oh@s:cym\&eng na na wir wir ! //
\glsurface ALN:  oh$^{C}_{E}$ na na wir wir !  //
\glauto \%aut  oh{\scriptsize .IM} neg{\scriptsize .PRT} neg{\scriptsize .PRT[or]who\_not.PRON.REL.NEG[or](n)or.CONJ[or]than.CONJ} true{\scriptsize .ADJ+SM} true{\scriptsize .ADJ+SM}   //
\glmanual \%gls  IM no no true true   //
\gleng oh, no, no, honestly! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat gwranda wan ! //
\glsurface ELE:  gwranda wan !  //
\glauto \%aut  listen{\scriptsize .V.2S.IMPER} weak{\scriptsize .ADJ+SM[or]pale.ADJ+SM}   //
\glmanual \%gls  listen{\scriptsize .2S.IMPER} now   //
\gleng listen now! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wsti mae Auntie@s:cym\&eng Carol@s:cym\&eng (y)n dod \# +/ . //
\glsurface ELE:  wsti mae Auntie$^{C}_{E}$ Carol$^{C}_{E}$ yn dod .  //
\glauto \%aut  know{\scriptsize .V.2S.PRES.SPOKEN} be{\scriptsize .V.3S.PRES} name name stative{\scriptsize .STAT} come{\scriptsize .V.INFIN}   //
\glmanual \%gls  know{\scriptsize .2S} be{\scriptsize .3S.PRES} Auntie Carol PRT come{\scriptsize .NONFIN}   //
\gleng you know, Auntie Carol's coming \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ALN:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, erbyn pump tydy ? //
\glsurface ELE:  erbyn pump tydy ?  //
\glauto \%aut  by{\scriptsize .PREP} five{\scriptsize .NUM} be{\scriptsize .V.3S.PRES.NEG}   //
\glmanual \%gls  by five be{\scriptsize .3S.PRES.NEG}   //
\gleng by five, isn't she ? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmhm@s:cym\&eng . //
\glsurface ALN:  mmhm$^{C}_{E}$ .  //
\glauto \%aut  mmhm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat right@s:cym\&eng . //
\glsurface ELE:  right$^{C}_{E}$ .  //
\glauto \%aut  right{\scriptsize .ADJ}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat tan chwech . //
\glsurface ELE:  tan chwech .  //
\glauto \%aut  until{\scriptsize .PREP} six{\scriptsize .NUM}   //
\glmanual \%gls  until six   //
\gleng till six. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat fedra i fynd efo Debbie@s:cym\&eng a Charlie@s:cym\&eng am chwech \# tan saith . //
\glsurface ELE:  fedra i fynd efo Debbie$^{C}_{E}$ a Charlie$^{C}_{E}$ am chwech tan saith .  //
\glauto \%aut  be\_able{\scriptsize .V.1S.PRES} I{\scriptsize .PRON.1S} go{\scriptsize .V.INFIN+SM} with{\scriptsize .PREP} name and{\scriptsize .CONJ} name for{\scriptsize .PREP} six{\scriptsize .NUM} until{\scriptsize .PREP} seven{\scriptsize .NUM}   //
\glmanual \%gls  can{\scriptsize .1S.NONPAST} PRON{\scriptsize .1S} go{\scriptsize .NONFIN} with Debbie and Charlie at six until seven   //
\gleng I can go with Debbie and Charlie at six, until seven. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng na dw i ddim isio mynd \# coelia di fi ! //
\glsurface ALN:  oh$^{C}_{E}$ na dw i ddim isio mynd coelia di fi !  //
\glauto \%aut  oh{\scriptsize .IM} neg{\scriptsize .PRT} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} want{\scriptsize .N.M.SG} go{\scriptsize .V.INFIN} believe{\scriptsize .V.2S.IMPER[or]believe.V.3S.PRES} you{\scriptsize .PRON.2S+SM} I{\scriptsize .PRON.1S+SM}   //
\glmanual \%gls  IM no be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} NEG want go{\scriptsize .NONFIN} believe{\scriptsize .2S.IMPER} PRON{\scriptsize .2S} PRON{\scriptsize .1S}   //
\gleng oh, no, I don't want to go, believe you me! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ oh@s:cym\&eng ti sure@s:cym\&eng ? //
\glsurface ELE:  oh$^{C}_{E}$ ti sure$^{C}_{E}$ ?  //
\glauto \%aut  oh{\scriptsize .IM} you{\scriptsize .PRON.2S} sure{\scriptsize .ADJ}   //
\glmanual \%gls  IM PRON{\scriptsize .2S} sure   //
\gleng oh, you sure? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat argian fawr na ! //
\glsurface ALN:  argian fawr na !  //
\glauto \%aut  good\_lord{\scriptsize .IM} big{\scriptsize .ADJ+SM} neg{\scriptsize .PRT}   //
\glmanual \%gls  lord big no   //
\gleng good grief, no! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat pam ? //
\glsurface ELE:  pam ?  //
\glauto \%aut  why?.ADV   //
\glmanual \%gls  why   //
\gleng why? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i (y)n ddigon hapus i chwarae (y)r stuff@s:cym\&eng am bres ond dw i (ddi)m isio gwrando (ar)no fo ! //
\glsurface ALN:  dw i yn ddigon hapus i chwarae yr stuff$^{C}_{E}$ am bres ond dw i ddim isio gwrando arno fo !  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} stative{\scriptsize .STAT} enough{\scriptsize .QUAN+SM} happy{\scriptsize .ADJ} to{\scriptsize .PREP} play{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} stuff{\scriptsize .SV.INFIN} for{\scriptsize .PREP} money{\scriptsize .N.M.SG+SM} but{\scriptsize .CONJ} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} want{\scriptsize .N.M.SG} listen{\scriptsize .V.INFIN} on\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT enough happy to play{\scriptsize .NONFIN} DET stuff for money but be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} NEG want listen{\scriptsize .NONFIN} on{\scriptsize .3SM} PRON{\scriptsize .3SM}   //
\gleng I'm happy enough to play the stuff for money, but I don't want to listen to it! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng [=! laughs] \&=laugh right@s:cym\&eng [?] o\_k@s:cym\&eng ! //
\glsurface ELE:  oh$^{C}_{E}$ right$^{C}_{E}$ o\_k$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM} right{\scriptsize .ADJ} OK{\scriptsize .IM}   //
\glmanual \%gls       //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng mae just@s:cym\&eng yn um@s:cym\&eng +. . . //
\glsurface ALN:  oh$^{C}_{E}$ mae just$^{C}_{E}$ yn um$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES} just{\scriptsize .ADV[or]just.ADJ} yn{\scriptsize .PRT} um{\scriptsize .IM}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} just PRT IM   //
\gleng oh, it's just \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=slurp e:r@s:cym\&eng style@s:cym\&eng mae (y)n chwarae (y)dy \&=mumble +. . . //
\glsurface ALN:  er$^{C}_{E}$ style$^{C}_{E}$ mae yn chwarae ydy .  //
\glauto \%aut  er{\scriptsize .IM} style{\scriptsize .N.SG} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} play{\scriptsize .V.INFIN} be{\scriptsize .V.3S.PRES}   //
\glmanual \%gls  IM style be{\scriptsize .3S.PRES} PRT play{\scriptsize .NONFIN} be{\scriptsize .3S.PRES}   //
\gleng er, the style he plays is \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae wneud jazz@s:cym\&eng straight@s:cym\&eng ond mae o (y)n ymylu ar \# be mae nhw (y)n galw (y)n trad@s:cym\&eng ["] . //
\glsurface ALN:  mae wneud jazz$^{C}_{E}$ straight$^{C}_{E}$ ond mae o yn ymylu ar be mae nhw yn galw yn trad$^{C}_{E}$ .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} make{\scriptsize .V.INFIN+SM} jazz{\scriptsize .N.SG} straight{\scriptsize .ADJ} but{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} border{\scriptsize .V.INFIN} on{\scriptsize .PREP} what{\scriptsize .INT} be{\scriptsize .V.3S.PRES} they{\scriptsize .PRON.3P} stative{\scriptsize .STAT} call{\scriptsize .V.INFIN} yn{\scriptsize .PRT} trad{\scriptsize .N.SG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} do{\scriptsize .NONFIN} jazz straight but be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT border{\scriptsize .NONFIN} on what be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL} PRT call{\scriptsize .NONFIN} PRT trad   //
\gleng he does straight jazz, but it borders on what they call "trad" . //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +^ trad@s:cym\&eng oh@s:cym\&eng right@s:cym\&eng . //
\glsurface ELE:  trad$^{C}_{E}$ oh$^{C}_{E}$ right$^{C}_{E}$ .  //
\glauto \%aut  trad{\scriptsize .N.SG} oh{\scriptsize .IM} right{\scriptsize .ADJ}   //
\glmanual \%gls       //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ac yn trad@s:cym\&eng mae gen ti un\_ai \# clarinet@s:cym\&eng \#\# sef yr offeryn gwaetha yn y byd \# yn [/] \# yn whine\_io@s:eng+cym \# just@s:cym\&eng \&=imit:clarinet neu mae gen ti drombone@s:cym\&eng . //
\glsurface ALN:  ac yn trad$^{C}_{E}$ mae gen ti un\_ai clarinet$^{C}_{E}$ sef yr offeryn gwaetha yn y byd yn yn whine\_io$^{E+}_{C}$ just$^{C}_{E}$ neu mae gen ti drombone$^{C}_{E}$ .  //
\glauto \%aut  and{\scriptsize .CONJ} yn{\scriptsize .PRT} trad{\scriptsize .N.SG} be{\scriptsize .V.3S.PRES} by{\scriptsize .PREP.SPOKEN} you{\scriptsize .PRON.2S} either{\scriptsize .ADV} clarinet{\scriptsize .N.SG} namely{\scriptsize .CONJ} the{\scriptsize .DET.DEF} instrument{\scriptsize .N.M.SG} bad{\scriptsize .ADJ.SUP} in{\scriptsize .PREP} the{\scriptsize .DET.DEF} world{\scriptsize .N.M.SG} yn{\scriptsize .PRT} yn{\scriptsize .PRT} whine{\scriptsize .N.SG+WV} just{\scriptsize .ADV[or]just.ADJ} or{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} by{\scriptsize .PREP.SPOKEN} you{\scriptsize .PRON.2S} trombone{\scriptsize .N.SG+SM}   //
\glmanual \%gls  and in trad be{\scriptsize .3S.PRES} with{\scriptsize .2S} PRON{\scriptsize .2S} either clarinet namely DET instrument worst in DET world PRT PRT whine{\scriptsize .NONFIN} just or be{\scriptsize .3S.PRES} with{\scriptsize .2S} PRON{\scriptsize .2S} trombone   //
\gleng and in trad, you've got either a clarinet, which is the worst instrument in the world, whining, like $<$clarinet impression$>$, or you've got a trombone. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ well@s:cym\&eng pwy sy (y)n chwarae hwnna te ? //
\glsurface ELE:  well$^{C}_{E}$ pwy sy yn chwarae hwnna te ?  //
\glauto \%aut  well{\scriptsize .ADV} who{\scriptsize .PRON} which\_is{\scriptsize .V.3SP.PRES.REL} stative{\scriptsize .STAT} play{\scriptsize .V.INFIN} that{\scriptsize .PRON.M.SG.SPOKEN} tea{\scriptsize .N.M.SG}   //
\glmanual \%gls  well who be{\scriptsize .PRES.REL} PRT play{\scriptsize .NONFIN} that then   //
\gleng well, who plays that then? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng does neb yn chwarae \&kl clarinet@s:cym\&eng efo Ed@s:cym\&eng . //
\glsurface ALN:  oh$^{C}_{E}$ does neb yn chwarae clarinet$^{C}_{E}$ efo Ed$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES.INDEF.NEG} anyone{\scriptsize .PRON} stative{\scriptsize .STAT} play{\scriptsize .V.INFIN} clarinet{\scriptsize .N.SG} with{\scriptsize .PREP} name   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES.NEG} nobody PRT play{\scriptsize .NONFIN} clarinet with Ed   //
\gleng oh, no one plays clarinet with Ed. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond \# mae gynno fo foy@s:cym\&eng yn chwarae trombone@s:cym\&eng . //
\glsurface ALN:  ond mae gynno fo foy$^{C}_{E}$ yn chwarae trombone$^{C}_{E}$ .  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} with\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S} boy{\scriptsize .N.SG+SM} stative{\scriptsize .STAT} play{\scriptsize .V.INFIN} trombone{\scriptsize .N.SG}   //
\glmanual \%gls  but be{\scriptsize .3S.PRES} with{\scriptsize .3SM} PRON{\scriptsize .3SM} boy PRT play{\scriptsize .NONFIN} trombone   //
\gleng but he's got a guy playing trombone. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ hmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o:h@s:cym\&eng ! //
\glsurface ELE:  oh$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat um@s:cym\&eng mae o \&=sigh +. . . //
\glsurface ALN:  um$^{C}_{E}$ mae o .  //
\glauto \%aut  um{\scriptsize .IM} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM}   //
\gleng um, it's \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yn trad@s:cym\&eng yn lle bod gen ti un person@s:cym\&eng yn cymryd y melody@s:cym\&eng \# neu yn chwarae (y)r blaen +/ . //
\glsurface ALN:  yn trad$^{C}_{E}$ yn lle bod gen ti un person$^{C}_{E}$ yn cymryd y melody$^{C}_{E}$ neu yn chwarae yr blaen .  //
\glauto \%aut  yn{\scriptsize .PRT} trad{\scriptsize .N.SG} in{\scriptsize .PREP} where{\scriptsize .INT} be{\scriptsize .V.INFIN} by{\scriptsize .PREP.SPOKEN} you{\scriptsize .PRON.2S} one{\scriptsize .NUM} person{\scriptsize .N.SG} stative{\scriptsize .STAT} take{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} melody{\scriptsize .N.SG} or{\scriptsize .CONJ} stative{\scriptsize .STAT} play{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} front{\scriptsize .N.M.SG}   //
\glmanual \%gls  in trad in place be{\scriptsize .NONFIN} with{\scriptsize .2S} PRON{\scriptsize .2S} one person PRT take{\scriptsize .NONFIN} DET melody or PRT play{\scriptsize .nonfin} DET front   //
\gleng in trad, instead of having one person taking the melody, or playing the front \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, mae gen ti dau neu dri yn chwarae (y)r un pryd . //
\glsurface ALN:  mae gen ti dau neu dri yn chwarae yr un pryd .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} by{\scriptsize .PREP.SPOKEN} you{\scriptsize .PRON.2S} two{\scriptsize .NUM.M} or{\scriptsize .CONJ} three{\scriptsize .NUM.M+SM} stative{\scriptsize .STAT} play{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} one{\scriptsize .NUM} when{\scriptsize .INT[or]time.N.M.SG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} with{\scriptsize .2S} PRON{\scriptsize .2S} two{\scriptsize .M} or three{\scriptsize .M} PRT play{\scriptsize .NONFIN} DET one time   //
\gleng  \dots you have two or three playing at the same time. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng mae yn flêr (fe)lly . //
\glsurface ELE:  oh$^{C}_{E}$ mae yn flêr felly .  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} untidy{\scriptsize .ADJ+SM} so{\scriptsize .ADV}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} PRT messy thus   //
\gleng oh, it's a mess then //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a \&ma maen nhw (y)n cwffio (y)n\_erbyn ei\_gilydd yn gerddorol . //
\glsurface ALN:  a maen nhw yn cwffio yn\_erbyn ei\_gilydd yn gerddorol .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3P.PRES.SPOKEN} they{\scriptsize .PRON.3P} stative{\scriptsize .STAT} fight{\scriptsize .V.INFIN} against{\scriptsize .PREP[or]against.PRON} each\_other{\scriptsize .PRON.3SP} stative{\scriptsize .STAT} musical{\scriptsize .ADJ+SM}   //
\glmanual \%gls  and be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL} PRT fight{\scriptsize .NONFIN} against each\_other PRT musical   //
\gleng and they fight against each other musically. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a just@s:cym\&eng $<$yn meddwl$>$ [?] +"/ . //
\glsurface ALN:  a just$^{C}_{E}$ yn meddwl .  //
\glauto \%aut  and{\scriptsize .CONJ} just{\scriptsize .ADV[or]just.ADJ} stative{\scriptsize .STAT} think{\scriptsize .V.INFIN}   //
\glmanual \%gls  and just PRT think{\scriptsize .NONFIN}   //
\gleng and [you] just think: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" come@s:eng on@s:eng hogiau ! //
\glsurface ALN:  come$^{E}$ on$^{E}$ hogiau !  //
\glauto \%aut  come{\scriptsize .V.INFIN} on{\scriptsize .PREP} lads{\scriptsize .N.M.PL}   //
\glmanual \%gls  come on boys   //
\gleng "come on, boys!" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" get a grip ! //
\glsurface ALN: [-eng] get a grip !  //
\glauto \%aut [-eng] get{\scriptsize .V.INFIN} a{\scriptsize .DET.INDEF} grip{\scriptsize .N.SG}   //
\glmanual \%gls [-eng]      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng right@s:cym\&eng . //
\glsurface ELE:  oh$^{C}_{E}$ right$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM} right{\scriptsize .ADJ}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat os ti sefyll yn y cefn yn mynd \&=imit:acoustic\_bass er@s:cym\&eng dw (ddi)m yn meindio hynna . //
\glsurface ALN:  os ti sefyll yn y cefn yn mynd er$^{C}_{E}$ dw ddim yn meindio hynna .  //
\glauto \%aut  if{\scriptsize .CONJ} you{\scriptsize .PRON.2S} stand{\scriptsize .V.INFIN} in{\scriptsize .PREP} the{\scriptsize .DET.DEF} back{\scriptsize .N.M.SG} stative{\scriptsize .STAT} go{\scriptsize .V.INFIN} er{\scriptsize .IM} be{\scriptsize .V.1S.PRES.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} mind{\scriptsize .V.INFIN} that{\scriptsize .PRON.SP.SPOKEN}   //
\glmanual \%gls  if PRON{\scriptsize .2S} stand{\scriptsize .NONFIN} in DET back PRT go{\scriptsize .NONFIN} IM be{\scriptsize .1S.PRES} NEG PRT mind{\scriptsize .NONFIN} that   //
\gleng if you're standing at the back going "doo doo doo doo doo", I don't mind that. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (be)cause@s:eng \# timod \# mae o (y)n job@s:cym\&eng . //
\glsurface ALN:  because$^{E}$ timod mae o yn job$^{C}_{E}$ .  //
\glauto \%aut  because{\scriptsize .CONJ} know{\scriptsize .V.2S.PRES.SPOKEN} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} yn{\scriptsize .PRT} job{\scriptsize .N.SG}   //
\glmanual \%gls  because know{\scriptsize .2S} be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT job   //
\gleng cause, you know, it's a job. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a mae o (y)n fwy o hwyl na \#\# +/ . //
\glsurface ALN:  a mae o yn fwy o hwyl na .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} more{\scriptsize .ADJ.COMP+SM} of{\scriptsize .PREP} fun{\scriptsize .N.F.SG} neg{\scriptsize .PRT}   //
\glmanual \%gls  and be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT more of fun than   //
\gleng and it's more fun than \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ++ swydd [?] . //
\glsurface ELE:  swydd .  //
\glauto \%aut  job{\scriptsize .N.F.SG}   //
\glmanual \%gls  job   //
\gleng  \dots a job. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=groan dysgu Frank@s:cym\&eng . //
\glsurface ALN:  dysgu Frank$^{C}_{E}$ .  //
\glauto \%aut  teach{\scriptsize .V.INFIN} name   //
\glmanual \%gls  teach{\scriptsize .NONFIN} Frank   //
\gleng teaching Frank. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ahh@s:cym\&eng ! //
\glsurface ELE:  ahh$^{C}_{E}$ !  //
\glauto \%aut  ahh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond \# dw i (dd)im actu(ally)@s:eng [?] isio mynd i wrando ar y stuff@s:cym\&eng . //
\glsurface ALN:  ond dw i ddim actually$^{E}$ isio mynd i wrando ar y stuff$^{C}_{E}$ .  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} actual{\scriptsize .ADJ+LY} want{\scriptsize .N.M.SG} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} listen{\scriptsize .V.INFIN+SM} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} stuff{\scriptsize .SV.INFIN}   //
\glmanual \%gls  but be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} NEG actually want go{\scriptsize .NONFIN} to listen{\scriptsize .NONFIN} on DET stuff   //
\gleng but I don't actually want to go and listen to the stuff. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o\_k@s:cym\&eng . //
\glsurface ELE:  o\_k$^{C}_{E}$ .  //
\glauto \%aut  OK{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a (ba)sai Ed@s:cym\&eng yn gwybod na mynd i wrando i chwerthin faswn i anyway@s:eng . //
\glsurface ALN:  a basai Ed$^{C}_{E}$ yn gwybod na mynd i wrando i chwerthin faswn i anyway$^{E}$ .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PLUPERF.SPOKEN} name stative{\scriptsize .STAT} know{\scriptsize .V.INFIN} (n)or{\scriptsize .CONJ} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} listen{\scriptsize .V.INFIN+SM} I{\scriptsize .PRON.1S} laugh{\scriptsize .V.INFIN} be{\scriptsize .V.1S.PLUPERF.SPOKEN+SM} I{\scriptsize .PRON.1S} anyway{\scriptsize .ADV}   //
\glmanual \%gls  and be{\scriptsize .3S.CONDIT} Ed PRT know{\scriptsize .NONFIN} PRT go{\scriptsize .NONFIN} to listen{\scriptsize .NONFIN} to laugh{\scriptsize .NONFIN} be{\scriptsize .1S.CONDIT} PRON{\scriptsize .1S} anyway   //
\gleng and Ed would know that I was going to listen for a laugh anyway. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=gasp \# basai ? //
\glsurface ELE:  basai ?  //
\glauto \%aut  be{\scriptsize .V.3S.PLUPERF.SPOKEN}   //
\glmanual \%gls  be{\scriptsize .3S.CONDIT}   //
\gleng would he? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng basai ! //
\glsurface ALN:  oh$^{C}_{E}$ basai !  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PLUPERF.SPOKEN}   //
\glmanual \%gls  IM be{\scriptsize .3S.CONDIT}   //
\gleng oh, yes! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng right@s:cym\&eng o\_k@s:cym\&eng . //
\glsurface ELE:  oh$^{C}_{E}$ right$^{C}_{E}$ o\_k$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM} right{\scriptsize .ADJ} OK{\scriptsize .IM}   //
\glmanual \%gls       //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ oh@s:cym\&eng mae (y)n dallt ! //
\glsurface ALN:  oh$^{C}_{E}$ mae yn dallt !  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} understand{\scriptsize .V.INFIN.SPOKEN}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} PRT understand{\scriptsize .NONFIN}   //
\gleng he understands! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (yn)dy [=! quietly] . //
\glsurface ELE:  yndy .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na Emma@s:cym\&eng oedd yn gobeithio mynd $<$hefo (e)i$>$ [?] \# plant hi heddiw [?] . //
\glsurface ELE:  na Emma$^{C}_{E}$ oedd yn gobeithio mynd hefo ei plant hi heddiw .  //
\glauto \%aut  neg{\scriptsize .PRT} name be{\scriptsize .V.3S.IMPERF} stative{\scriptsize .STAT} hope{\scriptsize .V.INFIN} go{\scriptsize .V.INFIN} with{\scriptsize .PREP+H} her{\scriptsize .ADJ.POSS} children{\scriptsize .N.M.PL} she{\scriptsize .PRON.F.3S} today{\scriptsize .ADV}   //
\glmanual \%gls  no Emma be{\scriptsize .3S.IMP} PRT hope{\scriptsize .NONFIN} go{\scriptsize .NONFIN} with POSS{\scriptsize .3SF} children PRON{\scriptsize .3SF} today   //
\gleng no, Emma was hoping to go with her children today //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat Emma@s:cym\&eng +. . . //
\glsurface ALN:  Emma$^{C}_{E}$ .  //
\glauto \%aut  name   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (y)n ysgol . //
\glsurface ELE:  yn ysgol .  //
\glauto \%aut  yn{\scriptsize .PRT} school{\scriptsize .N.F.SG}   //
\glmanual \%gls  in school   //
\gleng at school. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&ko honna oedda chdi (y)n weld yn Galeri@s:cym\&eng ac oedda chdi byth yn sure@s:cym\&eng pwy oedd hi . //
\glsurface ELE:  honna oedda chdi yn weld yn Galeri$^{C}_{E}$ ac oedda chdi byth yn sure$^{C}_{E}$ pwy oedd hi .  //
\glauto \%aut  claim{\scriptsize .V.2S.IMPER} be{\scriptsize .V.2S.PAST.NORTH} you{\scriptsize .PRON} stative{\scriptsize .STAT} see{\scriptsize .V.INFIN+SM} in{\scriptsize .PREP} name and{\scriptsize .CONJ} be{\scriptsize .V.2S.PAST.NORTH} you{\scriptsize .PRON} forever{\scriptsize .ADV} stative{\scriptsize .STAT} sure{\scriptsize .ADJ} who{\scriptsize .PRON} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  that be{\scriptsize .2S.IMP} PRON{\scriptsize .2S} PRT see{\scriptsize .NONFIN} in Galeri and be{\scriptsize .2S.IMP} PRON{\scriptsize .2S} never PRT sure who be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF}   //
\gleng ?remember, the one you'd see in Galeri, and you were never sure who she was. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ac $<$oedd hi gweitied \&\textipa{@}$>$ [///] oedd ei phlant hi (y)n cael gwersi (y)na . //
\glsurface ELE:  ac oedd hi gweitied oedd ei phlant hi yn cael gwersi yna .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} wait{\scriptsize .V.INFIN} be{\scriptsize .V.3S.IMPERF} her{\scriptsize .ADJ.POSS} children{\scriptsize .N.M.PL+AM} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} have{\scriptsize .V.INFIN} lessons{\scriptsize .N.F.PL} there{\scriptsize .ADV}   //
\glmanual \%gls  and be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT wait{\scriptsize .NONFIN} be{\scriptsize .3S.IMP} POSS{\scriptsize .3SF} children PRON{\scriptsize .3SF} PRT have{\scriptsize .NONFIN} lessons there  //
\gleng and she was waiting. . . her children were having lessons there. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ac oedd hi (y)n gweld chdi (y)na . //
\glsurface ELE:  ac oedd hi yn gweld chdi yna .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} see{\scriptsize .V.INFIN} you{\scriptsize .PRON} there{\scriptsize .ADV}   //
\glmanual \%gls  and be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT see{\scriptsize .NONFIN} PRON{\scriptsize .2S} there   //
\gleng and she used to see you there. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond oedda chdi byth yn sure@s:cym\&eng +. . . //
\glsurface ELE:  ond oedda chdi byth yn sure$^{C}_{E}$ .  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.2S.PAST.NORTH} you{\scriptsize .PRON} forever{\scriptsize .ADV} stative{\scriptsize .STAT} sure{\scriptsize .ADJ}   //
\glmanual \%gls  but be{\scriptsize .2S.IMP} PRON{\scriptsize .2S} never PRT sure   //
\gleng but you were never sure \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae nhw (y)n [/] \# yn eu harddegau . //
\glsurface ELE:  mae nhw yn yn eu harddegau .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} they{\scriptsize .PRON.3P} yn{\scriptsize .PRT} yn{\scriptsize .PRT} their{\scriptsize .ADJ.POSS} teens{\scriptsize .N.M.PL+H}   //
\glmanual \%gls  be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL} PRT in POSS{\scriptsize .3PL} teens   //
\gleng they're in their teens. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat un [?] yn chwarae \# pres . //
\glsurface ELE:  un yn chwarae pres .  //
\glauto \%aut  one{\scriptsize .NUM} stative{\scriptsize .STAT} play{\scriptsize .V.INFIN} money{\scriptsize .N.M.SG}   //
\glmanual \%gls  one PRT play{\scriptsize .NONFIN} brass   //
\gleng one plays brass. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be mae (y)n dysgu yn Cae\_Top@s:cym\&eng ? //
\glsurface ALN:  be mae yn dysgu yn Cae\_Top$^{C}_{E}$ ?  //
\glauto \%aut  what{\scriptsize .INT} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} teach{\scriptsize .V.INFIN} in{\scriptsize .PREP} name   //
\glmanual \%gls  what be{\scriptsize .3S.PRES} PRT teach{\scriptsize .NONFIN} in Cae\_Top   //
\gleng what, she teaches at Cae Top? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yndy . //
\glsurface ELE:  yndy .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng oedd hi yn $<$party@s:cym\&eng Freda@s:cym\&eng$>$ [?] doedd ? //
\glsurface ELE:  oh$^{C}_{E}$ oedd hi yn party$^{C}_{E}$ Freda$^{C}_{E}$ doedd ?  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} yn{\scriptsize .PRT} party{\scriptsize .N.SG} name be{\scriptsize .V.3S.IMPERF.NEG}   //
\glmanual \%gls  IM be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} in party Freda be{\scriptsize .3S.IMP.NEG}   //
\gleng oh, she was at Freda's party, wasn't she? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat Graham@s:cym\&eng a \# Emma@s:cym\&eng . //
\glsurface ELE:  Graham$^{C}_{E}$ a Emma$^{C}_{E}$ .  //
\glauto \%aut  name and{\scriptsize .CONJ} name   //
\glmanual \%gls  Graham and Emma   //
\gleng Graham and Emma. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd um@s:cym\&eng \#\#\# \&=dental\_click \# wneud lot@s:cym\&eng efo beiciau mynydd a \#\#\# ballu . //
\glsurface ELE:  oedd um$^{C}_{E}$ wneud lot$^{C}_{E}$ efo beiciau mynydd a ballu .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} um{\scriptsize .IM} make{\scriptsize .V.INFIN+SM} lot{\scriptsize .N.SG} with{\scriptsize .PREP} bikes{\scriptsize .N.M.PL} mountain{\scriptsize .N.M.SG} and{\scriptsize .CONJ} refuse{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} IM do{\scriptsize .NONFIN} lot with bikes mountain and stuff   //
\gleng she used to, um, do a lot of mountain biking and. . . stuff. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ti (dd)im yn cofio ? //
\glsurface ELE:  ti ddim yn cofio ?  //
\glauto \%aut  you{\scriptsize .PRON.2S} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} remember{\scriptsize .V.INFIN}   //
\glmanual \%gls  PRON{\scriptsize .2S} NEG PRT remember{\scriptsize .NONFIN}   //
\gleng you don't remember? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae [?] (y)n canu cloch . //
\glsurface ALN:  mae yn canu cloch .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} sing{\scriptsize .V.INFIN} bell{\scriptsize .N.F.SG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRT sing{\scriptsize .NONFIN} bell   //
\gleng rings a bell. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat sure@s:cym\&eng (fa)swn i nabod nhw (peta)swn i gweld nhw . //
\glsurface ALN:  sure$^{C}_{E}$ faswn i nabod nhw petaswn i gweld nhw .  //
\glauto \%aut  sure{\scriptsize .ADJ} be{\scriptsize .V.1S.PLUPERF.SPOKEN+SM} I{\scriptsize .PRON.1S} know{\scriptsize .V.INFIN} they{\scriptsize .PRON.3P} be{\scriptsize .V.1S.PLUPERF.HYP} I{\scriptsize .PRON.1S} see{\scriptsize .V.INFIN} they{\scriptsize .PRON.3P}   //
\glmanual \%gls  sure be{\scriptsize .1S.CONDIT} PRON{\scriptsize .1S} know{\scriptsize .NONFIN} PRON{\scriptsize .3PL} if\_be{\scriptsize .1S.CONDIT} PRON{\scriptsize .1S} see{\scriptsize .NONFIN} PRON{\scriptsize .3PL}   //
\gleng sure I'd know them if I saw them. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ $<$well@s:cym\&eng oedd$>$ [?] hi (y)n gweld chdi \# pan oedda chdi (y)n wneud Cofis\_Bach@s:cym\&eng . //
\glsurface ELE:  well$^{C}_{E}$ oedd hi yn gweld chdi pan oedda chdi yn wneud Cofis\_Bach$^{C}_{E}$ .  //
\glauto \%aut  well{\scriptsize .ADV} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} see{\scriptsize .V.INFIN} you{\scriptsize .PRON} when{\scriptsize .CONJ} be{\scriptsize .V.2S.PAST.NORTH} you{\scriptsize .PRON} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM} name   //
\glmanual \%gls  well be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT see{\scriptsize .NONFIN} PRON{\scriptsize .2S} when be{\scriptsize .2S.IMP} PRON{\scriptsize .2S} PRT do{\scriptsize .NONFIN} Cofis\_Bach   //
\gleng she saw you while you were doing "Cofis Bach" . //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat right@s:cym\&eng . //
\glsurface ALN:  right$^{C}_{E}$ .  //
\glauto \%aut  right{\scriptsize .ADJ}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ac oedd hi yn y car@s:cym\&eng weithiau ac oedd plant yn cael gwersi \&=sniff . //
\glsurface ELE:  ac oedd hi yn y car$^{C}_{E}$ weithiau ac oedd plant yn cael gwersi .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} in{\scriptsize .PREP} the{\scriptsize .DET.DEF} car{\scriptsize .N.SG} times{\scriptsize .N.F.PL+SM} and{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} children{\scriptsize .N.M.PL} stative{\scriptsize .STAT} have{\scriptsize .V.INFIN} lessons{\scriptsize .N.F.PL}   //
\glmanual \%gls  and be{\scriptsize .3S.IMP} PRO{\scriptsize .3SF} in DET car sometimes and be{\scriptsize .3S.IMP} children PRT have{\scriptsize .NONFIN} lessons   //
\gleng she was in the car sometimes, and the children were having lessons. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat iawn . //
\glsurface ALN:  iawn .  //
\glauto \%aut  OK{\scriptsize .ADV}   //
\glmanual \%gls  alright   //
\gleng alright. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ac oedd hi +"/ . //
\glsurface ELE:  ac oedd hi .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  and be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF}   //
\gleng and she was: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" ah@s:cym\&eng weles i Alan@s:cym\&eng neithiwr ! //
\glsurface ELE:  ah$^{C}_{E}$ weles i Alan$^{C}_{E}$ neithiwr !  //
\glauto \%aut  ah{\scriptsize .IM} see{\scriptsize .V.1S.PAST.SPOKEN} to{\scriptsize .PREP} name last{\scriptsize .night.ADV}   //
\glmanual \%gls  IM see{\scriptsize .1S.PAST} PRON{\scriptsize .1S} Alan last\_night   //
\gleng "ooh, I saw Alan last night!" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat neu $<$weles i dy ŵr di neithiwr$>$ ["] . //
\glsurface ELE:  neu weles i dy ŵr di neithiwr .  //
\glauto \%aut  or{\scriptsize .CONJ} see{\scriptsize .V.1S.PAST.SPOKEN} I{\scriptsize .PRON.1S} your{\scriptsize .ADJ.POSS} man{\scriptsize .N.M.SG+SM} you{\scriptsize .PRON.2S+SM} last{\scriptsize .night.ADV}   //
\glmanual \%gls  or see{\scriptsize .1S.PAST} PRON{\scriptsize .1S} POSS{\scriptsize .2S} husband PRON{\scriptsize .2S} last\_night   //
\gleng or "I saw your husband last night." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng na $<$dw i s(ure)@s:cym\&eng$>$ [//] timod $<$dw i sure@s:cym\&eng$>$ [/] dw i sure@s:cym\&eng wneith hi fwynhau o . //
\glsurface ALN:  oh$^{C}_{E}$ na dw i sure$^{C}_{E}$ timod dw i sure$^{C}_{E}$ dw i sure$^{C}_{E}$ wneith hi fwynhau o .  //
\glauto \%aut  oh{\scriptsize .IM} neg{\scriptsize .PRT} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} sure{\scriptsize .ADJ} know{\scriptsize .V.2S.PRES.SPOKEN} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} sure{\scriptsize .ADJ} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} sure{\scriptsize .ADJ} do{\scriptsize .V.3S.FUT+SM} she{\scriptsize .PRON.F.3S} enjoy{\scriptsize .V.INFIN+SM} he{\scriptsize .PRON.M.3S.SPOKEN}   //
\glmanual \%gls  IM no be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} sure know{\scriptsize .2S} be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} sure be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} sure do{\scriptsize .3S.NONPAST} PRON{\scriptsize .3SF} enjoy{\scriptsize .NONFIN} PRON{\scriptsize .3SM}   //
\gleng oh, no, I'm sure, you know, I'm sure she'll enjoy it. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ hmm@s:cym\&eng mmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ mmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM} mmm{\scriptsize .IM}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i just@s:cym\&eng yn er@s:cym\&eng +. . . //
\glsurface ALN:  dw i just$^{C}_{E}$ yn er$^{C}_{E}$ .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} just{\scriptsize .ADV[or]just.ADJ} yn{\scriptsize .PRT} er{\scriptsize .IM}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} just PRT IM   //
\gleng I'm just, er \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ah@s:cym\&eng well@s:cym\&eng mae [/] mae (y)r plentyn newydd gychwyn yn band@s:cym\&eng hŷn . //
\glsurface ELE:  ah$^{C}_{E}$ well$^{C}_{E}$ mae mae yr plentyn newydd gychwyn yn band$^{C}_{E}$ hŷn .  //
\glauto \%aut  ah{\scriptsize .IM} well{\scriptsize .ADV} be{\scriptsize .V.3S.PRES} be{\scriptsize .V.3S.PRES} the{\scriptsize .DET.DEF} child{\scriptsize .N.M.SG} new{\scriptsize .ADJ} start{\scriptsize .V.INFIN+SM} yn{\scriptsize .PRT} band{\scriptsize .N.SG} older{\scriptsize .ADJ}   //
\glmanual \%gls  IM well be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES} DET child new start{\scriptsize .NONFIN} in band older   //
\gleng oh, well, the child has just started in the senior band. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ah@s:cym\&eng well@s:cym\&eng . //
\glsurface ALN:  ah$^{C}_{E}$ well$^{C}_{E}$ .  //
\glauto \%aut  ah{\scriptsize .IM} well{\scriptsize .ADV}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat fydd [/] fydd [/] fydd band@s:cym\&eng Beaumaris@s:cym\&eng [?] i\_gyd yna . //
\glsurface ALN:  fydd fydd fydd band$^{C}_{E}$ Beaumaris$^{C}_{E}$ i\_gyd yna .  //
\glauto \%aut  be{\scriptsize .V.3S.FUT+SM} be{\scriptsize .V.3S.FUT+SM} be{\scriptsize .V.3S.FUT+SM} band{\scriptsize .N.SG} name all{\scriptsize .ADJ} there{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.FUT} be{\scriptsize .3S.FUT} be{\scriptsize .3S.FUT} band Beaumaris all there   //
\gleng the Beaumaris band will all be there //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ Beaumaris@s:cym\&eng . //
\glsurface ELE:  Beaumaris$^{C}_{E}$ .  //
\glauto \%aut  name   //
\glmanual \%gls  Beaumaris   //
\gleng Beaumaris. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a: \# mae o (we)di gwirioni ar bersonoliaeth Ed@s:cym\&eng . //
\glsurface ELE:  a mae o wedi gwirioni ar bersonoliaeth Ed$^{C}_{E}$ .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} after{\scriptsize .PREP} dote{\scriptsize .V.INFIN} on{\scriptsize .PREP} personality{\scriptsize .N.F.SG+SM} name   //
\glmanual \%gls  and be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT{\scriptsize .PAST} become\_silly{\scriptsize .NONFIN} on personality Ed   //
\gleng and he's quite infatuated by Ed's personality. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&s mae [/] $<$mae (y)n \# un \# \&g$>$ [//] mae (y)n gymeriad tydy ? //
\glsurface ELE:  mae mae yn un mae yn gymeriad tydy ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} one{\scriptsize .NUM} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} character{\scriptsize .N.M.SG+SM} be{\scriptsize .V.3S.PRES.NEG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES} PRT one be{\scriptsize .3S.PRES} PRT character be{\scriptsize .3S.PRES.NEG}   //
\gleng he's quite a character, isn't he? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yndy yndy . //
\glsurface ALN:  yndy yndy .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.EMPH} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES}   //
\gleng he is, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat annwyl iawn . //
\glsurface ELE:  annwyl iawn .  //
\glauto \%aut  dear{\scriptsize .ADJ} very{\scriptsize .ADV}   //
\glmanual \%gls  dear very   //
\gleng really nice. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat fydd y crew@s:cym\&eng iau i\_gyd yna (y)n mynd +"/ . //
\glsurface ALN:  fydd y crew$^{C}_{E}$ iau i\_gyd yna yn mynd .  //
\glauto \%aut  be{\scriptsize .V.3S.FUT+SM} the{\scriptsize .DET.DEF} crew{\scriptsize .N.SG} liver{\scriptsize .N.M.SG[or]yoke.N.F.SG} all{\scriptsize .ADJ} there{\scriptsize .ADV} stative{\scriptsize .STAT} go{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .3S.FUT} DET crew younger all there PRT go{\scriptsize .NONFIN}   //
\gleng the younger lot are all there going: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" wow@s:cym\&eng Ed@s:cym\&eng ti (y)dy (y)r un ! //
\glsurface ALN:  wow$^{C}_{E}$ Ed$^{C}_{E}$ ti ydy yr un !  //
\glauto \%aut  wow{\scriptsize .IM} name you{\scriptsize .PRON.2S} be{\scriptsize .V.3S.PRES} the{\scriptsize .DET.DEF} one{\scriptsize .NUM}   //
\glmanual \%gls  IM Ed PRON{\scriptsize .2S} be{\scriptsize .3S.PRES} DET one   //
\gleng "wow, Ed, you are the man!" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ mmm@s:cym\&eng a:h@s:cym\&eng ! //
\glsurface ELE:  mmm$^{C}_{E}$ ah$^{C}_{E}$ !  //
\glauto \%aut  mmm{\scriptsize .IM} ah{\scriptsize .IM}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a fydd crew@s:cym\&eng hŷn i\_gyd yn sefyll a mynd +"/ . //
\glsurface ALN:  a fydd crew$^{C}_{E}$ hŷn i\_gyd yn sefyll a mynd .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.FUT+SM} crew{\scriptsize .N.SG} older{\scriptsize .ADJ} all{\scriptsize .ADJ} stative{\scriptsize .STAT} stand{\scriptsize .V.INFIN} and{\scriptsize .CONJ} go{\scriptsize .V.INFIN}   //
\glmanual \%gls  and be{\scriptsize .3S.FUT} crew older all PRT stand{\scriptsize .NONFIN} and go{\scriptsize .NONFIN}   //
\gleng and the older lot will all be standing around going: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" \&=snigger Ed@s:cym\&eng be ti (y)n wneud +! ? //
\glsurface ALN:  Ed$^{C}_{E}$ be ti yn wneud ?  //
\glauto \%aut  name what{\scriptsize .INT} you{\scriptsize .PRON.2S} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  Ed what PRON{\scriptsize .2S} PRT do{\scriptsize .NONFIN}   //
\gleng "ha, Ed, what are you doing!?"" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng case@s:cym\&eng ! //
\glsurface ELE:  mmm$^{C}_{E}$ case$^{C}_{E}$ !  //
\glauto \%aut  mmm{\scriptsize .IM} case{\scriptsize .N.SG}   //
\glmanual \%gls      //
\gleng oh, what a case! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat lovely ! //
\glsurface ELE: [-eng] lovely !  //
\glauto \%aut [-eng] love{\scriptsize .V.INFIN+ADV}   //
\glmanual \%gls [-eng]    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond er@s:cym\&eng +/ . //
\glsurface ALN:  ond er$^{C}_{E}$ .  //
\glauto \%aut  but{\scriptsize .CONJ} er{\scriptsize .IM}   //
\glmanual \%gls  but IM   //
\gleng but, er \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng $<$na (ba)swn i licio$>$ [?] cael \# um@s:cym\&eng gwybod pa gaster@s:cym\&eng i roid ar y piano@s:cym\&eng . //
\glsurface ELE:  yeah$^{C}_{E}$ na baswn i licio cael um$^{C}_{E}$ gwybod pa gaster$^{C}_{E}$ i roid ar y piano$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV} neg{\scriptsize .PRT} be{\scriptsize .V.1S.PLUPERF.SPOKEN} I{\scriptsize .PRON.1S} like{\scriptsize .V.INFIN} have{\scriptsize .V.INFIN} um{\scriptsize .IM} know{\scriptsize .V.INFIN} which{\scriptsize .ADJ} gaster{\scriptsize .N.SG[or]caster.N.SG+SM} to{\scriptsize .PREP} give{\scriptsize .V.0.IMPERF+SM} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} piano{\scriptsize .N.SG}   //
\glmanual \%gls  yeah no be{\scriptsize .1S.CONDIT} PRON{\scriptsize .1S} like{\scriptsize .NONFIN} get{\scriptsize .NONFIN} IM know{\scriptsize .NONFIN} which caster to put{\scriptsize .NONFIN} on DET piano   //
\gleng yeah, I'd like to know which caster to put on the piano. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (be)cause@s:eng dw i gallu dynnu fo allan de ? //
\glsurface ELE:  because$^{E}$ dw i gallu dynnu fo allan de ?  //
\glauto \%aut  because{\scriptsize .CONJ} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} be\_able{\scriptsize .V.INFIN} take{\scriptsize .V.INFIN+SM} he{\scriptsize .PRON.M.3S} out{\scriptsize .ADV} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  because be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} can{\scriptsize .NONFIN} pull{\scriptsize .NONFIN} PRON{\scriptsize .3SM} out TAG   //
\gleng cause I can pull it out, right? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a wedyn mae Hugh@s:cym\&eng yn gorod gario fo nôl i\_fewn . //
\glsurface ELE:  a wedyn mae Hugh$^{C}_{E}$ yn gorod gario fo nôl i\_fewn .  //
\glauto \%aut  and{\scriptsize .CONJ} afterwards{\scriptsize .ADV} be{\scriptsize .V.3S.PRES} name stative{\scriptsize .STAT} have\_to{\scriptsize .V.INFIN} carry{\scriptsize .V.INFIN+SM} he{\scriptsize .PRON.M.3S} fetch{\scriptsize .V.INFIN} in{\scriptsize .ADV}   //
\glmanual \%gls  and then be{\scriptsize .3S.PRES} Hugh PRT must{\scriptsize .NONFIN} carry{\scriptsize .NONFIN} PRON{\scriptsize .3SM} back in   //
\gleng and then Hugh has to carry it back in. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat piano@s:cym\&eng (y)n drwm (dy)dy ? //
\glsurface ELE:  piano$^{C}_{E}$ yn drwm dydy ?  //
\glauto \%aut  piano{\scriptsize .N.SG} stative{\scriptsize .STAT} heavy{\scriptsize .ADJ+SM} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM}   //
\glmanual \%gls  piano PRT heavy be{\scriptsize .3S.PRES.NEG}   //
\gleng piano's heavy, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (y)dy Hugh@s:cym\&eng (y)n hogyn mawr cry (fe)lly ? //
\glsurface ALN:  ydy Hugh$^{C}_{E}$ yn hogyn mawr cry felly ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} name yn{\scriptsize .PRT} lad{\scriptsize .N.M.SG} big{\scriptsize .ADJ} strong{\scriptsize .ADJ} so{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} Hugh PRT boy big strong thus   //
\gleng is Hugh a big strong lad then? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng rugby@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ rugby$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM} rugby{\scriptsize .N.SG}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat rugby@s:cym\&eng ac [!] amaethu [?] ? //
\glsurface ALN:  rugby$^{C}_{E}$ ac amaethu ?  //
\glauto \%aut  rugby{\scriptsize .N.SG} and{\scriptsize .CONJ} cultivate{\scriptsize .V.INFIN}   //
\glmanual \%gls  rugby and farming   //
\gleng rugby AND farming? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie . //
\glsurface ELE:  ie .  //
\glauto \%aut  yes{\scriptsize .ADV}   //
\glmanual \%gls  yes   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae dipyn o gyfuniad dydy ? //
\glsurface ALN:  mae dipyn o gyfuniad dydy ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} little{\scriptsize .bit.N.M.SG+SM} of{\scriptsize .PREP} combination{\scriptsize .N.M.SG+SM} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} a\_bit of combination be{\scriptsize .3S.PRES.NEG}   //
\gleng that's quite a combination, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yndy . //
\glsurface ELE:  yndy .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a dw i meddwl bod o (y)n ddigon dewr i ddod efo ni $<$nos Sad(wrn)$>$ [//] nos Wener . //
\glsurface ELE:  a dw i meddwl bod o yn ddigon dewr i ddod efo ni nos Sadwrn nos Wener .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} think{\scriptsize .V.INFIN} be{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} enough{\scriptsize .QUAN+SM} brave{\scriptsize .ADJ} to{\scriptsize .PREP} come{\scriptsize .V.INFIN+SM} with{\scriptsize .PREP} we{\scriptsize .PRON.1P} night{\scriptsize .N.F.SG} Saturn{\scriptsize .N.M.SG} night{\scriptsize .N.F.SG} name   //
\glmanual \%gls  and be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} think{\scriptsize .NONFIN} be{\scriptsize .NONFIN} PRON{\scriptsize .3SM} PRT enough brave to come{\scriptsize .NONFIN} with PRON{\scriptsize .1PL} night Saturday night Friday   //
\gleng and I think he's brave enough to come with us on Friday night. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be (y)dy o mynd allan efo chi ? //
\glsurface ALN:  be ydy o mynd allan efo chi ?  //
\glauto \%aut  what{\scriptsize .INT} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} go{\scriptsize .V.INFIN} out{\scriptsize .ADV} with{\scriptsize .PREP} you{\scriptsize .PRON.2P}   //
\glmanual \%gls  what be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} go{\scriptsize .NONFIN} out with PRON{\scriptsize .2PL}   //
\gleng what, is he going out with you? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well@s:cym\&eng y syniad oedd +"/ . //
\glsurface ELE:  well$^{C}_{E}$ y syniad oedd .  //
\glauto \%aut  well{\scriptsize .ADV} the{\scriptsize .DET.DEF} idea{\scriptsize .N.M.SG} be{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  well DET idea be{\scriptsize .3S.IMP}   //
\gleng well, the idea was: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng oedd [?] Ian@s:cym\&eng am ddod \# am y tro cynta ers\_talwm iawn xxx . //
\glsurface ELE:  oh$^{C}_{E}$ oedd Ian$^{C}_{E}$ am ddod am y tro cynta ers\_talwm iawn .  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.IMPERF} name for{\scriptsize .PREP} come{\scriptsize .V.INFIN+SM} for{\scriptsize .PREP} the{\scriptsize .DET.DEF} turn{\scriptsize .N.M.SG} first{\scriptsize .ORD} for\_some\_time{\scriptsize .ADV} OK{\scriptsize .ADV}   //
\glmanual \%gls  IM be{\scriptsize .3S.IMP} Ian for come{\scriptsize .NONFIN} for DET time first for\_a\_while very   //
\gleng oh, Ian was going to come, for the first time in a very long while. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ a mae Ian@s:cym\&eng $<$yn jibio$>$ [?] yndy ? //
\glsurface ALN:  a mae Ian$^{C}_{E}$ yn jibio yndy ?  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} name stative{\scriptsize .STAT} jib{\scriptsize .V.INFIN} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  and be{\scriptsize .3S.PRES} Ian PRT jib{\scriptsize .NONFIN} PRT be{\scriptsize .3S.PRES}  //
\gleng and Ian is chickening out, is he? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na mae \# Jonny@s:cym\&eng yn gorod cael tynnu dannedd (y)na i\_gyd . //
\glsurface ELE:  na mae Jonny$^{C}_{E}$ yn gorod cael tynnu dannedd yna i\_gyd .  //
\glauto \%aut  neg{\scriptsize .PRT} be{\scriptsize .V.3S.PRES} name stative{\scriptsize .STAT} have\_to{\scriptsize .V.INFIN} have{\scriptsize .V.INFIN} take{\scriptsize .V.INFIN} tooth{\scriptsize .N.M.PL} there{\scriptsize .ADV} all{\scriptsize .ADJ}   //
\glmanual \%gls  no be{\scriptsize .3S.PRES} Jonny PRT must{\scriptsize .NONFIN} have{\scriptsize .NONFIN} pull{\scriptsize .NONFIN} teeth there all   //
\gleng no, Jonny's got to have some teeth pulled out, that's all. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng yeah@s:cym\&eng . //
\glsurface ALN:  oh$^{C}_{E}$ yeah$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM} yeah{\scriptsize .ADV}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=click wedyn +/ . //
\glsurface ELE:  wedyn .  //
\glauto \%aut  afterwards{\scriptsize .ADV}   //
\glmanual \%gls  then   //
\gleng then \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o\_k@s:cym\&eng well@s:cym\&eng o\_k@s:cym\&eng mae hynna ddigon teg . //
\glsurface ALN:  o\_k$^{C}_{E}$ well$^{C}_{E}$ o\_k$^{C}_{E}$ mae hynna ddigon teg .  //
\glauto \%aut  OK{\scriptsize .IM} well{\scriptsize .ADV} OK{\scriptsize .IM} be{\scriptsize .V.3S.PRES} that{\scriptsize .PRON.SP.SPOKEN} enough{\scriptsize .QUAN+SM} fair{\scriptsize .ADJ}   //
\glmanual \%gls  o\_k well o\_k be{\scriptsize .3S.PRES} that enough fair   //
\gleng ok, well, that's fair enough. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat xxx . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae lot@s:cym\&eng o ddannedd dydy ? //
\glsurface ELE:  mae lot$^{C}_{E}$ o ddannedd dydy ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} lot{\scriptsize .N.SG} of{\scriptsize .PREP} tooth{\scriptsize .N.M.PL+SM} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} lot of teeth be{\scriptsize .3S.PRES.NEG}   //
\gleng that's a lot of teeth, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae hynna (y)n lot@s:cym\&eng o ddannedd . //
\glsurface ALN:  mae hynna yn lot$^{C}_{E}$ o ddannedd .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} that{\scriptsize .PRON.SP.SPOKEN} yn{\scriptsize .PRT} lot{\scriptsize .N.SG} of{\scriptsize .PREP} tooth{\scriptsize .N.M.PL+SM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} that PRT lot of teeth   //
\gleng that is a lot of teeth. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o'n i (y)n sâl ar\_ôl un ! //
\glsurface ALN:  o'n i yn sâl ar\_ôl un !  //
\glauto \%aut  unk to{\scriptsize .PREP} stative{\scriptsize .STAT} ill{\scriptsize .ADJ} after{\scriptsize .PREP} one{\scriptsize .NUM}   //
\glmanual \%gls  be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} PRT sick after one   //
\gleng I was sick after one! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=laugh ie dyn ti de ! //
\glsurface ELE:  ie dyn ti de !  //
\glauto \%aut  yes{\scriptsize .ADV} man{\scriptsize .N.M.SG[or]draw.V.3S.PRES+SM} you{\scriptsize .PRON.2S} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  yes man PRON{\scriptsize .2S} TAG   //
\gleng yes, you're a man, aren't you! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat er dw i meddwl na (y)r annwyd oedd hynna sti ? //
\glsurface ALN:  er dw i meddwl na yr annwyd oedd hynna sti ?  //
\glauto \%aut  er{\scriptsize .IM} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} think{\scriptsize .V.INFIN} than{\scriptsize .CONJ} the{\scriptsize .DET.DEF} cold{\scriptsize .N.M.SG} be{\scriptsize .V.3S.IMPERF} that{\scriptsize .PRON.SP.SPOKEN} you\_know{\scriptsize .IM}   //
\glmanual \%gls  although be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} think{\scriptsize .NONFIN} PRT DET cold be{\scriptsize .1S.IMP} that know{\scriptsize .2S}   //
\gleng although, I think that was my cold, you know? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat i:e ie ! //
\glsurface ELE:  ie ie !  //
\glauto \%aut  yes{\scriptsize .ADV} yes{\scriptsize .ADV}   //
\glmanual \%gls  yes yes   //
\gleng yes, yes! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (be)cause@s:eng oedd [/] oedd [/] oedd Ken@s:cym\&eng yn gwaith heddiw . //
\glsurface ALN:  because$^{E}$ oedd oedd oedd Ken$^{C}_{E}$ yn gwaith heddiw .  //
\glauto \%aut  because{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} be{\scriptsize .V.3S.IMPERF} be{\scriptsize .V.3S.IMPERF} name yn{\scriptsize .PRT} work{\scriptsize .N.M.SG} today{\scriptsize .ADV}   //
\glmanual \%gls  because be{\scriptsize .3S.IMP} be{\scriptsize .3S.IMP} be{\scriptsize .3S.IMP} Ken in work today   //
\gleng cause Ken was at work today. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat i::e xxx ! //
\glsurface ELE:  ie !  //
\glauto \%aut  yes{\scriptsize .ADV}   //
\glmanual \%gls  yes   //
\gleng yeeeees!! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie deud ti ! //
\glsurface ALN:  ie deud ti !  //
\glauto \%aut  yes{\scriptsize .ADV} say{\scriptsize .V.INFIN.SPOKEN} you{\scriptsize .PRON.2S}   //
\glmanual \%gls  yes say{\scriptsize .2S.IMP} PRON{\scriptsize .2S}   //
\gleng yes, go on! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd Ken@s:cym\&eng heddiw yn deud oedd o (y)n teimlo (y)n grim@s:eng . //
\glsurface ALN:  oedd Ken$^{C}_{E}$ heddiw yn deud oedd o yn teimlo yn grim$^{E}$ .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} name today{\scriptsize .ADV} stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN} be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} feel{\scriptsize .V.INFIN} stative{\scriptsize .STAT} grim{\scriptsize .ADJ}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} Ken today PRT say{\scriptsize .NONFIN} be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} PRT feel{\scriptsize .NONFIN} PRT grim   //
\gleng Ken, today, was saying he felt grim. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a::h@s:cym\&eng ! //
\glsurface ELE:  ah$^{C}_{E}$ !  //
\glauto \%aut  ah{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ ac oedd o (we)di cysgu ar y llawr neithiwr . //
\glsurface ALN:  ac oedd o wedi cysgu ar y llawr neithiwr .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} after{\scriptsize .PREP} sleep{\scriptsize .V.INFIN} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} floor{\scriptsize .N.M.SG[or]ground.N.M.SG} last{\scriptsize .night.ADV}   //
\glmanual \%gls  and be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} PRT{\scriptsize .PAST} sleep{\scriptsize .NONFIN} on DET floor last\_night   //
\gleng and he'd slept on the floor last night. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a:::h@s:cym\&eng ! //
\glsurface ELE:  ah$^{C}_{E}$ !  //
\glauto \%aut  ah{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat so@s:cym\&eng $<$oedd gen Gail@s:cym\&eng$>$ [/] \# oedd gen Gail@s:cym\&eng \# wres o \# tri\_deg\_naw pwynt pedwar . //
\glsurface ALN:  so$^{C}_{E}$ oedd gen Gail$^{C}_{E}$ oedd gen Gail$^{C}_{E}$ wres o tri\_deg\_naw pwynt pedwar .  //
\glauto \%aut  so{\scriptsize .ADV} be{\scriptsize .V.3S.IMPERF} by{\scriptsize .PREP.SPOKEN} name be{\scriptsize .V.3S.IMPERF} by{\scriptsize .PREP.SPOKEN} name warmth{\scriptsize .N.M.SG+SM} he{\scriptsize .PRON.M.3S.SPOKEN} unk point{\scriptsize .N.M.SG} four{\scriptsize .NUM.M}   //
\glmanual \%gls  so be{\scriptsize .3S.IMP} with Gail be{\scriptsize .3S.IMP} with Gail heat of thirty\_nine point four   //
\gleng so Gail had a temperature of 39 .4 . //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ \&=laugh \# oh@s:cym\&eng na ! //
\glsurface ELE:  oh$^{C}_{E}$ na !  //
\glauto \%aut  oh{\scriptsize .IM} neg{\scriptsize .PRT}   //
\glmanual \%gls  IM no   //
\gleng haha. . . oh, no! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat gosh@s:eng be (y)dy hwnna mewn Fahrenheit@s:cym\&eng ? //
\glsurface ELE:  gosh$^{E}$ be ydy hwnna mewn Fahrenheit$^{C}_{E}$ ?  //
\glauto \%aut  gosh{\scriptsize .IM} what{\scriptsize .INT} be{\scriptsize .V.3S.PRES} that{\scriptsize .PRON.M.SG.SPOKEN} in{\scriptsize .PREP} name   //
\glmanual \%gls  gosh what be{\scriptsize .3S.PRES} that in Fahrenheit   //
\gleng gosh, what's that in Fahrenheit? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw (ddi)m yn (gwy)bod . //
\glsurface ALN:  dw ddim yn gwybod .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} NEG PRT know{\scriptsize .NONFIN}   //
\gleng I don't know. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond oedd o (ddi)m dros pedwar\_deg a dyna lle ti dechrau poeni . //
\glsurface ALN:  ond oedd o ddim dros pedwar\_deg a dyna lle ti dechrau poeni .  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} not{\scriptsize .ADV+SM} over{\scriptsize .PREP+SM} forty{\scriptsize .NUM} and{\scriptsize .CONJ} that{\scriptsize .is.ADV} where{\scriptsize .INT[or]place.N.M.SG} you{\scriptsize .PRON.2S} begin{\scriptsize .V.INFIN} worry{\scriptsize .V.INFIN}   //
\glmanual \%gls  but be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} NEG over forty and there where PRON{\scriptsize .2S} start{\scriptsize .NONFIN} worry{\scriptsize .NONFIN}   //
\gleng but it wasn't over forty, and that's where you start to worry. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=dental\_click . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond aethon nhw â hi i (y)sbyty . //
\glsurface ALN:  ond aethon nhw â hi i ysbyty .  //
\glauto \%aut  but{\scriptsize .CONJ} go{\scriptsize .V.3P.PAST.SPOKEN} they{\scriptsize .PRON.3P} with{\scriptsize .PREP} she{\scriptsize .PRON.F.3S} to{\scriptsize .PREP} hospital{\scriptsize .N.M.SG}   //
\glmanual \%gls  but go{\scriptsize .3PL.PAST} PRON{\scriptsize .3PL} with PRON{\scriptsize .3SF} to hospital   //
\gleng but they took her to hospital. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat gosh@s:eng mae (y)n uchel yndy ? //
\glsurface ELE:  gosh$^{E}$ mae yn uchel yndy ?  //
\glauto \%aut  gosh{\scriptsize .IM} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} high{\scriptsize .ADJ} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  gosh be{\scriptsize .3S.PRES} PRT high be{\scriptsize .3S.PRES}   //
\gleng gosh, that's high, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be oedd ta ? //
\glsurface ELE:  be oedd ta ?  //
\glauto \%aut  what{\scriptsize .INT} be{\scriptsize .V.3S.IMPERF} or{\scriptsize .CONJ}   //
\glmanual \%gls  what be{\scriptsize .3S.IMP} then   //
\gleng what was it, then? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat annwyd ? //
\glsurface ELE:  annwyd ?  //
\glauto \%aut  cold{\scriptsize .N.M.SG}   //
\glmanual \%gls  cold   //
\gleng a cold? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=slurp dw (ddi)m yn (gwy)bod . //
\glsurface ALN:  dw ddim yn gwybod .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} NEG PRT know{\scriptsize .NONFIN}   //
\gleng I don't know //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ryw fath o \# +/ . //
\glsurface ALN:  ryw fath o .  //
\glauto \%aut  some{\scriptsize .PREQ+SM} type{\scriptsize .N.M.SG+SM[or]bath.N.M.SG+SM} he{\scriptsize .PRON.M.3S.SPOKEN}   //
\glmanual \%gls  some kind of   //
\gleng some kind of \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ++ virus@s:cym\&eng \&=cough ? //
\glsurface ELE:  virus$^{C}_{E}$ ?  //
\glauto \%aut  virus{\scriptsize .N.SG}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, dwymyn . //
\glsurface ALN:  dwymyn .  //
\glauto \%aut  fever{\scriptsize .N.F.SG+SM}   //
\glmanual \%gls  fever   //
\gleng  \dots fever. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond oedden nhw ar yr \#\# Calpol@s:cym\&eng a \# Calprofen a +. . . //
\glsurface ALN:  ond oedden nhw ar yr Calpol$^{C}_{E}$ a Calprofen a .  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.3P.IMPERF.SPOKEN} they{\scriptsize .PRON.3P} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} name and{\scriptsize .CONJ} name and{\scriptsize .CONJ}   //
\glmanual \%gls  but be{\scriptsize .3PL.IMP} PRON{\scriptsize .3PL} on DET Calpol and Calprofen and   //
\gleng but they were on the Calpol and Calprofen and \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng bechod ! //
\glsurface ELE:  oh$^{C}_{E}$ bechod !  //
\glauto \%aut  oh{\scriptsize .IM} how{\scriptsize .sad.IM[or]sin.N.M.SG+SM}   //
\glmanual \%gls  IM sin   //
\gleng oh, poor them! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd hi (we)di dychryn ? //
\glsurface ELE:  oedd hi wedi dychryn ?  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} after{\scriptsize .PREP} frighten{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT{\scriptsize .PAST} frighten{\scriptsize .NONFIN}   //
\gleng were they frightened? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ \&=sniff \&=slurp . //
\glsurface ALN:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw (ddi)m [?] me(ddw)l bod nhw (y)n or hapus . //
\glsurface ALN:  dw ddim meddwl bod nhw yn or hapus .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} not{\scriptsize .ADV+SM} think{\scriptsize .V.INFIN} be{\scriptsize .V.INFIN} they{\scriptsize .PRON.3P} yn{\scriptsize .PRT} over-.PRT+SM happy{\scriptsize .ADJ}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} NEG think{\scriptsize .NONFIN} be{\scriptsize .NONFIN} PRON{\scriptsize .3PL} PRT over happy   //
\gleng I don't think they were too happy. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ be (y)dy hoed hi wan ? //
\glsurface ELE:  be ydy hoed hi wan ?  //
\glauto \%aut  what{\scriptsize .INT} be{\scriptsize .V.3S.PRES} age{\scriptsize .N.M.SG+H} she{\scriptsize .PRON.F.3S} weak{\scriptsize .ADJ+SM[or]pale.ADJ+SM}   //
\glmanual \%gls  what be{\scriptsize .3S.PRES} age PRON{\scriptsize .3SF} now   //
\gleng how old is she now? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae Gail@s:cym\&eng yn \# +/ . //
\glsurface ALN:  mae Gail$^{C}_{E}$ yn .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} name yn{\scriptsize .PRT}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} Gail PRT   //
\gleng Gail is \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ydy naw mis (rhyw)beth felly ? //
\glsurface ELE:  ydy naw mis rhywbeth felly ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} nine{\scriptsize .NUM} month{\scriptsize .N.M.SG} something{\scriptsize .N.M.SG} so{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} nine month something thus   //
\gleng is she nine months, something like that? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (rhyw)beth felly yndy . //
\glsurface ALN:  rhywbeth felly yndy .  //
\glauto \%aut  something{\scriptsize .N.M.SG} so{\scriptsize .ADV} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  something thus be{\scriptsize .3S.PRES}   //
\gleng something like that, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat so@s:cym\&eng wnaeth Ken@s:cym\&eng gysgu ar y llawr yn y stafell fyw . //
\glsurface ALN:  so$^{C}_{E}$ wnaeth Ken$^{C}_{E}$ gysgu ar y llawr yn y stafell fyw .  //
\glauto \%aut  so{\scriptsize .ADV} do{\scriptsize .V.3S.PAST+SM} name sleep{\scriptsize .V.INFIN+SM} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} floor{\scriptsize .N.M.SG[or]ground.N.M.SG} in{\scriptsize .PREP} the{\scriptsize .DET.DEF} room{\scriptsize .N.F.SG} live{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  so do{\scriptsize .3S.PAST} Ken sleep{\scriptsize .NONFIN} on DET floor in DET room living   //
\gleng so Ken slept on the floor in the living room. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wnaeth [?] Gail@s:cym\&eng cysgu efo \# Hayley@s:cym\&eng . //
\glsurface ALN:  wnaeth Gail$^{C}_{E}$ cysgu efo Hayley$^{C}_{E}$ .  //
\glauto \%aut  do{\scriptsize .V.3S.PAST+SM} name sleep{\scriptsize .V.INFIN} with{\scriptsize .PREP} name   //
\glmanual \%gls  do{\scriptsize .3S.PAST} Gail sleep{\scriptsize .NONFIN} with Hayley   //
\gleng Gail slept with Hayley. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a:h@s:cym\&eng wela i . //
\glsurface ELE:  ah$^{C}_{E}$ wela i .  //
\glauto \%aut  ah{\scriptsize .IM} see{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S}   //
\glmanual \%gls  IM see{\scriptsize .1S.NONPAST} PRON{\scriptsize .1S}   //
\gleng oh, I see. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a (we)dyn oedd hi sâl yn gwely . //
\glsurface ALN:  a wedyn oedd hi sâl yn gwely .  //
\glauto \%aut  and{\scriptsize .CONJ} afterwards{\scriptsize .ADV} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} ill{\scriptsize .ADJ} yn{\scriptsize .PRT} bed{\scriptsize .N.M.SG}   //
\glmanual \%gls  and then be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} sick in bed   //
\gleng and then she was sick in bed. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng na ! //
\glsurface ELE:  oh$^{C}_{E}$ na !  //
\glauto \%aut  oh{\scriptsize .IM} neg{\scriptsize .PRT}   //
\glmanual \%gls  IM no   //
\gleng oh, no! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ideal@s:cym\&eng o:h@s:cym\&eng ! //
\glsurface ALN:  ideal$^{C}_{E}$ oh$^{C}_{E}$ !  //
\glauto \%aut  ideal{\scriptsize .N.SG} oh{\scriptsize .IM}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng been@s:eng there@s:eng done@s:eng that@s:eng yeah@s:cym\&eng . //
\glsurface ELE:  oh$^{C}_{E}$ been$^{E}$ there$^{E}$ done$^{E}$ that$^{E}$ yeah$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM} been{\scriptsize .V.PASTPART} there{\scriptsize .ADV} done{\scriptsize .AV.PAST} that{\scriptsize .DEM.FAR[or]that.CONJ} yeah{\scriptsize .ADV}   //
\glmanual \%gls          //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=dental\_click melys melys goffadwriaeth te ? //
\glsurface ALN:  melys melys goffadwriaeth te ?  //
\glauto \%aut  sweet{\scriptsize .ADJ} sweet{\scriptsize .ADJ} memory{\scriptsize .N.F.SG+SM} tea{\scriptsize .N.M.SG}   //
\glmanual \%gls  sweet sweet remembrance TAG   //
\gleng sweet, sweet memories, eh? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng bechod ! //
\glsurface ELE:  oh$^{C}_{E}$ bechod !  //
\glauto \%aut  oh{\scriptsize .IM} how{\scriptsize .sad.IM[or]sin.N.M.SG+SM}   //
\glmanual \%gls  IM sin   //
\gleng oh, poor them! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" Mam@s:cym\&eng ["] \&di \&=imit:vomitting ! //
\glsurface ALN:  Mam$^{C}_{E}$ !  //
\glauto \%aut  name   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" oh@s:cym\&eng na ! //
\glsurface ALN:  oh$^{C}_{E}$ na !  //
\glauto \%aut  oh{\scriptsize .IM} neg{\scriptsize .PRT}   //
\glmanual \%gls  IM no   //
\gleng "oh, no!" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=laugh ry hwyr \&=laugh ! //
\glsurface ELE:  ry hwyr !  //
\glauto \%aut  too{\scriptsize .ADJ+SM} late{\scriptsize .ADJ}   //
\glmanual \%gls  too late   //
\gleng too late! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ rhy hwyr ! //
\glsurface ALN:  rhy hwyr !  //
\glauto \%aut  too{\scriptsize .ADJ} late{\scriptsize .ADJ}   //
\glmanual \%gls  too late   //
\gleng too late! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" oh@s:cym\&eng alla i gadw hyn ar y \# gobennydd . //
\glsurface ALN:  oh$^{C}_{E}$ alla i gadw hyn ar y gobennydd .  //
\glauto \%aut  oh{\scriptsize .IM} be\_able{\scriptsize .V.3S.PRES.SPOKEN+SM} I{\scriptsize .PRON.1S[or]to.PREP} keep{\scriptsize .V.INFIN+SM} this{\scriptsize .PRON.SP} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} pillow{\scriptsize .N.M.SG}   //
\glmanual \%gls  IM can{\scriptsize .1S.NONPAST} PRON{\scriptsize .1S} keep{\scriptsize .NONFIN} this on DET pillow   //
\gleng "oh, I can keep this on the pillow." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ o:h@s:cym\&eng dear@s:cym\&eng [?] ! //
\glsurface ELE:  oh$^{C}_{E}$ dear$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM} dear{\scriptsize .ADJ[or]tear.N.SG+SM}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" oh@s:cym\&eng na o:h@s:cym\&eng ! //
\glsurface ALN:  oh$^{C}_{E}$ na oh$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM} neg{\scriptsize .PRT} oh{\scriptsize .IM}   //
\glmanual \%gls  IM no IM   //
\gleng oh, no, eurgh! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond nad [=? na] oedd Ken@s:cym\&eng +. . . //
\glsurface ALN:  ond nad oedd Ken$^{C}_{E}$ .  //
\glauto \%aut  but{\scriptsize .CONJ} who\_not{\scriptsize .PRON.REL.NEG[or]carve.V.3S.PRES} be{\scriptsize .V.3S.IMPERF} name   //
\glmanual \%gls  but no be{\scriptsize .3S.IMP} Ken   //
\gleng but Ken was \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat just@s:cym\&eng \# y bore (y)ma \# ar\_draws yr awr ginio oedd o just@s:cym\&eng yn [/] \#\# yn +// . //
\glsurface ALN:  just$^{C}_{E}$ y bore yma ar\_draws yr awr ginio oedd o just$^{C}_{E}$ yn yn .  //
\glauto \%aut  just{\scriptsize .ADV[or]just.ADJ} the{\scriptsize .DET.DEF} morning{\scriptsize .N.M.SG} here{\scriptsize .ADV} across{\scriptsize .PREP} the{\scriptsize .DET.DEF} hour{\scriptsize .N.F.SG} dinner{\scriptsize .N.M.SG+SM} be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} just{\scriptsize .ADV[or]just.ADJ} yn{\scriptsize .PRT} yn{\scriptsize .PRT}   //
\glmanual \%gls  just DET morning here over DET hour lunch be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} just PRT PRT   //
\gleng just, this morning over the lunch hour he was just \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd ei drwyn yn [=? o (y)n] redeg . //
\glsurface ALN:  oedd ei drwyn yn redeg .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} his{\scriptsize .ADJ.POSS} nose{\scriptsize .N.M.SG+SM} stative{\scriptsize .STAT} decline{\scriptsize .V.INFIN+SM[or]conjugate.V.INFIN+SM[or]run.V.INFIN+SM}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} POSS{\scriptsize .3SM} nose PRT run{\scriptsize .NONFIN}   //
\gleng his nose was running //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd o (y)n shivery@s:eng . //
\glsurface ALN:  oedd o yn shivery$^{E}$ .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} shivery{\scriptsize .ADJ}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} PRT shivery   //
\gleng he was shivery. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat nage ! //
\glsurface ELE:  nage !  //
\glauto \%aut  no{\scriptsize .ADV}   //
\glmanual \%gls  no   //
\gleng no! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng dw i (ddi)m isio (y)r annwyd (y)ma de ? //
\glsurface ELE:  oh$^{C}_{E}$ dw i ddim isio yr annwyd yma de ?  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} want{\scriptsize .N.M.SG} the{\scriptsize .DET.DEF} cold{\scriptsize .N.M.SG} here{\scriptsize .ADV} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  IM be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} NEG want DET cold here TAG   //
\gleng oh, I don't want this cold. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ hmm@s:cym\&eng [?] . //
\glsurface ALN:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae o (y)n para (y)n +// . //
\glsurface ALN:  mae o yn para yn .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} last{\scriptsize .V.INFIN} yn{\scriptsize .PRT}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT last{\scriptsize .NONFIN} PRT   //
\gleng it lasts \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well@s:cym\&eng \# mae (y)n fyr ofnadwy . //
\glsurface ALN:  well$^{C}_{E}$ mae yn fyr ofnadwy .  //
\glauto \%aut  well{\scriptsize .ADV} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} short{\scriptsize .ADJ+SM} terrible{\scriptsize .ADJ}   //
\glmanual \%gls  well be{\scriptsize .3S.PRES} PRT short awful   //
\gleng well, it's incredibly short. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$well@s:cym\&eng yeah@s:cym\&eng$>$ [?] chdi (y)dy hynna . //
\glsurface ELE:  well$^{C}_{E}$ yeah$^{C}_{E}$ chdi ydy hynna .  //
\glauto \%aut  well{\scriptsize .ADV} yeah{\scriptsize .ADV} you{\scriptsize .PRON} be{\scriptsize .V.3S.PRES} that{\scriptsize .PRON.SP.SPOKEN}   //
\glmanual \%gls  well yeah PRON{\scriptsize .2S} be{\scriptsize .3S.PRES} that   //
\gleng well, yes, for you, that is. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng . //
\glsurface ALN:  yeah$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$gyn i deimlad$>$ [?] falle fydda i (y)n rhochian heno ? //
\glsurface ALN:  gyn i deimlad falle fydda i yn rhochian heno ?  //
\glauto \%aut  with{\scriptsize .PREP[or]before.PREP+SM} I{\scriptsize .PRON.1S} feeling{\scriptsize .N.M.SG+SM} maybe{\scriptsize .ADV.SPOKEN} be{\scriptsize .V.1S.FUT+SM} I{\scriptsize .PRON.1S} stative{\scriptsize .STAT} grunt{\scriptsize .V.INFIN} tonight{\scriptsize .ADV}   //
\glmanual \%gls  with PRON{\scriptsize .1S} feeling perhaps be{\scriptsize .1S.FUT} PRON{\scriptsize .1S} PRT grunt{\scriptsize .NONFIN} tonight   //
\gleng I've got a feeling I'll be snoring tonight //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ben dy hun ta boy@s:cym\&eng ! //
\glsurface ELE:  ben dy hun ta boy$^{C}_{E}$ !  //
\glauto \%aut  head{\scriptsize .N.M.SG+SM} your{\scriptsize .ADJ.POSS} self{\scriptsize .PRON.SG} or{\scriptsize .CONJ} boy{\scriptsize .N.SG}   //
\glmanual \%gls  head POSS{\scriptsize .2S} self then boy   //
\gleng on your own then, love! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng \# braidd de ? //
\glsurface ALN:  yeah$^{C}_{E}$ braidd de ?  //
\glauto \%aut  yeah{\scriptsize .ADV} rather{\scriptsize .ADV} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  yeah somewhat TAG   //
\gleng yes, it is a bit, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i (we)di ffeindio rhywun falle wneith gymryd y bass@s:cym\&eng arian oddi\_ar y nwylo fi . //
\glsurface ALN:  dw i wedi ffeindio rhywun falle wneith gymryd y bass$^{C}_{E}$ arian oddi\_ar y nwylo fi .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} after{\scriptsize .PREP} find{\scriptsize .V.INFIN} someone{\scriptsize .N.M.SG} maybe{\scriptsize .ADV.SPOKEN} do{\scriptsize .V.3S.FUT+SM} take{\scriptsize .V.INFIN+SM} the{\scriptsize .DET.DEF} bass{\scriptsize .N.SG[or]pass.ADJ+SM} money{\scriptsize .N.M.SG} from\_on{\scriptsize .PREP} the{\scriptsize .DET.DEF} hands{\scriptsize .N.F.PL+NM} I{\scriptsize .PRON.1S+SM}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT{\scriptsize .PAST} find{\scriptsize .NONFIN} somebody maybe do{\scriptsize .3S.NONPAST} take{\scriptsize .NONFIN} DET bass silver off POSS{\scriptsize .1S} hands PRON{\scriptsize .1S}   //
\gleng I've found someone who might take the silver bass off my hands. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=cheer ! //
\glsurface ELE:  !  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat just@s:cym\&eng bod o (y)n America@s:cym\&eng . //
\glsurface ALN:  just$^{C}_{E}$ bod o yn America$^{C}_{E}$ .  //
\glauto \%aut  just{\scriptsize .ADV[or]just.ADJ} be{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} in{\scriptsize .PREP} name   //
\glmanual \%gls  just be{\scriptsize .NONFIN} PRON{\scriptsize .3SM} in America   //
\gleng except that he's in America. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be y boy@s:cym\&eng neu yr bass@s:cym\&eng ? //
\glsurface ELE:  be y boy$^{C}_{E}$ neu yr bass$^{C}_{E}$ ?  //
\glauto \%aut  what{\scriptsize .INT} the{\scriptsize .DET.DEF} boy{\scriptsize .N.SG} or{\scriptsize .CONJ} the{\scriptsize .DET.DEF} bass{\scriptsize .N.SG[or]pass.ADJ+SM}   //
\glmanual \%gls  what DET boy or DET bass   //
\gleng what, the guy or the bass? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat y boy@s:cym\&eng . //
\glsurface ALN:  y boy$^{C}_{E}$ .  //
\glauto \%aut  the{\scriptsize .DET.DEF} boy{\scriptsize .N.SG}   //
\glmanual \%gls  DET boy   //
\gleng the guy. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat y \&b\textipa{@} bass@s:cym\&eng s(y) gynno fo . //
\glsurface ALN:  y bass$^{C}_{E}$ sy gynno fo .  //
\glauto \%aut  the{\scriptsize .DET.DEF} bass{\scriptsize .N.SG[or]pass.ADJ+SM} which\_is{\scriptsize .V.3SP.PRES.REL} with\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S}   //
\glmanual \%gls  DET bass be{\scriptsize .PRES.REL} with{\scriptsize .3SM} PRON{\scriptsize .3SM}   //
\gleng the bass he's got. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wnes i gynnig swopio bass@s:cym\&eng s(y) gen i \# efo fo am fass@s:cym\&eng s(y) gynno fo . //
\glsurface ALN:  wnes i gynnig swopio bass$^{C}_{E}$ sy gen i efo fo am fass$^{C}_{E}$ sy gynno fo .  //
\glauto \%aut  do{\scriptsize .V.1S.PAST.SPOKEN+SM} I{\scriptsize .PRON.1S} offer{\scriptsize .V.INFIN+SM} swap{\scriptsize .V.INFIN} bass{\scriptsize .N.SG[or]pass.ADJ+SM} which\_is{\scriptsize .V.3SP.PRES.REL} by{\scriptsize .PREP.SPOKEN} I{\scriptsize .PRON.1S} with{\scriptsize .PREP} he{\scriptsize .PRON.M.3S} for{\scriptsize .PREP} bass{\scriptsize .N.SG+SM[or]mass.N.SG+SM} which\_is{\scriptsize .V.3SP.PRES.REL} with\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S}   //
\glmanual \%gls  do{\scriptsize .1S.PAST} PRON{\scriptsize .1S} offer{\scriptsize .NONFIN} swap{\scriptsize .NONFIN} bass be{\scriptsize .PRES.REL} with{\scriptsize .1S} PRON{\scriptsize .1S} with PRON{\scriptsize .3SM} for bass be{\scriptsize .PRES.REL} with{\scriptsize .3SM} PRON{\scriptsize .3SM}   //
\gleng I offered to swap the bass I've got with him for a bass he's got. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ oh@s:cym\&eng right@s:cym\&eng . //
\glsurface ELE:  oh$^{C}_{E}$ right$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM} right{\scriptsize .ADJ}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a mae (y)na fwy o fynd ar $<$y basses@s:cym\&eng$>$ [//] y bass@s:cym\&eng arian (y)ma \# yn America@s:cym\&eng . //
\glsurface ALN:  a mae yna fwy o fynd ar y basses$^{C}_{E}$ y bass$^{C}_{E}$ arian yma yn America$^{C}_{E}$ .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} there{\scriptsize .ADV} more{\scriptsize .ADJ.COMP+SM} of{\scriptsize .PREP} go{\scriptsize .V.INFIN+SM} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} passe{\scriptsize .N.SG+SM+PL} the{\scriptsize .DET.DEF} bass{\scriptsize .N.SG[or]pass.ADJ+SM} money{\scriptsize .N.M.SG} here{\scriptsize .ADV} in{\scriptsize .PREP} name   //
\glmanual \%gls  and be{\scriptsize .3S.PRES} there more of go on DET basses DET bass silver here in America   //
\gleng and there's a lot more call for these silver basses in America. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$(doe)s@s:cym\&eng (yn)a \&vi ddim$>$ [?] mynd arnyn nhw o\_gwbl yn wlad yma . //
\glsurface ALN:  does$^{C}_{E}$ yna ddim mynd arnyn nhw o\_gwbl yn wlad yma .  //
\glauto \%aut  does{\scriptsize .SV.INFIN} there{\scriptsize .ADV} not{\scriptsize .ADV+SM} go{\scriptsize .V.INFIN} on\_them{\scriptsize .PREP+PRON.3P} they{\scriptsize .PRON.3P} at\_all{\scriptsize .ADV} stative{\scriptsize .STAT} country{\scriptsize .N.F.SG+SM} here{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.PRES.NEG} there NEG go on{\scriptsize .3PL} PRON{\scriptsize .3PL} at\_all in country here   //
\gleng there's no call for them of all in this country. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=sniff oh@s:cym\&eng right@s:cym\&eng . //
\glsurface ELE:  oh$^{C}_{E}$ right$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM} right{\scriptsize .ADJ}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat so@s:cym\&eng dw i (we)di methu gwerthu o yn wlad yma o\_gwbl . //
\glsurface ALN:  so$^{C}_{E}$ dw i wedi methu gwerthu o yn wlad yma o\_gwbl .  //
\glauto \%aut  so{\scriptsize .ADV} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} after{\scriptsize .PREP} fail{\scriptsize .V.INFIN} sell{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} country{\scriptsize .N.F.SG+SM} here{\scriptsize .ADV} at\_all{\scriptsize .ADV}   //
\glmanual \%gls  so be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT{\scriptsize .PAST} fail{\scriptsize .NONFIN} sell{\scriptsize .NONFIN} PRON{\scriptsize .3SM} in country here at\_all   //
\gleng so I've not been able to sell it in this country at all. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat so@s:cym\&eng mae well gen i gyfnewid o efo (y)r boy@s:cym\&eng (y)ma yn America@s:cym\&eng am bass@s:cym\&eng arall . //
\glsurface ALN:  so$^{C}_{E}$ mae well gen i gyfnewid o efo yr boy$^{C}_{E}$ yma yn America$^{C}_{E}$ am bass$^{C}_{E}$ arall .  //
\glauto \%aut  so{\scriptsize .ADV} be{\scriptsize .V.3S.PRES} better{\scriptsize .ADJ.COMP+SM} by{\scriptsize .PREP.SPOKEN} I{\scriptsize .PRON.1S} exchange{\scriptsize .V.INFIN+SM} he{\scriptsize .PRON.M.3S.SPOKEN} with{\scriptsize .PREP} the{\scriptsize .DET.DEF} boy{\scriptsize .N.SG} here{\scriptsize .ADV} in{\scriptsize .PREP} name for{\scriptsize .PREP} bass{\scriptsize .N.SG[or]pass.ADJ+SM} other{\scriptsize .ADJ}   //
\glmanual \%gls  so be{\scriptsize .3S.PRES} better with{\scriptsize .1S} PRON{\scriptsize .1S} exchange{\scriptsize .NONFIN} PRON{\scriptsize .3SM} with DET boy here in America for bass other   //
\gleng so I'd rather exchange it with this guy in America for another bass. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a un\_ai gwerthu hwnna (y)n wlad yma +. . . //
\glsurface ALN:  a un\_ai gwerthu hwnna yn wlad yma .  //
\glauto \%aut  and{\scriptsize .CONJ} either{\scriptsize .ADV} sell{\scriptsize .V.INFIN} that{\scriptsize .PRON.M.SG.SPOKEN} stative{\scriptsize .STAT} country{\scriptsize .N.F.SG+SM} here{\scriptsize .ADV}   //
\glmanual \%gls  and either sell{\scriptsize .NONFIN} that in country here   //
\gleng and either sell that in this country \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat pa liw dyn nhw ? //
\glsurface ELE:  pa liw dyn nhw ?  //
\glauto \%aut  which{\scriptsize .ADJ} colour{\scriptsize .N.M.SG+SM} man{\scriptsize .N.M.SG[or]draw.V.3S.PRES+SM} they{\scriptsize .PRON.3P}   //
\glmanual \%gls  what colour be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL}   //
\gleng what colour are they? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yr un lliw â: $<$(y)r Fender@s:cym\&eng$>$ [/] \# yr hen Fender@s:cym\&eng . //
\glsurface ALN:  yr un lliw â yr Fender$^{C}_{E}$ yr hen Fender$^{C}_{E}$ .  //
\glauto \%aut  the{\scriptsize .DET.DEF} one{\scriptsize .NUM} colour{\scriptsize .N.M.SG} with{\scriptsize .PREP} the{\scriptsize .DET.DEF} name the{\scriptsize .DET.DEF} old{\scriptsize .ADJ} name   //
\glmanual \%gls  DET one colour as DET Fender DET old Fender   //
\gleng the same colour as the old Fender. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng o\_k@s:cym\&eng ta . //
\glsurface ELE:  mmm$^{C}_{E}$ o\_k$^{C}_{E}$ ta .  //
\glauto \%aut  mmm{\scriptsize .IM} OK{\scriptsize .IM} or{\scriptsize .CONJ}   //
\glmanual \%gls  IM o\_k then   //
\gleng mmm, ok then. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yr un math o beth just@s:cym\&eng bod gynno fo bumed tant yno [=? arno] fo . //
\glsurface ALN:  yr un math o beth just$^{C}_{E}$ bod gynno fo bumed tant yno fo .  //
\glauto \%aut  the{\scriptsize .DET.DEF} one{\scriptsize .NUM} type{\scriptsize .N.M.SG} of{\scriptsize .PREP} what{\scriptsize .INT[or]thing.N.M.SG+SM} just{\scriptsize .ADV[or]just.ADJ} be{\scriptsize .V.INFIN} with\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S} fifth{\scriptsize .ORD+SM} string{\scriptsize .N.M.SG} there{\scriptsize .ADV} he{\scriptsize .PRON.M.3S}   //
\glmanual \%gls  DET one kind of thing just be{\scriptsize .NONFIN} with{\scriptsize .3SM} PRON{\scriptsize .3SM} fifth string in{\scriptsize .3SM} PRON{\scriptsize .3SM}   //
\gleng the same kind of thing, except it's got the fifth string. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (y)dy o swnio (y)n guitar@s:cym\&eng bach right@s:cym\&eng nice@s:cym\&eng ? //
\glsurface ELE:  ydy o swnio yn guitar$^{C}_{E}$ bach right$^{C}_{E}$ nice$^{C}_{E}$ ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} sound{\scriptsize .V.INFIN} yn{\scriptsize .PRT} guitar{\scriptsize .N.SG} small{\scriptsize .ADJ} right{\scriptsize .ADJ} nice{\scriptsize .ADJ}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} sound{\scriptsize .NONFIN} PRT guitar little right nice   //
\gleng does it sound like a nice little guitar? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ xxx . //
\glsurface ALN:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae (y)r un peth â (y)r un coch oedd gen i wnes i swopio am yr un du . //
\glsurface ALN:  mae yr un peth â yr un coch oedd gen i wnes i swopio am yr un du .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} the{\scriptsize .DET.DEF} one{\scriptsize .NUM} thing{\scriptsize .N.M.SG} with{\scriptsize .PREP} the{\scriptsize .DET.DEF} one{\scriptsize .NUM} red{\scriptsize .ADJ} be{\scriptsize .V.3S.IMPERF} by{\scriptsize .PREP.SPOKEN} I{\scriptsize .PRON.1S} do{\scriptsize .V.1S.PAST.SPOKEN+SM} I{\scriptsize .PRON.1S} swap{\scriptsize .V.INFIN} for{\scriptsize .PREP} the{\scriptsize .DET.DEF} one{\scriptsize .NUM} black{\scriptsize .ADJ[or]side.N.M.SG+SM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} DET one thing as DET one red be{\scriptsize .3S.IMP} with{\scriptsize .1S} PRON{\scriptsize .1S} do{\scriptsize .1S.PAST} PRON{\scriptsize .1S} swap{\scriptsize .NONFIN} for DET one black   //
\gleng it's the same as the red one I used to have which I swapped for the black one. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$oh@s:cym\&eng nefi blue@s:cym\&eng$>$ [=! under\_breath] ! //
\glsurface ELE:  oh$^{C}_{E}$ nefi blue$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM} heavens{\scriptsize .IM} blue{\scriptsize .N.SG}   //
\glmanual \%gls  IM heavens blue   //
\gleng oh, good heavens! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i (ddi)m yn gallu dal i\_fyny efo hynna ! //
\glsurface ELE:  dw i ddim yn gallu dal i\_fyny efo hynna !  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} be\_able{\scriptsize .V.INFIN} hold{\scriptsize .V.INFIN} up{\scriptsize .ADV[or]up.ADV} with{\scriptsize .PREP} that{\scriptsize .PRON.SP.SPOKEN}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} NEG PRT can{\scriptsize .NONFIN} hold{\scriptsize .NONFIN} up with that   //
\gleng I can't keep up with that! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na mae ddigon teg . //
\glsurface ALN:  na mae ddigon teg .  //
\glauto \%aut  neg{\scriptsize .PRT} be{\scriptsize .V.3S.PRES} enough{\scriptsize .QUAN+SM} fair{\scriptsize .ADJ}   //
\glmanual \%gls  no be{\scriptsize .3S.PRES} enough fair   //
\gleng no, that's fair enough. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i (we)di stopio trio esbonio fo i chdi . //
\glsurface ALN:  dw i wedi stopio trio esbonio fo i chdi .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} after{\scriptsize .PREP} stop{\scriptsize .V.INFIN} try{\scriptsize .V.INFIN} explain{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S} to{\scriptsize .PREP} you{\scriptsize .PRON}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT{\scriptsize .PAST} stop{\scriptsize .NONFIN} try{\scriptsize .NONFIN} explain{\scriptsize .NONFIN} PRON{\scriptsize .3SM} to PRON{\scriptsize .2S}   //
\gleng I've stopped trying to explain it to you. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat argian fawr ! //
\glsurface ELE:  argian fawr !  //
\glauto \%aut  good\_lord{\scriptsize .IM} big{\scriptsize .ADJ+SM}   //
\glmanual \%gls  lord big   //
\gleng good grief! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a $<$mae o (we)di [/] (we)di$>$ [/] mae o (we)di mynegi diddordeb yn be s(y) gen i . //
\glsurface ALN:  a mae o wedi wedi mae o wedi mynegi diddordeb yn be sy gen i .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} after{\scriptsize .PREP} after{\scriptsize .PREP} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} after{\scriptsize .PREP} express{\scriptsize .V.INFIN} interest{\scriptsize .N.M.SG} yn{\scriptsize .PRT} what{\scriptsize .INT} which\_is{\scriptsize .V.3SP.PRES.REL} by{\scriptsize .PREP.SPOKEN} I{\scriptsize .PRON.1S}   //
\glmanual \%gls  and be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT{\scriptsize .PAST} PRT{\scriptsize .PAST} be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT{\scriptsize .PAST} express{\scriptsize .NONFIN} interest in what be{\scriptsize .PRES.REL} with{\scriptsize .1S} PRON{\scriptsize .1S}   //
\gleng and he's expressed interest in what I've got. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ac os alla i +// . //
\glsurface ALN:  ac os alla i .  //
\glauto \%aut  and{\scriptsize .CONJ} if{\scriptsize .CONJ} be\_able{\scriptsize .V.3S.PRES.SPOKEN+SM} I{\scriptsize .PRON.1S[or]to.PREP}   //
\glmanual \%gls  and if can{\scriptsize .1S.NONPAST} PRON{\scriptsize .1S}   //
\gleng and if I can \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat os wneith o gymryd o mae o mynd i fod yn [/] \# yn well ffordd i fi wneud na [/] \#\#\# na cadw beth s(y) gen i . //
\glsurface ALN:  os wneith o gymryd o mae o mynd i fod yn yn well ffordd i fi wneud na na cadw beth sy gen i .  //
\glauto \%aut  if{\scriptsize .CONJ} do{\scriptsize .V.3S.FUT+SM} of{\scriptsize .PREP} take{\scriptsize .V.INFIN+SM} he{\scriptsize .PRON.M.3S.SPOKEN} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} be{\scriptsize .V.INFIN+SM} yn{\scriptsize .PRT} stative{\scriptsize .STAT} better{\scriptsize .ADJ.COMP+SM} road{\scriptsize .N.F.SG} to{\scriptsize .PREP} I{\scriptsize .PRON.1S+SM} make{\scriptsize .V.INFIN+SM} neg{\scriptsize .PRT} (n)or{\scriptsize .CONJ} keep{\scriptsize .V.INFIN} what{\scriptsize .INT} which\_is{\scriptsize .V.3SP.PRES.REL} by{\scriptsize .PREP.SPOKEN} I{\scriptsize .PRON.1S}   //
\glmanual \%gls  if do{\scriptsize .3S.NONPAST} PRON{\scriptsize .3SM} take{\scriptsize .NONFIN} PRON{\scriptsize .3SM} be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} go{\scriptsize .NONFIN} to be{\scriptsize .NONFIN} PRT PRT better way to PRON{\scriptsize .1S} do{\scriptsize .NONFIN} than than keep{\scriptsize .NONFIN} what be{\scriptsize .PRES.REL} with{\scriptsize .1S} PRON{\scriptsize .1S}   //
\gleng if he'll take it, it'll be a better way for me to do it than keeping what I've got. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$dw i$>$ [/] $<$dw i$>$ [/] $<$dw i$>$ [/] dw i (ddi)m yn mynd i ddefnyddio (y)r un arian . //
\glsurface ALN:  dw i dw i dw i dw i ddim yn mynd i ddefnyddio yr un arian .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} use{\scriptsize .V.INFIN+SM} the{\scriptsize .DET.DEF} one{\scriptsize .NUM} money{\scriptsize .N.M.SG}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} NEG PRT go{\scriptsize .NONFIN} to use{\scriptsize .NONFIN} DET one silver   //
\gleng I'm not going to use the silver one. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ mmhm@s:cym\&eng . //
\glsurface ELE:  mmhm$^{C}_{E}$ .  //
\glauto \%aut  mmhm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat right@s:cym\&eng ! //
\glsurface ELE:  right$^{C}_{E}$ !  //
\glauto \%aut  right{\scriptsize .ADJ}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae (y)n nice@s:cym\&eng ond \# \&=grunt +/ . //
\glsurface ALN:  mae yn nice$^{C}_{E}$ ond .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} nice{\scriptsize .ADJ} but{\scriptsize .CONJ}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRT nice but   //
\gleng it's nice, but \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ mmhm@s:cym\&eng . //
\glsurface ELE:  mmhm$^{C}_{E}$ .  //
\glauto \%aut  mmhm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=sigh ti (y)n colli diddordeb dwyt ? //
\glsurface ALN:  ti yn colli diddordeb dwyt ?  //
\glauto \%aut  you{\scriptsize .PRON.2S} stative{\scriptsize .STAT} lose{\scriptsize .V.INFIN} interest{\scriptsize .N.M.SG} be{\scriptsize .V.2S.PRES.NEG}   //
\glmanual \%gls  PRON{\scriptsize .2S} PRT lose{\scriptsize .NONFIN} interest be{\scriptsize .2S.PRES.NEG}   //
\gleng you're losing interest, aren't you? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ \&=laugh ers [=! laughing] i chdi werthu (y)r un Scarborough@s:cym\&eng (y)na yndw ! //
\glsurface ELE:  ers i chdi werthu yr un Scarborough$^{C}_{E}$ yna yndw !  //
\glauto \%aut  since{\scriptsize .ADJ} to{\scriptsize .PREP} you{\scriptsize .PRON} sell{\scriptsize .V.INFIN+SM} the{\scriptsize .DET.DEF} one{\scriptsize .NUM} name there{\scriptsize .ADV} be{\scriptsize .V.1S.PRES.EMPH}   //
\glmanual \%gls  since to PRON{\scriptsize .2S} sell{\scriptsize .NONFIN} DET one Scarborough there be{\scriptsize .1S.PRES}   //
\gleng haha, ever since you sold that Scarborough one, yes I am! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o\_k@s:cym\&eng . //
\glsurface ALN:  o\_k$^{C}_{E}$ .  //
\glauto \%aut  OK{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o'n i (y)n meddwl (ba)swn i just@s:cym\&eng yn cadw chdi \# up@s:eng to@s:eng date@s:eng efo be o'n i (y)n wneud de ? //
\glsurface ALN:  o'n i yn meddwl baswn i just$^{C}_{E}$ yn cadw chdi up$^{E}$ to$^{E}$ date$^{E}$ efo be o'n i yn wneud de ?  //
\glauto \%aut  unk to{\scriptsize .PREP} stative{\scriptsize .STAT} think{\scriptsize .V.INFIN} be{\scriptsize .V.1S.PLUPERF.SPOKEN} I{\scriptsize .PRON.1S} just{\scriptsize .ADV[or]just.ADJ} stative{\scriptsize .STAT} keep{\scriptsize .V.INFIN} you{\scriptsize .PRON} up{\scriptsize .ADV} to{\scriptsize .PREP} date{\scriptsize .N.SG} with{\scriptsize .PREP} what{\scriptsize .INT} unk to{\scriptsize .PREP} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} PRT think{\scriptsize .NONFIN} be{\scriptsize .1S.CONDIT} PRON{\scriptsize .1S} just PRT keep{\scriptsize .NONFIN} PRON{\scriptsize .2S} up to date with what be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} PRT do{\scriptsize .NONFIN} TAG   //
\gleng I just thought I'd keep you up to date with what I was doing, ok? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ia ia ia ia . //
\glsurface ELE:  ia ia ia ia .  //
\glauto \%aut  yes{\scriptsize .ADV} yes{\scriptsize .ADV} yes{\scriptsize .ADV} yes{\scriptsize .ADV}   //
\glmanual \%gls  yes yes yes yes   //
\gleng yes yes, yes yes //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng yeah@s:cym\&eng yeah@s:cym\&eng ! //
\glsurface ALN:  yeah$^{C}_{E}$ yeah$^{C}_{E}$ yeah$^{C}_{E}$ !  //
\glauto \%aut  yeah{\scriptsize .ADV} yeah{\scriptsize .ADV} yeah{\scriptsize .ADV}   //
\glmanual \%gls       //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng sti be ddes i\_fewn gynnau \# um@s:cym\&eng \#\# i ddeud wrth Debbie@s:cym\&eng bod (y)n amser gwely de ? //
\glsurface ELE:  oh$^{C}_{E}$ sti be ddes i\_fewn gynnau um$^{C}_{E}$ i ddeud wrth Debbie$^{C}_{E}$ bod yn amser gwely de ?  //
\glauto \%aut  oh{\scriptsize .IM} you\_know{\scriptsize .IM} what{\scriptsize .INT} come{\scriptsize .V.1S.PAST.SPOKEN+SM} in{\scriptsize .ADV} light{\scriptsize .V.INFIN+SM} um{\scriptsize .IM} to{\scriptsize .PREP} say{\scriptsize .V.INFIN.SPOKEN+SM} by{\scriptsize .PREP} name be{\scriptsize .V.INFIN} yn{\scriptsize .PRT} time{\scriptsize .N.M.SG} bed{\scriptsize .N.M.SG} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  IM know{\scriptsize .2S} what come{\scriptsize .1S.PAST} in just\_now IM to say{\scriptsize .NONFIN} to Debbie be{\scriptsize .NONFIN} PRT time bed TAG   //
\gleng oh, you know what, I came in just now, to tell Debbie it was bedtime, right? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng . //
\glsurface ALN:  yeah$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dyna lle oedd hi (e)fo Max\_N@s:cym\&eng ymlaen \# yn wneud um@s:cym\&eng \# symudiadau \# dawnsio \# disco@s:cym\&eng (y)ma . //
\glsurface ELE:  dyna lle oedd hi efo Max\_N$^{C}_{E}$ ymlaen yn wneud um$^{C}_{E}$ symudiadau dawnsio disco$^{C}_{E}$ yma .  //
\glauto \%aut  that{\scriptsize .is.ADV} where{\scriptsize .INT[or]place.N.M.SG} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} with{\scriptsize .PREP} name forward{\scriptsize .ADV} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM} um{\scriptsize .IM} movements{\scriptsize .N.M.PL} dance{\scriptsize .V.INFIN} disco{\scriptsize .N.SG} here{\scriptsize .ADV}   //
\glmanual \%gls  there place be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} with Max\_N on PRT do{\scriptsize .NONFIN} IM movements dance{\scriptsize .NONFIN} disco here   //
\gleng that's where she was with "Max N" on, doing these disco dance movements. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ALN:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat edrych right@s:cym\&eng dda hefyd \&=laugh ! //
\glsurface ELE:  edrych right$^{C}_{E}$ dda hefyd !  //
\glauto \%aut  look{\scriptsize .V.2S.IMPER} right{\scriptsize .ADJ} good{\scriptsize .ADJ+SM} also{\scriptsize .ADV}   //
\glmanual \%gls  look{\scriptsize .NONFIN} right good too   //
\gleng looked quite good too! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$ti (we)di$>$ [//] $<$(y)dy hi$>$ [?] [//] wedi dangos routine@s:cym\&eng mae (we)di ddysgu yn +. . . //
\glsurface ALN:  ti wedi ydy hi wedi dangos routine$^{C}_{E}$ mae wedi ddysgu yn .  //
\glauto \%aut  you{\scriptsize .PRON.2S} after{\scriptsize .PREP} be{\scriptsize .V.3S.PRES} she{\scriptsize .PRON.F.3S} after{\scriptsize .PREP} show{\scriptsize .V.INFIN} routine{\scriptsize .N.SG} be{\scriptsize .V.3S.PRES} after{\scriptsize .PREP} teach{\scriptsize .V.INFIN+SM} yn{\scriptsize .PRT}   //
\glmanual \%gls  PRON{\scriptsize .2S} PRT{\scriptsize .PAST} be{\scriptsize .3S.PRES} PRON{\scriptsize .3SF} PRT{\scriptsize .PAST} show{\scriptsize .NONFIN} routine be{\scriptsize .3S.PRES} PRT{\scriptsize .PAST} learn{\scriptsize .NONFIN} in   //
\gleng has she shown you the routine she learnt in \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na . //
\glsurface ELE:  na .  //
\glauto \%aut  neg{\scriptsize .PRT}   //
\glmanual \%gls  no   //
\gleng no. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw [/] dw i heb weld dim\_byd . //
\glsurface ALN:  dw dw i heb weld dim\_byd .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} without{\scriptsize .PREP} see{\scriptsize .V.INFIN+SM} anything{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} without see{\scriptsize .NONFIN} nothing   //
\gleng I've not seen anything. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng mae [//] well@s:cym\&eng mae (we)di dangos um@s:cym\&eng [?] ambell i beth . //
\glsurface ELE:  oh$^{C}_{E}$ mae well$^{C}_{E}$ mae wedi dangos um$^{C}_{E}$ ambell i beth .  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES} well{\scriptsize .ADV} be{\scriptsize .V.3S.PRES} after{\scriptsize .PREP} show{\scriptsize .V.INFIN} um{\scriptsize .IM} occasional{\scriptsize .PREQ} to{\scriptsize .PREP} what{\scriptsize .INT}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} well be{\scriptsize .3S.PRES} PRT{\scriptsize .PAST} show{\scriptsize .NONFIN} IM some to thing   //
\gleng oh, well, she's shown me some things. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond [?] na dw i (ddi)m (we)di gweld \# y dawns fach na na . //
\glsurface ELE:  ond na dw i ddim wedi gweld y dawns fach na na .  //
\glauto \%aut  but{\scriptsize .CONJ} neg{\scriptsize .PRT} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} after{\scriptsize .PREP} see{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} dance{\scriptsize .N.F.SG} small{\scriptsize .ADJ+SM} neg{\scriptsize .PRT} neg{\scriptsize .PRT}   //
\glmanual \%gls  but no be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} NEG PRT{\scriptsize .PAST} see{\scriptsize .NONFIN} DET dance little no no   //
\gleng but no, I haven't seen the little dance, no, no. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ (ba)sai (y)n ddoniol ! //
\glsurface ALN:  basai yn ddoniol !  //
\glauto \%aut  be{\scriptsize .V.3S.PLUPERF.SPOKEN} stative{\scriptsize .STAT} funny{\scriptsize .ADJ+SM}   //
\glmanual \%gls  be{\scriptsize .3S.CONDIT} PRT funny   //
\gleng would be funny! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat doniol ? //
\glsurface ELE:  doniol ?  //
\glauto \%aut  funny{\scriptsize .ADJ}   //
\glmanual \%gls  funny   //
\gleng funny? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well@s:cym\&eng na ddim +// . //
\glsurface ALN:  well$^{C}_{E}$ na ddim .  //
\glauto \%aut  well{\scriptsize .ADV} neg{\scriptsize .PRT} nothing{\scriptsize .N.M.SG+SM[or]not.ADV+SM}   //
\glmanual \%gls  well no NEG   //
\gleng well, no, not \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o\_k@s:cym\&eng (ba)sai fo [?] (y)n ddiddorol . //
\glsurface ALN:  o\_k$^{C}_{E}$ basai fo yn ddiddorol .  //
\glauto \%aut  OK{\scriptsize .IM} be{\scriptsize .V.3S.PLUPERF.SPOKEN} he{\scriptsize .PRON.M.3S} stative{\scriptsize .STAT} interesting{\scriptsize .ADJ+SM}   //
\glmanual \%gls  o\_k be{\scriptsize .3S.CONDIT} PRON{\scriptsize .3SM} PRT interesting   //
\gleng ok, it would be interesting. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng basai basai . //
\glsurface ELE:  oh$^{C}_{E}$ basai basai .  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PLUPERF.SPOKEN} be{\scriptsize .V.3S.PLUPERF.SPOKEN}   //
\glmanual \%gls  IM be{\scriptsize .3S.CONDIT} be{\scriptsize .3S.CONDIT}   //
\gleng oh, yes, it would. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (ba)set ti medru mynd i mynd i nôl hi (y)n fuan \# dydd Iau ac eistedd yn gefn y neuadd . //
\glsurface ALN:  baset ti medru mynd i mynd i nôl hi yn fuan dydd Iau ac eistedd yn gefn y neuadd .  //
\glauto \%aut  be{\scriptsize .V.2S.PLUPERF} you{\scriptsize .PRON.2S} be{\scriptsize .V.INFIN} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} fetch{\scriptsize .V.INFIN} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} soon{\scriptsize .ADJ+SM} day{\scriptsize .N.M.SG} Jupiter{\scriptsize .N.M.SG} and{\scriptsize .CONJ} sit{\scriptsize .V.3S.PRES[or]sit.V.INFIN} stative{\scriptsize .STAT} back{\scriptsize .N.M.SG+SM} the{\scriptsize .DET.DEF} hall{\scriptsize .N.F.SG}   //
\glmanual \%gls  be{\scriptsize .2S.CONDIT} PRON{\scriptsize .2S} can{\scriptsize .NONFIN} go{\scriptsize .NONFIN} to fetch{\scriptsize .NONFIN} PRON{\scriptsize .3SF} PRT early day Thursday and sit{\scriptsize .NONFIN} in back DET hall     //
\gleng you could go early to pick her up on Thursday and sit at the back of the hall. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat neu ti (y)n meddwl (ba)sai hynna (y)n freak\_o@s:eng+cym (y)r athrawes allan ? //
\glsurface ALN:  neu ti yn meddwl basai hynna yn freak\_o$^{E+}_{C}$ yr athrawes allan ?  //
\glauto \%aut  or{\scriptsize .CONJ} you{\scriptsize .PRON.2S} stative{\scriptsize .STAT} think{\scriptsize .V.INFIN} be{\scriptsize .V.3S.PLUPERF.SPOKEN} that{\scriptsize .PRON.SP.SPOKEN} stative{\scriptsize .STAT} freak{\scriptsize .N.SG+WV[or]break.SV.INFIN+SM+WV} the{\scriptsize .DET.DEF} teacher{\scriptsize .N.F.SG} out{\scriptsize .ADV}   //
\glmanual \%gls  or PRON{\scriptsize .2S} PRT think{\scriptsize .NONFIN} be{\scriptsize .3S.CONDIT} that PRT freak{\scriptsize .NONFIN} DET teacher out   //
\gleng or do you think that would freak out the teacher? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ah@s:cym\&eng dw i (y)n sure@s:cym\&eng (ba)swn [=? (peta)swn] i +// . //
\glsurface ELE:  ah$^{C}_{E}$ dw i yn sure$^{C}_{E}$ baswn i .  //
\glauto \%aut  ah{\scriptsize .IM} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} stative{\scriptsize .STAT} sure{\scriptsize .ADJ} be{\scriptsize .V.1S.PLUPERF.SPOKEN} I{\scriptsize .PRON.1S}   //
\glmanual \%gls  IM be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT sure be{\scriptsize .1S.CONDIT} PRON{\scriptsize .1S}   //
\gleng oh, I'm sure I'd \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng ti (y)n gallu gweld o (y)r drws dwyt ? //
\glsurface ELE:  oh$^{C}_{E}$ ti yn gallu gweld o yr drws dwyt ?  //
\glauto \%aut  oh{\scriptsize .IM} you{\scriptsize .PRON.2S} stative{\scriptsize .STAT} be\_able{\scriptsize .V.INFIN} see{\scriptsize .V.INFIN} of{\scriptsize .PREP} the{\scriptsize .DET.DEF} door{\scriptsize .N.M.SG} be{\scriptsize .V.2S.PRES.NEG}   //
\glmanual \%gls  IM PRON{\scriptsize .2S} PRT can{\scriptsize .NONFIN} see{\scriptsize .NONFIN} from DET door be{\scriptsize .2S.PRES.NEG}   //
\gleng oh, you can see from the door can't you? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wyt . //
\glsurface ALN:  wyt .  //
\glauto \%aut  be{\scriptsize .V.2S.PRES}   //
\glmanual \%gls  be{\scriptsize .2S.PRES}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat er (peta)set ti yna (ba)sai hi (dd)im yn dawnsio na (ba)sai ? //
\glsurface ALN:  er petaset ti yna basai hi ddim yn dawnsio na basai ?  //
\glauto \%aut  er{\scriptsize .IM} be{\scriptsize .V.2S.PLUPERF.HYP} you{\scriptsize .PRON.2S} there{\scriptsize .ADV} be{\scriptsize .V.3S.PLUPERF.SPOKEN} she{\scriptsize .PRON.F.3S} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} dance{\scriptsize .V.INFIN} neg{\scriptsize .PRT} be{\scriptsize .V.3S.PLUPERF.SPOKEN}   //
\glmanual \%gls  although if\_be{\scriptsize .2S.CONDIT} PRON{\scriptsize .2S} there be{\scriptsize .3S.CONDIT} PRON{\scriptsize .3SF} NEG PRT dance{\scriptsize .NONFIN} NEG be{\scriptsize .3S.CONDIT}   //
\gleng although, if you were there she wouldn't dance, would she? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na na oh@s:cym\&eng na $<$dyna pam$>$ [/] \# dyna pam +. . . //
\glsurface ELE:  na na oh$^{C}_{E}$ na dyna pam dyna pam .  //
\glauto \%aut  neg{\scriptsize .PRT} neg{\scriptsize .PRT[or]who\_not.PRON.REL.NEG[or](n)or.CONJ[or]than.CONJ} oh{\scriptsize .IM} neg{\scriptsize .PRT} that{\scriptsize .is.ADV} why?.ADV that{\scriptsize .is.ADV} why?.ADV   //
\glmanual \%gls  no no IM no there why there why   //
\gleng no, no, oh, no, that's why \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat timod weles hi [?] (y)n dawnsio gynnau . //
\glsurface ELE:  timod weles hi yn dawnsio gynnau .  //
\glauto \%aut  know{\scriptsize .V.2S.PRES.SPOKEN} see{\scriptsize .V.1S.PAST.SPOKEN} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} dance{\scriptsize .V.INFIN} light{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  know{\scriptsize .2S} see{\scriptsize .1S.PAST} PRON{\scriptsize .3SF} PRT dance{\scriptsize .NONFIN} just\_now   //
\gleng you know, I saw her dancing just now. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ac oedd hi (y)n sbïo at yr \# tân twel . //
\glsurface ELE:  ac oedd hi yn sbïo at yr tân twel .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} look{\scriptsize .V.INFIN} to{\scriptsize .PREP} the{\scriptsize .DET.DEF} fire{\scriptsize .N.M.SG} see{\scriptsize .V.2S.PRES.SPOKEN}   //
\glmanual \%gls  and be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT look{\scriptsize .NONFIN} towards DET fire see{\scriptsize .2S}   //
\gleng and she was looking towards the fire, you see. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat doe(dd) (ddi)m yn dallt bod fi (y)n sbïo arni nag oedd ? //
\glsurface ELE:  doedd ddim yn dallt bod fi yn sbïo arni nag oedd ?  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF.NEG} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} understand{\scriptsize .V.INFIN.SPOKEN} be{\scriptsize .V.INFIN} I{\scriptsize .PRON.1S+SM} stative{\scriptsize .STAT} look{\scriptsize .V.INFIN} on\_her{\scriptsize .PREP+PRON.F.3S} than{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} NEG PRT understand{\scriptsize .NONFIN} be{\scriptsize .NONFIN} PRON{\scriptsize .1S} PRT look{\scriptsize .NONFIN} on{\scriptsize .3SF} NEG be{\scriptsize .3S.IMP}   //
\gleng she didn't realise I was looking at her, did she //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat so@s:cym\&eng oedd hi (y)n mynd amdani yn ei byd bach ei hun ? //
\glsurface ALN:  so$^{C}_{E}$ oedd hi yn mynd amdani yn ei byd bach ei hun ?  //
\glauto \%aut  so{\scriptsize .ADV} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} go{\scriptsize .V.INFIN} for\_her{\scriptsize .PREP+PRON.F.3S} yn{\scriptsize .PRT} her{\scriptsize .ADJ.POSS[or]his.ADJ.POSS} world{\scriptsize .N.M.SG} small{\scriptsize .ADJ} her{\scriptsize .ADJ.POSS[or]his.ADJ.POSS[or]go.V.2S.PRES} self{\scriptsize .PRON.SG}   //
\glmanual \%gls  so be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT go{\scriptsize .NONFIN} for{\scriptsize .3SF} in POSS{\scriptsize .3S} world little POSS{\scriptsize .3S} self   //
\gleng so she was just going for it, in her own little world? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd . //
\glsurface ELE:  oedd .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  be{\scriptsize .3S.IMP}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd . //
\glsurface ELE:  oedd .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  be{\scriptsize .3S.IMP}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat digon del . //
\glsurface ELE:  digon del .  //
\glauto \%aut  enough{\scriptsize .QUAN} pretty{\scriptsize .ADJ}   //
\glmanual \%gls  enough cute   //
\gleng quite cute. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat rhythmic@s:cym\&eng de ? //
\glsurface ELE:  rhythmic$^{C}_{E}$ de ?  //
\glauto \%aut  rhythmic{\scriptsize .ADJ} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  rhythmic TAG   //
\gleng rhythmic, right? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng $<$o'n i (y)n$>$ [//] $<$ces i$>$ [//] $<$o'n i (y)n$>$ [//] ces i lun ohoni (y)n canu xxx $<$yn\_do efo$>$ [?] +// . //
\glsurface ALN:  oh$^{C}_{E}$ o'n i yn ces i o'n i yn ces i lun ohoni yn canu yn\_do efo .  //
\glauto \%aut  oh{\scriptsize .IM} unk to{\scriptsize .PREP} yn{\scriptsize .PRT} get{\scriptsize .V.1S.PAST.SPOKEN} I{\scriptsize .PRON.1S} unk to{\scriptsize .PREP} yn{\scriptsize .PRT} get{\scriptsize .V.1S.PAST.SPOKEN} I{\scriptsize .PRON.1S} picture{\scriptsize .N.M.SG+SM} from\_her{\scriptsize .PREP+PRON.F.3S} stative{\scriptsize .STAT} sing{\scriptsize .V.INFIN} wasn't\_it{\scriptsize .IM} with{\scriptsize .PREP}   //
\glmanual \%gls  IM be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} PRT get{\scriptsize .1S.PAST} PRON{\scriptsize .1S} be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} PRT get{\scriptsize .1S.PAST} PRON{\scriptsize .1S} picture of{\scriptsize .3SF} PRT sing{\scriptsize .NONFIN} yes with   //
\gleng oh, I got a picture of her singing, didn't I, with \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat just@s:cym\&eng roid y cefn y video@s:cym\&eng dros y drws $<$fel bod hi (dd)im$>$ [?] y ngweld i . //
\glsurface ALN:  just$^{C}_{E}$ roid y cefn y video$^{C}_{E}$ dros y drws fel bod hi ddim y ngweld i .  //
\glauto \%aut  just{\scriptsize .ADV[or]just.ADJ} give{\scriptsize .V.0.IMPERF+SM} the{\scriptsize .DET.DEF} back{\scriptsize .N.M.SG} the{\scriptsize .DET.DEF} video{\scriptsize .N.SG} over{\scriptsize .PREP+SM} the{\scriptsize .DET.DEF} door{\scriptsize .N.M.SG} like{\scriptsize .CONJ} be{\scriptsize .V.INFIN} she{\scriptsize .PRON.F.3S} nothing{\scriptsize .N.M.SG+SM[or]not.ADV+SM} the{\scriptsize .DET.DEF} see{\scriptsize .V.INFIN+NM} to{\scriptsize .PREP}   //
\glmanual \%gls  just put{\scriptsize .NONFIN} DET back DET video over DET door as be{\scriptsize .NONFIN} PRON{\scriptsize .3SF} NEG POSS{\scriptsize .1S} see{\scriptsize .NONFIN} PRON{\scriptsize .1S}   //
\gleng just putting the back of the video over the door so she couldn't see me. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a:h@s:cym\&eng yeah@s:cym\&eng ! //
\glsurface ELE:  ah$^{C}_{E}$ yeah$^{C}_{E}$ !  //
\glauto \%aut  ah{\scriptsize .IM} yeah{\scriptsize .ADV}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat argian gobeithio bod yr um@s:cym\&eng \#\# stuff@s:cym\&eng dal (we)di recordio yn hwn de . //
\glsurface ALN:  argian gobeithio bod yr um$^{C}_{E}$ stuff$^{C}_{E}$ dal wedi recordio yn hwn de .  //
\glauto \%aut  good\_lord{\scriptsize .IM} hope{\scriptsize .V.INFIN} be{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} um{\scriptsize .IM} stuff{\scriptsize .SV.INFIN} hold{\scriptsize .V.INFIN} after{\scriptsize .PREP} record{\scriptsize .V.INFIN} in{\scriptsize .PREP} this{\scriptsize .PRON.M.SG} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  lord hope{\scriptsize .NONFIN} be{\scriptsize .NONFIN} DET IM stuff still PRT{\scriptsize .PAST} record{\scriptsize .NONFIN} in this TAG   //
\gleng blimey, I hope the stuff has still recorded on this, eh? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ \&=gasp . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=dental\_click ah@s:cym\&eng ti (y)n medd(wl) chdi (we)di colli hwnnw (he)fyd ? //
\glsurface ELE:  ah$^{C}_{E}$ ti yn meddwl chdi wedi colli hwnnw hefyd ?  //
\glauto \%aut  ah{\scriptsize .IM} you{\scriptsize .PRON.2S} stative{\scriptsize .STAT} think{\scriptsize .V.INFIN} you{\scriptsize .PRON} after{\scriptsize .PREP} lose{\scriptsize .V.INFIN} that{\scriptsize .PRON.M.SG} also{\scriptsize .ADV}   //
\glmanual \%gls  IM PRON{\scriptsize .2S} PRT think{\scriptsize .NONFIN} PRON{\scriptsize .2S} PRT{\scriptsize .PAST} lose{\scriptsize .NONFIN} that too   //
\gleng oh, you think you've lost that too? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw (ddi)m yn gwybod . //
\glsurface ALN:  dw ddim yn gwybod .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} NEG PRT know{\scriptsize .NONFIN}   //
\gleng I don't know. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat daria $<$fedri di roid$>$ [?] ar cyfrifiadur neu (rhyw)beth ? //
\glsurface ELE:  daria fedri di roid ar cyfrifiadur neu rhywbeth ?  //
\glauto \%aut  drat{\scriptsize .IM} be{\scriptsize .V.2S.PRES+SM} you{\scriptsize .PRON.2S+SM} give{\scriptsize .V.0.IMPERF+SM} on{\scriptsize .PREP} computer{\scriptsize .N.M.SG} or{\scriptsize .CONJ} something{\scriptsize .N.M.SG}   //
\glmanual \%gls  bother can{\scriptsize .2S.NONPAST} PRON{\scriptsize .2S} put{\scriptsize .NONFIN} on computer or something   //
\gleng bother, can you put it on the computer or something? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$fydd raid fi$>$ [?] wneud bydd ? //
\glsurface ALN:  fydd raid fi wneud bydd ?  //
\glauto \%aut  be{\scriptsize .V.3S.FUT+SM} necessity{\scriptsize .N.M.SG+SM} I{\scriptsize .PRON.1S+SM} make{\scriptsize .V.INFIN+SM} be{\scriptsize .V.3S.FUT}   //
\glmanual \%gls  be{\scriptsize .3S.FUT} necessity PRON{\scriptsize .1S} do{\scriptsize .NONFIN} be{\scriptsize .3S.FUT}   //
\gleng I'll have to, won't I? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat tria wneud yeah@s:cym\&eng . //
\glsurface ELE:  tria wneud yeah$^{C}_{E}$ .  //
\glauto \%aut  try{\scriptsize .V.2S.IMPER} make{\scriptsize .V.INFIN+SM} yeah{\scriptsize .ADV}   //
\glmanual \%gls  try{\scriptsize .2S.IMPER} do{\scriptsize .NONFIN} yeah   //
\gleng try and do it, yeah. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a:h@s:cym\&eng (ba)sai (hwn)na (y)n bity@s:cym\&eng ! //
\glsurface ELE:  ah$^{C}_{E}$ basai hwnna yn bity$^{C}_{E}$ !  //
\glauto \%aut  ah{\scriptsize .IM} be{\scriptsize .V.3S.PLUPERF.SPOKEN} that{\scriptsize .PRON.M.SG.SPOKEN} stative{\scriptsize .STAT} pity{\scriptsize .N.SG+SM}   //
\glmanual \%gls  IM be{\scriptsize .3S.CONDIT} that PRT pity   //
\gleng oh, that would be a pity. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be (y)dy stuff@s:cym\&eng darllen $<$i fod dim$>$ [?] yna ? //
\glsurface ELE:  be ydy stuff$^{C}_{E}$ darllen i fod dim yna ?  //
\glauto \%aut  what{\scriptsize .INT} be{\scriptsize .V.3S.PRES} stuff{\scriptsize .SV.INFIN} read{\scriptsize .V.INFIN} to{\scriptsize .PREP} be{\scriptsize .V.INFIN+SM} nothing{\scriptsize .N.M.SG[or]not.ADV} there{\scriptsize .ADV}   //
\glmanual \%gls  what be{\scriptsize .3S.PRES} stuff reading to be{\scriptsize .NONFIN} NEG there   //
\gleng what, is the reading stuff supposed to not be there. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yndy . //
\glsurface ALN:  yndy .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=dental\_click . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mi ddylai fod yna \# ond um@s:cym\&eng +. . . //
\glsurface ALN:  mi ddylai fod yna ond um$^{C}_{E}$ .  //
\glauto \%aut  aff{\scriptsize .PRT} ought{\scriptsize .V.3S.IMPERF+SM} be{\scriptsize .V.INFIN+SM} there{\scriptsize .ADV} but{\scriptsize .CONJ} um{\scriptsize .IM}   //
\glmanual \%gls  PRT should{\scriptsize .3S.CONDIT} be{\scriptsize .NONFIN} there but IM   //
\gleng it should be there, but, um \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng dyna ti (y)n gael $<$am olchi$>$ [?] dy \# phone@s:cym\&eng $<$yn injan olchi$>$ [?] de ? //
\glsurface ELE:  oh$^{C}_{E}$ dyna ti yn gael am olchi dy phone$^{C}_{E}$ yn injan olchi de ?  //
\glauto \%aut  oh{\scriptsize .IM} that{\scriptsize .is.ADV} you{\scriptsize .PRON.2S} stative{\scriptsize .STAT} have{\scriptsize .V.INFIN+SM} for{\scriptsize .PREP} wash{\scriptsize .V.INFIN+SM} your{\scriptsize .ADJ.POSS} phone{\scriptsize .N.SG} yn{\scriptsize .PRT} engine{\scriptsize .N.F.SG} wash{\scriptsize .V.INFIN+SM} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  IM there PRON{\scriptsize .2S} PRT get{\scriptsize .NONFIN} for wash{\scriptsize .NONFIN} POSS{\scriptsize .2S} phone in engine wash{\scriptsize .NONFIN} TAG   //
\gleng oh, that's what you get for washing your phone in the washing machine, eh? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie yeah@s:cym\&eng yeah@s:cym\&eng ! //
\glsurface ALN:  ie yeah$^{C}_{E}$ yeah$^{C}_{E}$ !  //
\glauto \%aut  yes{\scriptsize .ADV} yeah{\scriptsize .ADV} yeah{\scriptsize .ADV}   //
\glmanual \%gls  yes yeah yeah   //
\gleng yeah, yeah! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=sniff . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dyna dw i gael am gael trowsus efo \# pumed pocket@s:cym\&eng yno fo de ? //
\glsurface ALN:  dyna dw i gael am gael trowsus efo pumed pocket$^{C}_{E}$ yno fo de ?  //
\glauto \%aut  that{\scriptsize .is.ADV} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} have{\scriptsize .V.INFIN+SM} for{\scriptsize .PREP} have{\scriptsize .V.INFIN+SM} trousers{\scriptsize .N.M.SG} with{\scriptsize .PREP} fifth{\scriptsize .ORD} pocket{\scriptsize .N.SG} there{\scriptsize .ADV} he{\scriptsize .PRON.M.3S} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  there be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} get{\scriptsize .NONFIN} for get{\scriptsize .NONFIN} trousers with fifth pocket in{\scriptsize .3SM} PRON{\scriptsize .3SM} TAG   //
\gleng that's what I get for having trousers with a fifth pocket in them, eh? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng ! //
\glsurface ELE:  hmm$^{C}_{E}$ !  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well ti beidio cael trowsus efo pocket@s:cym\&eng yno (fe)lly (y)ndy . //
\glsurface ELE:  well ti beidio cael trowsus efo pocket$^{C}_{E}$ yno felly yndy .  //
\glauto \%aut  better{\scriptsize .ADJ.COMP+SM} you{\scriptsize .PRON.2S} stop{\scriptsize .V.INFIN+SM} have{\scriptsize .V.INFIN} trousers{\scriptsize .N.M.SG} with{\scriptsize .PREP} pocket{\scriptsize .N.SG} there{\scriptsize .ADV} so{\scriptsize .ADV} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  better PRON{\scriptsize .2S} desist{\scriptsize .NONFIN} get{\scriptsize .NONFIN} trousers with pocket in{\scriptsize .3SM} thus be{\scriptsize .3S.PRES}   //
\gleng better you don't get trousers with a pocket in them, then, isn't it //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well o lawer . //
\glsurface ELE:  well o lawer .  //
\glauto \%aut  better{\scriptsize .ADJ.COMP+SM} of{\scriptsize .PREP} many{\scriptsize .QUAN+SM}   //
\glmanual \%gls  better of much   //
\gleng better by far. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie +// . //
\glsurface ELE:  ie .  //
\glauto \%aut  yes{\scriptsize .ADV}   //
\glmanual \%gls  yes   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na +. . . //
\glsurface ELE:  na .  //
\glauto \%aut  neg{\scriptsize .PRT}   //
\glmanual \%gls  no   //
\gleng no. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng gaeth hi hwyl . //
\glsurface ELE:  oh$^{C}_{E}$ gaeth hi hwyl .  //
\glauto \%aut  oh{\scriptsize .IM} captive{\scriptsize .ADJ+SM} she{\scriptsize .PRON.F.3S} fun{\scriptsize .N.F.SG}   //
\glmanual \%gls  IM have{\scriptsize .3S.PAST} PRON{\scriptsize .3SF} fun   //
\gleng oh, she had fun. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd hi (y)n deu(d) (wr)tha fi gynnau yn gegin . //
\glsurface ELE:  oedd hi yn deud wrtha fi gynnau yn gegin .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN} to\_me{\scriptsize .PREP+PRON.1S} I{\scriptsize .PRON.1S+SM} light{\scriptsize .V.INFIN+SM} stative{\scriptsize .STAT} kitchen{\scriptsize .N.F.SG+SM}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT say{\scriptsize .NONFIN} to{\scriptsize .1S} PRON{\scriptsize .1S} just\_now in kitchen   //
\gleng she was telling me just now in the kitchen. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd hi (y)n deud um@s:cym\&eng \#\# efo eu sioe nhw y Dolig +. . . //
\glsurface ELE:  oedd hi yn deud um$^{C}_{E}$ efo eu sioe nhw y Dolig .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN} um{\scriptsize .IM} with{\scriptsize .PREP} their{\scriptsize .ADJ.POSS} show{\scriptsize .N.F.SG} they{\scriptsize .PRON.3P} the{\scriptsize .DET.DEF} name   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT say{\scriptsize .NONFIN} IM with POSS{\scriptsize .3PL} show PRON{\scriptsize .3PL} DET Christmas   //
\gleng she was saying that with their Christmas show \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ac oedd hi gofyn os oedden ni (y)n +// . //
\glsurface ELE:  ac oedd hi gofyn os oedden ni yn .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} ask{\scriptsize .V.INFIN} if{\scriptsize .CONJ} be{\scriptsize .V.3P.IMPERF.SPOKEN} we{\scriptsize .PRON.1P} yn{\scriptsize .PRT}   //
\glmanual \%gls  and be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} ask{\scriptsize .NONFIN} if be{\scriptsize .3PL.IMP} PRON{\scriptsize .3PL} PRT   //
\gleng and she was asking whether we were \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" dan ni (y)n wneud sioe yn Theatr\_Gwynedd@s:cym\&eng a dan ni (y)n wneud sioe yn (y)r ysgol . //
\glsurface ELE:  dan ni yn wneud sioe yn Theatr\_Gwynedd$^{C}_{E}$ a dan ni yn wneud sioe yn yr ysgol .  //
\glauto \%aut  be{\scriptsize .V.1P.PRES.SPOKEN} we{\scriptsize .PRON.1P} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM} show{\scriptsize .N.F.SG} in{\scriptsize .PREP} name and{\scriptsize .CONJ} be{\scriptsize .V.1P.PRES.SPOKEN} we{\scriptsize .PRON.1P} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM} show{\scriptsize .N.F.SG} in{\scriptsize .PREP} the{\scriptsize .DET.DEF} school{\scriptsize .N.F.SG}   //
\glmanual \%gls  be{\scriptsize .1PL.PRES} PRON{\scriptsize .1PL} PRT do{\scriptsize .NONFIN} show in Theatr\_Gwynedd and be{\scriptsize .1PL.PRES} PRON{\scriptsize .1PL} PRT do{\scriptsize .NONFIN} show in DET school   //
\gleng "we're doing a show in Theatr Gwynedd, and we're doing a show at school" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat meddai hi +" . //
\glsurface ELE:  meddai hi .  //
\glauto \%aut  say{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  say{\scriptsize .3S.PAST} PRON{\scriptsize .3SF}   //
\gleng she said. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat i [?] (y)n deud +"/ . //
\glsurface ELE:  i yn deud .  //
\glauto \%aut  to{\scriptsize .PREP} stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN}   //
\glmanual \%gls  PRON{\scriptsize .1S} PRT say{\scriptsize .NONFIN}   //
\gleng I said: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" na un sioe . //
\glsurface ELE:  na un sioe .  //
\glauto \%aut  neg{\scriptsize .PRT} one{\scriptsize .NUM} show{\scriptsize .N.F.SG}   //
\glmanual \%gls  no one show   //
\gleng no, one show. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$meddai hi$>$ [?] +"/ . //
\glsurface ELE:  meddai hi .  //
\glauto \%aut  say{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  say{\scriptsize .3S.PAST} PRON{\scriptsize .3SF}   //
\gleng she said. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" yn Theatr\_Gwynedd@s:cym\&eng . //
\glsurface ELE:  yn Theatr\_Gwynedd$^{C}_{E}$ .  //
\glauto \%aut  in{\scriptsize .PREP} name   //
\glmanual \%gls  in Theatr\_Gwynedd   //
\gleng in Theatr Gwynedd. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" well@s:cym\&eng blwyddyn dau sy (y)n cael \# er@s:cym\&eng actio \# a chanu . //
\glsurface ELE:  well$^{C}_{E}$ blwyddyn dau sy yn cael er$^{C}_{E}$ actio a chanu .  //
\glauto \%aut  well{\scriptsize .ADV} year{\scriptsize .N.F.SG} two{\scriptsize .NUM.M} which\_is{\scriptsize .V.3SP.PRES.REL} stative{\scriptsize .STAT} have{\scriptsize .V.INFIN} er{\scriptsize .IM} act{\scriptsize .V.INFIN} and{\scriptsize .CONJ} sing{\scriptsize .V.INFIN+AM}   //
\glmanual \%gls  well year two{\scriptsize .M} be{\scriptsize .PRES.REL} PRT get{\scriptsize .NONFIN} IM act{\scriptsize .NONFIN} and sing{\scriptsize .NONFIN}   //
\gleng "it's year two that get to act and sing." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat meddai hi +"/ . //
\glsurface ELE:  meddai hi .  //
\glauto \%aut  say{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  say{\scriptsize .3S.PAST} PRON{\scriptsize .3SF}   //
\gleng she said. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wrth\_gwrs mae hi ar dop@s:cym\&eng y babanod wan yndy ? //
\glsurface ALN:  wrth\_gwrs mae hi ar dop$^{C}_{E}$ y babanod wan yndy ?  //
\glauto \%aut  of\_course{\scriptsize .ADV} be{\scriptsize .V.3S.PRES} she{\scriptsize .PRON.F.3S} on{\scriptsize .PREP} top{\scriptsize .N.SG+SM} the{\scriptsize .DET.DEF} babies{\scriptsize .N.M.PL} weak{\scriptsize .ADJ+SM[or]pale.ADJ+SM} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  of\_course be{\scriptsize .3S.PRES} PRON{\scriptsize .3SF} on top DET infants now be{\scriptsize .3S.PRES}   //
\gleng of course, she's at the top of infants now, isn't she? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yndy ! //
\glsurface ELE:  yndy !  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES}   //
\gleng yes! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a wnes i feddwl +"/ . //
\glsurface ELE:  a wnes i feddwl .  //
\glauto \%aut  and{\scriptsize .CONJ} do{\scriptsize .V.1S.PAST.SPOKEN+SM} I{\scriptsize .PRON.1S} think{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  and do{\scriptsize .1S.PAST} PRON{\scriptsize .1S} think   //
\gleng and I thought: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" uh\_oh@s:cym\&eng \&=laugh ! //
\glsurface ELE:  uh\_oh$^{C}_{E}$ !  //
\glauto \%aut  uh\_oh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng "uh-oh!" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" oh@s:cym\&eng ti (y)n meddwl fy(ddi) di (y)n canu ? //
\glsurface ELE:  oh$^{C}_{E}$ ti yn meddwl fyddi di yn canu ?  //
\glauto \%aut  oh{\scriptsize .IM} you{\scriptsize .PRON.2S} stative{\scriptsize .STAT} think{\scriptsize .V.INFIN} be{\scriptsize .V.2S.FUT+SM} you{\scriptsize .PRON.2S+SM} stative{\scriptsize .STAT} sing{\scriptsize .V.INFIN}   //
\glmanual \%gls  IM PRON{\scriptsize .2S} PRT think{\scriptsize .NONFIN} be{\scriptsize .2S.FUT} PRON{\scriptsize .2S} PRT sing{\scriptsize .NONFIN}   //
\gleng "oh, do you think you'll be singing?" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat meddai fi de +" . //
\glsurface ELE:  meddai fi de .  //
\glauto \%aut  say{\scriptsize .V.3S.IMPERF} I{\scriptsize .PRON.1S+SM} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  say{\scriptsize .1S.PAST} PRON{\scriptsize .1S} TAG   //
\gleng I said, right. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" well@s:cym\&eng ! //
\glsurface ELE:  well$^{C}_{E}$ !  //
\glauto \%aut  well{\scriptsize .ADV}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat meddai hi +" . //
\glsurface ELE:  meddai hi .  //
\glauto \%aut  say{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  say{\scriptsize .3S.PAST} PRON{\scriptsize .3SF}   //
\gleng she said. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat um@s:cym\&eng \# +. . . //
\glsurface ELE:  um$^{C}_{E}$ .  //
\glauto \%aut  um{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" mae (y)r athrawon yn deud bod fi (y)n canu mewn tune@s:cym\&eng . //
\glsurface ELE:  mae yr athrawon yn deud bod fi yn canu mewn tune$^{C}_{E}$ .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} the{\scriptsize .DET.DEF} teachers{\scriptsize .N.M.PL} stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN} be{\scriptsize .V.INFIN} I{\scriptsize .PRON.1S+SM} stative{\scriptsize .STAT} sing{\scriptsize .V.INFIN} in{\scriptsize .PREP} tune{\scriptsize .N.SG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} DET teachers PRT say{\scriptsize .NONFIN} be{\scriptsize .NONFIN} PRON{\scriptsize .1S} PRT sing{\scriptsize .NONFIN} in tune   //
\gleng "the teachers say I sing in tune." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a mae (y)na rei (w)chi [?] (y)n canu bob \&d +// . //
\glsurface ELE:  a mae yna rei wchi yn canu bob .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} there{\scriptsize .ADV} some{\scriptsize .PRON+SM} know{\scriptsize .V.2P.PRES.SPOKEN} stative{\scriptsize .STAT} sing{\scriptsize .V.INFIN} each{\scriptsize .PREQ+SM}   //
\glmanual \%gls  and be{\scriptsize .3S.PRES} there some know{\scriptsize .2PL} PRT sing{\scriptsize .NONFIN} every   //
\gleng and there are some, you know, who sing every \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wedyn dyma hi dechrau canu wedyn \# bob +. . . //
\glsurface ELE:  wedyn dyma hi dechrau canu wedyn bob .  //
\glauto \%aut  afterwards{\scriptsize .ADV} this{\scriptsize .is.V} she{\scriptsize .PRON.F.3S} begin{\scriptsize .V.INFIN} sing{\scriptsize .V.INFIN} afterwards{\scriptsize .ADV} each{\scriptsize .PREQ+SM}   //
\glmanual \%gls  after here PRON{\scriptsize .3SF} PRT start{\scriptsize .NONFIN} sing{\scriptsize .NONFIN} after every  //
\gleng then she started to sing then, every \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ALN:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat sut [?] de dw (ddi)m yn gwybod . //
\glsurface ELE:  sut de dw ddim yn gwybod .  //
\glauto \%aut  how{\scriptsize .INT} isn't\_it{\scriptsize .IM[or]south.N.M.SG[or]right.N.M.SG[or]tea.N.M.SG+SM} be{\scriptsize .V.1S.PRES.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN}   //
\glmanual \%gls  how TAG be{\scriptsize .1S.PRES} NEG PRT know{\scriptsize .NONFIN}   //
\gleng how, I don't know. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=sniff gawn ni weld \# be ddigwyddith . //
\glsurface ELE:  gawn ni weld be ddigwyddith .  //
\glauto \%aut  get{\scriptsize .V.1P.PRES+SM} we{\scriptsize .PRON.1P} see{\scriptsize .V.INFIN+SM} what{\scriptsize .INT} happen{\scriptsize .V.3S.FUT.SPOKEN+SM}   //
\glmanual \%gls  get{\scriptsize .1PL.NONPAST} PRON{\scriptsize .1PL} see{\scriptsize .NONFIN} what happen{\scriptsize .3S.NONPAST}   //
\gleng we'll see what happens. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng mae (y)n medru dydy ond +/ . //
\glsurface ALN:  oh$^{C}_{E}$ mae yn medru dydy ond .  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} be{\scriptsize .V.INFIN} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} but{\scriptsize .CONJ}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} PRT can{\scriptsize .NONFIN} be{\scriptsize .3S.PRES.NEG} but   //
\gleng oh, she can do it, can't she, but \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yndy yndy . //
\glsurface ELE:  yndy yndy .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.EMPH} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES}   //
\gleng yes, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wneith hi wneud yr Eisteddfod@s:cym\&eng ohoni ? //
\glsurface ALN:  wneith hi wneud yr Eisteddfod$^{C}_{E}$ ohoni ?  //
\glauto \%aut  do{\scriptsize .V.3S.FUT+SM} she{\scriptsize .PRON.F.3S} make{\scriptsize .V.INFIN+SM} the{\scriptsize .DET.DEF} name from\_her{\scriptsize .PREP+PRON.F.3S}   //
\glmanual \%gls  do{\scriptsize .3S.NONPAST} PRON{\scriptsize .3SF} do{\scriptsize .NONFIN} DET Eisteddfod of{\scriptsize .3SF}   //
\gleng will she do the Eisteddfod, of it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yn (y)r ysgol ella wneith hi sti . //
\glsurface ELE:  yn yr ysgol ella wneith hi sti .  //
\glauto \%aut  in{\scriptsize .PREP} the{\scriptsize .DET.DEF} school{\scriptsize .N.F.SG} maybe{\scriptsize .ADV.SPOKEN} do{\scriptsize .V.3S.FUT+SM} she{\scriptsize .PRON.F.3S} you\_know{\scriptsize .IM}   //
\glmanual \%gls  in DET school maybe do{\scriptsize .3S.NONPAST} PRON{\scriptsize .3SF} know{\scriptsize .2S}   //
\gleng she might do in school, you know. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae hi (we)di magu dipyn bach mwy o hyder erbyn hyn do ? //
\glsurface ELE:  mae hi wedi magu dipyn bach mwy o hyder erbyn hyn do ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} she{\scriptsize .PRON.F.3S} after{\scriptsize .PREP} rear{\scriptsize .V.INFIN} little{\scriptsize .bit.N.M.SG+SM} small{\scriptsize .ADJ} more{\scriptsize .ADJ.COMP} of{\scriptsize .PREP} confidence{\scriptsize .N.M.SG} by{\scriptsize .PREP} this{\scriptsize .PRON.SP} yes{\scriptsize .ADV.PAST}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRON{\scriptsize .3SF} PRT{\scriptsize .PAST} nurture{\scriptsize .NONFIN} a\_bit little more of confidence by this yes   //
\gleng she's built up a little bit more confidence by now, hasn't she? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat do . //
\glsurface ALN:  do .  //
\glauto \%aut  yes{\scriptsize .ADV.PAST}   //
\glmanual \%gls  yes   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat os wneith hi lwyddo yn Theatr\_Gwynedd@s:cym\&eng $<$fydd (y)na (dd)im$>$ [?] stopio hi . //
\glsurface ALN:  os wneith hi lwyddo yn Theatr\_Gwynedd$^{C}_{E}$ fydd yna ddim stopio hi .  //
\glauto \%aut  if{\scriptsize .CONJ} do{\scriptsize .V.3S.FUT+SM} she{\scriptsize .PRON.F.3S} succeed{\scriptsize .V.INFIN+SM} in{\scriptsize .PREP} name be{\scriptsize .V.3S.FUT+SM} there{\scriptsize .ADV} not{\scriptsize .ADV+SM} stop{\scriptsize .V.INFIN} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  if do{\scriptsize .3S.NONPAST} PRON{\scriptsize .3SF} succeed{\scriptsize .NONFIN} in Theatr\_Gwynedd be{\scriptsize .3S.FUT} there NEG stop{\scriptsize .NONFIN} PRON{\scriptsize .3SF}   //
\gleng if she does well in Theatr Gwynedd, there'll be no stopping her. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat fydd hi (y)n canu trwy (y)r adeg . //
\glsurface ALN:  fydd hi yn canu trwy yr adeg .  //
\glauto \%aut  be{\scriptsize .V.3S.FUT+SM} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} sing{\scriptsize .V.INFIN} through{\scriptsize .PREP} the{\scriptsize .DET.DEF} time{\scriptsize .N.F.SG}   //
\glmanual \%gls  be{\scriptsize .3S.FUT} PRON{\scriptsize .3SF} PRT sing{\scriptsize .NONFIN} through DET time   //
\gleng she'll be singing all the time. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat um@s:cym\&eng +. . . //
\glsurface ELE:  um$^{C}_{E}$ .  //
\glauto \%aut  um{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat gawn ni weld . //
\glsurface ALN:  gawn ni weld .  //
\glauto \%aut  get{\scriptsize .V.1P.PRES+SM} we{\scriptsize .PRON.1P} see{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  get{\scriptsize .1PL.NONPAST} PRON{\scriptsize .1PL} see{\scriptsize .NONFIN}   //
\gleng we'll see. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie . //
\glsurface ELE:  ie .  //
\glauto \%aut  yes{\scriptsize .ADV}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd [?] Charlie@s:cym\&eng (we)di mwynhau heddiw oedd ? //
\glsurface ELE:  oedd Charlie$^{C}_{E}$ wedi mwynhau heddiw oedd ?  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} name after{\scriptsize .PREP} enjoy{\scriptsize .V.INFIN} today{\scriptsize .ADV} be{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} Charlie PRT{\scriptsize .PAST} enjoy{\scriptsize .NONFIN} today be{\scriptsize .3S.IMP}   //
\gleng Charlie enjoyed today, did he? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd . //
\glsurface ALN:  oedd .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  be{\scriptsize .3S.IMP}   //
\gleng he did. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$oedd o (y)n$>$ [///] \# mi ddeudodd o bod o ddim isio wneud o . //
\glsurface ALN:  oedd o yn mi ddeudodd o bod o ddim isio wneud o .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} yn{\scriptsize .PRT} aff{\scriptsize .PRT} say{\scriptsize .V.3S.PAST+SM} he{\scriptsize .PRON.M.3S.SPOKEN} be{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} not{\scriptsize .ADV+SM} want{\scriptsize .N.M.SG} make{\scriptsize .V.INFIN+SM} he{\scriptsize .PRON.M.3S.SPOKEN}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} PRT PRT say{\scriptsize .3S.PAST} PRON{\scriptsize .3SM} be{\scriptsize .NONFIN} PRON{\scriptsize .3SM} NEG want do{\scriptsize .NONFIN} PRON{\scriptsize .3SM}   //
\gleng he said he didn't want to do it. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a wedyn mi benderfynodd o heddiw bod isio \&r trio . //
\glsurface ALN:  a wedyn mi benderfynodd o heddiw bod isio trio .  //
\glauto \%aut  and{\scriptsize .CONJ} afterwards{\scriptsize .ADV} aff{\scriptsize .PRT} decide{\scriptsize .V.3S.PAST+SM} he{\scriptsize .PRON.M.3S.SPOKEN} today{\scriptsize .ADV} be{\scriptsize .V.INFIN} want{\scriptsize .N.M.SG} try{\scriptsize .V.INFIN}   //
\glmanual \%gls  and then PRT decide{\scriptsize .3S.PAST} PRON{\scriptsize .3SM} today be{\scriptsize .NONFIN} want try{\scriptsize .NONFIN}   //
\gleng and then he decided today that he wants to try. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$oedd o$>$ [?] [///] bod isio trio fo . //
\glsurface ALN:  oedd o bod isio trio fo .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} be{\scriptsize .V.INFIN} want{\scriptsize .N.M.SG} try{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} be{\scriptsize .NONFIN} want try{\scriptsize .NONFIN} PRON{\scriptsize .3SM}   //
\gleng that he wants to try it. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat er\_mwyn chdi a fi . //
\glsurface ALN:  er\_mwyn chdi a fi .  //
\glauto \%aut  for\_the\_sake\_of{\scriptsize .PREP} you{\scriptsize .PRON} and{\scriptsize .CONJ} I{\scriptsize .PRON.1S+SM}   //
\glmanual \%gls  for PRON{\scriptsize .2S} and PRON{\scriptsize .1S}   //
\gleng for our sake. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng blodyn ! //
\glsurface ELE:  oh$^{C}_{E}$ blodyn !  //
\glauto \%aut  oh{\scriptsize .IM} flower{\scriptsize .N.M.SG}   //
\glmanual \%gls  IM flower   //
\gleng oh, what a dear! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=deep\_breath . //
\glsurface ALN:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=slurp gobeithio bod o (ddi)m yn roi pressure@s:eng arno fo ei hun . //
\glsurface ELE:  gobeithio bod o ddim yn roi pressure$^{E}$ arno fo ei hun .  //
\glauto \%aut  hope{\scriptsize .V.INFIN} be{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} give{\scriptsize .V.INFIN+SM} pressure{\scriptsize .N.SG} on\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S} her{\scriptsize .ADJ.POSS[or]his.ADJ.POSS} self{\scriptsize .PRON.SG}   //
\glmanual \%gls  hope{\scriptsize .NONFIN} be{\scriptsize .NONFIN} PRON{\scriptsize .3SM} NEG PRT put{\scriptsize .NONFIN} pressure on{\scriptsize .3SM} PRON{\scriptsize .3SM} POSS{\scriptsize .3S} self   //
\gleng I hope he's not putting pressure on himself. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a wedyn $<$wnaeth um@s:cym\&eng$>$ [//] aeth o fewn . //
\glsurface ALN:  a wedyn wnaeth um$^{C}_{E}$ aeth o fewn .  //
\glauto \%aut  and{\scriptsize .CONJ} afterwards{\scriptsize .ADV} do{\scriptsize .V.3S.PAST+SM} um{\scriptsize .IM} go{\scriptsize .V.3S.PAST} he{\scriptsize .PRON.M.3S.SPOKEN} in{\scriptsize .PREP+SM}   //
\glmanual \%gls  and then do{\scriptsize .3S.PAST} IM go{\scriptsize .3S.PAST} PRON{\scriptsize .3SM} in   //
\gleng and then he went in. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a wnaeth Izzy@s:cym\&eng \# ddeud +"/ . //
\glsurface ALN:  a wnaeth Izzy$^{C}_{E}$ ddeud .  //
\glauto \%aut  and{\scriptsize .CONJ} do{\scriptsize .V.3S.PAST+SM} name say{\scriptsize .V.INFIN.SPOKEN+SM}   //
\glmanual \%gls  and do{\scriptsize .3S.PAST} Izzy say{\scriptsize .NONFIN}   //
\gleng and Izzy said: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" oh@s:cym\&eng gobeithio bod gen ti lais fath â dy fam@s:cym\&eng ddim llais fath â dy dad@s:cym\&eng ! //
\glsurface ALN:  oh$^{C}_{E}$ gobeithio bod gen ti lais fath â dy fam$^{C}_{E}$ ddim llais fath â dy dad$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM} hope{\scriptsize .V.INFIN} be{\scriptsize .V.INFIN} by{\scriptsize .PREP.SPOKEN} you{\scriptsize .PRON.2S} voice{\scriptsize .N.M.SG+SM} type{\scriptsize .N.M.SG+SM} as{\scriptsize .CONJ} your{\scriptsize .ADJ.POSS} fam{\scriptsize .N.SG[or]mam.N.SG+SM} not{\scriptsize .ADV+SM} voice{\scriptsize .N.M.SG} type{\scriptsize .N.M.SG+SM} as{\scriptsize .CONJ} your{\scriptsize .ADJ.POSS} dad{\scriptsize .N.SG[or]tad.N.SG+SM}   //
\glmanual \%gls  IM hope{\scriptsize .NONFIN} be{\scriptsize .NONFIN} with{\scriptsize .2S} PRON{\scriptsize .2S} voice kind with POSS{\scriptsize .2S} mother NEG voice kind with POSS{\scriptsize .2S} father   //
\gleng "oh, hopefully you've got a voice like your mother's, not a voice like your father's!" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=gasp na gobeithio bod gynno fo lais fath â dy dad@s:cym\&eng a dim fath â dy fam ! //
\glsurface ELE:  na gobeithio bod gynno fo lais fath â dy dad$^{C}_{E}$ a dim fath â dy fam !  //
\glauto \%aut  neg{\scriptsize .PRT} hope{\scriptsize .V.INFIN} be{\scriptsize .V.INFIN} with\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S} voice{\scriptsize .N.M.SG+SM} type{\scriptsize .N.M.SG+SM} as{\scriptsize .CONJ} your{\scriptsize .ADJ.POSS} dad{\scriptsize .N.SG[or]tad.N.SG+SM} and{\scriptsize .CONJ} not{\scriptsize .ADV} type{\scriptsize .N.M.SG+SM} as{\scriptsize .CONJ} your{\scriptsize .ADJ.POSS} mother{\scriptsize .N.F.SG+SM}   //
\glmanual \%gls  no hope{\scriptsize .NONFIN} be{\scriptsize .NONFIN} with{\scriptsize .3SM} PRON{\scriptsize .3SM} voice kind with POSS{\scriptsize .2S} father and NEG kind with POSS{\scriptsize .2S} mother   //
\gleng oh, no, hopefully he's got a voice like your father and not like your mother! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$do'n i (ddi)m isio deud \&n$>$ [///] wnes i (ddi)m deud dim\_byd . //
\glsurface ALN:  do'n i ddim isio deud wnes i ddim deud dim\_byd .  //
\glauto \%aut  unk to{\scriptsize .PREP} not{\scriptsize .ADV+SM} want{\scriptsize .N.M.SG} say{\scriptsize .V.INFIN.SPOKEN} do{\scriptsize .V.1S.PAST.SPOKEN+SM} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} say{\scriptsize .V.INFIN.SPOKEN} anything{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .1S.IMP.NEG} PRON{\scriptsize .1S} NEG want say{\scriptsize .NONFIN} do{\scriptsize .1S.PAST} PRON{\scriptsize .1S} NEG say{\scriptsize .NONFIN} nothing   //
\gleng I didn't want to say. . . I said nothing. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wnes [?] [/] wnes [?] i (ddi)m gair . //
\glsurface ALN:  wnes wnes i ddim gair .  //
\glauto \%aut  do{\scriptsize .V.1S.PAST.SPOKEN+SM} do{\scriptsize .V.1S.PAST.SPOKEN+SM} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} word{\scriptsize .N.M.SG}   //
\glmanual \%gls  do{\scriptsize .1S.PAST} do{\scriptsize .1S.PAST} PRON{\scriptsize .1S} NEG word   //
\gleng I didn't say a word. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng [=! laughs] \&=laugh ! //
\glsurface ELE:  oh$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond oedd gynni hi wên mawr ar ei gwyneb . //
\glsurface ALN:  ond oedd gynni hi wên mawr ar ei gwyneb .  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} with\_her{\scriptsize .PREP+PRON.F.3S} she{\scriptsize .PRON.F.3S} smile{\scriptsize .N.F.SG+SM} big{\scriptsize .ADJ} on{\scriptsize .PREP} her{\scriptsize .ADJ.POSS[or]his.ADJ.POSS} face{\scriptsize .N.M.SG}   //
\glmanual \%gls  but be{\scriptsize .3S.IMP} with{\scriptsize .3SF} PRON{\scriptsize .3SF} smile big on POSS{\scriptsize .3S} face   //
\gleng but she had a big smile on her face. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yn deud wrth Charlie@s:cym\&eng neu deud wrthat ti oedd o ? //
\glsurface ELE:  yn deud wrth Charlie$^{C}_{E}$ neu deud wrthat ti oedd o ?  //
\glauto \%aut  stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN} by{\scriptsize .PREP} name or{\scriptsize .CONJ} say{\scriptsize .V.INFIN.SPOKEN} to\_you{\scriptsize .PREP+PRON.2S} you{\scriptsize .PRON.2S} be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN}   //
\glmanual \%gls  PRT say{\scriptsize .NONFIN} to Charlie or say{\scriptsize .NONFIN} to{\scriptsize .2S} PRON{\scriptsize .2S} be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM}   //
\gleng was it said to Charlie or to you? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$deud wrth \&g$>$ [///] er@s:cym\&eng oedd hi yn deud wrth Charlie@s:cym\&eng Charlie@s:cym\&eng ddeu(d) (wr)tha fi . //
\glsurface ALN:  deud wrth er$^{C}_{E}$ oedd hi yn deud wrth Charlie$^{C}_{E}$ Charlie$^{C}_{E}$ ddeud wrtha fi .  //
\glauto \%aut  say{\scriptsize .V.INFIN.SPOKEN} by{\scriptsize .PREP} er{\scriptsize .IM} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN} by{\scriptsize .PREP} name name say{\scriptsize .V.INFIN.SPOKEN+SM} to\_me{\scriptsize .PREP+PRON.1S} I{\scriptsize .PRON.1S+SM}   //
\glmanual \%gls  say{\scriptsize .NONFIN} to IM be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT say{\scriptsize .NONFIN} to Charlie Charlie say{\scriptsize .NONFIN} to{\scriptsize .1S} PRON{\scriptsize .1S}   //
\gleng she'd said it to Charlie, Charlie said it to me. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ \&=gasp . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat case@s:cym\&eng ! //
\glsurface ELE:  case$^{C}_{E}$ !  //
\glauto \%aut  case{\scriptsize .N.SG}   //
\glmanual \%gls     //
\gleng what a case! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng well@s:cym\&eng dan ni gwybod hynny dydan ? //
\glsurface ALN:  oh$^{C}_{E}$ well$^{C}_{E}$ dan ni gwybod hynny dydan ?  //
\glauto \%aut  oh{\scriptsize .IM} well{\scriptsize .ADV} be{\scriptsize .V.1P.PRES.SPOKEN} we{\scriptsize .PRON.1P} know{\scriptsize .V.INFIN} that{\scriptsize .PRON.SP} be{\scriptsize .V.1P.PRES.NEG[or]be.V.1P.PRES.NEG+SM}   //
\glmanual \%gls  IM well be{\scriptsize .1PL.PRES} PRON{\scriptsize .1PL} know{\scriptsize .NONFIN} that be{\scriptsize .1PL.PRES.NEG}   //
\gleng oh, well, we know that, don't we? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ \&=gasp . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd Brian@s:cym\&eng yn deud bod hi yn [/] \#\# yn yr ysgol lle oedd o (y)n arfer dysgu . //
\glsurface ALN:  oedd Brian$^{C}_{E}$ yn deud bod hi yn yn yr ysgol lle oedd o yn arfer dysgu .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} name stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN} be{\scriptsize .V.INFIN} she{\scriptsize .PRON.F.3S} yn{\scriptsize .PRT} in{\scriptsize .PREP} the{\scriptsize .DET.DEF} school{\scriptsize .N.F.SG} where{\scriptsize .INT[or]place.N.M.SG} be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} use{\scriptsize .V.INFIN} teach{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} Brian PRT say{\scriptsize .NONFIN} be{\scriptsize .NONFIN} PRON{\scriptsize .3SF} in in DET school where be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} PRT use{\scriptsize .NONFIN} learn{\scriptsize .NONFIN}   //
\gleng Brian was saying that she was at the school where he used to teach. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie yeah@s:cym\&eng . //
\glsurface ELE:  ie yeah$^{C}_{E}$ .  //
\glauto \%aut  yes{\scriptsize .ADV} yeah{\scriptsize .ADV}   //
\glmanual \%gls  yes yeah   //
\gleng yes, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a hi $<$yn arfer$>$ [?] dysgu . //
\glsurface ALN:  a hi yn arfer dysgu .  //
\glauto \%aut  and{\scriptsize .CONJ} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} use{\scriptsize .V.INFIN} teach{\scriptsize .V.INFIN}   //
\glmanual \%gls  and PRON{\scriptsize .3SF} PRT use{\scriptsize .NONFIN} learn{\scriptsize .NONFIN}   //
\gleng and she used to teach. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$oedd hi \# um@s:cym\&eng$>$ [/] oedd hi mor mor wahanol i weddill y staff@s:cym\&eng . //
\glsurface ALN:  oedd hi um$^{C}_{E}$ oedd hi mor mor wahanol i weddill y staff$^{C}_{E}$ .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} um{\scriptsize .IM} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} so{\scriptsize .ADV} so{\scriptsize .ADV} different{\scriptsize .ADJ+SM} to{\scriptsize .PREP} remnant{\scriptsize .N.M.SG+SM[or]remainder.N.M.SG+SM} the{\scriptsize .DET.DEF} staff{\scriptsize .N.SG}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} IM be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} so so different to remainder DET staff   //
\gleng she was so very different from the rest of the staff. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat just@s:cym\&eng gwbl gwbl gwbl wahanol . //
\glsurface ALN:  just$^{C}_{E}$ gwbl gwbl gwbl wahanol .  //
\glauto \%aut  just{\scriptsize .ADV[or]just.ADJ} all{\scriptsize .ADJ+SM} all{\scriptsize .ADJ+SM} all{\scriptsize .ADJ+SM} different{\scriptsize .ADJ+SM}   //
\glmanual \%gls  just all all all different   //
\gleng just totally totally totally different. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat sefyll allan yn llwyr . //
\glsurface ALN:  sefyll allan yn llwyr .  //
\glauto \%aut  stand{\scriptsize .V.INFIN} out{\scriptsize .ADV} stative{\scriptsize .STAT} complete{\scriptsize .ADJ}   //
\glmanual \%gls  stand{\scriptsize .NONFIN} out PRT entire   //
\gleng stood out entirely. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yn enwedig pan oedd [?] hi mynd allan efo nhw . //
\glsurface ALN:  yn enwedig pan oedd hi mynd allan efo nhw .  //
\glauto \%aut  stative{\scriptsize .STAT} especially{\scriptsize .ADJ} when{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} go{\scriptsize .V.INFIN} out{\scriptsize .ADV} with{\scriptsize .PREP} they{\scriptsize .PRON.3P}   //
\glmanual \%gls  PRT special when be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} go{\scriptsize .NONFIN} out with PRON{\scriptsize .3PL}   //
\gleng especially when she went out with them. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yep@s:cym\&eng ! //
\glsurface ELE:  yep$^{C}_{E}$ !  //
\glauto \%aut  yep{\scriptsize .ADV}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd ddim quite@s:cym\&eng mor wyllt â Jenny@s:cym\&eng chwaith . //
\glsurface ALN:  oedd ddim quite$^{C}_{E}$ mor wyllt â Jenny$^{C}_{E}$ chwaith .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} not{\scriptsize .ADV+SM} quite{\scriptsize .ADV} as{\scriptsize .ADJ} wild{\scriptsize .ADJ+SM} as{\scriptsize .CONJ} name neither{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} NEG quite so wild as Jenny either   //
\gleng she wasn't quite as wild as Jenny either. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be +! ? //
\glsurface ELE:  be ?  //
\glauto \%aut  what{\scriptsize .INT}   //
\glmanual \%gls  what   //
\gleng what!? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae hi (we)di stopio mynd allan efo nhw . //
\glsurface ALN:  mae hi wedi stopio mynd allan efo nhw .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} she{\scriptsize .PRON.F.3S} after{\scriptsize .PREP} stop{\scriptsize .V.INFIN} go{\scriptsize .V.INFIN} out{\scriptsize .ADV} with{\scriptsize .PREP} they{\scriptsize .PRON.3P}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRON{\scriptsize .3SF} PRT{\scriptsize .PAST} stop{\scriptsize .NONFIN} go{\scriptsize .NONFIN} out with PRON{\scriptsize .3PL}   //
\gleng she's stopped going out with them. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ar\_ôl iddi cael ei \# gwahodd \# i adael un \# lle oedden nhw . //
\glsurface ALN:  ar\_ôl iddi cael ei gwahodd i adael un lle oedden nhw .  //
\glauto \%aut  after{\scriptsize .PREP} to\_her{\scriptsize .PREP+PRON.F.3S} have{\scriptsize .V.INFIN} her{\scriptsize .ADJ.POSS[or]his.ADJ.POSS} invite{\scriptsize .V.INFIN} to{\scriptsize .PREP} leave{\scriptsize .V.INFIN+SM} one{\scriptsize .NUM} where{\scriptsize .INT[or]place.N.M.SG} be{\scriptsize .V.3P.IMPERF.SPOKEN} they{\scriptsize .PRON.3P}   //
\glmanual \%gls  after to{\scriptsize .3SF} have{\scriptsize .NONFIN} POSS{\scriptsize .3S} invite{\scriptsize .NONFIN} to leave{\scriptsize .NONFIN} one place be{\scriptsize .3PL.IMP} PRON{\scriptsize .3PL}   //
\gleng after she got invited to leave one place they were at. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ooh@s:cym\&eng ! //
\glsurface ELE:  ooh$^{C}_{E}$ !  //
\glauto \%aut  ooh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond er@s:cym\&eng +/ . //
\glsurface ALN:  ond er$^{C}_{E}$ .  //
\glauto \%aut  but{\scriptsize .CONJ} er{\scriptsize .IM}   //
\glmanual \%gls  but IM   //
\gleng but, er \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$maen nhw (y)n llanast (y)na$>$ [?] . //
\glsurface ELE:  maen nhw yn llanast yna .  //
\glauto \%aut  be{\scriptsize .V.3P.PRES.SPOKEN} they{\scriptsize .PRON.3P} yn{\scriptsize .PRT} mess{\scriptsize .N.M.SG} there{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL} PRT mess there   //
\gleng they're in a mess there. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd hi (y)n deud $<$bod hi (y)n um@s:cym\&eng$>$ [//] \#\#\# bod Izzy@s:cym\&eng yn [/] yn [/] yn sefyll allan . //
\glsurface ALN:  oedd hi yn deud bod hi yn um$^{C}_{E}$ bod Izzy$^{C}_{E}$ yn yn yn sefyll allan .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN} be{\scriptsize .V.INFIN} she{\scriptsize .PRON.F.3S} yn{\scriptsize .PRT} um{\scriptsize .IM} be{\scriptsize .V.INFIN} name yn{\scriptsize .PRT} yn{\scriptsize .PRT} stative{\scriptsize .STAT} stand{\scriptsize .V.INFIN} out{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT say{\scriptsize .NONFIN} be{\scriptsize .NONFIN} PRON{\scriptsize .3SF} PRT IM be{\scriptsize .NONFIN} Izzy PRT PRT PRT stand{\scriptsize .NONFIN} out   //
\gleng she was saying that Izzy stood out. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmhm@s:cym\&eng . //
\glsurface ELE:  mmhm$^{C}_{E}$ .  //
\glauto \%aut  mmhm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat fath â rywun gwbl gwbl wahanol i gweddill y staff@s:cym\&eng . //
\glsurface ALN:  fath â rywun gwbl gwbl wahanol i gweddill y staff$^{C}_{E}$ .  //
\glauto \%aut  type{\scriptsize .N.M.SG+SM} with{\scriptsize .PREP} someone{\scriptsize .N.M.SG+SM} all{\scriptsize .ADJ+SM} all{\scriptsize .ADJ+SM} different{\scriptsize .ADJ+SM} to{\scriptsize .PREP} remnant{\scriptsize .N.M.SG[or]remainder.N.M.SG} the{\scriptsize .DET.DEF} staff{\scriptsize .N.SG}   //
\glmanual \%gls  kind with somebody all all different to remainder DET staff   //
\gleng like someone totally totally different from the rest of the staff. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$oedden ni (y)n gwybod hyn de$>$ [?] ? //
\glsurface ELE:  oedden ni yn gwybod hyn de ?  //
\glauto \%aut  be{\scriptsize .V.3P.IMPERF.SPOKEN} we{\scriptsize .PRON.1P} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN} this{\scriptsize .PRON.SP} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  be{\scriptsize .1PL.IMP} PRON{\scriptsize .1PL} PRT know{\scriptsize .NONFIN} this TAG   //
\gleng we knew that, right? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedden ond (ba)set ti (y)n disgwyl falle mewn ysgol uwchradd bod (y)na \#\# +/ . //
\glsurface ALN:  oedden ond baset ti yn disgwyl falle mewn ysgol uwchradd bod yna .  //
\glauto \%aut  be{\scriptsize .V.3P.IMPERF.SPOKEN} but{\scriptsize .CONJ} be{\scriptsize .V.2S.PLUPERF} you{\scriptsize .PRON.2S} stative{\scriptsize .STAT} expect{\scriptsize .V.INFIN} maybe{\scriptsize .ADV.SPOKEN} in{\scriptsize .PREP} school{\scriptsize .N.F.SG} superior{\scriptsize .ADJ} be{\scriptsize .V.INFIN} there{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .1PL.IMP} but be{\scriptsize .2S.CONDIT} PRON{\scriptsize .2S} PRT expect{\scriptsize .NONFIN} maybe in school secondary be{\scriptsize .NONFIN} there   //
\gleng yes, but you'd expect that maybe in a secondary school there'd be \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng . //
\glsurface ELE:  yeah$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, mwy nag un person@s:cym\&eng . //
\glsurface ALN:  mwy nag un person$^{C}_{E}$ .  //
\glauto \%aut  more{\scriptsize .ADJ.COMP} than{\scriptsize .CONJ} one{\scriptsize .NUM} person{\scriptsize .N.SG}   //
\glmanual \%gls  more than one person   //
\gleng more than one person. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ah@s:cym\&eng mae hi ar ben ei hun ! //
\glsurface ELE:  ah$^{C}_{E}$ mae hi ar ben ei hun !  //
\glauto \%aut  ah{\scriptsize .IM} be{\scriptsize .V.3S.PRES} she{\scriptsize .PRON.F.3S} on{\scriptsize .PREP} head{\scriptsize .N.M.SG+SM} her{\scriptsize .ADJ.POSS[or]his.ADJ.POSS[or]go.V.2S.PRES} self{\scriptsize .PRON.SG}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} PRON{\scriptsize .3SF} on head POSS{\scriptsize .3S} self   //
\gleng ah, she's on her own! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well@s:cym\&eng mae (y)na hynny . //
\glsurface ALN:  well$^{C}_{E}$ mae yna hynny .  //
\glauto \%aut  well{\scriptsize .ADV} be{\scriptsize .V.3S.PRES} there{\scriptsize .ADV} that{\scriptsize .PRON.SP}   //
\glmanual \%gls  well be{\scriptsize .3S.PRES} there that   //
\gleng well, there is that. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a wir +// . //
\glsurface ELE:  a wir .  //
\glauto \%aut  and{\scriptsize .CONJ} true{\scriptsize .ADJ+SM}   //
\glmanual \%gls  and true   //
\gleng and really \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd o (we)di canu \# +/ . //
\glsurface ELE:  oedd o wedi canu .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} after{\scriptsize .PREP} sing{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} PRT{\scriptsize .PAST} sing{\scriptsize .NONFIN}   //
\gleng he had sung. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie oedd . //
\glsurface ALN:  ie oedd .  //
\glauto \%aut  yes{\scriptsize .ADV} be{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  yes be{\scriptsize .3S.IMP}   //
\gleng yes, he had. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, yn [/] yn gry meddai fo . //
\glsurface ELE:  yn yn gry meddai fo .  //
\glauto \%aut  yn{\scriptsize .PRT} stative{\scriptsize .STAT} strong{\scriptsize .ADJ+SM} say{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S}   //
\glmanual \%gls  PRT PRT strong say{\scriptsize .3S.PAST} PRON{\scriptsize .3SM}   //
\gleng strongly, he said. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd o (we)di cael ryw fath o ymateb +. . . //
\glsurface ELE:  oedd o wedi cael ryw fath o ymateb .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} after{\scriptsize .PREP} have{\scriptsize .V.INFIN} some{\scriptsize .PREQ+SM} type{\scriptsize .N.M.SG+SM[or]bath.N.M.SG+SM} of{\scriptsize .PREP} respond{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} PRT{\scriptsize .PAST} have{\scriptsize .NONFIN} some kind of response   //
\gleng he got some sort of response \dots . //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng $<$wnaeth o (ddi)m$>$ [?] +/ . //
\glsurface ALN:  mmm$^{C}_{E}$ wnaeth o ddim .  //
\glauto \%aut  mmm{\scriptsize .IM} do{\scriptsize .V.3S.PAST+SM} he{\scriptsize .PRON.M.3S.SPOKEN} not{\scriptsize .ADV+SM}   //
\glmanual \%gls  IM do{\scriptsize .3S.PAST} PRON{\scriptsize .3SM} NEG   //
\gleng mmm, he didn't \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat poeni dw i \# de bod o: yn meddwl mai trio am \# rhan Joseph@s:cym\&eng mae o . //
\glsurface ELE:  poeni dw i de bod o yn meddwl mai trio am rhan Joseph$^{C}_{E}$ mae o .  //
\glauto \%aut  worry{\scriptsize .V.INFIN} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} isn't\_it{\scriptsize .IM[or]south.N.M.SG[or]right.N.M.SG[or]tea.N.M.SG+SM} be{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} think{\scriptsize .V.INFIN} that\_it\_is{\scriptsize .CONJ.FOCUS} try{\scriptsize .V.INFIN} for{\scriptsize .PREP} part{\scriptsize .N.F.SG} name be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN}   //
\glmanual \%gls  worry{\scriptsize .NONFIN} be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} TAG be{\scriptsize .NONFIN} PRON{\scriptsize .3SM} PRT think{\scriptsize .NONFIN} PRT try{\scriptsize .NONFIN} for part Joseph be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM}   //
\gleng I'm worried, right, that he thinks he's trying for the part of Joseph. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ALN:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ti (y)n gwybod ? //
\glsurface ELE:  ti yn gwybod ?  //
\glauto \%aut  you{\scriptsize .PRON.2S} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN}   //
\glmanual \%gls  PRON{\scriptsize .2S} PRT know{\scriptsize .NONFIN}   //
\gleng you know? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ac o'n i (y)n trio deud +"/ . //
\glsurface ELE:  ac o'n i yn trio deud .  //
\glauto \%aut  and{\scriptsize .CONJ} unk to{\scriptsize .PREP} stative{\scriptsize .STAT} try{\scriptsize .V.INFIN} say{\scriptsize .V.INFIN.SPOKEN}   //
\glmanual \%gls  and be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} PRT try{\scriptsize .NONFIN} say{\scriptsize .NONFIN}   //
\gleng and I was trying to say: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" oh@s:cym\&eng mae raid mai blwyddyn chwech fyddan nhw (y)n rhoid yn Joseph@s:cym\&eng . //
\glsurface ELE:  oh$^{C}_{E}$ mae raid mai blwyddyn chwech fyddan nhw yn rhoid yn Joseph$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES} necessity{\scriptsize .N.M.SG+SM} that\_it\_is{\scriptsize .CONJ.FOCUS} year{\scriptsize .N.F.SG} six{\scriptsize .NUM} be{\scriptsize .V.3P.FUT.SPOKEN+SM} they{\scriptsize .PRON.3P} yn{\scriptsize .PRT} give{\scriptsize .V.0.IMPERF} in{\scriptsize .PREP} name   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} necessity PRT year six be{\scriptsize .3PL.FUT} PRON{\scriptsize .3PL} PRT put{\scriptsize .NONFIN} PRT Joseph   //
\gleng "oh, surely it will be year six they give Joseph." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat meddai fi +" . //
\glsurface ELE:  meddai fi .  //
\glauto \%aut  say{\scriptsize .V.3S.IMPERF} I{\scriptsize .PRON.1S+SM}   //
\glmanual \%gls  say{\scriptsize .1S.PAST} PRON{\scriptsize .1S}   //
\gleng I said. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" y prif ran . //
\glsurface ELE:  y prif ran .  //
\glauto \%aut  the{\scriptsize .DET.DEF} principal{\scriptsize .PREQ} part{\scriptsize .N.F.SG+SM}   //
\glmanual \%gls  DET main part   //
\gleng "the main part." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat naci ["] meddai fo +"/ . //
\glsurface ELE:  naci meddai fo .  //
\glauto \%aut  no{\scriptsize .ADV} say{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S}   //
\glmanual \%gls  no say{\scriptsize .3S.PAST} PRON{\scriptsize .3SM}   //
\gleng "no", he said: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" pump a chwech . //
\glsurface ELE:  pump a chwech .  //
\glauto \%aut  five{\scriptsize .NUM} and{\scriptsize .CONJ} six{\scriptsize .NUM}   //
\glmanual \%gls  five and six   //
\gleng "five and six." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ofn sy gen i iddo gael ei siomi (fe)lly timod . //
\glsurface ELE:  ofn sy gen i iddo gael ei siomi felly timod .  //
\glauto \%aut  fear{\scriptsize .N.M.SG} which\_is{\scriptsize .V.3SP.PRES.REL} by{\scriptsize .PREP.SPOKEN} I{\scriptsize .PRON.1S} to\_him{\scriptsize .PREP+PRON.M.3S} have{\scriptsize .V.INFIN+SM} her{\scriptsize .ADJ.POSS[or]his.ADJ.POSS} disappoint{\scriptsize .V.INFIN} so{\scriptsize .ADV} know{\scriptsize .V.2S.PRES.SPOKEN}   //
\glmanual \%gls  fear be{\scriptsize .PRES.REL} with{\scriptsize .1S} PRON{\scriptsize .1S} to{\scriptsize .3SM} get{\scriptsize .NONFIN} POSS{\scriptsize .3S} disappoint{\scriptsize .NONFIN} thus know{\scriptsize .2S}   //
\gleng I'm afraid of him being disappointed, you know. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat er mab um@s:cym\&eng +. . . //
\glsurface ALN:  er mab um$^{C}_{E}$ .  //
\glauto \%aut  er{\scriptsize .IM} son{\scriptsize .N.M.SG} um{\scriptsize .IM}   //
\glmanual \%gls  although son IM   //
\gleng although, the son of \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat xxx . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be (y)dy enw hi ? //
\glsurface ALN:  be ydy enw hi ?  //
\glauto \%aut  what{\scriptsize .INT} be{\scriptsize .V.3S.PRES} name{\scriptsize .N.M.SG} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  what be{\scriptsize .3S.PRES} name PRON{\scriptsize .3SF}   //
\gleng what's her name? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat beth Hayley@s:cym\&eng ? //
\glsurface ELE:  beth Hayley$^{C}_{E}$ ?  //
\glauto \%aut  what{\scriptsize .INT} name   //
\glmanual \%gls  what Hayley   //
\gleng what, Hayley? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat Hayley@s:cym\&eng . //
\glsurface ALN:  Hayley$^{C}_{E}$ .  //
\glauto \%aut  name   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat gymaint o Hayleys@s:cym\&eng ! //
\glsurface ALN:  gymaint o Hayleys$^{C}_{E}$ !  //
\glauto \%aut  so{\scriptsize .much.ADJ+SM} from{\scriptsize .PREP} name   //
\glmanual \%gls  so\_much of Hayleys   //
\gleng so many Hayleys! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat alla i (ddi)m cadw fyny . //
\glsurface ALN:  alla i ddim cadw fyny .  //
\glauto \%aut  be\_able{\scriptsize .V.3S.PRES.SPOKEN+SM} I{\scriptsize .PRON.1S[or]to.PREP} not{\scriptsize .ADV+SM} keep{\scriptsize .V.INFIN} above{\scriptsize .ADV}   //
\glmanual \%gls  can{\scriptsize .1S.NONPAST} PRON{\scriptsize .1S} NEG keep{\scriptsize .NONFIN} up   //
\gleng I can't keep up. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well@s:cym\&eng peth ydy os oes gynno fo \# natur a ffordd o ei gwmpas o (we)di arfer a dipyn o hyder well@s:cym\&eng \# rhywun o flwyddyn chwech geith o p(a) (y)r un bynnag de ? //
\glsurface ELE:  well$^{C}_{E}$ peth ydy os oes gynno fo natur a ffordd o ei gwmpas o wedi arfer a dipyn o hyder well$^{C}_{E}$ rhywun o flwyddyn chwech geith o pa yr un bynnag de ?  //
\glauto \%aut  well{\scriptsize .ADV} thing{\scriptsize .N.M.SG} be{\scriptsize .V.3S.PRES} if{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES.INDEF[or]age.N.F.SG} with\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S} nature{\scriptsize .N.F.SG} and{\scriptsize .CONJ} road{\scriptsize .N.F.SG} of{\scriptsize .PREP} his{\scriptsize .ADJ.POSS} pair{\scriptsize .of.compasses.N.M.SG+SM[or]scope.N.M.SG+SM} he{\scriptsize .PRON.M.3S.SPOKEN} after{\scriptsize .PREP} use{\scriptsize .V.INFIN} and{\scriptsize .CONJ} little{\scriptsize .bit.N.M.SG+SM} of{\scriptsize .PREP} confidence{\scriptsize .N.M.SG} well{\scriptsize .ADV} someone{\scriptsize .N.M.SG} of{\scriptsize .PREP} year{\scriptsize .N.F.SG+SM} six{\scriptsize .NUM} get{\scriptsize .V.3S.PRES+SM} he{\scriptsize .PRON.M.3S.SPOKEN} which{\scriptsize .ADJ} the{\scriptsize .DET.DEF} one{\scriptsize .NUM} -ever{\scriptsize .ADJ} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  well thing be{\scriptsize .3S.PRES} if be{\scriptsize .3S.PRES} with{\scriptsize .3SM} PRON{\scriptsize .3SM} nature and way of POSS{\scriptsize .3SM} compass PRON{\scriptsize .3SM} PRT{\scriptsize .PAST} use{\scriptsize .NONFIN} and a\_bit of confidence well someone from year six have{\scriptsize .3S.NONPAST} PRON{\scriptsize .3SM} which DET one ever TAG   //
\gleng well, the thing is, if he's got a certain manner and way about him, used to it and a bit of confidence, well, it's someone from year six that's going to get it anyway, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$(ba)swn i (y)n \&gw$>$ [/] $<$(ba)swn i (y)n gweld bod$>$ [/] \# (ba)swn \&=sigh [/] \#\# (ba)swn i (y)n gweld fyddai chwarae drums@s:cym\&eng yn \#\# well i Charlie@s:cym\&eng . //
\glsurface ALN:  baswn i yn baswn i yn gweld bod baswn baswn i yn gweld fyddai chwarae drums$^{C}_{E}$ yn well i Charlie$^{C}_{E}$ .  //
\glauto \%aut  be{\scriptsize .V.1S.PLUPERF.SPOKEN} I{\scriptsize .PRON.1S} yn{\scriptsize .PRT} be{\scriptsize .V.1S.PLUPERF.SPOKEN} I{\scriptsize .PRON.1S} stative{\scriptsize .STAT} see{\scriptsize .V.INFIN} be{\scriptsize .V.INFIN} be{\scriptsize .V.1S.PLUPERF.SPOKEN} be{\scriptsize .V.1S.PLUPERF.SPOKEN} I{\scriptsize .PRON.1S} stative{\scriptsize .STAT} see{\scriptsize .V.INFIN} be{\scriptsize .V.3S.COND+SM} play{\scriptsize .V.INFIN} drum{\scriptsize .N.SG+PL} stative{\scriptsize .STAT} better{\scriptsize .ADJ.COMP+SM} to{\scriptsize .PREP} name   //
\glmanual \%gls  be{\scriptsize .1S.CONDIT} PRON{\scriptsize .1S} PRT be{\scriptsize .1S.CONDIT} PRON{\scriptsize .1S} PRT see{\scriptsize .NONFIN} be{\scriptsize .NONFIN} be{\scriptsize .1S.CONDIT} be{\scriptsize .1S.CONDIT} PRON{\scriptsize .1S} PRT see{\scriptsize .NONFIN} be{\scriptsize .3S.CONDIT} play{\scriptsize .NONFIN} drums PRT better for Charlie   //
\gleng I'd reckon that playing the drums would be better for Charlie. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ ia ella (ba)sai fo \# mwynhau hynny . //
\glsurface ELE:  ia ella basai fo mwynhau hynny .  //
\glauto \%aut  yes{\scriptsize .ADV} maybe{\scriptsize .ADV.SPOKEN} be{\scriptsize .V.3S.PLUPERF.SPOKEN} he{\scriptsize .PRON.M.3S} enjoy{\scriptsize .V.INFIN} that{\scriptsize .PRON.SP}   //
\glmanual \%gls  yes maybe be{\scriptsize .3S.CONDIT} PRON{\scriptsize .3SM} enjoy{\scriptsize .NONFIN} that   //
\gleng yes, he might enjoy that. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond timod \&=sigh $<$rhaid fo$>$ [?] gael y cyfle $<$os (y)dy o isio$>$ [?] wan de ? //
\glsurface ELE:  ond timod rhaid fo gael y cyfle os ydy o isio wan de ?  //
\glauto \%aut  but{\scriptsize .CONJ} know{\scriptsize .V.2S.PRES.SPOKEN} necessity{\scriptsize .N.M.SG} he{\scriptsize .PRON.M.3S} have{\scriptsize .V.INFIN+SM} the{\scriptsize .DET.DEF} opportunity{\scriptsize .N.M.SG} if{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} want{\scriptsize .N.M.SG} weak{\scriptsize .ADJ+SM[or]pale.ADJ+SM} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  but know{\scriptsize .2S} necessity PRON{\scriptsize .3SM} have{\scriptsize .NONFIN} DET opportunity if be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} want now TAG   //
\gleng but, you know, he has to have the chance if he wants to now, eh? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ oh@s:cym\&eng ie \# yeah@s:cym\&eng yeah@s:cym\&eng . //
\glsurface ALN:  oh$^{C}_{E}$ ie yeah$^{C}_{E}$ yeah$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM} yes{\scriptsize .ADV} yeah{\scriptsize .ADV} yeah{\scriptsize .ADV}   //
\glmanual \%gls  IM yes yeah yeah   //
\gleng oh yes, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond hefyd os (y)dy o (y)n wneud hynna $<$mae o (y)n$>$ [/] \# mae (y)n marcio fath â drummer@s:cym\&eng yr ysgol am y blwyddyn a hanner nesa dydy ? //
\glsurface ALN:  ond hefyd os ydy o yn wneud hynna mae o yn mae yn marcio fath â drummer$^{C}_{E}$ yr ysgol am y blwyddyn a hanner nesa dydy ?  //
\glauto \%aut  but{\scriptsize .CONJ} also{\scriptsize .ADV} if{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM} that{\scriptsize .PRON.SP.SPOKEN} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} in{\scriptsize .PREP} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} mark{\scriptsize .V.INFIN} type{\scriptsize .N.M.SG+SM} with{\scriptsize .PREP} drummer{\scriptsize .N.SG} the{\scriptsize .DET.DEF} school{\scriptsize .N.F.SG} for{\scriptsize .PREP} the{\scriptsize .DET.DEF} year{\scriptsize .N.F.SG} and{\scriptsize .CONJ} half{\scriptsize .N.M.SG} next{\scriptsize .ADJ.SUP.SPOKEN} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM}   //
\glmanual \%gls  but also if be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT do{\scriptsize .NONFIN} that be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT be{\scriptsize .3S.PRES} PRT mark{\scriptsize .NONFIN} kind with drummer DET school for DET year and half next be{\scriptsize .3S.PRES.NEG}   //
\gleng but also, if he does that, he's marked out as the school drummer for the next year and a half, isn't he? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yndy yndy . //
\glsurface ELE:  yndy yndy .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.EMPH} be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES}   //
\gleng he is, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat sy (y)n defnyddiol i [/] i w brofiad o . //
\glsurface ALN:  sy yn defnyddiol i i brofiad o .  //
\glauto \%aut  which\_is{\scriptsize .V.3SP.PRES.REL} stative{\scriptsize .STAT} useful{\scriptsize .ADJ} to{\scriptsize .PREP} to{\scriptsize .PREP} experience{\scriptsize .N.M.SG+SM} he{\scriptsize .PRON.M.3S.SPOKEN}   //
\glmanual \%gls  be{\scriptsize .PRES.REL} PRT useful for for POSS{\scriptsize .3SM} experience PRON{\scriptsize .3SM}  //
\gleng which is useful for his experience. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat efallai bod fi (y)n gweld pethau fath â chwaraewyr yn hytrach +// . //
\glsurface ALN:  efallai bod fi yn gweld pethau fath â chwaraewyr yn hytrach .  //
\glauto \%aut  perhaps{\scriptsize .CONJ} be{\scriptsize .V.INFIN} I{\scriptsize .PRON.1S+SM} stative{\scriptsize .STAT} see{\scriptsize .V.INFIN} things{\scriptsize .N.M.PL} type{\scriptsize .N.M.SG+SM} with{\scriptsize .PREP} players{\scriptsize .N.M.PL[or]players.N.M.PL} stative{\scriptsize .STAT} rather{\scriptsize .ADV}   //
\glmanual \%gls  maybe be{\scriptsize .NONFIN} PRON{\scriptsize .1S} PRT see{\scriptsize .NONFIN} things kind with players PRT rather   //
\gleng maybe I'm looking at things like a player, rather than \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ ydy . //
\glsurface ELE:  ydy .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES}   //
\glmanual \%gls  be{\scriptsize .3S.PRES}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie dw i gwybod . //
\glsurface ALN:  ie dw i gwybod .  //
\glauto \%aut  yes{\scriptsize .ADV} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} know{\scriptsize .V.INFIN}   //
\glmanual \%gls  yes be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} know{\scriptsize .NONFIN}   //
\gleng yes, I know. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wsti mae o isio bod yn rhan o (y)r \#\# +/ . //
\glsurface ELE:  wsti mae o isio bod yn rhan o yr .  //
\glauto \%aut  know{\scriptsize .V.2S.PRES.SPOKEN} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} want{\scriptsize .N.M.SG} be{\scriptsize .V.INFIN} yn{\scriptsize .PRT} part{\scriptsize .N.F.SG} of{\scriptsize .PREP} the{\scriptsize .DET.DEF}   //
\glmanual \%gls  know{\scriptsize .2S} be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} want be{\scriptsize .NONFIN} PRT part of DET   //
\gleng you see, he wants to be a part of the \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=sigh . //
\glsurface ALN:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, crew@s:cym\&eng dydy ? //
\glsurface ELE:  crew$^{C}_{E}$ dydy ?  //
\glauto \%aut  crew{\scriptsize .N.SG} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM}   //
\glmanual \%gls  crew be{\scriptsize .3S.PRES.NEG}   //
\gleng  \dots gang, doesn't he? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yndy . //
\glsurface ALN:  yndy .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3S.PRES}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dyna sy (y)n bwysig iddo fo wan de ? //
\glsurface ELE:  dyna sy yn bwysig iddo fo wan de ?  //
\glauto \%aut  that{\scriptsize .is.ADV} which\_is{\scriptsize .V.3SP.PRES.REL} stative{\scriptsize .STAT} important{\scriptsize .ADJ+SM} to\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S} weak{\scriptsize .ADJ+SM[or]pale.ADJ+SM} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  there be{\scriptsize .PRES.REL} PRT important to{\scriptsize .3SM} PRON{\scriptsize .3SM} now TAG   //
\gleng that's what's important to him now, right? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng yeah@s:cym\&eng . //
\glsurface ALN:  yeah$^{C}_{E}$ yeah$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV} yeah{\scriptsize .ADV}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$dw xxx$>$ [//] dw i (y)n gweld pethau (y)n wahanol dydw ? //
\glsurface ALN:  dw dw i yn gweld pethau yn wahanol dydw ?  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} stative{\scriptsize .STAT} see{\scriptsize .V.INFIN} things{\scriptsize .N.M.PL} stative{\scriptsize .STAT} different{\scriptsize .ADJ+SM} be{\scriptsize .V.1S.PRES.NEG}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT see{\scriptsize .NONFIN} things PRT different be{\scriptsize .1S.PRES.NEG}   //
\gleng I see things differently, don't I? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well@s:cym\&eng profiad ydy o i\_gyd de ond \# wsti \&m \&=sigh +/ . //
\glsurface ELE:  well$^{C}_{E}$ profiad ydy o i\_gyd de ond wsti .  //
\glauto \%aut  well{\scriptsize .ADV} experience{\scriptsize .N.M.SG} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} all{\scriptsize .ADJ} isn't\_it{\scriptsize .IM[or]south.N.M.SG[or]right.N.M.SG[or]tea.N.M.SG+SM} but{\scriptsize .CONJ} know{\scriptsize .V.2S.PRES.SPOKEN}   //
\glmanual \%gls  well experience be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} all TAG but know{\scriptsize .2S}   //
\gleng well, it's all experience, isn't it, but, you know \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well@s:cym\&eng os (dy)dy o ddim yn cael rhan \# +/ . //
\glsurface ALN:  well$^{C}_{E}$ os dydy o ddim yn cael rhan .  //
\glauto \%aut  well{\scriptsize .ADV} if{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} he{\scriptsize .PRON.M.3S.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} have{\scriptsize .V.INFIN} part{\scriptsize .N.F.SG}   //
\glmanual \%gls  well if be{\scriptsize .3S.PRES.NEG} PRON{\scriptsize .3SM} NEG PRT get{\scriptsize .NONFIN} part   //
\gleng well, if he doesn't get a part \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng ia fyddai wrth ei fodd (peta)sai fo (y)n cael cyfle sure@s:cym\&eng . //
\glsurface ELE:  oh$^{C}_{E}$ ia fyddai wrth ei fodd petasai fo yn cael cyfle sure$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM} yes{\scriptsize .ADV} be{\scriptsize .V.3S.COND+SM} by{\scriptsize .PREP} his{\scriptsize .ADJ.POSS} means{\scriptsize .N.M.SG+SM} be{\scriptsize .V.3S.PLUPERF.HYP} he{\scriptsize .PRON.M.3S} stative{\scriptsize .STAT} have{\scriptsize .V.INFIN} opportunity{\scriptsize .N.M.SG} sure{\scriptsize .ADJ}   //
\glmanual \%gls  IM yes be{\scriptsize .3S.CONDIT} by POSS{\scriptsize .3SM} pleasure if\_be{\scriptsize .3S.CONDIT} PRON{\scriptsize .3SM} PRT get{\scriptsize .NONFIN} opportunity sure   //
\gleng oh, yes, he'd love it if he got the chance, of course. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat call . //
\glsurface ALN:  call .  //
\glauto \%aut  sane{\scriptsize .ADJ}   //
\glmanual \%gls  wise   //
\gleng wise. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wna i ddeud +"/ . //
\glsurface ALN:  wna i ddeud .  //
\glauto \%aut  do{\scriptsize .V.13S.PRES+SM} I{\scriptsize .PRON.1S[or]to.PREP} say{\scriptsize .V.INFIN.SPOKEN+SM}   //
\glmanual \%gls  do{\scriptsize .1S.NONPAST} PRON{\scriptsize .1S} say{\scriptsize .NONFIN}   //
\gleng I'll say: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" gwranda Izzy@s:cym\&eng \# (ba)swn i (ddi)m meindio (peta)sai fo (y)n drymio . //
\glsurface ALN:  gwranda Izzy$^{C}_{E}$ baswn i ddim meindio petasai fo yn drymio .  //
\glauto \%aut  listen{\scriptsize .V.2S.IMPER} name be{\scriptsize .V.1S.PLUPERF.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} mind{\scriptsize .V.INFIN} be{\scriptsize .V.3S.PLUPERF.HYP} he{\scriptsize .PRON.M.3S} stative{\scriptsize .STAT} drum{\scriptsize .V.INFIN}   //
\glmanual \%gls  listen{\scriptsize .2S.IMPER} Izzy be{\scriptsize .1S.CONDIT} PRON{\scriptsize .1S} NEG mind{\scriptsize .NONFIN} if\_be{\scriptsize .3S.CONDIT} PRON{\scriptsize .3S} PRT drum{\scriptsize .NONFIN}   //
\gleng "listen Izzy, I wouldn't mind if he drummed." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (e)ll(i) [?] ti fyw hebddo fo . //
\glsurface ALN:  elli ti fyw hebddo fo .  //
\glauto \%aut  be\_able{\scriptsize .V.2S.PRES+SM} you{\scriptsize .PRON.2S} live{\scriptsize .V.INFIN+SM} without\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S}   //
\glmanual \%gls  can{\scriptsize .2S.NONPAST} PRON{\scriptsize .2S} live{\scriptsize .NONFIN} without{\scriptsize .3SM} PRON{\scriptsize .3SM}   //
\gleng you can live without him. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wedyn (a)llai fo ddysgu dau neu dri fill@s:eng gan Dean@s:cym\&eng ac er@s:cym\&eng \#\# fydd o (y)n barod wedyn . //
\glsurface ALN:  wedyn allai fo ddysgu dau neu dri fill$^{E}$ gan Dean$^{C}_{E}$ ac er$^{C}_{E}$ fydd o yn barod wedyn .  //
\glauto \%aut  afterwards{\scriptsize .ADV} be\_able{\scriptsize .V.3S.IMPERF+SM} he{\scriptsize .PRON.M.3S} teach{\scriptsize .V.INFIN+SM} two{\scriptsize .NUM.M} or{\scriptsize .CONJ} three{\scriptsize .NUM.M+SM} fill{\scriptsize .SV.INFIN} by{\scriptsize .PREP[or]can.N.M.SG+SM} name and{\scriptsize .CONJ} er{\scriptsize .IM} be{\scriptsize .V.3S.FUT+SM} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} ready{\scriptsize .ADJ+SM} afterwards{\scriptsize .ADV}   //
\glmanual \%gls  then can{\scriptsize .3S.CONDIT} PRON{\scriptsize .3SM} learn{\scriptsize .NONFIN} two{\scriptsize .M} or three{\scriptsize .M} fill from Dean and IM be{\scriptsize .3S.FUT} PRON{\scriptsize .3SM} PRT ready after   //
\gleng then he can learn two or three fills off Dean, and he'll be ready then. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd um@s:cym\&eng +. . . //
\glsurface ALN:  oedd um$^{C}_{E}$ .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} um{\scriptsize .IM}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} IM   //
\gleng there was \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat es i fewn i neuadd bore (y)ma . //
\glsurface ALN:  es i fewn i neuadd bore yma .  //
\glauto \%aut  go{\scriptsize .V.1S.PAST.SPOKEN} I{\scriptsize .PRON.1S} in{\scriptsize .PREP+SM} to{\scriptsize .PREP} hall{\scriptsize .N.F.SG} morning{\scriptsize .N.M.SG} here{\scriptsize .ADV}   //
\glmanual \%gls  go{\scriptsize .1S.PAST} PRON{\scriptsize .1S} in to hall morning here   //
\gleng I went into the hall this morning. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o'n i (y)n dysgu (y)r ail flwyddyn perfformio ac oedd Anna@s:cym\&eng (y)n cofrestru pawb . //
\glsurface ALN:  o'n i yn dysgu yr ail flwyddyn perfformio ac oedd Anna$^{C}_{E}$ yn cofrestru pawb .  //
\glauto \%aut  unk to{\scriptsize .PREP} stative{\scriptsize .STAT} teach{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} second{\scriptsize .ORD} year{\scriptsize .N.F.SG+SM} perform{\scriptsize .V.INFIN} and{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} name stative{\scriptsize .STAT} register{\scriptsize .V.INFIN} everyone{\scriptsize .PRON}   //
\glmanual \%gls  be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} PRT learn{\scriptsize .NONFIN} DET second year performance and be{\scriptsize .3S.IMP} Anna PRT register{\scriptsize .NONFIN} everybody   //
\gleng I was teaching second year performance, and Anna was registering everyone. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat and@s:eng [?] um@s:cym\&eng +. . . //
\glsurface ALN:  and$^{E}$ um$^{C}_{E}$ .  //
\glauto \%aut  and{\scriptsize .CONJ} um{\scriptsize .IM}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=mumble \&=mumble . //
\glsurface ALN:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" Izzy@s:cym\&eng ! //
\glsurface ALN:  Izzy$^{C}_{E}$ !  //
\glauto \%aut  name   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" xxx . //
\glsurface ALN:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" mae tu allan yn car@s:cym\&eng . //
\glsurface ALN:  mae tu allan yn car$^{C}_{E}$ .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} side{\scriptsize .N.M.SG} out{\scriptsize .ADV} yn{\scriptsize .PRT} car{\scriptsize .N.SG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} side out in car   //
\gleng "she's outside in the car." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" be mae (y)n wneud fan (y)na ? //
\glsurface ALN:  be mae yn wneud fan yna ?  //
\glauto \%aut  what{\scriptsize .INT} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM} place{\scriptsize .N.MF.SG+SM} there{\scriptsize .ADV}   //
\glmanual \%gls  what be{\scriptsize .3S.PRES} PRT do{\scriptsize .NONFIN} place there   //
\gleng "what's she doing there?" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" oedd o (y)n edrych fath â cael snog@s:eng pan wnes i basio ! //
\glsurface ALN:  oedd o yn edrych fath â cael snog$^{E}$ pan wnes i basio !  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} look{\scriptsize .V.INFIN} type{\scriptsize .N.M.SG+SM} as{\scriptsize .CONJ} have{\scriptsize .V.INFIN} snog{\scriptsize .SV.INFIN} when{\scriptsize .CONJ} do{\scriptsize .V.1S.PAST.SPOKEN+SM} I{\scriptsize .PRON.1S} pass{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} PRT look{\scriptsize .NONFIN} kind with have{\scriptsize .NONFIN} snog when do{\scriptsize .1S.PAST} PRON{\scriptsize .1S} pass{\scriptsize .NONFIN}   //
\gleng "it looked like having a snog when I passed!" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng ! //
\glsurface ELE:  hmm$^{C}_{E}$ !  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" be ? //
\glsurface ALN:  be ?  //
\glauto \%aut  what{\scriptsize .INT}   //
\glmanual \%gls  what   //
\gleng "what?" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" mae (rhyw)beth mae pobl ifanc yn wneud Anna@s:cym\&eng . //
\glsurface ALN:  mae rhywbeth mae pobl ifanc yn wneud Anna$^{C}_{E}$ .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} something{\scriptsize .N.M.SG} be{\scriptsize .V.3S.PRES} people{\scriptsize .N.F.SG} young{\scriptsize .ADJ} stative{\scriptsize .STAT} make{\scriptsize .V.INFIN+SM} name   //
\glmanual \%gls  be{\scriptsize .3S.PRES} something be{\scriptsize .3S.PRES} people young PRT do{\scriptsize .NONFIN} Anna   //
\gleng "it's something young people do, Anna." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" oh@s:cym\&eng [=! gasps] ! //
\glsurface ALN:  oh$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng "oh!" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat meddai hi +" . //
\glsurface ALN:  meddai hi .  //
\glauto \%aut  say{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  say{\scriptsize .3S.PAST} PRON{\scriptsize .3SF}   //
\gleng she said. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=laugh . //
\glsurface ALN:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=laugh . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" stuff@s:cym\&eng yna i\_gyd tu ôl i fi wan ! //
\glsurface ALN:  stuff$^{C}_{E}$ yna i\_gyd tu ôl i fi wan !  //
\glauto \%aut  stuff{\scriptsize .SV.INFIN} there{\scriptsize .ADV} all{\scriptsize .ADJ} side{\scriptsize .N.M.SG} track{\scriptsize .N.M.SG[or]rear.ADJ} to{\scriptsize .PREP} I{\scriptsize .PRON.1S+SM} weak{\scriptsize .ADJ+SM[or]pale.ADJ+SM}   //
\glmanual \%gls  stuff there all side back to PRON{\scriptsize .1S} now   //
\gleng "all that stuff is behind me now!" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o\_k@s:cym\&eng . //
\glsurface ELE:  o\_k$^{C}_{E}$ .  //
\glauto \%aut  OK{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=sniff . //
\glsurface ALN:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (dy)dy (ddi)m yn sôn am gael un arall ta ? //
\glsurface ELE:  dydy ddim yn sôn am gael un arall ta ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} mention{\scriptsize .V.INFIN} for{\scriptsize .PREP} have{\scriptsize .V.INFIN+SM} one{\scriptsize .NUM} other{\scriptsize .ADJ} or{\scriptsize .CONJ}   //
\glmanual \%gls  be{\scriptsize .3S.PRES.NEG} PRT talk{\scriptsize .NONFIN} about get{\scriptsize .NONFIN} one other then    //
\gleng she's not talking about having another one, then? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat er@s:cym\&eng na . //
\glsurface ALN:  er$^{C}_{E}$ na .  //
\glauto \%aut  er{\scriptsize .IM} neg{\scriptsize .PRT}   //
\glmanual \%gls  IM no   //
\gleng er, no. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na(g) (y)dy wir ? //
\glsurface ELE:  nag ydy wir ?  //
\glauto \%aut  than{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} true{\scriptsize .ADJ+SM}   //
\glmanual \%gls  NEG be{\scriptsize .3S.PRES} true   //
\gleng is she not? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat bob awr neithiwr . //
\glsurface ALN:  bob awr neithiwr .  //
\glauto \%aut  each{\scriptsize .PREQ+SM} hour{\scriptsize .N.F.SG} last{\scriptsize .night.ADV}   //
\glmanual \%gls  every hour last\_night   //
\gleng every hour last night. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat bob awr neithiwr . //
\glsurface ALN:  bob awr neithiwr .  //
\glauto \%aut  each{\scriptsize .PREQ+SM} hour{\scriptsize .N.F.SG} last{\scriptsize .night.ADV}   //
\glmanual \%gls  every hour last\_night   //
\gleng every hour last night. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd o (y)n effro ? //
\glsurface ELE:  oedd o yn effro ?  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} awake{\scriptsize .ADJ}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} awake    //
\gleng was he awake? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd . //
\glsurface ALN:  oedd .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  be{\scriptsize .3S.IMP}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o:h@s:cym\&eng (dy)dy o (ddi)m yn dda (fe)lly ? //
\glsurface ELE:  oh$^{C}_{E}$ dydy o ddim yn dda felly ?  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} he{\scriptsize .PRON.M.3S.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} good{\scriptsize .ADJ+SM} so{\scriptsize .ADV}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES.NEG} PRON{\scriptsize .3SM} NEG PRT good thus   //
\gleng oh, he's not well then? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw (ddi)m yn gwybod be sy ond doedd hi (dd)im yn impressed@s:eng \&=laugh . //
\glsurface ALN:  dw ddim yn gwybod be sy ond doedd hi ddim yn impressed$^{E}$ .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN} what{\scriptsize .INT} which\_is{\scriptsize .V.3SP.PRES.REL} but{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF.NEG} she{\scriptsize .PRON.F.3S} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} impress{\scriptsize .V.INFIN+AV}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} NEG PRT know{\scriptsize .NONFIN} what be{\scriptsize .PRES.REL} but be{\scriptsize .3S.IMP.NEG} PRON{\scriptsize .3SF} NEG PRT impressed   //
\gleng I don't know what's the matter, but she wasn't impressed! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng bech:od ! //
\glsurface ELE:  oh$^{C}_{E}$ bechod !  //
\glauto \%aut  oh{\scriptsize .IM} how{\scriptsize .sad.IM[or]sin.N.M.SG+SM}   //
\glmanual \%gls  IM sin   //
\gleng ah, bless! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=cough \&=cough \&=cough . //
\glsurface ALN:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng mae o (y)n tynny am \# ddeuflwydd ? //
\glsurface ELE:  oh$^{C}_{E}$ mae o yn tynny am ddeuflwydd ?  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} yn{\scriptsize .PRT} unk for{\scriptsize .PREP} two\_years{\scriptsize .N.F.SG+SM}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} PRT draw{\scriptsize .NONFIN} for two\_years   //
\gleng oh, he's getting on for two years? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat blwydd a hanner ? //
\glsurface ELE:  blwydd a hanner ?  //
\glauto \%aut  year{\scriptsize .N.F.SG} and{\scriptsize .CONJ} half{\scriptsize .N.M.SG}   //
\glmanual \%gls  year and half   //
\gleng 18 months? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ rywbeth felly . //
\glsurface ALN:  rywbeth felly .  //
\glauto \%aut  something{\scriptsize .N.M.SG+SM} so{\scriptsize .ADV}   //
\glmanual \%gls  something thus   //
\gleng something like that. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat deuflwydd yndy dydy ? //
\glsurface ELE:  deuflwydd yndy dydy ?  //
\glauto \%aut  two\_years{\scriptsize .N.F.SG} be{\scriptsize .V.3S.PRES.EMPH} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM}   //
\glmanual \%gls  two\_years be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES.NEG}   //
\gleng it is two years, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$wnaeth Kate@s:cym\&eng ddeud$>$ [/] \# $<$oedd Kate@s:cym\&eng xxx$>$ [//] xxx Kate@s:cym\&eng yn tagu neithiwr . //
\glsurface ALN:  wnaeth Kate$^{C}_{E}$ ddeud oedd Kate$^{C}_{E}$ Kate$^{C}_{E}$ yn tagu neithiwr .  //
\glauto \%aut  do{\scriptsize .V.3S.PAST+SM} name say{\scriptsize .V.INFIN.SPOKEN+SM} be{\scriptsize .V.3S.IMPERF} name name stative{\scriptsize .STAT} choke{\scriptsize .V.INFIN} last{\scriptsize .night.ADV}   //
\glmanual \%gls  do{\scriptsize .3S.PAST} Kate say{\scriptsize .NONFIN} be{\scriptsize .3S.IMP} Kate x Kate PRT choke{\scriptsize .NONFIN} last\_night  //
\gleng Kate was saying. . . Kate was coughing last night. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wnaeth hi ddeud +"/ . //
\glsurface ALN:  wnaeth hi ddeud .  //
\glauto \%aut  do{\scriptsize .V.3S.PAST+SM} she{\scriptsize .PRON.F.3S} say{\scriptsize .V.INFIN.SPOKEN+SM}   //
\glmanual \%gls  do{\scriptsize .3S.PAST} PRON{\scriptsize .3SF} say{\scriptsize .NONFIN}   //
\gleng she said: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" well@s:cym\&eng o'n i fyny rhwng \# tri a hanner (we)di pump yn watsio ryw film@s:cym\&eng (be)cause@s:eng bod fi methu cysgu yn tagu . //
\glsurface ALN:  well$^{C}_{E}$ o'n i fyny rhwng tri a hanner wedi pump yn watsio ryw film$^{C}_{E}$ because$^{E}$ bod fi methu cysgu yn tagu .  //
\glauto \%aut  well{\scriptsize .ADV} unk to{\scriptsize .PREP} above{\scriptsize .ADV} between{\scriptsize .PREP} three{\scriptsize .NUM.M} and{\scriptsize .CONJ} half{\scriptsize .N.M.SG} after{\scriptsize .PREP} five{\scriptsize .NUM} stative{\scriptsize .STAT} watch{\scriptsize .V.INFIN} some{\scriptsize .PREQ+SM} film{\scriptsize .N.SG} because{\scriptsize .CONJ} be{\scriptsize .V.INFIN} I{\scriptsize .PRON.1S+SM} fail{\scriptsize .V.INFIN} sleep{\scriptsize .V.INFIN} stative{\scriptsize .STAT} choke{\scriptsize .V.INFIN}   //
\glmanual \%gls  well be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} up between three{\scriptsize .M} and half after five PRT watch{\scriptsize .NONFIN} some film because be{\scriptsize .NONFIN} PRON{\scriptsize .1S} fail{\scriptsize .NONFIN} sleep{\scriptsize .NONFIN} PRT choke{\scriptsize .NONFIN}   //
\gleng "well, I was up between 3 and 5:30 watching some film, cause I couldn't sleep for coughing." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ hmm@s:cym\&eng hmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM} hmm{\scriptsize .IM}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a ddeudodd Anna@s:cym\&eng +"/ . //
\glsurface ALN:  a ddeudodd Anna$^{C}_{E}$ .  //
\glauto \%aut  and{\scriptsize .CONJ} say{\scriptsize .V.3S.PAST+SM} name   //
\glmanual \%gls  and say{\scriptsize .3S.PAST} Anna   //
\gleng and Anna said: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" gwranda unrhyw bryd ti fyny nos \# just@s:cym\&eng ffonia fi o\_k@s:cym\&eng ? //
\glsurface ALN:  gwranda unrhyw bryd ti fyny nos just$^{C}_{E}$ ffonia fi o\_k$^{C}_{E}$ ?  //
\glauto \%aut  listen{\scriptsize .V.2S.IMPER} any{\scriptsize .ADJ} mind{\scriptsize .N.M.SG[or]when.INT+SM[or]time.N.M.SG+SM} you{\scriptsize .PRON.2S} above{\scriptsize .ADV} night{\scriptsize .N.F.SG} just{\scriptsize .ADV[or]just.ADJ} phone{\scriptsize .V.2S.IMPER[or]phone.V.3S.PRES} I{\scriptsize .PRON.1S+SM} OK{\scriptsize .IM}   //
\glmanual \%gls  listen{\scriptsize .2S.IMPER} any time PRON{\scriptsize .2S} up night just phone{\scriptsize .2S.IMPER} PRON{\scriptsize .1S} o\_k   //
\gleng "listen, any time you're up at night, just phone me, ok?" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=laugh . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" dw i (y)n bound@s:cym\&eng o fod ar [?] effro ! //
\glsurface ALN:  dw i yn bound$^{C}_{E}$ o fod ar effro !  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} stative{\scriptsize .STAT} bound{\scriptsize .SV.INFIN[or]pound.SV.INFIN+SM} of{\scriptsize .PREP} be{\scriptsize .V.INFIN+SM} on{\scriptsize .PREP} awake{\scriptsize .ADJ}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT bound of be{\scriptsize .NONFIN} on awake   //
\gleng "I'm bound to be awake!" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$oh@s:cym\&eng bechod$>$ [=! quietly] ! //
\glsurface ELE:  oh$^{C}_{E}$ bechod !  //
\glauto \%aut  oh{\scriptsize .IM} how{\scriptsize .sad.IM[or]sin.N.M.SG+SM}   //
\glmanual \%gls  IM sin   //
\gleng oh, bless! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na (dy)dy (ddi)m yn swnio fel bod hi (y)n mynd am ail un ar hyn o bryd . //
\glsurface ALN:  na dydy ddim yn swnio fel bod hi yn mynd am ail un ar hyn o bryd .  //
\glauto \%aut  neg{\scriptsize .PRT} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} sound{\scriptsize .V.INFIN} like{\scriptsize .CONJ} be{\scriptsize .V.INFIN} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} go{\scriptsize .V.INFIN} for{\scriptsize .PREP} second{\scriptsize .ORD} one{\scriptsize .NUM} on{\scriptsize .PREP} this{\scriptsize .PRON.SP} of{\scriptsize .PREP} mind{\scriptsize .N.M.SG[or]when.INT+SM[or]time.N.M.SG+SM}   //
\glmanual \%gls  no be{\scriptsize .3S.PRES.NEG} NEG PRT sound{\scriptsize .NONFIN} like be{\scriptsize .NONFIN} PRON{\scriptsize .3SF} PRT go{\scriptsize .NONFIN} for second one on this of time   //
\gleng no, it doesn't sound like she's trying for a second one at the moment. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond um@s:cym\&eng \# wedi deud hynny oedd hi (y)n gwadu \# bob\_dim \# mynd am y cyntaf . //
\glsurface ALN:  ond um$^{C}_{E}$ wedi deud hynny oedd hi yn gwadu bob\_dim mynd am y cyntaf .  //
\glauto \%aut  but{\scriptsize .CONJ} um{\scriptsize .IM} after{\scriptsize .PREP} say{\scriptsize .V.INFIN.SPOKEN} that{\scriptsize .PRON.SP} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} deny{\scriptsize .V.INFIN} everything{\scriptsize .PRON} go{\scriptsize .V.INFIN} for{\scriptsize .PREP} the{\scriptsize .DET.DEF} first{\scriptsize .ORD}   //
\glmanual \%gls  but IM PRT{\scriptsize .PAST} say{\scriptsize .NONFIN} that be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT deny{\scriptsize .NONFIN} everything go{\scriptsize .NONFIN} for DET first   //
\gleng but, having said that, she was denying everything trying for the first one. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a wedyn yn sydyn \# wnaethon ni glocio . //
\glsurface ALN:  a wedyn yn sydyn wnaethon ni glocio .  //
\glauto \%aut  and{\scriptsize .CONJ} afterwards{\scriptsize .ADV} stative{\scriptsize .STAT} sudden{\scriptsize .ADJ} do{\scriptsize .V.3P.PAST.SPOKEN+SM} we{\scriptsize .PRON.1P} clock{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  and then PRT sudden do{\scriptsize .1PL.PAST} PRON{\scriptsize .1PL} clock{\scriptsize .NONFIN}   //
\gleng and then suddenly we twigged. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ ie ie . //
\glsurface ELE:  ie ie .  //
\glauto \%aut  yes{\scriptsize .ADV} yes{\scriptsize .ADV}   //
\glmanual \%gls  yes yes   //
\gleng yes, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ac er@s:cym\&eng \&=belch +. . . //
\glsurface ALN:  ac er$^{C}_{E}$ .  //
\glauto \%aut  and{\scriptsize .CONJ} er{\scriptsize .IM}   //
\glmanual \%gls  and IM   //
\gleng and, er \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng esgusodwch fi pobl y tâp ! //
\glsurface ALN:  oh$^{C}_{E}$ esgusodwch fi pobl y tâp !  //
\glauto \%aut  oh{\scriptsize .IM} excuse{\scriptsize .V.2P.PRES[or]excuse.V.2P.IMPER} I{\scriptsize .PRON.1S+SM} people{\scriptsize .N.F.SG} the{\scriptsize .DET.DEF} tape{\scriptsize .N.M.SG}   //
\glmanual \%gls  IM excuse{\scriptsize .2PL.IMPER} PRON{\scriptsize .1S} people DET tape   //
\gleng oh, excuse me, people hearing this tape! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat um@s:cym\&eng \# dyma Andrew@s:cym\&eng yn ei chyhuddo hi . //
\glsurface ALN:  um$^{C}_{E}$ dyma Andrew$^{C}_{E}$ yn ei chyhuddo hi .  //
\glauto \%aut  um{\scriptsize .IM} this{\scriptsize .is.V} name stative{\scriptsize .STAT} her{\scriptsize .ADJ.POSS} accuse{\scriptsize .V.INFIN+AM} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  IM here Andrew PRT POSS{\scriptsize .3SF} accuse{\scriptsize .NONFIN} PRON{\scriptsize .3SF}   //
\gleng um, then Andrew accused her. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat aeth [?] hi just@s:cym\&eng sefyll (y)na a cochi (y)n llwyr . //
\glsurface ALN:  aeth hi just$^{C}_{E}$ sefyll yna a cochi yn llwyr .  //
\glauto \%aut  go{\scriptsize .V.3S.PAST} she{\scriptsize .PRON.F.3S} just{\scriptsize .ADV[or]just.ADJ} stand{\scriptsize .V.INFIN} there{\scriptsize .ADV} and{\scriptsize .CONJ} blush{\scriptsize .V.INFIN} stative{\scriptsize .STAT} complete{\scriptsize .ADJ}   //
\glmanual \%gls  go{\scriptsize .3S.PAST} PRON{\scriptsize .3SF} just stand{\scriptsize .NONFIN} there and redden{\scriptsize .NONFIN} PRT total   //
\gleng she just stood there, and blushed all over. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=laugh . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ofnadwy o ddoniol ! //
\glsurface ALN:  ofnadwy o ddoniol !  //
\glauto \%aut  terrible{\scriptsize .ADJ} of{\scriptsize .PREP} funny{\scriptsize .ADJ+SM}   //
\glmanual \%gls  awful of funny   //
\gleng incredibly funny! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng ! //
\glsurface ELE:  oh$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$be wnaeth$>$ [?] Louise@s:cym\&eng ta ? //
\glsurface ELE:  be wnaeth Louise$^{C}_{E}$ ta ?  //
\glauto \%aut  what{\scriptsize .INT} do{\scriptsize .V.3S.PAST+SM} name or{\scriptsize .CONJ}   //
\glmanual \%gls  what do{\scriptsize .3S.PAST} Louise then   //
\gleng what did Louise do then? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (y)dy (we)di \&ni [//] cynhyrchu mwy o (y)r pethau \# chutneys@s:cym\&eng (y)ma \# jams@s:eng (y)ma ? //
\glsurface ELE:  ydy wedi cynhyrchu mwy o yr pethau chutneys$^{C}_{E}$ yma jams$^{E}$ yma ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} after{\scriptsize .PREP} produce{\scriptsize .V.INFIN} more{\scriptsize .ADJ.COMP} of{\scriptsize .PREP} the{\scriptsize .DET.DEF} things{\scriptsize .N.M.PL} chutney{\scriptsize .N.SG+PL} here{\scriptsize .ADV} jam{\scriptsize .SV.INFIN+PV} here{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRT{\scriptsize .PAST} produce{\scriptsize .NONFIN} more of DET things chutneys here jams here   //
\gleng has she produced more of these chutney things, these jam things? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=sigh mae angen i fi anfon y rysáit gwreiddiol ati hi . //
\glsurface ALN:  mae angen i fi anfon y rysáit gwreiddiol ati hi .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} need{\scriptsize .N.M.SG} to{\scriptsize .PREP} I{\scriptsize .PRON.1S+SM} send{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} recipe{\scriptsize .N.F.SG} original{\scriptsize .ADJ} to\_her{\scriptsize .PREP+PRON.F.3S} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} need to PRON{\scriptsize .1S} send{\scriptsize .NONFIN} DET recipe original to{\scriptsize .3SF} PRON{\scriptsize .3SF}   //
\gleng I need to send her the original recipe. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae (we)di rhoi mwy o vinegar@s:cym\&eng i\_fewn . //
\glsurface ALN:  mae wedi rhoi mwy o vinegar$^{C}_{E}$ i\_fewn .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} after{\scriptsize .PREP} give{\scriptsize .V.INFIN} more{\scriptsize .ADJ.COMP} of{\scriptsize .PREP} vinegar{\scriptsize .N.SG} in{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRT{\scriptsize .PAST} put{\scriptsize .NONFIN} more of vinegar in   //
\gleng she put more vinegar in. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond oedd hi meddwl bod \# un mam@s:cym\&eng yn fwy rich@s:eng oherwydd \#\# bod o (we)di cael amser i aeddfedu . //
\glsurface ALN:  ond oedd hi meddwl bod un mam$^{C}_{E}$ yn fwy rich$^{E}$ oherwydd bod o wedi cael amser i aeddfedu .  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} think{\scriptsize .V.INFIN} be{\scriptsize .V.INFIN} one{\scriptsize .NUM} mam{\scriptsize .N.SG} stative{\scriptsize .STAT} more{\scriptsize .ADJ.COMP+SM} rich{\scriptsize .ADJ} because{\scriptsize .CONJ} be{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} after{\scriptsize .PREP} have{\scriptsize .V.INFIN} time{\scriptsize .N.M.SG} to{\scriptsize .PREP} mature{\scriptsize .V.INFIN}   //
\glmanual \%gls  but be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} think{\scriptsize .NONFIN} be{\scriptsize .NONFIN} one mother PRT more rich because be{\scriptsize .NONFIN} PRON{\scriptsize .3SM} PRT{\scriptsize .PAST} have{\scriptsize .NONFIN} time to mature{\scriptsize .NONFIN}   //
\gleng but she thought that Mam's one was more rich because it had had time to mature. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well@s:cym\&eng dy fam wastad yn deud tydy \# i gadw fo am (y)chydig fisoedd . //
\glsurface ELE:  well$^{C}_{E}$ dy fam wastad yn deud tydy i gadw fo am ychydig fisoedd .  //
\glauto \%aut  well{\scriptsize .ADV} your{\scriptsize .ADJ.POSS} mother{\scriptsize .N.F.SG+SM} flat{\scriptsize .ADJ+SM} stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN} be{\scriptsize .V.3S.PRES.NEG} I{\scriptsize .PRON.1S[or]to.PREP} keep{\scriptsize .V.INFIN+SM} he{\scriptsize .PRON.M.3S} for{\scriptsize .PREP} a\_little{\scriptsize .QUAN} months{\scriptsize .N.M.PL+SM}   //
\glmanual \%gls  well POSS{\scriptsize .2S} mother always PRT say{\scriptsize .NONFIN} be{\scriptsize .3S.PRES.NEG} to keep{\scriptsize .NONFIN} PRON{\scriptsize .3SM} for a\_bit months   //
\gleng well, your mother always says, doesn't she, to keep it for a few months. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ALN:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng ti (dd)im yn cofio oe(dd) hi (y)n sgwennu arno fo doedd ? //
\glsurface ELE:  oh$^{C}_{E}$ ti ddim yn cofio oedd hi yn sgwennu arno fo doedd ?  //
\glauto \%aut  oh{\scriptsize .IM} you{\scriptsize .PRON.2S} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} remember{\scriptsize .V.INFIN} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} write{\scriptsize .V.INFIN} on\_him{\scriptsize .PREP+PRON.M.3S} he{\scriptsize .PRON.M.3S} be{\scriptsize .V.3S.IMPERF.NEG}   //
\glmanual \%gls  IMP PRON{\scriptsize .2S} NEG PRT remember{\scriptsize .NONFIN} be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT write{\scriptsize .NONFIN} on{\scriptsize .3SM} PRON{\scriptsize .3SM} be{\scriptsize .3S.IMP.NEG}   //
\gleng oh, don't you remember, she used to write on it, didn't she? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" peidiwch â \#\# bwyta am \# +// . //
\glsurface ELE:  peidiwch â bwyta am .  //
\glauto \%aut  stop{\scriptsize .V.2P.IMPER} with{\scriptsize .PREP} eat{\scriptsize .V.INFIN} for{\scriptsize .PREP}   //
\glmanual \%gls  desist{\scriptsize .2PL.IMPER} with eat{\scriptsize .NONFIN} for   //
\gleng "do not eat for \dots " //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ oedd . //
\glsurface ALN:  oedd .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  be{\scriptsize .3S.IMP}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, timod +"/ . //
\glsurface ELE:  timod .  //
\glauto \%aut  know{\scriptsize .V.2S.PRES.SPOKEN}   //
\glmanual \%gls  know{\scriptsize .2S}   //
\gleng you know: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" dau dri mis . //
\glsurface ELE:  dau dri mis .  //
\glauto \%aut  two{\scriptsize .NUM.M} three{\scriptsize .NUM.M+SM} month{\scriptsize .N.M.SG}   //
\glmanual \%gls  two{\scriptsize .M} three{\scriptsize .M} month   //
\gleng "two, three months." //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ALN:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond hefyd y siwgwr [/] siwgwr tywyll tywyll yeah@s:cym\&eng ? //
\glsurface ELE:  ond hefyd y siwgwr siwgwr tywyll tywyll yeah$^{C}_{E}$ ?  //
\glauto \%aut  but{\scriptsize .CONJ} also{\scriptsize .ADV} the{\scriptsize .DET.DEF} sugar{\scriptsize .N.M.SG} sugar{\scriptsize .N.M.SG} dark{\scriptsize .ADJ} dark{\scriptsize .ADJ} yeah{\scriptsize .ADV}   //
\glmanual \%gls  but also DET sugar sugar dark dark yeah   //
\gleng but also, the really dark sugar, right? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ia . //
\glsurface ALN:  ia .  //
\glauto \%aut  yes{\scriptsize .ADV}   //
\glmanual \%gls  yes   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i bron â gorffen yr un +// . //
\glsurface ALN:  dw i bron â gorffen yr un .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} breast{\scriptsize .N.F.SG[or]almost.ADV} as{\scriptsize .CONJ[or]with.PREP[or]go.V.3S.PRES} complete{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} one{\scriptsize .NUM}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} nearly with finish{\scriptsize .NONFIN} DET one   //
\gleng I've nearly finished the one \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$oedd yr un$>$ [///] o'n i (y)n licio (y)r un \&o oedd hi (we)di wneud . //
\glsurface ALN:  oedd yr un o'n i yn licio yr un oedd hi wedi wneud .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} the{\scriptsize .DET.DEF} one{\scriptsize .NUM} unk to{\scriptsize .PREP} stative{\scriptsize .STAT} like{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} one{\scriptsize .NUM} be{\scriptsize .V.3S.IMPERF} she{\scriptsize .PRON.F.3S} after{\scriptsize .PREP} make{\scriptsize .V.INFIN+SM}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} DET one be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} PRT like{\scriptsize .NONFIN} DET one be{\scriptsize .3S.IMP} PRON{\scriptsize .3SF} PRT{\scriptsize .PAST} do{\scriptsize .NONFIN}   //
\gleng the one \dots I liked the one she'd made. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$oedd o (y)n$>$ [//] \# oe(dd) (y)n debyg iawn i \# afal (we)di stiwio really@s:eng doedd ? //
\glsurface ALN:  oedd o yn oedd yn debyg iawn i afal wedi stiwio really$^{E}$ doedd ?  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} he{\scriptsize .PRON.M.3S.SPOKEN} yn{\scriptsize .PRT} be{\scriptsize .V.3S.IMPERF} stative{\scriptsize .STAT} similar{\scriptsize .ADJ+SM} very{\scriptsize .ADV} to{\scriptsize .PREP} apple{\scriptsize .N.M.SG} after{\scriptsize .PREP} stew{\scriptsize .V.INFIN} real{\scriptsize .ADJ+LY} be{\scriptsize .V.3S.IMPERF.NEG}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} PRON{\scriptsize .3SM} PRT be{\scriptsize .3S.IMP} PRT similar very to apple PRT{\scriptsize .PAST} stew{\scriptsize .NONFIN} really be{\scriptsize .3S.IMP.NEG}   //
\gleng it was very similar to stewed apple really, wasn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd . //
\glsurface ELE:  oedd .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  be{\scriptsize .3S.IMP}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat gen i afalau gan Margaret@s:cym\&eng cofia . //
\glsurface ELE:  gen i afalau gan Margaret$^{C}_{E}$ cofia .  //
\glauto \%aut  by{\scriptsize .PREP.SPOKEN} I{\scriptsize .PRON.1S} apple{\scriptsize .N.M.PL} by{\scriptsize .PREP[or]can.N.M.SG+SM} name remember{\scriptsize .V.2S.IMPER[or]remember.V.3S.PRES}   //
\glmanual \%gls  with{\scriptsize .1S} PRON{\scriptsize .1S} apples from Margaret remember{\scriptsize .2S.IMPER}   //
\gleng I've got some apples from Margaret, remember. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ALN:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wneud (rhyw)beth efo nhw . //
\glsurface ELE:  wneud rhywbeth efo nhw .  //
\glauto \%aut  make{\scriptsize .V.INFIN+SM} something{\scriptsize .N.M.SG} with{\scriptsize .PREP} they{\scriptsize .PRON.3P}   //
\glmanual \%gls  do{\scriptsize .NONFIN} something with PRON{\scriptsize .3PL}   //
\gleng do something with them. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat nos Wener . //
\glsurface ALN:  nos Wener .  //
\glauto \%aut  night{\scriptsize .N.F.SG} name   //
\glmanual \%gls  night Friday   //
\gleng Friday night. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw (dd)im adref $<$nos Wener$>$ [?] . //
\glsurface ELE:  dw ddim adref nos Wener .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} not{\scriptsize .ADV+SM} homewards{\scriptsize .ADV} night{\scriptsize .N.F.SG} name   //
\glmanual \%gls  be{\scriptsize .1S.PRES} NEG home night Friday   //
\gleng I'm not home on Friday night. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng $<$wnawn ni$>$ [?] coginio (fe)lly . //
\glsurface ALN:  oh$^{C}_{E}$ wnawn ni coginio felly .  //
\glauto \%aut  oh{\scriptsize .IM} do{\scriptsize .V.1P.PRES+SM} we{\scriptsize .PRON.1P} cook{\scriptsize .V.INFIN} so{\scriptsize .ADV}   //
\glmanual \%gls  IM do{\scriptsize .1PL.NONPAST} PRON{\scriptsize .1PL} cook{\scriptsize .NONFIN} thus   //
\gleng oh, we'll cook, then. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat huh@s:cym\&eng ? //
\glsurface ELE:  huh$^{C}_{E}$ ?  //
\glauto \%aut  huh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wnawn ni coginio . //
\glsurface ALN:  wnawn ni coginio .  //
\glauto \%aut  do{\scriptsize .V.1P.PRES+SM} we{\scriptsize .PRON.1P} cook{\scriptsize .V.INFIN}   //
\glmanual \%gls  do{\scriptsize .1PL.NONPAST} PRON{\scriptsize .1PL} cook{\scriptsize .NONFIN}   //
\gleng we'll cook. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o'n i mynd i ddeud +"/ . //
\glsurface ALN:  o'n i mynd i ddeud .  //
\glauto \%aut  unk to{\scriptsize .PREP} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} say{\scriptsize .V.INFIN.SPOKEN+SM}   //
\glmanual \%gls  be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} go{\scriptsize .NONFIN} to say{\scriptsize .NONFIN}   //
\gleng I was going to say: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" be wnawn ni nos Wener ? //
\glsurface ALN:  be wnawn ni nos Wener ?  //
\glauto \%aut  what{\scriptsize .INT} do{\scriptsize .V.1P.PRES+SM} we{\scriptsize .PRON.1P} night{\scriptsize .N.F.SG} name   //
\glmanual \%gls  what do{\scriptsize .1PL.NONPAST} PRON{\scriptsize .1PL} night Friday   //
\gleng "what will we do on Friday night?" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond wrth\_gwrs dan ni allan dydan ? //
\glsurface ALN:  ond wrth\_gwrs dan ni allan dydan ?  //
\glauto \%aut  but{\scriptsize .CONJ} of\_course{\scriptsize .ADV} be{\scriptsize .V.1P.PRES.SPOKEN} we{\scriptsize .PRON.1P} out{\scriptsize .ADV} be{\scriptsize .V.1P.PRES.NEG[or]be.V.1P.PRES.NEG+SM}   //
\glmanual \%gls  but of\_course be{\scriptsize .1PL.PRES} PRON{\scriptsize .1PL} out be{\scriptsize .1PL.PRES.NEG}   //
\gleng but of co urse, we're out, aren't we //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ oh@s:cym\&eng \# dach [//] well@s:cym\&eng dach chi allan $<$rhan fwya$>$ [?] nos Wener . //
\glsurface ELE:  oh$^{C}_{E}$ dach well$^{C}_{E}$ dach chi allan rhan fwya nos Wener .  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.2P.PRES.SPOKEN} well{\scriptsize .ADV} be{\scriptsize .V.2P.PRES.SPOKEN} you{\scriptsize .PRON.2P} out{\scriptsize .ADV} part{\scriptsize .N.F.SG} biggest{\scriptsize .ADJ.SUP+SM} night{\scriptsize .N.F.SG} name   //
\glmanual \%gls  IM be{\scriptsize .2PL.PRES} well be{\scriptsize .2PL.PRES} PRON{\scriptsize .2PL} out part biggest night Friday   //
\gleng well, you're out mostly on Friday night. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat Charlie@s:cym\&eng gael [?] gwers am bedwar . //
\glsurface ALN:  Charlie$^{C}_{E}$ gael gwers am bedwar .  //
\glauto \%aut  name have{\scriptsize .V.INFIN+SM} lesson{\scriptsize .N.F.SG} for{\scriptsize .PREP} four{\scriptsize .NUM.M+SM}   //
\glmanual \%gls  Charlie have{\scriptsize .NONFIN} lesson for four   //
\gleng Charlie has a lesson at 4 . //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wedyn mae gynnon ni ymarfer am chwarter i saith . //
\glsurface ALN:  wedyn mae gynnon ni ymarfer am chwarter i saith .  //
\glauto \%aut  afterwards{\scriptsize .ADV} be{\scriptsize .V.3S.PRES} light{\scriptsize .V.1P.PAST.SPOKEN+SM} we{\scriptsize .PRON.1P} practice{\scriptsize .V.INFIN} for{\scriptsize .PREP} quarter{\scriptsize .N.M.SG} to{\scriptsize .PREP} seven{\scriptsize .NUM}   //
\glmanual \%gls  then be{\scriptsize .3S.PRES} with{\scriptsize .1PL} PRON{\scriptsize .1PL} practice for quarter to seven   //
\gleng then we have a practice at 6:45 . //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat chwarter i saith mae (y)n dechrau +! ? //
\glsurface ELE:  chwarter i saith mae yn dechrau ?  //
\glauto \%aut  quarter{\scriptsize .N.M.SG} to{\scriptsize .PREP} seven{\scriptsize .NUM} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} begin{\scriptsize .V.INFIN}   //
\glmanual \%gls  quarter to seven be{\scriptsize .3S.PRES} PRT start{\scriptsize .NONFIN}   //
\gleng it starts at 6:45!? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie . //
\glsurface ALN:  ie .  //
\glauto \%aut  yes{\scriptsize .ADV}   //
\glmanual \%gls  yes   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat fydd raid fi ddeud +"/ . //
\glsurface ALN:  fydd raid fi ddeud .  //
\glauto \%aut  be{\scriptsize .V.3S.FUT+SM} necessity{\scriptsize .N.M.SG+SM} I{\scriptsize .PRON.1S+SM} say{\scriptsize .V.INFIN.SPOKEN+SM}   //
\glmanual \%gls  be{\scriptsize .3S.FUT} necessity PRON{\scriptsize .1S} say{\scriptsize .NONFIN}   //
\gleng I'll have to say: //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +" gwrandwch \# dw i o (y)ma am chwarter i wyth ar yr hwyra o\_k@s:cym\&eng hogiau ? //
\glsurface ALN:  gwrandwch dw i o yma am chwarter i wyth ar yr hwyra o\_k$^{C}_{E}$ hogiau ?  //
\glauto \%aut  listen{\scriptsize .V.2P.IMPER} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} he{\scriptsize .PRON.M.3S.SPOKEN[or]from.PREP[or]of.PREP} here{\scriptsize .ADV} for{\scriptsize .PREP} quarter{\scriptsize .N.M.SG} to{\scriptsize .PREP} eight{\scriptsize .NUM} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} late{\scriptsize .ADJ.SUP} OK{\scriptsize .IM} lads{\scriptsize .N.M.PL}   //
\glmanual \%gls  listen{\scriptsize .2PL.IMPER} be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} from here for quarter to eight on DET latest o\_k boys   //
\gleng "listen, I'm out of here by 7:45 at the latest, ok guys?" //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat fydda i just@s:cym\&eng yn mynd . //
\glsurface ALN:  fydda i just$^{C}_{E}$ yn mynd .  //
\glauto \%aut  be{\scriptsize .V.1S.FUT+SM} I{\scriptsize .PRON.1S} just{\scriptsize .ADV[or]just.ADJ} stative{\scriptsize .STAT} go{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .1S.FUT} PRON{\scriptsize .1S} just PRT go{\scriptsize .NONFIN}   //
\gleng I'll just go. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat pam bod o (y)n hwyr yn dechrau ? //
\glsurface ELE:  pam bod o yn hwyr yn dechrau ?  //
\glauto \%aut  why?.ADV be{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} late{\scriptsize .ADJ} stative{\scriptsize .STAT} begin{\scriptsize .V.INFIN}   //
\glmanual \%gls  when be{\scriptsize .NONFIN} PRON{\scriptsize .3SM} PRT late PRT start{\scriptsize .NONFIN}   //
\gleng why is it starting late? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat chwarter i saith mae (y)n dechrau . //
\glsurface ALN:  chwarter i saith mae yn dechrau .  //
\glauto \%aut  quarter{\scriptsize .N.M.SG} to{\scriptsize .PREP} seven{\scriptsize .NUM} be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} begin{\scriptsize .V.INFIN}   //
\glmanual \%gls  quarter to seven be{\scriptsize .3S.PRES} PRT start{\scriptsize .NONFIN}   //
\gleng 6:45 it starts. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dyna [/] dyna pryd maen nhw yn ymarfer . //
\glsurface ALN:  dyna dyna pryd maen nhw yn ymarfer .  //
\glauto \%aut  that{\scriptsize .is.ADV} that{\scriptsize .is.ADV} when{\scriptsize .INT[or]time.N.M.SG} be{\scriptsize .V.3P.PRES.SPOKEN} they{\scriptsize .PRON.3P} stative{\scriptsize .STAT} practice{\scriptsize .V.INFIN}   //
\glmanual \%gls  there there when be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL} PRT practise{\scriptsize .NONFIN}   //
\gleng that's when they practise. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng . //
\glsurface ELE:  oh$^{C}_{E}$ .  //
\glauto \%aut  oh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat fedrwch chi (ddi)m newid o i hanner (we)di chwech tan hanner (we)di saith . //
\glsurface ELE:  fedrwch chi ddim newid o i hanner wedi chwech tan hanner wedi saith .  //
\glauto \%aut  be{\scriptsize .V.2P.PRES+SM} you{\scriptsize .PRON.2P} not{\scriptsize .ADV+SM} change{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} to{\scriptsize .PREP} half{\scriptsize .N.M.SG} after{\scriptsize .PREP} six{\scriptsize .NUM} until{\scriptsize .PREP} half{\scriptsize .N.M.SG} after{\scriptsize .PREP} seven{\scriptsize .NUM}   //
\glmanual \%gls  can{\scriptsize .2PL.NONPAST} PRON{\scriptsize .2PL} NEG change{\scriptsize .NONFIN} PRON{\scriptsize .3SM} to half after six until half after seven   //
\gleng can't you change it to 6:30 till 7:30? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae (y)n well na pump tydy ? //
\glsurface ALN:  mae yn well na pump tydy ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} better{\scriptsize .ADJ.COMP+SM} neg{\scriptsize .PRT[or]who\_not.PRON.REL.NEG[or](n)or.CONJ[or]than.CONJ} five{\scriptsize .NUM} be{\scriptsize .V.3S.PRES.NEG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRT better than five be{\scriptsize .3S.PRES.NEG}   //
\gleng it's better than 5, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat neu newid o hanner (we)di chwech tan hanner (we)di saith . //
\glsurface ELE:  neu newid o hanner wedi chwech tan hanner wedi saith .  //
\glauto \%aut  or{\scriptsize .CONJ} change{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} half{\scriptsize .N.M.SG} after{\scriptsize .PREP} six{\scriptsize .NUM} until{\scriptsize .PREP} half{\scriptsize .N.M.SG} after{\scriptsize .PREP} seven{\scriptsize .NUM}   //
\glmanual \%gls  or change{\scriptsize .NONFIN} PRON{\scriptsize .3SM} half after six until half after seven   //
\gleng or change it to 6:30 till 7:30 . //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$dim yn gwybod$>$ [?] . //
\glsurface ALN:  dim yn gwybod .  //
\glauto \%aut  not{\scriptsize .ADV} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN}   //
\glmanual \%gls  NEG PRT know{\scriptsize .NONFIN}   //
\gleng don't know. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (be)cause@s:eng [?] fydd Debbie@s:cym\&eng efo chdi . //
\glsurface ELE:  because$^{E}$ fydd Debbie$^{C}_{E}$ efo chdi .  //
\glauto \%aut  because{\scriptsize .CONJ} be{\scriptsize .V.3S.FUT+SM} name with{\scriptsize .PREP} you{\scriptsize .PRON}   //
\glmanual \%gls  because be{\scriptsize .3S.IMP} Debbie with PRON{\scriptsize .2S}   //
\gleng cause Debbie will be with you. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ALN:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat os na \&\textipa{S} (y)dy hi isio mynd at dy dad@s:cym\&eng . //
\glsurface ELE:  os na ydy hi isio mynd at dy dad$^{C}_{E}$ .  //
\glauto \%aut  if{\scriptsize .CONJ} neg{\scriptsize .PRT} be{\scriptsize .V.3S.PRES} she{\scriptsize .PRON.F.3S} want{\scriptsize .N.M.SG} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} your{\scriptsize .ADJ.POSS} dad{\scriptsize .N.SG[or]tad.N.SG+SM}   //
\glmanual \%gls  if NEG be{\scriptsize .3S.PRES} PRON{\scriptsize .3SF} want go{\scriptsize .NONFIN} to POSS{\scriptsize .2S} father   //
\gleng if she doesn't want to go to your dad's. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm mi allasai gallai . //
\glsurface ALN:  mmm mi allasai gallai .  //
\glauto \%aut  unk I{\scriptsize .PRON.1S[or]aff.PRT} be\_able{\scriptsize .V.3S.PLUPERF+SM} be\_able{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  IM PRT can{\scriptsize .3S.CONDIT} can{\scriptsize .3S.CONDIT}   //
\gleng mmm, she could, couldn't she //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ella (ba)sai dad@s:cym\&eng yn mwynhau dod fyny roid hi gwely . //
\glsurface ALN:  ella basai dad$^{C}_{E}$ yn mwynhau dod fyny roid hi gwely .  //
\glauto \%aut  maybe{\scriptsize .ADV.SPOKEN} be{\scriptsize .V.3S.PLUPERF.SPOKEN} dad{\scriptsize .N.SG[or]tad.N.SG+SM} stative{\scriptsize .STAT} enjoy{\scriptsize .V.INFIN} come{\scriptsize .V.INFIN} above{\scriptsize .ADV} give{\scriptsize .V.0.IMPERF+SM} she{\scriptsize .PRON.F.3S} bed{\scriptsize .N.M.SG}   //
\glmanual \%gls  maybe be{\scriptsize .3S.CONDIT} dad PRT enjoy{\scriptsize .NONFIN} come{\scriptsize .NONFIN} up put{\scriptsize .NONFIN} PRON{\scriptsize .3SF} bed   //
\gleng Dad might enjoy coming up to put her to bed. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well@s:cym\&eng ceith ddarllen story@s:cym\&eng hir iddi ceith ? //
\glsurface ALN:  well$^{C}_{E}$ ceith ddarllen story$^{C}_{E}$ hir iddi ceith ?  //
\glauto \%aut  well{\scriptsize .ADV} get{\scriptsize .V.3S.PRES} read{\scriptsize .V.INFIN+SM} story{\scriptsize .N.SG} long{\scriptsize .ADJ} to\_her{\scriptsize .PREP+PRON.F.3S} get{\scriptsize .V.3S.PRES}   //
\glmanual \%gls  well get{\scriptsize .3S.NONPAST} read{\scriptsize .NONFIN} story long to{\scriptsize .3SF} get{\scriptsize .3S.NONPAST}   //
\gleng well, he can read her a long story, can't he? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae \&maɨs +/ . //
\glsurface ALN:  mae .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES}   //
\glmanual \%gls  be{\scriptsize .3S.PRES}   //
\gleng there's \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (ba)sai mwy diddorol iddi na ddod i capel basai ? //
\glsurface ELE:  basai mwy diddorol iddi na ddod i capel basai ?  //
\glauto \%aut  be{\scriptsize .V.3S.PLUPERF.SPOKEN} more{\scriptsize .ADJ.COMP} interesting{\scriptsize .ADJ} to\_her{\scriptsize .PREP+PRON.F.3S} (n)or{\scriptsize .CONJ} come{\scriptsize .V.INFIN+SM} to{\scriptsize .PREP} chapel{\scriptsize .N.M.SG} be{\scriptsize .V.3S.PLUPERF.SPOKEN}   //
\glmanual \%gls  be{\scriptsize .3S.CONDIT} more interesting to{\scriptsize .3SF} than come{\scriptsize .NONFIN} to chapel be{\scriptsize .3S.CONDIT}   //
\gleng it would be more interesting for her than going to chapel, wouldn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o ran $<$(ba)sai \&x$>$ [///] chwarae gwmpas fasai de ? //
\glsurface ELE:  o ran basai chwarae gwmpas fasai de ?  //
\glauto \%aut  of{\scriptsize .PREP} part{\scriptsize .N.F.SG+SM} be{\scriptsize .V.3S.PLUPERF.SPOKEN} play{\scriptsize .V.INFIN} pair{\scriptsize .of.compasses.N.M.SG+SM[or]scope.N.M.SG+SM} be{\scriptsize .V.3S.PLUPERF.SPOKEN+SM} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  on part be{\scriptsize .3S.CONDIT} play{\scriptsize .NONFIN} around be{\scriptsize .3S.CONDIT} TAG   //
\gleng in that, she'd just be playing around, wouldn't she //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ar\_ôl tro dwytha . //
\glsurface ELE:  ar\_ôl tro dwytha .  //
\glauto \%aut  after{\scriptsize .PREP} turn{\scriptsize .N.M.SG} last{\scriptsize .ADJ}   //
\glmanual \%gls  after time previous   //
\gleng after last time. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat er dw (ddi)m yn dallt pam maen [?] nhw (y)n ymarfer yn +// . //
\glsurface ALN:  er dw ddim yn dallt pam maen nhw yn ymarfer yn .  //
\glauto \%aut  er{\scriptsize .IM} be{\scriptsize .V.1S.PRES.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} understand{\scriptsize .V.INFIN.SPOKEN} why?.ADV be{\scriptsize .V.3P.PRES.SPOKEN} they{\scriptsize .PRON.3P} stative{\scriptsize .STAT} practice{\scriptsize .V.INFIN} yn{\scriptsize .PRT}   //
\glmanual \%gls  although be{\scriptsize .1S.PRES} NEG PRT understand{\scriptsize .NONFIN} why be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL} PRT practise{\scriptsize .NONFIN} PRT   //
\gleng although, I don't get why they practise \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$mae nhw$>$ [/] maen nhw (y)n dal i ymarfer ddwy waith . //
\glsurface ALN:  mae nhw maen nhw yn dal i ymarfer ddwy waith .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} they{\scriptsize .PRON.3P} be{\scriptsize .V.3P.PRES.SPOKEN} they{\scriptsize .PRON.3P} stative{\scriptsize .STAT} hold{\scriptsize .V.INFIN} I{\scriptsize .PRON.1S} practice{\scriptsize .V.INFIN} two{\scriptsize .NUM.F+SM} time{\scriptsize .N.F.SG+SM}   //
\glmanual \%gls  be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL} be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL} PRT continue{\scriptsize .NONFIN} to practise{\scriptsize .NONFIN} two{\scriptsize .F} time   //
\gleng they still practise twice. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (dy)dy o (ddi)m fel bod ni (y)n trio \#\# um@s:cym\&eng \# tynnu \# chwaraewyr newydd i\_fewn ar hyn o bryd chwaith . //
\glsurface ALN:  dydy o ddim fel bod ni yn trio um$^{C}_{E}$ tynnu chwaraewyr newydd i\_fewn ar hyn o bryd chwaith .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} he{\scriptsize .PRON.M.3S.SPOKEN} not{\scriptsize .ADV+SM} like{\scriptsize .CONJ} be{\scriptsize .V.INFIN} we{\scriptsize .PRON.1P} stative{\scriptsize .STAT} try{\scriptsize .V.INFIN} um{\scriptsize .IM} take{\scriptsize .V.INFIN} players{\scriptsize .N.M.PL[or]players.N.M.PL} new{\scriptsize .ADJ} in{\scriptsize .ADV} on{\scriptsize .PREP} this{\scriptsize .PRON.SP} of{\scriptsize .PREP} mind{\scriptsize .N.M.SG[or]when.INT+SM[or]time.N.M.SG+SM} neither{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.PRES.NEG} PRON{\scriptsize .3SM} NEG like be{\scriptsize .1PL.PRES} PRON{\scriptsize .1PL} PRT try{\scriptsize .NONFIN} IM pull{\scriptsize .NONFIN} players new in at this of time either   //
\gleng it's not as though we were trying to pull in new players at the moment either. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae pawb sy yn y ddau band@s:cym\&eng yn brofiadol erbyn hyn . //
\glsurface ALN:  mae pawb sy yn y ddau band$^{C}_{E}$ yn brofiadol erbyn hyn .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} everyone{\scriptsize .PRON} which\_is{\scriptsize .V.3SP.PRES.REL} in{\scriptsize .PREP} the{\scriptsize .DET.DEF} two{\scriptsize .NUM.M+SM} band{\scriptsize .N.SG} stative{\scriptsize .STAT} experienced{\scriptsize .ADJ+SM} by{\scriptsize .PREP} this{\scriptsize .PRON.SP}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} everyone be{\scriptsize .PRES.REL} in DET two{\scriptsize .M} band PRT experienced by this   //
\gleng everyone who's in both bands is experienced by now. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond \# dw i just@s:cym\&eng yn cau ngheg . //
\glsurface ALN:  ond dw i just$^{C}_{E}$ yn cau ngheg .  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} just{\scriptsize .ADV[or]just.ADJ} stative{\scriptsize .STAT} close{\scriptsize .V.INFIN} mouth{\scriptsize .N.F.SG+NM}   //
\glmanual \%gls  but be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} just PRT close{\scriptsize .NONFIN} mouth{\scriptsize .POSSD.1S}   //
\gleng but, I just keep my mouth shut. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmhm@s:cym\&eng . //
\glsurface ELE:  mmhm$^{C}_{E}$ .  //
\glauto \%aut  mmhm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (be)cause@s:eng os dw i mynegi barn \#\#\# $<$ella fydd raid mi$>$ [?] wneud rywbeth . //
\glsurface ALN:  because$^{E}$ os dw i mynegi barn ella fydd raid mi wneud rywbeth .  //
\glauto \%aut  because{\scriptsize .CONJ} if{\scriptsize .CONJ} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} express{\scriptsize .V.INFIN} opinion{\scriptsize .N.F.SG[or]judge.V.2S.IMPER} maybe{\scriptsize .ADV.SPOKEN} be{\scriptsize .V.3S.FUT+SM} necessity{\scriptsize .N.M.SG+SM} I{\scriptsize .PRON.1S[or]aff.PRT} make{\scriptsize .V.INFIN+SM} something{\scriptsize .N.M.SG+SM}   //
\glmanual \%gls  because if be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT express{\scriptsize .NONFIN} opinion maybe be{\scriptsize .3S.FUT} necessity PRON{\scriptsize .1S} do{\scriptsize .NONFIN} something  //
\gleng cause if I express an opinion, I might have to do something. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat well@s:cym\&eng ie ! //
\glsurface ELE:  well$^{C}_{E}$ ie !  //
\glauto \%aut  well{\scriptsize .ADV} yes{\scriptsize .ADV}   //
\glmanual \%gls  well yes   //
\gleng well, yes! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a mae gen i ddigon ar y mhlât yn barod . //
\glsurface ALN:  a mae gen i ddigon ar y mhlât yn barod .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} by{\scriptsize .PREP.SPOKEN} I{\scriptsize .PRON.1S} enough{\scriptsize .QUAN+SM} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} plate{\scriptsize .N.M.SG+NM} stative{\scriptsize .STAT} ready{\scriptsize .ADJ+SM}   //
\glmanual \%gls  and be{\scriptsize .3S.PRES} with{\scriptsize .1S} PRON{\scriptsize .1S} enough on POSS{\scriptsize .1S} plate PRT ready   //
\gleng and I have enough on my plate already. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat sbïa mae (y)na rywun (we)di baeddu +. . . //
\glsurface ELE:  sbïa mae yna rywun wedi baeddu .  //
\glauto \%aut  look{\scriptsize .V.2S.IMPER} be{\scriptsize .V.3S.PRES} there{\scriptsize .ADV} someone{\scriptsize .N.M.SG+SM} after{\scriptsize .PREP} soil{\scriptsize .V.INFIN}   //
\glmanual \%gls  look{\scriptsize .2S.IMPER} be{\scriptsize .3S.PRES} there someone PRT{\scriptsize .PAST} soil{\scriptsize .NONFIN}   //
\gleng look, someone's soiled. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat tsiocled yeah@s:cym\&eng ? //
\glsurface ALN:  tsiocled yeah$^{C}_{E}$ ?  //
\glauto \%aut  chocolate{\scriptsize .N.M.SG} yeah{\scriptsize .ADV}   //
\glmanual \%gls  chocolate yes   //
\gleng chocolate, yeah? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$wnei di$>$ [=? o'n i (we)di] golchi (r)hein ? //
\glsurface ELE:  wnei di golchi rhein ?  //
\glauto \%aut  do{\scriptsize .V.2S.PRES+SM} you{\scriptsize .PRON.2S+SM} wash{\scriptsize .V.INFIN} these{\scriptsize .PRON}   //
\glmanual \%gls  do{\scriptsize .2S.NONPAST} PRON{\scriptsize .2S} wash{\scriptsize .NONFIN} PRON{\scriptsize .3SF}   //
\gleng could you wash it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng de ? //
\glsurface ELE:  yeah$^{C}_{E}$ de ?  //
\glauto \%aut  yeah{\scriptsize .ADV} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  yeah TAG   //
\gleng yes? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat bound@s:cym\&eng o ddigwydd dydy ? //
\glsurface ALN:  bound$^{C}_{E}$ o ddigwydd dydy ?  //
\glauto \%aut  bound{\scriptsize .SV.INFIN[or]pound.SV.INFIN+SM} of{\scriptsize .PREP} happen{\scriptsize .V.INFIN+SM} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM}   //
\glmanual \%gls  bound of happen{\scriptsize .NONFIN} be{\scriptsize .3S.PRES.NEG}   //
\gleng bound to happen, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be ? //
\glsurface ELE:  be ?  //
\glauto \%aut  what{\scriptsize .INT}   //
\glmanual \%gls  what   //
\gleng what? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat bound@s:cym\&eng o ddigwydd dydy ? //
\glsurface ALN:  bound$^{C}_{E}$ o ddigwydd dydy ?  //
\glauto \%aut  bound{\scriptsize .SV.INFIN[or]pound.SV.INFIN+SM} of{\scriptsize .PREP} happen{\scriptsize .V.INFIN+SM} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM}   //
\glmanual \%gls  bound of happen{\scriptsize .NONFIN} be{\scriptsize .3S.PRES.NEG}   //
\gleng bound to happen, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng (we)di [=? dw i] gobeithio (ba)sen nhw [?] (we)di para ychydig bach glanach yn hirach . //
\glsurface ELE:  yeah$^{C}_{E}$ wedi gobeithio basen nhw wedi para ychydig bach glanach yn hirach .  //
\glauto \%aut  yeah{\scriptsize .ADV} after{\scriptsize .PREP} hope{\scriptsize .V.INFIN} be{\scriptsize .V.3P.PLUPERF.SPOKEN} they{\scriptsize .PRON.3P} after{\scriptsize .PREP} last{\scriptsize .V.INFIN} a\_little{\scriptsize .QUAN} small{\scriptsize .ADJ} clean{\scriptsize .ADJ.COMP} stative{\scriptsize .STAT} long{\scriptsize .ADJ.COMP}   //
\glmanual \%gls  yeah PRT{\scriptsize .PAST} hope{\scriptsize .NONFIN} be{\scriptsize .3PL.CONDIT} PRT{\scriptsize .PAST} last{\scriptsize .NONFIN} a\_bit little cleaner PRT longer    //
\gleng yeah, I'd hoped they'd have lasted a bit cleaner for longer. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat efo breichiau mor lydan \# maen nhw just@s:cym\&eng roi bwyd arnyn nhw yn\_dydyn ? //
\glsurface ALN:  efo breichiau mor lydan maen nhw just$^{C}_{E}$ roi bwyd arnyn nhw yn\_dydyn ?  //
\glauto \%aut  with{\scriptsize .PREP} arms{\scriptsize .N.F.PL} so{\scriptsize .ADV} wide{\scriptsize .ADJ+SM[or]broad.ADJ+SM} be{\scriptsize .V.3P.PRES.SPOKEN} they{\scriptsize .PRON.3P} just{\scriptsize .ADV[or]just.ADJ} give{\scriptsize .V.INFIN+SM} food{\scriptsize .N.M.SG} on\_them{\scriptsize .PREP+PRON.3P} they{\scriptsize .PRON.3P} aren't\_they{\scriptsize .IM}   //
\glmanual \%gls  with arms so wide be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL} just put{\scriptsize .NONFIN} food on{\scriptsize .3PL} PRON{\scriptsize .3PL} be{\scriptsize .3PL.NONPAST.NEG}   //
\gleng with such wide arms, they just put food on them don't they? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yndyn . //
\glsurface ELE:  yndyn .  //
\glauto \%aut  be{\scriptsize .V.3P.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .3PL.PRES}   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat just@s:cym\&eng bwrdd (y)dy o de ? //
\glsurface ALN:  just$^{C}_{E}$ bwrdd ydy o de ?  //
\glauto \%aut  just{\scriptsize .ADV[or]just.ADJ} table{\scriptsize .N.M.SG} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  just table be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} TAG   //
\gleng it's just a table, isn't it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat hmm@s:cym\&eng yeah@s:cym\&eng . //
\glsurface ELE:  hmm$^{C}_{E}$ yeah$^{C}_{E}$ .  //
\glauto \%aut  hmm{\scriptsize .IM} yeah{\scriptsize .ADV}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat bach dyn nhw sti . //
\glsurface ALN:  bach dyn nhw sti .  //
\glauto \%aut  small{\scriptsize .ADJ} man{\scriptsize .N.M.SG[or]draw.V.3S.PRES+SM} they{\scriptsize .PRON.3P} you\_know{\scriptsize .IM}   //
\glmanual \%gls  little be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL} know{\scriptsize .2S}   //
\gleng they're little, you know. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ia llythyr Nicky@s:cym\&eng bore (y)ma \# $<$neu \&d Olivia@s:cym\&eng$>$ [?] . //
\glsurface ELE:  ia llythyr Nicky$^{C}_{E}$ bore yma neu Olivia$^{C}_{E}$ .  //
\glauto \%aut  yes{\scriptsize .ADV} letter{\scriptsize .N.M.SG} name morning{\scriptsize .N.M.SG} here{\scriptsize .ADV} or{\scriptsize .CONJ} name   //
\glmanual \%gls  yes letter Nicky morning here or Olivia   //
\gleng yes, Nicky's letter this morning. . . or Olivia's. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dwn i (ddi)m pwy sy (we)di bod wrthi . //
\glsurface ELE:  dwn i ddim pwy sy wedi bod wrthi .  //
\glauto \%aut  know{\scriptsize .V.1S.PRES.NEG} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} who{\scriptsize .PRON} which\_is{\scriptsize .V.3SP.PRES.REL} after{\scriptsize .PREP} be{\scriptsize .V.INFIN} to\_her{\scriptsize .PREP+PRON.F.3S}   //
\glmanual \%gls  know{\scriptsize .1S.NONPAST.NEG} PRON{\scriptsize .1S} NEG who be{\scriptsize .PRES.REL} PRT{\scriptsize .PAST} be{\scriptsize .NONFIN} at{\scriptsize .3SF}   //
\gleng I don't know who's been at it. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat y syniad gwreiddiol oedd bod ni mynd ar drain@s:cym\&eng i Rhyd\_Ddu@s:cym\&eng . //
\glsurface ELE:  y syniad gwreiddiol oedd bod ni mynd ar drain$^{C}_{E}$ i Rhyd\_Ddu$^{C}_{E}$ .  //
\glauto \%aut  the{\scriptsize .DET.DEF} idea{\scriptsize .N.M.SG} original{\scriptsize .ADJ} be{\scriptsize .V.3S.IMPERF} be{\scriptsize .V.INFIN} we{\scriptsize .PRON.1P} go{\scriptsize .V.INFIN} on{\scriptsize .PREP} drain{\scriptsize .N.SG[or]train.SV.INFIN+SM} to{\scriptsize .PREP} name   //
\glmanual \%gls  DET idea original be{\scriptsize .3S.IMP} be{\scriptsize .NONFIN} PRON{\scriptsize .1PL} go{\scriptsize .NONFIN} on train to Rhyd\_Ddu   //
\gleng the original idea was that we'd go by train to Rhyd Ddu. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a [?] rŵan \# dan ni mynd o\_amgylch gwahanol lefydd \# llenyddol barddonol . //
\glsurface ELE:  a rŵan dan ni mynd o\_amgylch gwahanol lefydd llenyddol barddonol .  //
\glauto \%aut  and{\scriptsize .CONJ} now{\scriptsize .ADV} be{\scriptsize .V.1P.PRES.SPOKEN} we{\scriptsize .PRON.1P} go{\scriptsize .V.INFIN} around{\scriptsize .PREP} different{\scriptsize .ADJ} places{\scriptsize .N.M.PL+SM} literary{\scriptsize .ADJ} poetic{\scriptsize .ADJ}   //
\glmanual \%gls  and now be{\scriptsize .1PL.PRES} PRON{\scriptsize .1PL} PRT go{\scriptsize .NONFIN} around different places literary poeticol  //
\gleng and now we're going around various poetical, literary sites. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a dan ni (y)n gorod \# um@s:cym\&eng \# ffeindio (rhyw)beth am y person@s:cym\&eng (y)na . //
\glsurface ELE:  a dan ni yn gorod um$^{C}_{E}$ ffeindio rhywbeth am y person$^{C}_{E}$ yna .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.1P.PRES.SPOKEN} we{\scriptsize .PRON.1P} stative{\scriptsize .STAT} have\_to{\scriptsize .V.INFIN} um{\scriptsize .IM} find{\scriptsize .V.INFIN} something{\scriptsize .N.M.SG} for{\scriptsize .PREP} the{\scriptsize .DET.DEF} person{\scriptsize .N.SG} there{\scriptsize .ADV}   //
\glmanual \%gls  and be{\scriptsize .1PL.PRES} PRON{\scriptsize .1PL} PRT must{\scriptsize .NONFIN} IM find{\scriptsize .NONFIN} something about DET person there   //
\gleng and we've got to find out something about that person. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i (we)di cael Emily\_Huws@s:cym\&eng yn Bontnewydd@s:cym\&eng \# yn [=? y] Newborough@s:eng . //
\glsurface ELE:  dw i wedi cael Emily\_Huws$^{C}_{E}$ yn Bontnewydd$^{C}_{E}$ yn Newborough$^{E}$ .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} after{\scriptsize .PREP} have{\scriptsize .V.INFIN} name in{\scriptsize .PREP} name in{\scriptsize .PREP} name   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} PRT{\scriptsize .PAST} get{\scriptsize .NONFIN} Emily\_Huws in Bontnewydd in Newborough{\scriptsize .}   //
\gleng I've got Emily Huws, in Bontnewydd, in the Newborough. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$dw i gorod$>$ [?] darllen rywbeth o ei gwaith hi . //
\glsurface ELE:  dw i gorod darllen rywbeth o ei gwaith hi .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} have\_to{\scriptsize .V.INFIN} read{\scriptsize .V.INFIN} something{\scriptsize .N.M.SG+SM} of{\scriptsize .PREP} her{\scriptsize .ADJ.POSS} time{\scriptsize .N.F.SG[or]work.N.M.SG} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} must{\scriptsize .NONFIN} read{\scriptsize .NONFIN} something of POSS{\scriptsize .3SF} work PRON{\scriptsize .3SF}   //
\gleng I've got to read some of her work. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae Ceridwen@s:cym\&eng yn gorod wneud (ry)wbeth task@s:cym\&eng yn fan (y)na (fe)lly . //
\glsurface ELE:  mae Ceridwen$^{C}_{E}$ yn gorod wneud rywbeth task$^{C}_{E}$ yn fan yna felly .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} name stative{\scriptsize .STAT} have\_to{\scriptsize .V.INFIN} make{\scriptsize .V.INFIN+SM} something{\scriptsize .N.M.SG+SM} task{\scriptsize .N.SG} stative{\scriptsize .STAT} place{\scriptsize .N.MF.SG+SM} there{\scriptsize .ADV} so{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} Ceridwen PRT must{\scriptsize .NONFIN} do{\scriptsize .NONFIN} something task in place there thus   //
\gleng Ceridwen has to do something, a task there, like. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (we)dyn fyddan ni mynd i Rhos:gadfan@s:cym\&eng . //
\glsurface ELE:  wedyn fyddan ni mynd i Rhosgadfan$^{C}_{E}$ .  //
\glauto \%aut  afterwards{\scriptsize .ADV} be{\scriptsize .V.3P.FUT.SPOKEN+SM} we{\scriptsize .PRON.1P} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} name   //
\glmanual \%gls  then be{\scriptsize .1PL.FUT} PRON{\scriptsize .1PL} go{\scriptsize .NONFIN} to Rhosgadfan   //
\gleng then we're going to Rhosgadfan. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae (y)n(a) dafarn fan (y)na hefyd does ? //
\glsurface ELE:  mae yna dafarn fan yna hefyd does ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} there{\scriptsize .ADV} tavern{\scriptsize .N.MF.SG+SM} place{\scriptsize .N.MF.SG+SM} there{\scriptsize .ADV} also{\scriptsize .ADV} be{\scriptsize .V.3S.PRES.INDEF.NEG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} there pub place there too be{\scriptsize .3S.PRES.NEG}   //
\gleng there's a pub there too, isn't there? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng ri:ght@s:cym\&eng ! //
\glsurface ALN:  oh$^{C}_{E}$ right$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM} right{\scriptsize .ADJ}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (we)dyn mae (y)na ryw saith neu wyth o lefydd . //
\glsurface ELE:  wedyn mae yna ryw saith neu wyth o lefydd .  //
\glauto \%aut  afterwards{\scriptsize .ADV} be{\scriptsize .V.3S.PRES} there{\scriptsize .ADV} some{\scriptsize .PREQ+SM} seven{\scriptsize .NUM} or{\scriptsize .CONJ} eight{\scriptsize .NUM} of{\scriptsize .PREP} places{\scriptsize .N.M.PL+SM}   //
\glmanual \%gls  then be{\scriptsize .3S.PRES} there some seven or eight of places   //
\gleng then there are about seven or eight places. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat so@s:cym\&eng mae nhw (y)n benderfynol \# xxx sesh@s:cym\&eng (fe)lly ? //
\glsurface ALN:  so$^{C}_{E}$ mae nhw yn benderfynol sesh$^{C}_{E}$ felly ?  //
\glauto \%aut  so{\scriptsize .ADV} be{\scriptsize .V.3S.PRES} they{\scriptsize .PRON.3P} stative{\scriptsize .STAT} decisive{\scriptsize .ADJ+SM} session{\scriptsize .N.SG} so{\scriptsize .ADV}   //
\glmanual \%gls  so be{\scriptsize .3PL.PRES} PRON{\scriptsize .3PL} PRT determined drinking\_session thus   //
\gleng so they're determined xxx a sesh then ? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng yeah@s:cym\&eng dwn i (ddi)m . //
\glsurface ELE:  oh$^{C}_{E}$ yeah$^{C}_{E}$ dwn i ddim .  //
\glauto \%aut  oh{\scriptsize .IM} yeah{\scriptsize .ADV} know{\scriptsize .V.1S.PRES.NEG} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM}   //
\glmanual \%gls  IM yeah know{\scriptsize .1S.NONPAST.NEG} PRON{\scriptsize .1S} NEG   //
\gleng oh, yeah, don't know. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a wedyn um@s:cym\&eng \#\# +/ . //
\glsurface ELE:  a wedyn um$^{C}_{E}$ .  //
\glauto \%aut  and{\scriptsize .CONJ} afterwards{\scriptsize .ADV} um{\scriptsize .IM}   //
\glmanual \%gls  and then IM   //
\gleng and then, um \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat pa ddyddiad ? //
\glsurface ALN:  pa ddyddiad ?  //
\glauto \%aut  which{\scriptsize .ADJ} date{\scriptsize .N.M.SG+SM}   //
\glmanual \%gls  which date   //
\gleng what date? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, swper ym Molly's@s:cym\&eng . //
\glsurface ELE:  swper ym Molly's$^{C}_{E}$ .  //
\glauto \%aut  supper{\scriptsize .N.MF.SG} in{\scriptsize .PREP} name   //
\glmanual \%gls  dinner in Molly''s   //
\gleng dinner at Molly's. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat um:@s:cym\&eng pedwerydd o Dachwedd . //
\glsurface ELE:  um$^{C}_{E}$ pedwerydd o Dachwedd .  //
\glauto \%aut  um{\scriptsize .IM} fourth{\scriptsize .ORD.M} of{\scriptsize .PREP} November{\scriptsize .N.M.SG+SM}   //
\glmanual \%gls  IM fourth of November   //
\gleng um, the fourth of November. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat diwedd hanner tymor . //
\glsurface ELE:  diwedd hanner tymor .  //
\glauto \%aut  end{\scriptsize .N.M.SG} half{\scriptsize .N.M.SG} season{\scriptsize .N.M.SG}   //
\glmanual \%gls  end half term   //
\gleng end of half term. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat y dydd Sadwrn ola . //
\glsurface ELE:  y dydd Sadwrn ola .  //
\glauto \%aut  the{\scriptsize .DET.DEF} day{\scriptsize .N.M.SG} Saturn{\scriptsize .N.M.SG} last{\scriptsize .ADJ}   //
\glmanual \%gls  DET day Saturday last   //
\gleng the last Saturday. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat so@s:cym\&eng ti (y)n mynd allan y pnawn dydd Sadwrn olaf ? //
\glsurface ALN:  so$^{C}_{E}$ ti yn mynd allan y pnawn dydd Sadwrn olaf ?  //
\glauto \%aut  so{\scriptsize .ADV} you{\scriptsize .PRON.2S} stative{\scriptsize .STAT} go{\scriptsize .V.INFIN} out{\scriptsize .ADV} the{\scriptsize .DET.DEF} afternoon{\scriptsize .N.M.SG} day{\scriptsize .N.M.SG} Saturn{\scriptsize .N.M.SG} last{\scriptsize .ADJ}   //
\glmanual \%gls  so PRON{\scriptsize .2S} PRT go{\scriptsize .NONFIN} out DET afternoon day Saturday last   //
\gleng so you're going out on the afternoon of the last Saturday? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$a (y)r nos$>$ [=! yawns] . //
\glsurface ELE:  a yr nos .  //
\glauto \%aut  and{\scriptsize .CONJ} the{\scriptsize .DET.DEF} night{\scriptsize .N.F.SG}   //
\glmanual \%gls  and DET night   //
\gleng and the night. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ideal@s:cym\&eng ! //
\glsurface ALN:  ideal$^{C}_{E}$ !  //
\glauto \%aut  ideal{\scriptsize .N.SG}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat xxx nôl o Llundain dydd Gwener tydan ? //
\glsurface ELE:  nôl o Llundain dydd Gwener tydan ?  //
\glauto \%aut  back{\scriptsize .ADV[or]fetch.V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} London{\scriptsize .N.F.SG} day{\scriptsize .N.M.SG} name be{\scriptsize .V.1P.PRES.NEG}   //
\glmanual \%gls  back from London day Friday be{\scriptsize .1PL.PRES.NEG}   //
\gleng [ \dots ] back from London on Friday, aren't we? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yndan yndan . //
\glsurface ALN:  yndan yndan .  //
\glauto \%aut  be{\scriptsize .V.1P.PRES.EMPH} be{\scriptsize .V.1P.PRES.EMPH}   //
\glmanual \%gls  be{\scriptsize .1PL.PRES} be{\scriptsize .1PL.PRES}   //
\gleng we are, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na mae Cymru yn chwarae pnawn dydd Sadwrn . //
\glsurface ALN:  na mae Cymru yn chwarae pnawn dydd Sadwrn .  //
\glauto \%aut  neg{\scriptsize .PRT} be{\scriptsize .V.3S.PRES} Wales{\scriptsize .N.F.SG} stative{\scriptsize .STAT} play{\scriptsize .V.INFIN} afternoon{\scriptsize .N.M.SG} day{\scriptsize .N.M.SG} Saturn{\scriptsize .N.M.SG}   //
\glmanual \%gls  no be{\scriptsize .3S.PRES} Wales PRT play{\scriptsize .NONFIN} afternoon day Saturday   //
\gleng no, Wales are playing on Saturday afternoon. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yn\_erbyn pwy ? //
\glsurface ELE:  yn\_erbyn pwy ?  //
\glauto \%aut  against{\scriptsize .PREP[or]against.PRON} who{\scriptsize .PRON}   //
\glmanual \%gls  against who   //
\gleng against who? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat Australia@s:cym\&eng . //
\glsurface ALN:  Australia$^{C}_{E}$ .  //
\glauto \%aut  name   //
\glmanual \%gls  Australia   //
\gleng Australia. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng $<$o'n i (ddi)m gw(ybod) (hy)nny$>$ [?] . //
\glsurface ELE:  oh$^{C}_{E}$ o'n i ddim gwybod hynny .  //
\glauto \%aut  oh{\scriptsize .IM} unk to{\scriptsize .PREP} not{\scriptsize .ADV+SM} know{\scriptsize .V.INFIN} that{\scriptsize .PRON.SP}   //
\glmanual \%gls  IM be{\scriptsize .1S.IMP} PRON{\scriptsize .1S} NEG know{\scriptsize .NONFIN} that   //
\gleng oh, I didn't know that. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yn lle ? //
\glsurface ELE:  yn lle ?  //
\glauto \%aut  in{\scriptsize .PREP} where{\scriptsize .INT[or]place.N.M.SG}   //
\glmanual \%gls  in where   //
\gleng where? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yng Nghaerdydd . //
\glsurface ALN:  yng Nghaerdydd .  //
\glauto \%aut  in{\scriptsize .PREP} Cardiff{\scriptsize .NAME+NM}   //
\glmanual \%gls  in Cardiff   //
\gleng in Cardiff. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=dental\_click oh@s:cym\&eng ! //
\glsurface ELE:  oh$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae gen i gig@s:cym\&eng yn y dyddiadur . //
\glsurface ALN:  mae gen i gig$^{C}_{E}$ yn y dyddiadur .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} by{\scriptsize .PREP.SPOKEN} I{\scriptsize .PRON.1S} gig{\scriptsize .N.SG[or]cig.N.SG+SM} in{\scriptsize .PREP} the{\scriptsize .DET.DEF} diary{\scriptsize .N.M.SG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} with{\scriptsize .1S} PRON{\scriptsize .1S} gig in DET diary   //
\gleng I've got a gig in the diary. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be ar y nos Sadwrn ? //
\glsurface ELE:  be ar y nos Sadwrn ?  //
\glauto \%aut  what{\scriptsize .INT} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} night{\scriptsize .N.F.SG} Saturn{\scriptsize .N.M.SG}   //
\glmanual \%gls  what on DET night Saturday   //
\gleng what, on the Saturday night? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ar nos Sadwrn y pedwerydd . //
\glsurface ALN:  ar nos Sadwrn y pedwerydd .  //
\glauto \%aut  on{\scriptsize .PREP} night{\scriptsize .N.F.SG} Saturn{\scriptsize .N.M.SG} the{\scriptsize .DET.DEF} fourth{\scriptsize .ORD.M}   //
\glmanual \%gls  on night Saturday DET fourth   //
\gleng on Saturday night on the fourth. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat nag oes ! //
\glsurface ELE:  nag oes !  //
\glauto \%aut  than{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES.INDEF}   //
\glmanual \%gls  NEG be{\scriptsize .3S.PRES}   //
\gleng there isn't! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oes . //
\glsurface ALN:  oes .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.INDEF}   //
\glmanual \%gls  be{\scriptsize .3S.PRES}   //
\gleng there is. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ $<$wnest di ddeud$>$ [?] bod o (y)n rhydd . //
\glsurface ELE:  wnest di ddeud bod o yn rhydd .  //
\glauto \%aut  do{\scriptsize .V.2S.PAST.SPOKEN+SM} you{\scriptsize .PRON.2S+SM} say{\scriptsize .V.INFIN.SPOKEN+SM} be{\scriptsize .V.INFIN} he{\scriptsize .PRON.M.3S.SPOKEN} stative{\scriptsize .STAT} free{\scriptsize .ADJ[or]give.V.3S.PRES}   //
\glmanual \%gls  do{\scriptsize .2S.PAST} PRON{\scriptsize .2S} say{\scriptsize .NONFIN} be{\scriptsize .NONFIN} PRON{\scriptsize .3SM} PRT free   //
\gleng you said it was free. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat gig@s:cym\&eng be ? //
\glsurface ELE:  gig$^{C}_{E}$ be ?  //
\glauto \%aut  gig{\scriptsize .N.SG[or]cig.N.SG+SM} what{\scriptsize .INT}   //
\glmanual \%gls  gig what   //
\gleng what gig? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$Acoustique@s:cym\&eng yn$>$ [?] Sandy\_Beach@s:cym\&eng . //
\glsurface ALN:  Acoustique$^{C}_{E}$ yn Sandy\_Beach$^{C}_{E}$ .  //
\glauto \%aut  name in{\scriptsize .PREP} name   //
\glmanual \%gls       //
\gleng Acoustique and Sandy Beach. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=dental\_click \# a:h@s:cym\&eng ! //
\glsurface ELE:  ah$^{C}_{E}$ !  //
\glauto \%aut  ah{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond dw i (ddi)m yn gwybod os (y)dy o (y)mlaen . //
\glsurface ALN:  ond dw i ddim yn gwybod os ydy o ymlaen .  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN} if{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} he{\scriptsize .PRON.M.3S.SPOKEN} forward{\scriptsize .ADV}   //
\glmanual \%gls  but be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} NEG PRT know{\scriptsize .NONFIN} if be{\scriptsize .3S.PRES} PRON{\scriptsize .3SM} forward   //
\gleng but I don't know whether it's on. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mi grybyllwyd o oes pys yn\_ôl . //
\glsurface ALN:  mi grybyllwyd o oes pys yn\_ôl .  //
\glauto \%aut  aff{\scriptsize .PRT} mention{\scriptsize .V.0.PAST+SM} he{\scriptsize .PRON.M.3S.SPOKEN[or]from.PREP[or]of.PREP} be{\scriptsize .V.3S.PRES.INDEF[or]age.N.F.SG} pea{\scriptsize .N.F.PL} back{\scriptsize .ADV}   //
\glmanual \%gls  PRT suggest{\scriptsize .IMPERSONAL.PAST} PRON{\scriptsize .3SM} age peas back   //
\gleng it was mentioned ages ago //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i (ddi)m (we)di clywed dim\_byd un ffordd neu llall . //
\glsurface ALN:  dw i ddim wedi clywed dim\_byd un ffordd neu llall .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} after{\scriptsize .PREP} hear{\scriptsize .V.INFIN} anything{\scriptsize .ADV} one{\scriptsize .NUM} road{\scriptsize .N.F.SG} or{\scriptsize .CONJ} other{\scriptsize .PRON}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} NEG PRT{\scriptsize .PAST} hear{\scriptsize .NONFIN} nothing one way or other   //
\gleng I've not heard anything one way or the other. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ oh@s:cym\&eng ! //
\glsurface ELE:  oh$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$wna i$>$ [/] wna i dext\_io@s:eng+cym Tracy@s:cym\&eng wedyn . //
\glsurface ALN:  wna i wna i dext\_io$^{E+}_{C}$ Tracy$^{C}_{E}$ wedyn .  //
\glauto \%aut  do{\scriptsize .V.13S.PRES+SM} I{\scriptsize .PRON.1S[or]to.PREP} do{\scriptsize .V.13S.PRES+SM} I{\scriptsize .PRON.1S[or]to.PREP} text{\scriptsize .N.SG+SM+WV} name afterwards{\scriptsize .ADV}   //
\glmanual \%gls  do{\scriptsize .1S.NONPAST} PRON{\scriptsize .1S} do{\scriptsize .1S.NONPAST} PRON{\scriptsize .1S} text{\scriptsize .NONFIN} Tracy after   //
\gleng I'll text Tracy afterwards. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wna i ffonio hi . //
\glsurface ALN:  wna i ffonio hi .  //
\glauto \%aut  do{\scriptsize .V.13S.PRES+SM} I{\scriptsize .PRON.1S[or]to.PREP} phone{\scriptsize .V.INFIN} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  do{\scriptsize .1S.NONPAST} PRON{\scriptsize .1S} phone{\scriptsize .NONFIN} PRON{\scriptsize .3SF}   //
\gleng I'll phone her. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wnaeth hi (ddi)m neidio dydd [?] Sadwrn . //
\glsurface ALN:  wnaeth hi ddim neidio dydd Sadwrn .  //
\glauto \%aut  do{\scriptsize .V.3S.PAST+SM} she{\scriptsize .PRON.F.3S} not{\scriptsize .ADV+SM} jump{\scriptsize .V.INFIN} day{\scriptsize .N.M.SG} Saturn{\scriptsize .N.M.SG}   //
\glmanual \%gls  do{\scriptsize .3S.PAST} PRON{\scriptsize .3SF} NEG jump{\scriptsize .NONFIN} day Saturday   //
\gleng she didn't jump on Saturday. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat wnaeth hi ddim ? //
\glsurface ELE:  wnaeth hi ddim ?  //
\glauto \%aut  do{\scriptsize .V.3S.PAST+SM} she{\scriptsize .PRON.F.3S} not{\scriptsize .ADV+SM}   //
\glmanual \%gls  do{\scriptsize .3S.PAST} PRON{\scriptsize .3SF} NEG   //
\gleng she didn't? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oedd (y)na ormod o wynt na gormod o wynt . //
\glsurface ALN:  oedd yna ormod o wynt na gormod o wynt .  //
\glauto \%aut  be{\scriptsize .V.3S.IMPERF} there{\scriptsize .ADV} too\_much{\scriptsize .QUANT+SM} of{\scriptsize .PREP} wind{\scriptsize .N.M.SG+SM} neg{\scriptsize .PRT[or]who\_not.PRON.REL.NEG[or](n)or.CONJ[or]than.CONJ} too\_much{\scriptsize .QUANT} of{\scriptsize .PREP} wind{\scriptsize .N.M.SG+SM}   //
\glmanual \%gls  be{\scriptsize .3S.IMP} there too\_much of wind no too\_much of wind   //
\gleng there was too much wind, no, too much wind. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng phew@s:cym\&eng ! //
\glsurface ELE:  oh$^{C}_{E}$ phew$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM} phew{\scriptsize .IM}   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat pumed ar hugain o:: Dachwedd . //
\glsurface ALN:  pumed ar hugain o Dachwedd .  //
\glauto \%aut  fifth{\scriptsize .ORD} on{\scriptsize .PREP} twenty{\scriptsize .NUM+H} of{\scriptsize .PREP} November{\scriptsize .N.M.SG+SM}   //
\glmanual \%gls  fifth on twenty of November   //
\gleng 25th of November. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ew nefi blue@s:cym\&eng ! //
\glsurface ELE:  ew nefi blue$^{C}_{E}$ !  //
\glauto \%aut  oh{\scriptsize .IM} heavens{\scriptsize .IM} blue{\scriptsize .N.SG}   //
\glmanual \%gls  IM heavens blue   //
\gleng oh, good heavens! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a deud i mi wan de . //
\glsurface ELE:  a deud i mi wan de .  //
\glauto \%aut  and{\scriptsize .CONJ} say{\scriptsize .V.INFIN.SPOKEN} to{\scriptsize .PREP} I{\scriptsize .PRON.1S} weak{\scriptsize .ADJ+SM[or]pale.ADJ+SM} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  and say{\scriptsize .2S.IMPER} to PRON{\scriptsize .1S} now TAG   //
\gleng and tell me now, right. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ydy [/] \#\# (y)dy Queenie@s:cym\&eng [?] (we)di [/] (we)di bod yn [//] nôl yn \#\# Prince's\_Drive@s:eng neu +. . . //
\glsurface ELE:  ydy ydy Queenie$^{C}_{E}$ wedi wedi bod yn nôl yn Prince's\_Drive$^{E}$ neu .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} be{\scriptsize .V.3S.PRES} name after{\scriptsize .PREP} after{\scriptsize .PREP} be{\scriptsize .V.INFIN} stative{\scriptsize .STAT} fetch{\scriptsize .V.INFIN} in{\scriptsize .PREP} name or{\scriptsize .CONJ}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES} Queenie PRT{\scriptsize .PAST} PRT{\scriptsize .PAST} be{\scriptsize .NONFIN} in back in Prince''s\_Drive or   //
\gleng has Queenie been back to Prince's Drive, or \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ddim (we)di siarad yn ddiweddar . //
\glsurface ALN:  ddim wedi siarad yn ddiweddar .  //
\glauto \%aut  nothing{\scriptsize .N.M.SG+SM[or]not.ADV+SM} after{\scriptsize .PREP} talk{\scriptsize .V.INFIN} stative{\scriptsize .STAT} recent{\scriptsize .ADJ+SM}   //
\glmanual \%gls  NEG PRT{\scriptsize .PAST} talk{\scriptsize .NONFIN} PRT recent   //
\gleng haven't talked lately. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=sniff mmhm@s:cym\&eng . //
\glsurface ELE:  mmhm$^{C}_{E}$ .  //
\glauto \%aut  mmhm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae (y)n deud bod hi (y)n cael er@s:cym\&eng +. . . //
\glsurface ALN:  mae yn deud bod hi yn cael er$^{C}_{E}$ .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} say{\scriptsize .V.INFIN.SPOKEN} be{\scriptsize .V.INFIN} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} have{\scriptsize .V.INFIN} er{\scriptsize .IM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRT say{\scriptsize .NONFIN} be{\scriptsize .NONFIN} PRON{\scriptsize .3SF} PRT get{\scriptsize .NONFIN} IM   //
\gleng she says she's getting, er \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae isio symud eto . //
\glsurface ALN:  mae isio symud eto .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} want{\scriptsize .N.M.SG} move{\scriptsize .V.INFIN} again{\scriptsize .ADV}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} want move{\scriptsize .NONFIN} again   //
\gleng she wants to move again. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be o Penmon@s:cym\&eng ? //
\glsurface ELE:  be o Penmon$^{C}_{E}$ ?  //
\glauto \%aut  what{\scriptsize .INT} from{\scriptsize .PREP} name   //
\glmanual \%gls  what from Penmon   //
\gleng what, from Penmon? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ia . //
\glsurface ALN:  ia .  //
\glauto \%aut  yes{\scriptsize .ADV}   //
\glmanual \%gls  yes   //
\gleng yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat pam ? //
\glsurface ELE:  pam ?  //
\glauto \%aut  why?.ADV   //
\glmanual \%gls  why   //
\gleng why? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae gynni fancy@s:cym\&eng $<$mynd xxx$>$ [//] mynd i fyw ym Mharis@s:cym\&eng am gyfnod . //
\glsurface ALN:  mae gynni fancy$^{C}_{E}$ mynd mynd i fyw ym Mharis$^{C}_{E}$ am gyfnod .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} with\_her{\scriptsize .PREP+PRON.F.3S[or]gouge.V.2S.PRES+SM[or]light.V.2S.PRES+SM} fancy{\scriptsize .ADJ} go{\scriptsize .V.INFIN} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} live{\scriptsize .V.INFIN+SM} in{\scriptsize .PREP} name for{\scriptsize .PREP} period{\scriptsize .N.M.SG+SM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} with{\scriptsize .3SF} fancy go{\scriptsize .NONFIN} go{\scriptsize .NONFIN} to live{\scriptsize .NONFIN} in Paris for period   //
\gleng she fancies going to live in Paris for a while. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat right@s:cym\&eng . //
\glsurface ELE:  right$^{C}_{E}$ .  //
\glauto \%aut  right{\scriptsize .ADJ}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat fel ag yr wyt ti de ? //
\glsurface ALN:  fel ag yr wyt ti de ?  //
\glauto \%aut  like{\scriptsize .CONJ} with{\scriptsize .PREP} that{\scriptsize .CONJ[or]the.DET.DEF[or]drive.V.3S.PRES+SM} be{\scriptsize .V.2S.PRES} you{\scriptsize .PRON.2S} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  like as PRT be{\scriptsize .2S.PRES} PRON{\scriptsize .2S} TAG   //
\gleng as you do, eh? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o\_k@s:cym\&eng be mae isio wneud ym Mharis@s:cym\&eng ? //
\glsurface ELE:  o\_k$^{C}_{E}$ be mae isio wneud ym Mharis$^{C}_{E}$ ?  //
\glauto \%aut  OK{\scriptsize .IM} what{\scriptsize .INT} be{\scriptsize .V.3S.PRES} want{\scriptsize .N.M.SG} make{\scriptsize .V.INFIN+SM} in{\scriptsize .PREP} name   //
\glmanual \%gls  o\_k what be{\scriptsize .3S.PRES} want do{\scriptsize .NONFIN} in Paris   //
\gleng ok, what does she want to do in Paris? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat pwy a ŵyr ? //
\glsurface ALN:  pwy a ŵyr ?  //
\glauto \%aut  who{\scriptsize .PRON} and{\scriptsize .CONJ} grandchild{\scriptsize .N.M.SG}   //
\glmanual \%gls  who PRT know{\scriptsize .3S.NONPAST}   //
\gleng who knows? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (oe)s gynni hi waith $<$actio ar y funud$>$ [?] ? //
\glsurface ELE:  oes gynni hi waith actio ar y funud ?  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.INDEF[or]age.N.F.SG} with\_her{\scriptsize .PREP+PRON.F.3S} she{\scriptsize .PRON.F.3S} time{\scriptsize .N.F.SG+SM[or]work.N.M.SG+SM} act{\scriptsize .V.INFIN} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} minute{\scriptsize .N.M.SG+SM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} with{\scriptsize .3SF} PRON{\scriptsize .3SF} work acting on DET minute   //
\gleng has she got acting work at the moment? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oes oes . //
\glsurface ALN:  oes oes .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.INDEF} be{\scriptsize .V.3S.PRES.INDEF}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES}   //
\gleng yes, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mae (y)n gweithio \# trwy (y)r adeg ar hyn o bryd . //
\glsurface ALN:  mae yn gweithio trwy yr adeg ar hyn o bryd .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES} stative{\scriptsize .STAT} work{\scriptsize .V.INFIN} through{\scriptsize .PREP} the{\scriptsize .DET.DEF} time{\scriptsize .N.F.SG} on{\scriptsize .PREP} this{\scriptsize .PRON.SP} of{\scriptsize .PREP} mind{\scriptsize .N.M.SG[or]when.INT+SM[or]time.N.M.SG+SM}   //
\glmanual \%gls  be{\scriptsize .3S.PRES} PRT work{\scriptsize .NONFIN} through DET time on this of time   //
\gleng she's working all the time at the moment. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond yn lle (fe)lly ? //
\glsurface ELE:  ond yn lle felly ?  //
\glauto \%aut  but{\scriptsize .CONJ} in{\scriptsize .PREP} where{\scriptsize .INT[or]place.N.M.SG} so{\scriptsize .ADV}   //
\glmanual \%gls  but in where thus   //
\gleng right, but where? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat er@s:cym\&eng (e)fo Frân\_Wen@s:cym\&eng ar hyn o bryd mynd round@s:cym\&eng ysgolion . //
\glsurface ALN:  er$^{C}_{E}$ efo Frân\_Wen$^{C}_{E}$ ar hyn o bryd mynd round$^{C}_{E}$ ysgolion .  //
\glauto \%aut  er{\scriptsize .IM} with{\scriptsize .PREP} name on{\scriptsize .PREP} this{\scriptsize .PRON.SP} of{\scriptsize .PREP} mind{\scriptsize .N.M.SG[or]when.INT+SM[or]time.N.M.SG+SM} go{\scriptsize .V.INFIN} round{\scriptsize .ADJ} schools{\scriptsize .N.F.PL}   //
\glmanual \%gls  IM with Frân\_Wen on this of time go{\scriptsize .NONFIN} round schools   //
\gleng with Frân Wen [theatre company] at the moment, going round schools. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o:h@s:cym\&eng do (dd)eudodd hi (he)fyd do . //
\glsurface ELE:  oh$^{C}_{E}$ do ddeudodd hi hefyd do .  //
\glauto \%aut  oh{\scriptsize .IM} come{\scriptsize .V.1S.PRES.SPOKEN[or]yes.ADV.PAST[or]roof.N.M.SG+SM} say{\scriptsize .V.3S.PAST+SM} she{\scriptsize .PRON.F.3S} also{\scriptsize .ADV} yes{\scriptsize .ADV.PAST}   //
\glmanual \%gls  IM yes say{\scriptsize .3S.PAST} PRON{\scriptsize .3SF} also yes   //
\gleng oh, yes, she said so, too, yes //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat so@s:cym\&eng $<$dw i meddwl$>$ [/] dw i meddwl mai bwriad hi (y)dy mae gynni hi gyfnod \#\# +/ . //
\glsurface ALN:  so$^{C}_{E}$ dw i meddwl dw i meddwl mai bwriad hi ydy mae gynni hi gyfnod .  //
\glauto \%aut  so{\scriptsize .ADV} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} think{\scriptsize .V.INFIN} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} think{\scriptsize .V.INFIN} that\_it\_is{\scriptsize .CONJ.FOCUS} intention{\scriptsize .N.M.SG} she{\scriptsize .PRON.F.3S} be{\scriptsize .V.3S.PRES} be{\scriptsize .V.3S.PRES} with\_her{\scriptsize .PREP+PRON.F.3S} she{\scriptsize .PRON.F.3S} period{\scriptsize .N.M.SG+SM}   //
\glmanual \%gls  so be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} think{\scriptsize .NONFIN} be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} think{\scriptsize .NONFIN} PRT intention PRON{\scriptsize .3SF} be{\scriptsize .3S.PRES} be{\scriptsize .3S.PRES} with{\scriptsize .3SF} PRON{\scriptsize .3SF} period   //
\gleng so I think the plan is, she has a period \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ++ diwaith (fe)lly . //
\glsurface ELE:  diwaith felly .  //
\glauto \%aut  unemployed{\scriptsize .ADJ} so{\scriptsize .ADV}   //
\glmanual \%gls  unemployed thus   //
\gleng  \dots without work. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, diwaith yn dod i\_fyny . //
\glsurface ALN:  diwaith yn dod i\_fyny .  //
\glauto \%aut  unemployed{\scriptsize .ADJ} stative{\scriptsize .STAT} come{\scriptsize .V.INFIN} up{\scriptsize .ADV[or]up.ADV}   //
\glmanual \%gls  unemployed PRT come{\scriptsize .NONFIN} up   //
\gleng  \dots without work coming up. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng (y)dy am fynd i Baris@s:cym\&eng adeg hynny ? //
\glsurface ELE:  mmm$^{C}_{E}$ ydy am fynd i Baris$^{C}_{E}$ adeg hynny ?  //
\glauto \%aut  mmm{\scriptsize .IM} be{\scriptsize .V.3S.PRES} for{\scriptsize .PREP} go{\scriptsize .V.INFIN+SM} to{\scriptsize .PREP} name time{\scriptsize .N.F.SG} that{\scriptsize .ADJ.DEM.SP}   //
\glmanual \%gls  IM be{\scriptsize .3S.PRES} for go{\scriptsize .NONFIN} to Paris time that   //
\gleng she's going to go to Paris then ? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i meddwl . //
\glsurface ALN:  dw i meddwl .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} think{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} think{\scriptsize .NONFIN}   //
\gleng I think so. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dw i (dd)im yn gwybod . //
\glsurface ALN:  dw i ddim yn gwybod .  //
\glauto \%aut  be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN}   //
\glmanual \%gls  be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} NEG PRT know{\scriptsize .NONFIN}   //
\gleng I don't know. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat alla i (dd)im cadw i\_fyny (e)fo hi . //
\glsurface ALN:  alla i ddim cadw i\_fyny efo hi .  //
\glauto \%aut  be\_able{\scriptsize .V.3S.PRES.SPOKEN+SM} I{\scriptsize .PRON.1S[or]to.PREP} not{\scriptsize .ADV+SM} keep{\scriptsize .V.INFIN} up{\scriptsize .ADV[or]up.ADV} with{\scriptsize .PREP} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  can{\scriptsize .1S.NONPAST} PRON{\scriptsize .1S} NEG keep{\scriptsize .NONFIN} up with PRON{\scriptsize .3SF}   //
\gleng I can't keep up with her. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat $<$mmm@s:cym\&eng na$>$ [=! quietly] . //
\glsurface ELE:  mmm$^{C}_{E}$ na .  //
\glauto \%aut  mmm{\scriptsize .IM} neg{\scriptsize .PRT}   //
\glmanual \%gls  IM no   //
\gleng no. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ella fasai hi (y)n gwarch(od) +// . //
\glsurface ELE:  ella fasai hi yn gwarchod .  //
\glauto \%aut  maybe{\scriptsize .ADV.SPOKEN} be{\scriptsize .V.3S.PLUPERF.SPOKEN+SM} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} protect{\scriptsize .V.INFIN}   //
\glmanual \%gls  maybe be{\scriptsize .3S.CONDIT} PRON{\scriptsize .3SF} PRT guard{\scriptsize .NONFIN}   //
\gleng maybe she'd look after \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng na $<$Acoustique@s:cym\&eng oedd hwnna de$>$ [?] . //
\glsurface ELE:  oh$^{C}_{E}$ na Acoustique$^{C}_{E}$ oedd hwnna de .  //
\glauto \%aut  oh{\scriptsize .IM} neg{\scriptsize .PRT} name be{\scriptsize .V.3S.IMPERF} that{\scriptsize .PRON.M.SG.SPOKEN} isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  IM no Acoustique be{\scriptsize .3S.IMP} that TAG   //
\gleng oh no, that was Acoustique, wasn't it //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat \&=lateral\_clicks ella fasai hi (y)n gwarchod ar yr wythfed ar hugain sti ? //
\glsurface ELE:  ella fasai hi yn gwarchod ar yr wythfed ar hugain sti ?  //
\glauto \%aut  maybe{\scriptsize .ADV.SPOKEN} be{\scriptsize .V.3S.PLUPERF.SPOKEN+SM} she{\scriptsize .PRON.F.3S} stative{\scriptsize .STAT} protect{\scriptsize .V.INFIN} on{\scriptsize .PREP} the{\scriptsize .DET.DEF} eighth{\scriptsize .ORD} on{\scriptsize .PREP} twenty{\scriptsize .NUM+H} you\_know{\scriptsize .IM}   //
\glmanual \%gls  maybe be{\scriptsize .3S.CONDIT} PRON{\scriptsize .3SF} PRT guard{\scriptsize .NONFIN} on DET eighth on twenty know{\scriptsize .2S}   //
\gleng maybe she'd babysit on the 28th, you know? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be sy (y)n digwydd yr wythfed ar hugain eto ? //
\glsurface ALN:  be sy yn digwydd yr wythfed ar hugain eto ?  //
\glauto \%aut  what{\scriptsize .INT} which\_is{\scriptsize .V.3SP.PRES.REL} stative{\scriptsize .STAT} happen{\scriptsize .V.INFIN} the{\scriptsize .DET.DEF} eighth{\scriptsize .ORD} on{\scriptsize .PREP} twenty{\scriptsize .NUM+H} again{\scriptsize .ADV}   //
\glmanual \%gls  what be{\scriptsize .PRES.REL} PRT happen{\scriptsize .NONFIN} DET eighth on twenty again   //
\gleng what's happening on the 28th, again? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat xx pen\_blwydd [?] . //
\glsurface ELE:  pen\_blwydd .  //
\glauto \%aut  unk   //
\glmanual \%gls  birthday   //
\gleng xx birthday. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng ia . //
\glsurface ALN:  oh$^{C}_{E}$ ia .  //
\glauto \%aut  oh{\scriptsize .IM} yes{\scriptsize .ADV}   //
\glmanual \%gls  IM yes   //
\gleng oh, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng os (y)dy gwmpas . //
\glsurface ALN:  yeah$^{C}_{E}$ os ydy gwmpas .  //
\glauto \%aut  yeah{\scriptsize .ADV} if{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} pair{\scriptsize .of.compasses.N.M.SG+SM[or]scope.N.M.SG+SM}   //
\glmanual \%gls  yeah if be{\scriptsize .3S.PRES} around   //
\gleng yeah, if she's around. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat de [?] . //
\glsurface ELE:  de .  //
\glauto \%aut  isn't\_it{\scriptsize .IM}   //
\glmanual \%gls  TAG   //
\gleng isn't it. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat sure@s:cym\&eng y byddai hi . //
\glsurface ALN:  sure$^{C}_{E}$ y byddai hi .  //
\glauto \%aut  sure{\scriptsize .ADJ} that{\scriptsize .PRON.REL[or]the.DET.DEF} be{\scriptsize .V.3S.COND} she{\scriptsize .PRON.F.3S}   //
\glmanual \%gls  sure PRT be{\scriptsize .3S.CONDIT} PRON{\scriptsize .3SF}   //
\gleng she probably would. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (ba)sen nhw wrth eu boddau . //
\glsurface ELE:  basen nhw wrth eu boddau .  //
\glauto \%aut  be{\scriptsize .V.3P.PLUPERF.SPOKEN} they{\scriptsize .PRON.3P} by{\scriptsize .PREP} their{\scriptsize .ADJ.POSS} pleasure{\scriptsize .N.M.PL}   //
\glmanual \%gls  be{\scriptsize .3PL.CONDIT} PRON{\scriptsize .3PL} by POSS{\scriptsize .3PL} liking   //
\gleng they'd love it //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng . //
\glsurface ALN:  yeah$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat neu mi allen ni dympio +// . //
\glsurface ALN:  neu mi allen ni dympio .  //
\glauto \%aut  or{\scriptsize .CONJ} I{\scriptsize .PRON.1S[or]aff.PRT} be\_able{\scriptsize .V.3P.IMPER.SPOKEN+SM} we{\scriptsize .PRON.1P} dump{\scriptsize .V.INFIN}   //
\glmanual \%gls  or PRT can{\scriptsize .1PL.CONDIT} PRON{\scriptsize .1PL} dump{\scriptsize .NONFIN}   //
\gleng or we could dump \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat allen ni ofyn i +. . . //
\glsurface ALN:  allen ni ofyn i .  //
\glauto \%aut  be\_able{\scriptsize .V.3P.IMPER.SPOKEN+SM} we{\scriptsize .PRON.1P} ask{\scriptsize .V.INFIN+SM} I{\scriptsize .PRON.1S}   //
\glmanual \%gls  can{\scriptsize .1PL.CONDIT} PRON{\scriptsize .1PL} ask{\scriptsize .NONFIN} to   //
\gleng we could ask \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dympio nhw ? //
\glsurface ELE:  dympio nhw ?  //
\glauto \%aut  dump{\scriptsize .V.INFIN} they{\scriptsize .PRON.3P}   //
\glmanual \%gls  dump{\scriptsize .NONFIN} PRON{\scriptsize .3PL}   //
\gleng dump them? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat dim dympio ! //
\glsurface ELE:  dim dympio !  //
\glauto \%aut  not{\scriptsize .ADV} dump{\scriptsize .V.INFIN}   //
\glmanual \%gls  NEG dump{\scriptsize .NONFIN}   //
\gleng not dump! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat yeah@s:cym\&eng yeah@s:cym\&eng doedd [?] [/] oedd [///] o\_k@s:cym\&eng . //
\glsurface ALN:  yeah$^{C}_{E}$ yeah$^{C}_{E}$ doedd oedd o\_k$^{C}_{E}$ .  //
\glauto \%aut  yeah{\scriptsize .ADV} yeah{\scriptsize .ADV} be{\scriptsize .V.3S.IMPERF.NEG} be{\scriptsize .V.3S.IMPERF} OK{\scriptsize .IM}   //
\glmanual \%gls  yeah yeah be{\scriptsize .3S.IMP.NEG} be{\scriptsize .3S.IMP} o\_k   //
\gleng yeah, yeah, ok. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat gair anghywir . //
\glsurface ALN:  gair anghywir .  //
\glauto \%aut  word{\scriptsize .N.M.SG} incorrect{\scriptsize .ADJ}   //
\glmanual \%gls  word incorrect   //
\gleng wrong word. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mi allen ni adael [!!] \#\# y canhwyllau ein llygaid \#\# ar um@s:cym\&eng +. . . //
\glsurface ALN:  mi allen ni adael y canhwyllau ein llygaid ar um$^{C}_{E}$ .  //
\glauto \%aut  I{\scriptsize .PRON.1S[or]aff.PRT} be\_able{\scriptsize .V.3P.IMPER.SPOKEN+SM} we{\scriptsize .PRON.1P} leave{\scriptsize .V.INFIN+SM} the{\scriptsize .DET.DEF} candles{\scriptsize .N.F.PL} our{\scriptsize .ADJ.POSS} eyes{\scriptsize .N.M.PL} on{\scriptsize .PREP} um{\scriptsize .IM}   //
\glmanual \%gls  PRT can{\scriptsize .1PL.CONDIT} PRON{\scriptsize .1PL} leave{\scriptsize .NONFIN} DET candles POSS{\scriptsize .1PL} eyes on IM   //
\gleng we could leave the apples of our eyes with \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat sorry@s:cym\&eng dw i dal yn yr meddylfryd dympio . //
\glsurface ALN:  sorry$^{C}_{E}$ dw i dal yn yr meddylfryd dympio .  //
\glauto \%aut  sorry{\scriptsize .ADJ} be{\scriptsize .V.1S.PRES.SPOKEN} I{\scriptsize .PRON.1S} hold{\scriptsize .V.INFIN} in{\scriptsize .PREP} the{\scriptsize .DET.DEF} state{\scriptsize .of.mind.N.M.SG} dump{\scriptsize .V.INFIN}   //
\glmanual \%gls  sorry be{\scriptsize .1S.PRES} PRON{\scriptsize .1S} continue{\scriptsize .NONFIN} in DET frame\_of\_mind dump{\scriptsize .NONFIN}   //
\gleng sorry, I'm still in the dumping frame of mind //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, dympio nhw ar um@s:cym\&eng \# Adam\_Adams@s:cym\&eng . //
\glsurface ALN:  dympio nhw ar um$^{C}_{E}$ Adam\_Adams$^{C}_{E}$ .  //
\glauto \%aut  dump{\scriptsize .V.INFIN} they{\scriptsize .PRON.3P} on{\scriptsize .PREP} um{\scriptsize .IM} name   //
\glmanual \%gls  dump{\scriptsize .NONFIN} PRON{\scriptsize .3PL} on Adam\_Adams    //
\gleng dump them on Adam Adams. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ \&=laugh . //
\glsurface ELE:  .  //
\glauto \%aut    //
\glmanual \%gls    //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (be)cause@s:eng ddeudodd Mark@s:cym\&eng (ba)sen nhw medru mynd $<$i fan (y)no bysai$>$ [?] ? //
\glsurface ALN:  because$^{E}$ ddeudodd Mark$^{C}_{E}$ basen nhw medru mynd i fan yno bysai ?  //
\glauto \%aut  because{\scriptsize .CONJ} say{\scriptsize .V.3S.PAST+SM} name be{\scriptsize .V.3P.PLUPERF.SPOKEN} they{\scriptsize .PRON.3P} be{\scriptsize .V.INFIN} go{\scriptsize .V.INFIN} to{\scriptsize .PREP} van{\scriptsize .N.F.SG[or]place.N.MF.SG+SM} there{\scriptsize .ADV} finger{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  because say{\scriptsize .3S.PAST} Mark be{\scriptsize .3PL.CONDIT} PRON{\scriptsize .3PL} can{\scriptsize .NONFIN} go{\scriptsize .NONFIN} to place there be{\scriptsize .3S.CONDIT}   //
\gleng cause Mark said they could go there, couldn't they? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat o:h@s:cym\&eng do ! //
\glsurface ELE:  oh$^{C}_{E}$ do !  //
\glauto \%aut  oh{\scriptsize .IM} come{\scriptsize .V.1S.PRES.SPOKEN[or]yes.ADV.PAST[or]roof.N.M.SG+SM}   //
\glmanual \%gls  IM yes   //
\gleng oh, yes. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a mae ei [?] mam@s:cym\&eng (we)di deud +/ . //
\glsurface ALN:  a mae ei mam$^{C}_{E}$ wedi deud .  //
\glauto \%aut  and{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES} her{\scriptsize .ADJ.POSS[or]his.ADJ.POSS} mam{\scriptsize .N.SG} after{\scriptsize .PREP} say{\scriptsize .V.INFIN.SPOKEN}   //
\glmanual \%gls  and be{\scriptsize .3S.PRES} POSS{\scriptsize .3SF} mother PRT{\scriptsize .PAST} say{\scriptsize .NONFIN}   //
\gleng and her Mum said \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +$<$ fasai hi (y)n +// . //
\glsurface ELE:  fasai hi yn .  //
\glauto \%aut  be{\scriptsize .V.3S.PLUPERF.SPOKEN+SM} she{\scriptsize .PRON.F.3S} yn{\scriptsize .PRT}   //
\glmanual \%gls  be{\scriptsize .3S.CONDIT} PRON{\scriptsize .3SF} PRT   //
\gleng she'd \dots  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat huh@s:cym\&eng ? //
\glsurface ELE:  huh$^{C}_{E}$ ?  //
\glauto \%aut  huh{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng what? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat +, +$<$ $<$(ba)sai Nathan@s:cym\&eng medru \&m$>$ [//] (fa)sai Charlie@s:cym\&eng xxx Nathan@s:cym\&eng hefyd bysai ? //
\glsurface ALN:  basai Nathan$^{C}_{E}$ medru fasai Charlie$^{C}_{E}$ Nathan$^{C}_{E}$ hefyd bysai ?  //
\glauto \%aut  be{\scriptsize .V.3S.PLUPERF.SPOKEN} name be{\scriptsize .V.INFIN} be{\scriptsize .V.3S.PLUPERF.SPOKEN+SM} name name also{\scriptsize .ADV} finger{\scriptsize .V.3S.IMPERF}   //
\glmanual \%gls  be{\scriptsize .3S.CONDIT} Nathan can{\scriptsize .NONFIN} be{\scriptsize .3S.CONDIT} Charlie Nathan also be{\scriptsize .3S.CONDIT}   //
\gleng Nathan could \dots Charlie could [ \dots ] Nathan, couldn't he? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat be am Debbie@s:cym\&eng ? //
\glsurface ELE:  be am Debbie$^{C}_{E}$ ?  //
\glauto \%aut  what{\scriptsize .INT} for{\scriptsize .PREP} name   //
\glmanual \%gls  what for Debbie   //
\gleng what about Debbie? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat oh@s:cym\&eng (ba)sen ni (y)n ffeindio rywun . //
\glsurface ALN:  oh$^{C}_{E}$ basen ni yn ffeindio rywun .  //
\glauto \%aut  oh{\scriptsize .IM} be{\scriptsize .V.3P.PLUPERF.SPOKEN} we{\scriptsize .PRON.1P} stative{\scriptsize .STAT} find{\scriptsize .V.INFIN} someone{\scriptsize .N.M.SG+SM}   //
\glmanual \%gls  IM be{\scriptsize .1PL.CONDIT} PRON{\scriptsize .1PL} PRT find{\scriptsize .NONFIN} someone   //
\gleng oh, we'd find someone. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat mmm@s:cym\&eng . //
\glsurface ELE:  mmm$^{C}_{E}$ .  //
\glauto \%aut  mmm{\scriptsize .IM}   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat Pauline@s:cym\&eng . //
\glsurface ALN:  Pauline$^{C}_{E}$ .  //
\glauto \%aut  name   //
\glmanual \%gls     //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ie ella . //
\glsurface ELE:  ie ella .  //
\glauto \%aut  yes{\scriptsize .ADV} maybe{\scriptsize .ADV.SPOKEN}   //
\glmanual \%gls  yes maybe   //
\gleng yes, maybe. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat neu [///] ie \# neu Rachel@s:cym\&eng \# neu \# Sandra@s:cym\&eng . //
\glsurface ELE:  neu ie neu Rachel$^{C}_{E}$ neu Sandra$^{C}_{E}$ .  //
\glauto \%aut  or{\scriptsize .CONJ} yes{\scriptsize .ADV} or{\scriptsize .CONJ} name or{\scriptsize .CONJ} name   //
\glmanual \%gls  or yes or Rachel or Sandra   //
\gleng or. . . yes, or Rachel or Sandra. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ond [?] fydda i (ddi)m yma na (fy)dda ? //
\glsurface ELE:  ond fydda i ddim yma na fydda ?  //
\glauto \%aut  but{\scriptsize .CONJ} be{\scriptsize .V.1S.FUT+SM} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} here{\scriptsize .ADV} neg{\scriptsize .PRT} be{\scriptsize .V.1S.FUT+SM}   //
\glmanual \%gls  but be{\scriptsize .1S.FUT} PRON{\scriptsize .1S} NEG here NEG be{\scriptsize .1S.FUT}   //
\gleng but I won't be here, will I ? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na so@s:cym\&eng (dy)dy o (ddi)m broblem@s:cym\&eng chdi na(g) (y)dy ? //
\glsurface ALN:  na so$^{C}_{E}$ dydy o ddim broblem$^{C}_{E}$ chdi nag ydy ?  //
\glauto \%aut  neg{\scriptsize .PRT} so{\scriptsize .ADV} be{\scriptsize .V.3S.PRES.NEG.SPOKEN[or]be.V.3S.PRES.NEG+SM} he{\scriptsize .PRON.M.3S.SPOKEN} not{\scriptsize .ADV+SM} problem{\scriptsize .N.SG+SM} you{\scriptsize .PRON} than{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES}   //
\glmanual \%gls  no so be{\scriptsize .3S.PRES.NEG} PRON{\scriptsize .3SM} NEG problem PRON{\scriptsize .2S} NEG be{\scriptsize .3S.PRES}   //
\gleng no, so it's not a problem for you, is it? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat na(g) (y)dy . //
\glsurface ELE:  nag ydy .  //
\glauto \%aut  than{\scriptsize .CONJ} be{\scriptsize .V.3S.PRES}   //
\glmanual \%gls  NEG be{\scriptsize .3S.PRES}   //
\gleng no. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ha@s:cym\&eng ha@s:cym\&eng ideal@s:cym\&eng ! //
\glsurface ALN:  ha$^{C}_{E}$ ha$^{C}_{E}$ ideal$^{C}_{E}$ !  //
\glauto \%aut  ha{\scriptsize .IM} ha{\scriptsize .IM} ideal{\scriptsize .N.SG}   //
\glmanual \%gls  IM IM ideal   //
\gleng ha-ha, ideal! //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat a Andrew@s:cym\&eng [?] . //
\glsurface ELE:  a Andrew$^{C}_{E}$ .  //
\glauto \%aut  and{\scriptsize .CONJ} name   //
\glmanual \%gls      //
\gleng  //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat lle fasai hwn yn digwydd ? //
\glsurface ELE:  lle fasai hwn yn digwydd ?  //
\glauto \%aut  where{\scriptsize .INT} be{\scriptsize .V.3S.PLUPERF.SPOKEN+SM} this{\scriptsize .PRON.M.SG} stative{\scriptsize .STAT} happen{\scriptsize .V.INFIN}   //
\glmanual \%gls  where be{\scriptsize .3S.CONDIT} this PRT happen{\scriptsize .NONFIN}   //
\gleng where would this be happening? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat ym Mangor@s:cym\&eng ? //
\glsurface ELE:  ym Mangor$^{C}_{E}$ ?  //
\glauto \%aut  in{\scriptsize .PREP} name   //
\glmanual \%gls  in Bangor   //
\gleng in Bangor? //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat (doe)s gen i (dd)im syniad . //
\glsurface ALN:  does gen i ddim syniad .  //
\glauto \%aut  be{\scriptsize .V.3S.PRES.INDEF.NEG} by{\scriptsize .PREP.SPOKEN} I{\scriptsize .PRON.1S} not{\scriptsize .ADV+SM} idea{\scriptsize .N.M.SG}   //
\glmanual \%gls  be{\scriptsize .3S.PRES.NEG} with{\scriptsize .1S} PRON{\scriptsize .1S} NEG idea   //
\gleng I've no idea. //
\endgl
\xe

\ex
\begingl[lingstyle=gergl]
\glchat right@s:cym\&eng dw (ddi)m yn gwybod digon am y pethau (y)ma yli . //
\glsurface ELE:  right$^{C}_{E}$ dw ddim yn gwybod digon am y pethau yma yli .  //
\glauto \%aut  right{\scriptsize .ADJ} be{\scriptsize .V.1S.PRES.SPOKEN} not{\scriptsize .ADV+SM} stative{\scriptsize .STAT} know{\scriptsize .V.INFIN} enough{\scriptsize .QUAN} for{\scriptsize .PREP} the{\scriptsize .DET.DEF} things{\scriptsize .N.M.PL} here{\scriptsize .ADV} you\_know{\scriptsize .IM}   //
\glmanual \%gls  right be{\scriptsize .1S.PRES} NEG PRT know{\scriptsize .NONFIN} enough about DET things here see{\scriptsize .2S.IMPER}   //
\gleng right, I don't know enough about this, see. //
\endgl
\xe


\vfill

\begin{flushright}
Typeset using \LaTeX\ (\textit{latex-project.org}) on \longdate\today
\end{flushright}

\end{document}