
\documentclass[12pt]{article}
\usepackage{expex,epltxfn,tipa}

\begin{document}

\lingset{aboveexskip=0pt,glspace=1em plus .5em minus.2em,
   glrightskip=0pt plus 4em,aboveglftskip=0pt,glneveryline={\sl},
   glstyle=nlevel,everygl={\hangindent=1.2em \hangafter=1}}
% \sl used at top level rather than \it, because there is no italic
% family in tipa
\def\@{\textipa{@}}

\hsize=5.2in
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent {\it Illustrative running glossed text}
\bigskip

\noindent The text is a story in the Erromangan language, found in Chapter 9 of
Terry Crowley's `An Erromangan (Sye) Grammar'.

\exdisplay[aboveexskip=2ex]
\allocglfns{A}
\begingl
Yoco-nompi[\textsc{1sg:fut-mr:}do]
stori\glfnmark{A}[story]
gi[\sc obl]
lakih.[rat]
\glft I will tell a story about the rat.
\endgl
\glfntext{Corrected editorially to:
\exdisplay[aboveexskip=1ex,belowexskip=0pt]
\begingl
Yoco-nompi[\textsc{1sg:fut-mr:}do]
uvuvu[story]
\glft I will tell a story
\endgl\xe
}
\xe

\exdisplay
\begingl
Lakih[rat]
-etti[\textsc{3sg:distpast-br:}give.birth]
ra[\sc loc]
{novkilyenau $(<$novkilye-n}[dried.leaf-\sc const]
nau$)$.[bamboo]
\glft The rat gave birth in the dried leaves of the bamboo.
\endgl
\xe

\exdisplay\begingl
M-ante[\textsc{sg:es-mr:hab}]
{mandgi $(<$ m-and\textipa{@}g-i$\mskip2mu)$}[\textsc{sg:es-mr:}hear-\textsc{const}]
nam[talk]
orog-orog[big-\sc redup]
u-ntemne[\textsc{loc}-village]
\glft And she would always hear loud talk in the village.
\endgl
\xe

\exdisplay
\begingl
Ovoteme[\textsc{pl}:person]
rum-ante[\textsc{3pl:pasthab-mr:}stat]
u-ntemne[\textsc{loc}-village]
rum-nam[\textsc{3pl:pasthab-mr:}talk]
orog-orog.[big-\textsc{redup}]
\glft The people in the village would speak loudly.
\endgl
\xe

\exdisplay
\allocglfns{A}
\begingl
Yandyoc\glfnmark{A}[\textsc{3sg:pasthab-mr:}pick.up]
nitni[child:\textsc{3sg}]
m-ampai[\textsc{sg:es-mr:}take]
me-hac[\textsc{sg:es-}ascend]
m-aglu:[\textsc{sg:es-mr:}say]
\glft She would pick up her child and take them up and say.
\endgl
\glfntext{%
The sequence {\it yandyoc nitni\/} would appear in more general
speech as:
\exdisplay[aboveexskip=1ex,belowexskip=0pt]
\begingl Yem-andyok-i[\textsc{3sg:pasthab-mr:}pick.up-\textsc{const}]
nitni[child:\textsc{3sg}]
\glft She would pick up her child.
\endgl
\xe}
\xe

\exdisplay
\def\schwa{{\textipa @}}%
\def\({{\rm )}}%
\allocglfns{A}
\begingl
Is,[hey]
{ucohpe $(<$u-oc\@h-pe$\mskip1mu )$\glfnmark{A}}%
   [\textsc{2pl:imp-br:}look\textsc{-prec}]
nitu-g[child\textsc{-1sg}]
\glft Hey, you all look at my child first!
\endgl
\glfntext{Here again, the construct suffix that is present
with younger speakers is absent.  This results in the
underlying schwa being realized as {\it o} rather than zero.
The sequence {\it uchope nitug\/} would appear in the speech
of most people as:
\exdisplay\begingl
uchipe\thinspace(u-oc\@h-i-per)%
   [\textsc{2pl:imp-br:}look\textsc{-const-prec}]
nitu-g[child\textsc{-1sg}]
\endgl\xe
}

\xe

\end{document}

